### 2023-06-05 docs/ms-draft.Rmd
added files related to the ms document

---
title: "Probabilistic bioregion models suggest seabird assemblages are linked to the main, fast-changing oceanic current off eastern Australia"
# "Probabilistic bioregion models suggest seabird assemblages are linked to large scale oceanography off eastern Australia"
# author: Nicholas W Daudt
# date: '`r format(Sys.Date())`'
author:
  - Nicholas W. Daudt:
      email: nicholaswdaudt@gmail.com
      institute: [MarSciUO]
      correspondence: true
  - Eric J. Woehler:
      institute: [BirdLife, IMAS]
      correspondence: false
  - Matthew R. Schofield:
      institute: [MathStatUO]
      correspondence: false
  - Robert O. Smith:
      institute: [MarSciUO]
      correspondence: false
  - Leandro Bugoni:
      institute: [FURG]
      correspondence: false
  - William J. Rayment:
      institute: [MarSciUO]
      correspondence: false
institute:
  - MarSciUO: Department of Marine Science, University of Otago, Aotearoa New Zealand
  - BirdLife: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
  - IMAS: Institute Marine and Antarctic Studies, University of Tasmania, Hobart, Australia
  - MathStatUO: Department of Mathematics and Statistics, University of Otago, Aotearoa New Zealand
  - FURG: Institute of Biological Sciences, Universidade Federal do Rio Grande--FURG, Rio Grande, RS, Brazil
abstract: |
  **Aim**: To identify assemblages of seabird species at large scale, probabilistically, and related them to environmental conditions off eastern Australia, a world seabird hotspot.
  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and counts). RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within it but different from other regions. These groups also vary according to covariates, on which we included oceanographic static and climatic, as well as physiographic layers.
  **Results**: All models suggested two large scale assemblages ('northern' and 'southern' assemblages), except for the abundance-autumn model which indicated three groups. Model results seem to agree in a biogeographic barrier at ~32°S, where the East Australian Current mostly dissipates to the east and loses force. Furthermore, sea surface temperature and, in one case, sea surface salinity, were selected in all final models, further indicating close relationship between seabird assemblages and water masses. Results from both data types, presence-absence and counts resulted in similar spatial and species profile patterns.
  **Conclusion**: Two seabird assemblages were clearly identified from RCP models, suggesting the persistence of these groups at temporal and spatial scales. Given the rapid change that the East Australian Current is experiencing, and its importance on driving seabirds' distributions, the methods applied in our study could be replicated in future analyses to assess possible spatial/temporal changes in seabird assemblages, including on its structure.
output: 
  # html_document: default
  # word_document:
  #   reference_docx: "../docs/docx-template.docx"
  # officedown::rdocx_document:
  #   reference_docx: "../docs/docx-template.docx"
  # rticles::sim_article:
  #   keep_tex: TRUE
  bookdown::word_document2:
    pandoc_args:
      - --lua-filter=../docs/scholarly-metadata.lua
      - --lua-filter=../docs/author-info-blocks.lua
      - --reference-doc=../docs/docx-template.docx 
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

<!-- YAML for bookdown+lua edited following: -->

<!-- https://www.r-bloggers.com/2019/08/rmarkdown-template-that-manages-academic-affiliations-docx-or-pdf-output/ -->

<!-- 'rticles::sim_article' example? -->

<!-- https://github.com/emitanaka/combine2019/tree/master/output/articles/article_template_sim_article -->

```{r setup, include = FALSE, echo = FALSE}
library(gt)
# library(officedown)
# library(officer)
```

\pagebreak

# Introduction

Bioregionalization is the discipline that identifies, delimits, characterises, and names biogeographical regions [@vilhena2015; @morrone2018]. Although there is no agreement on the naming systems between proposed regionalizations [but see @ebach2008], the hierarchical nature of bioregions is well recognized [@vilhena2015; @kreft2010]. Thus, bioregions are usually described within geographically larger classifications, such as 'provinces' or 'realms' [@spalding2007; @costello2017]. The delimitation of bioregions is made through identifying sets of species that are endemic or with high levels of co-occurrence, distinct enough from other regions, which is interpreted as a natural representation of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010]. In particular, the identification of bioregions allows the representation of the distribution of multiple species at a given time and space as well as to relate these regions to environmental characteristics. By doing so, bioregions can be interpreted as species assemblages [@lyons2017; @hill2017]. Given these features, bioregions have a fundamental role in conservation biogeography [@whittaker2005], by being used as management units and even shedding light into conservation gaps [@spalding2007].

Seabird assemblages have been described and related to environmental characteristics worldwide. Considering the hierarchical nature of oceanographic processes and its resulting features (**Haury78**), it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987]. In the marine environment, large scale oceanographic features (water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e. the occurrence of species and its relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014]. On the other hand, at smaller scales, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014]. Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013]. Noticeably, sea surface temperature is a key driver of distribution of many seabird species, at individual- and assemblage-levels, as well as at large (water masses) and smaller (fronts) scales [@dunlop1988; @ribic1997; @quillfeldt2015]. Nonetheless, many regions still lack information regarding multiple seabird species distribution and how they assemble in space and time.

One such area that could benefit from better understanding of seabird distributions is eastern Australia, considered a world seabird hotspot [@karpouzi2007; @ramírez2017]. Surprisingly, however, there are not many studies on seabird distribution relative to the region's relevance for this group [@mott2018; @bernard2021]. Off northwest Australia, at-sea surveys, multivariate analyses [@dunlop1988] and ensemble species distribution models [@lavers2014] of species assemblages have shown that preferential habitat use of tropical species are related to sea surface temperature and salinity, as well as indicated the importance of the continental shelf for some species. In particular, @dunlop1988 suggests that salinity is the main variable influencing species distribution in oceanic waters whereas temperature would be the key factor on the continental shelf. On the other hand, off eastern Australia, there is a scarcity of studies at larger scales. @reid2000 and @evans2021a analysed seabird assemblages and their relationship with oceanography and prey data, however at coarse scale (tens of km), off southeast Tasmania. At a multi-species, assemblage-level, the latter studies seem to be the only ones on eastern Australia; other studies were mostly descriptive species accounts at local scale [@norman1992distribution; @blaber1994distribution], or focused on a single species [@priddel2014].

The East Australian Current (EAC) is the dominant current in eastern Australia, the major western boundary current in the South Pacific [@ganachaud2014]. The EAC has been extensively studied by physical oceanographers, although its link between oceanography and biological processes is still in its infancy. The dynamic features provided by EAC, such as eddies and upwellings, can likely influence marine biota [@suthers2011; @scales2014]. While most of the biological studies done within this system were focused on plankton entrapment within the cores of eddies [@garcia2022; @firme2023; @hassler2011], there is a growing body of work focusing on fisheries [@hobday2010; @young2011; @brieva2015]. These studies emphasize how the assemblage structure and probability of occurrence of a given taxa is driven by EAC. In particular, the potential biological consequence triggered by climate changes is the tropicalization of the biota, given the EAC is strengthening and getting warmer [@cai2005; @poloczanska2007; @hobday2010; @phillips2022], as shown elsewhere [@hastings2020]. Notwithstanding, the relationship between oceanography and marine megafauna along the EAC is poorly known [@suthers2011], albeit they are considered environmental sentinels [@hazen2019]. Some progress has been made using shark and ray movement data [@jaine2014; @niella2020; @niella2022], and satellite tracking have also demonstrated that increased sea surface temperature in EAC decreases penguin foraging success [@carroll2016]. However, information at the assemblage-level is still scarce [but see @evans2021a].

Given east Australia's importance for seabirds and the recent attention on how climate changes in the EAC can affect marine biota [@poloczanska2007], an assessment of seabird spatial distribution at large scale is timely. Seabirds can serve as sentinels for tracking EAC ecological changes in many facets [e.g. @sydeman2012], including by their distribution. Surprisingly, though, seabird distribution is under-represented in climate change studies, despite being pointed as a priority topic [@grémillet2009; @sydeman2012]. In this study, we aim to identify assemblages of seabird species probabilistically and related them to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique. We used the mixture-model Region of Common Profiles [RCP; @foster2013], which is under the best current available tools for describing bioregions [@hill2020; @woolley2020]. RCP is model-based approach, which include a series of advantages over traditional algorithm-based statistics [@warton2015; @hill2020]. East Australia supports a high diversity of seabirds and offers wide environmental gradients, from the tropics to the subantarctics. Therefore, we expect to find different assemblages relating to different environmental conditions. In addition, we provide a repeatable framework on which future studies can build up to assess how EAC changes will influence seabird assemblage distributions.

# Methods

## Study area

The EAC originates from the South Equatorial Current (SEC), a tropical, nutrient-poor water mass. When the SEC's North Caledonian Jet hits the Queensland Plateau (\~18°S), it splits into a north branch running towards the Gulf of Papua and a south branch, which marks the start of the EAC [@kessler2013; @ganachaud2014]. The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig. 1a**; @kessler2013; @ganachaud2014). The EAC interacts with topographical features on the seafloor, from which oceanographic features and processes arise, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019]. The EAC is mainly attached to the shelf break, but at \~32°S, it spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; typically mentioned as the 'Tasman Front'). Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig. 1a**), possibly carrying eddy-like features as a transient current [@oke2019].

## Seabird and environmental data sources

Seabird data were gathered aboard RV Investigator and RV Falkor during 15 voyages, between 2016--2021, over \~37° of latitude (**Fig. 1b**; **Table SX**). One observer continuously recorded seabirds within a 300-m forward quadrant (eye height **18-m, Eric?** a.s.l) throughout daylight hours while the ship was underway at a constant speed. Species, the number of birds and behaviour were logged directly onto a data entry system through a ship-based web server. Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species) **(Eric: is there any ref for the protocol?)**.

```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records by season on top of the 1° latitude × 1° longitude grid cells (b). In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue color. The EAC system originates from the South Equatorial Current (SEC) and its branches, North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), drawn in red. The EAC 'eastern extension' is an eddy field (blue shaded area) often mentioned as the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et al. (2019); for colour references see the online version.", fig.align = "center", echo = FALSE}

knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
```

To explain and predict species distribution and biogeography, a thoughtful choice of covariates is important [@barbet-massin2014]. We used seven oceanographic variables and calculated three climatic layers from them, and three physiographical variables (**Table 1**). These variables are the most used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of scales from large-scale water-mass properties (e.g. sea surface temperature) to increased predictability of prey availability in mesoscale oceanographic features (e.g. fronts and eddies; @weimerskirch2007; @scales2014). Seabird records were matched with the best open-access, remote-sensed or interpolated data available, and distance from the coast was calculated in R software (see details below).

```{r table1, echo = FALSE}

data.frame(
  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
  Abbreviation = c("sst", "sstgrad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke", "clim_ekesd", "clim_sstgrad", "bat", "slope", "dist_coast"),
  Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "cf", "cf", "cf", "0.25° × 0.25°", "cf", "cf", "NA"),
  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
  Source = c("Huang et al. (2021)", "Derived from Huang et al. (2021)", "E.U. Copernicus Marine Service Information", "E.U. Copernicus Marine Service Information", "ERDDAP/NOAA", "Derived from ERDDAP/NOAA", "IMOS/AODN", "Derived from IMOS/AODN", "Derived from IMOS/AODN", "Derived from Huang et al. (2021)", "GEBCO", "Derived from GEBCO", "Calculated in R (see main text)")
) |>
  gt::gt() |>
  ## SST & SSTgrad
  gt::tab_footnote(footnote = 
                     gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
                   locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
  ## SSS & MLD
  gt::tab_footnote(footnote = 
                     gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
                   locations = gt::cells_body(columns = Source, rows = 3:4)) |>
  ## CHL8 & log10-CHL8
  gt::tab_footnote(footnote = 
                     gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
                   locations = gt::cells_body(columns = Source, rows = 5:6)) |>
  ## EKE and derivations
  gt::tab_footnote(footnote = 
                     gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
                   locations = gt::cells_body(columns = Source, rows = 7:9)) |>
  ## GEBCO & slope
  gt::tab_footnote(footnote = 
                     gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
                   locations = gt::cells_body(columns = Source, rows = 11:12)) |>
  ## Improve cols names
  gt::cols_label(
    Variable = gt::md("**Variable**"),
    Abbreviation = gt::md("**Abbreviation**"),
    Units = gt::md("**Units**"),
    `Spatial.resolution` = gt::md("**Spatial resolution**"), 
    `Temporal.resolution` = gt::md("**Temporal resolution**"),
    Source = gt::md("**Source**")) |>
  ## Table title
  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources"))

```

## Data treatment

Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022], using packages **X, Y, ..., Z**. Code and (gridded) data are available in **Daudt et al. 2023 (Zenodo repo)**.

Before data extraction, a few environmental layers needed to be calculated. Spatial and temporal persistence of features such as eddy kinetic energy (EKE) and thermal fronts (SSTgrad) may influence seabird movement and distribution by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a]. Therefore, for EKE and SSTgrad, we calculated climatic, seasonal average layers based on 10-year data (2011--2021), as well as EKE standard deviation (EKEsd) over the same period. EKE was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ (**ROB: Ref here?**). Sea surface temperature gradient (SSTgrad) was calculated as the spatial gradient by identifying the maximum value for each neighbour cell and averaging them (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011]. EKE and SSTgrad calculations were done layer-by-layer to match seabird data on the finest scale possible. The slope was calculated using `raster::terrain()`, and distance from the coast was calculated through `sf::st_distance()` using Australia and Tasmania as mainland (ignoring smaller islands). The most proximate value for each seabird record was extracted from each environmental layer using built-in functions from `raster`, `terra`, and `sf` packages.

Seabirds are highly mobile animals, and many species migrate during non-breeding periods [@biology2002]. Therefore, seasonal differences in species composition were expected [e.g. @reid2023], as well as in their probability of occurrence and total numbers. Given that, we aggregated data by 1° latitude $\times$ 1° longitude grid cells (c. 110 $\times$ 110 km) and season (**Fig. 1b**), by summing seabirds from the same species. Birds following the boat were previously excluded from analyses. Considering our aim to test if presence-absence (P/A) and abundance (count) data would result in similar assemblage patterns, we built two species matrices, where the P/A matrix transformed values \>1 to 1. In addition, to avoid spurious correlations with assemblage/environment and numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12; autumn = 21; winter = 20; spring = 33). During data aggregation, the environmental data were averaged. Before modelling, environmental data were scaled and centred.

## Statistical modelling

We used the mixture model Region of Common Profiles (RCP) to identify and describe seabird assemblages. RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates biotic (multi-species) and abiotic (environmental) data [@woolley2020; @hill2020]. The model defines 'regions' of similar environmental spaces in which the probability of observing a set of species 'profiles' is constant but different from other regions [@foster2013; @foster2017]. As such, RCPs can be directly interpreted as assemblages, and its species profiles as the assemblage structure. RCP memberships are latent (unobserved) and vary as a function of environmental data, assuming sites within each RCP have constant species profiles. Given a set of environmental variables, the model also enables predicting assemblage probabilities onto sites where no biotic data exist. Additionally, as a feature of model-based approaches, RCPs allow the opportunity to compare models through information criteria, present uncertainty around the results, and check for model assumptions through checking distribution of residuals [@warton2015; @hill2020].

Before fitting models, covariates were checked for collinearity within each season using the modelling dataset. We removed variables $\left|x\right| >0.7$ [@dormann2013], favouring direct measures over the 'climatic' layers we calculated (e.g. if SSTgrad and 'climatic' SSTgrad were correlated, we chose SSTgrad). Thus, 'full models' differed between seasons as the covariates kept vary for each season, although most covariates were common.

Due to seabird migration, seasons could act as 'sampling artifacts' [@foster2017], biasing results [@ponti2023]. Given the difference in species between seasons that met our occurrence threshold, we decided to run RCP models separately by season and data type (P/A, count). RCPs were fitted using the `ecomix` package. All models follow the workflow: (i) finding the best number of groups that describes the data, (ii) re-fit the model by fixing the best number of groups and going through covariates selection, (iii) checking residuals, (iv) bootstrapping the best model to get uncertainty around parameter estimates, and (v) producing results. For steps (i) and (ii), we used Bayesian Information Criterion (BIC) as the metric for model selection, and diagnostics (iii) were assessed visually.

To identify the best number of groups that describe the data, we ran full models (all covariates) with 100 random start numbers [see @foster2013], between 2--7 groups, and chose the number of RCPs that minimized BIC. Then, to assess the best set of environmental covariates, we fixed the number of groups and fitted models for every possible linear, additive combination of covariates specified in the full model. We chose the best model as the one with the lowest BIC. Randomized quantile residual plots modified for mixture models were used to verify model assumptions [@dunstan2013]. To obtain estimates of uncertainty for model parameters (confidence intervals), we used the default settings of `ecomix::regional_mix.bootstrap()`, which applies Bayesian bootstrap 1000 times. A custom code was written to generate partial plots for exploring the influence of covariates on RCP membership.

We built a seasonal 10-year (2011--2021) average layer for predicting model results for each environmental covariate. These variables were then averaged within 1° latitude $\times$ 1° longitude grid cells, by season, between 9--47°S and 141--161°E. Worth noticing that climatic EKEmean and SSTgrad were processed beforehand as they may provide persistent oceanographic features (described above) and thus represent the same value as EKE and SSTgrad. Data for prediction were scaled and centred using the same coefficients from the modelling dataset.

Additionally, to investigate if the amount of information used to run the RCP models was representative, we ran rarefaction curves based on the P/A matrix and the assigned RCP P/A group for each sample (grid). Rarefaction curves were fit using the `iNEXT` package [@chao2014; @hsieh2016], and results are presented based on alpha diversity (i.e. species richness).

<!-- summer \~ bat+slope+eke+sst+sss+mld+sst_grad -->

<!-- autumn \~ bat+slope+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad -->

<!-- winter \~ bat+slope+eke+sst+sss+mld+sst_grad+clim_sst_grad -->

<!-- spring \~ bat+slope+dist_coast+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad+clim_eke_mean -->

# Results

## Summary of species richness and numbers

A total of 10,261 seabird records were obtained. They summed to 142,646 individuals from 80 species, belonging to 11 families and five orders. The five most abundant species summed 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*), whereas the five most frequent species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig. S1 - FO/nOCC/FN**). An increase in species richness towards the south was clear in all but the winter season (**Fig S2. - spp richness/total abund**). Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig. S2**). However, one cell in the Coral Sea summed to numbers of the same order of magnitude as off Tasmania in summer.

## RCP models

All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3. - multifit plots**). The best fitting models generally had **1--3** covariates selected, although the abundance-autumn model had five covariates (**Table 2 - full-model/selected model**). For the P/A-autumn model, we have deliberately chosen the 'second' best model based on parsimony---the top model had six covariates but was \<1 BIC unit lower than the second model with only two covariates. All the best models selected SST or SSS, variables essential for characterising water masses. The best models show a good fit with no apparent deviances in assumptions (**Fig. S4 - residuals**).

Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell. When predicting RCP onto space, patterns of point-predictions suggest a biogeographic change around \~34°S, with both P/A and abundance models pointing to similar patterns (**Fig. 2**). These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig. 3**, **Fig. S5**), showing that around \~34°S the 'northern' RCP tends to have higher probabilities in coastal regions, whereas the 'southern' RCP spreads northward offshore. The P/A-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig. 2**, **Fig. S5**).

```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions for seabird RCP models, by season, based on presence/absence data (top row) and abundance data (bottom row) off eastern Australia. Each cell was assigned to the RCP group with highest probability (i.e. hard-classes).", fig.align = "center", echo = FALSE}

knitr::include_graphics("../results/Fig2-point-predictions.png")
```

```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probabilities of RCP membership in in spring presence/absence (a) and abundance (b) models.", fig.align = "center", echo = FALSE}

knitr::include_graphics("../results/Fig3-probability-maps-spring.png")
```

*Relationship with environmental covariates (highlight SST) --**waiting for 'partial plots' (Fig. 4, Fig. S6)***

The broad pattern of two large-scale assemblages is reflected in their species composition. Although there are overlaps between groups, both P/A and abundance models captured typical tropical species in the northern RCP (e.g. boobies *Sula* spp., lesser frigatebird *Fregata ariel*) and typical Southern Ocean species in the southern RCP (e.g. mollymawks *Thalassarche* spp., fairy prion *Pachyptila turtur*) with higher occurrence/number probabilities in their usual habitat (**Fig. 5**, **Fig. S7**). The P/A-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs. However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross *Thalassarche cauta* and Australasian gannet *Morus serrator* standing out with high probabilities of occurrence.

# Discussion

This study shows that seabird assemblages off eastern Australia relate to oceanography and how the East Australian Current (EAC) may be playing a key role in structuring seabird distribution at a large scale. Our results are based on a cutting-edge multivariate statistical model [@hill2020], which brings advantages over classical methods for describing species assemblages. For instance, it includes uncertainty around model parameters that allow representing the fuzziness between biogeographical boundaries in a probabilistic framework, which can be monitored given the noticeable anomalies in EAC due to climate change [@phillips2022; @morrone2023biogeographical]. Our results draw an important baseline for seabird assemblages on a world hotspot for this group. **Furthermore, data and code are available to allow reproducibility and future applications to verify spatial and temporal persistence of the described assemblages.**

## Limitations {#limitations}

Although the results suggest consistent patterns, we should point out some caveats of our analyses. Seabird data were gathered opportunistically while taking part on several projects, whose sampling design were not planned to account for seabird spatial ecology specifically. As a result, there are spatial discontinuities on the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors more robustly. Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution to date [@mott2018; @bernard2021]. Dedicated seabird effort could thus benefit future studies advancing seabird research in this region, as well as improving patterns described here. By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at a finer scale. In addition, ideally, we could have had more information within each grid cell by season (our sampling units). Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8 - rarefaction curves**).

## Seabird bioregions off eastern Australia

Spatial patterns clearly showed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia. There is a noticeable link with EAC dynamics, although the weak statistical relationship between RCPs and environmental covariates---RCPs boundaries are closely related with where the EAC turns into its 'eastern extension' and the eddy field spawns [@oke2019]. However, the 'northern/southern' assemblage pattern does not agree with popular, large scale regionalization proposals, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007] province's and ecoregion's, and neither with the Large Marine Ecosystems [LME; @sherman1994]. The MEOW scheme was designed using coastal and shelf biota and it was meant to represent habitats up to the 200-m isobath, which clearly is not representative enough of seabirds' environment. The LME scheme is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management areas and thus a 'political' component on it, which seabirds do not seem to agree with. @longhurst2007 designated only one biogeochemical province for east Australia based on productivity patterns, which may work well at larger scales but seems oversimplified specifically for this region.

The described pattern, on the other hand, agree with the marine biogeographic units proposed by @costello2017, who points out a boundary at \~34°S between their 'Coral Sea' and 'South Australia' realms. The latter regionalization used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus encompassing more complexity of the biogeographic boundaries in the marine environment. Other biological data also suggests 'northern/southern' assemblages off eastern Australia, including tuna and billfish [also highly mobile top predators; @reygondeau2012], mesopelagic fauna [which includes seabird prey; @sutton2017a] and deep-sea invertebrates [@ohara2020] and fish [@ohara2020a]. All these studies focused mainly on describing biological assemblages, whereas the ones mentioned previously focused on the environment. Worth mentioning, the Longhurst' provinces had gone through an update including SST, SSS, bathymetry (the same variables kept by our selected best RCP models) and a temporal component to its equation [@reygondeau2013]. The updated provinces are easily comparable to the seabird assemblages described by us, although still not a perfect match. Australia's bioregionalization has a long story of disagreements [@ebach2012history], however, the inherently link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.

The relationship between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997]. Therefore, it is not surprising that at the large scale these two assemblages emerged off eastern Australia. However, the spatial consistency of these patterns throughout seasons is remarkable. Resident species seem to characterise these assemblages. The shy albatross [@hedd2001; @mason2018] and the fairy prions [@fromant2022] in the southern assemblage, and boobies [@bunce2015; @miller2018] and the tropical wedge-tailed shearwater *Ardenna pacifica* [@weimerskirch2020; @mcduie2015] in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species. Given that assemblage changes over time are driven by few species [@gotelli2022] and that those species are likely residents [@souza2023], the relationship between resident seabirds and the EAC system should be closely monitored. For instance, coastal seabirds such as the little penguin *Eudyptula minor* are signalling the effects climate change might have in changing seabird behaviour and decreasing foraging success [@carroll2016], which ultimately influence their breeding success and population dynamics [@johnson2021]. Therefore, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring seabird responses to the fast-changing EAC.

The seasonal models also captured the influence of migratory species, particularly in the southern assemblage. For instance, summer and autumn models reflected the abundant presence of the short-tailed shearwater *Ardenna tenuirostris* before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively. Given the geographic location of the southern assemblage, however, we believe subantarctic taxa were not well represented in our models due to their low number of occurrences (**Fig. S1 - FO/nOCC/FN**; hence they were removed before modelling). If increased, dedicated seabird effort is carried out, the data could reflect the seasonal northward pulse of these taxa into the southern assemblage (e.g. albatrosses *Diomedea* spp. and *Phoebetria* spp.), given that more effort translates as higher likelihood of recording these species. In turn, including these taxa onto the models could point towards a more stable 'three assemblage pattern' (as revealed by the P/A-autumn model) in all but summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**, instead of the 'southern/northern' assemblages described here.

<!-- Although we did not predict the models onto a future projection, a note on possible climate change effects on seabird off eastern Australia must be done.  -->

## Fitting RCP models to high mobile animals

To the best of our knowledge, this is the first study applying RCPs on high mobile animals such as seabirds. RCP models have been previously applied to plants [@lyons2017], and sessile and **avg?** mobile animals, such as macrobenthic organisms [@pantiukhin2021] and demersal fish [@hill2017]. Differences in species energy and physiological requirements can led animals to using larger areas given their needs and, commonly, pelagic seabirds have daily movements of about hundreds of kilometres per day [e.g. @mcduie2015]. During their movements, seabirds can wander through a variety of habitats and gradients including, for instance, shallow/deep water and poor/productive waters. As such, capturing high mobile species distribution and their relationship with environmental parameters is not a trivial task, and likely harder than for sessile organisms.

While remote tracking studies are advancing our understanding on how seabirds use their environment at fine-, individual-scale level [@tremblay2009], they fail to perceive how groups of species (assemblages) relate to the environment and each other on a shared seascape. Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007]. On the other hand, the spatial scale on which assemblages relate to their environment is also a field of debate [@hunt1987; @ribic1997; @scales2017]. Understanding the biology of the studied group is thus essential to set the appropriate scale of analysis. Our choice of 1° latitude $\times$ 1° longitude was mainly influenced by data availability and the trade-off between number of species per grid/season (see also \@ref(limitations)); nonetheless, the chosen spatial scale is similar with seabird distribution analyses elsewhere [@harris2007; @santora2013; @sojitra2022]. Therefore, for describing seabird assemblages at a large scale, we believe the chosen spatial scale was a good fit, and models described **expected?** patterns of species composition.

RCP models based on presence-absence and abundance (count) data could complement each other. Studies that previously applied RCPs opted by one or another type of data as input. @lyons2017, @hill2017, and @pantiukhin2021 used presence-absence data to run their models, whereas @lee2019good and @receveur2020 used abundance data. Information contained in abundance datasets are inherently better than simplified presence-absence ones. However, the use of different data input in multivariate, biodiversity assessments and assemblage descriptions have shown contrasting patterns at different spatial scales. At larger scales, presence-absence data to equate (relative) abundance in terms of patterns described, while at smaller scales abundance data may highlight local conditions and therefore reveal different patterns, within the broad picture [@marchant1990; @anderson2005; @melo2005; @waters2008marine]. If tied up with biological/ecological knowledge of a particular group, the information contained in both presence-absence and abundance datasets could be leverage to better interpret the resulted patterns. For example, seabirds will likely occur (*presence-absence data*) in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance (*count/abundance data*) are related to fine scale processes and prey availability [@hunt1987; @weimerskirch2007]. As such, the make-up of assemblages can be described through P/A models and its species probability of occurrence within RCPs at large scales, whereas the relationship with environmental variables at finer scales could improve our understanding on how seabirds relate to their environment and where feeding/aggregation hotspots might be, likely nested within the large scale bioregions.

# Conclusions

We applied a robust, repeatable method to assess seabird assemblages off eastern Australia for the first time. The models point toward a consistent pattern, throughout seasons, of a 'northern' and a 'southern' assemblage, where the spatial division clearly links with the EAC behaviour and SST seems to be the key variable discriminating these groups. As the EAC is predicted to strengthen [@poloczanska2007] as well as extreme weather events such as marine heatwaves to happen more often [@oliver2018], seabirds will likely respond to these changes. Southern ocean albatrosses off southeast Australia, for instance, have shown 'early stages of a southward range shift' due to environmental variability [@sojitra2022]. Resident species such as the little penguins and wedge-tailed shearwater are changing their behaviour and decrease their foraging success when SST is higher [@carroll2016; @peck2004]. Given that eastern Australia is a biodiversity and abundance hotspot for seabirds, it is imperative to set a monitoring program on which the assemblages and species profiles presented here could be used as the baseline for understanding possible effects of a warming ocean on seabird distribution. In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the EAC eastern extension/eddy-spawning area, may be fundamental to effectively monitoring EAC strengthening and its impact on the marine biota.

# References

### 2023-06-05 scripts/05_data_EDA.R
tidying up EDA script and saving FigS1 and FigS2

@@ -81,9 +81,9 @@ boxplot_sum.all_season <-
   theme(axis.title = element_text(size = 8),
         axis.text = element_text(size = 6))
 
-ggsave(boxplot_sum.all_season, 
-       filename = "./EDA/EDA_boxplot_total-count-seasons.png",
-       height = 7, width = 7, units = "cm", dpi = 300)
+# ggsave(boxplot_sum.all_season, 
+#        filename = "./EDA/EDA_boxplot_total-count-seasons.png",
+#        height = 7, width = 7, units = "cm", dpi = 300)
 
 ## Number of birds by species/grid/season -----------------------####
 df2 <- 
@@ -106,9 +106,9 @@ boxplot_spp <-
         axis.text.x = element_text(size = 6),
         strip.text = element_text(size = 8))
 
-ggsave(boxplot_spp, 
-       filename = "./EDA/EDA_boxplot_spp-seasons.png",
-       height = 13, width = 20, units = "cm", dpi = 300)
+# ggsave(boxplot_spp, 
+#        filename = "./EDA/EDA_boxplot_spp-seasons.png",
+#        height = 13, width = 20, units = "cm", dpi = 300)
 
 ## Number of occurrences, Frequency of Occurrence, Numeric Frequency by grid/season -------- ####
 
@@ -186,7 +186,7 @@ plot_nOCC_FO_NF <-
         strip.text = element_text(size = 8))
 
 ggsave(plot_nOCC_FO_NF, 
-       filename = "./EDA/EDA_spp-nOCC-FO-NF-seasons.png",
+       filename = "./results/FigS1_spp-nOCC-FO-NF-seasons.png",
        height = 14, width = 20, units = "cm", dpi = 300)
 
 ## String vector w/ spp names to remove from models & how many spp/season was modelled -----------####
@@ -269,6 +269,8 @@ ggsave(effort_raw_seasons,
        filename = "./EDA/EDA_map-effort-raw.png",
        height = 10, width = 8, units = "cm", dpi = 300)
 
+save("effort_raw_seasons", file = "./EDA/EDA_map-effort-raw.rda")
+
 # Save a map to scheme oceanography in Inkscape 
 ggsave((ggplot() + 
           geom_sf(data = world_sf) +
@@ -285,7 +287,7 @@ ggsave((ggplot() +
                                             pad_x = unit(6.2, "cm"),
                                             pad_y = unit(0.4, "cm"),
[…]

### 2023-06-05 scripts/08_2_modelling_fit-RCPs-NegBin.R
fixed colouring bug in point-pred and started script 09 to build final figs

@@ -656,8 +656,25 @@ for (season_vec in seasons_vec) {
   
 }
 
+## Given the order the RCPs were assigned for 'spring' during modelling,
+# I needed to change 'RCP1' to 'RCP2' and vice-versa, to agree with the 
+# colouring pattern from other seasons
+
+plotRCPs_all <- 
+  plotRCPs_all %>% 
+  dplyr::mutate(final_RCP = as.character(final_RCP)) %>% 
+  dplyr::mutate(final_RCP = ifelse(season == "spring" & final_RCP == "RCP1",
+                                   "RCPx", final_RCP)) %>% 
+  dplyr::mutate(final_RCP = ifelse(season == "spring" & final_RCP == "RCP2",
+                                   "RCP1", final_RCP)) %>% 
+  dplyr::mutate(final_RCP = ifelse(season == "spring" & final_RCP == "RCPx",
+                                   "RCP2", final_RCP))
+
 plotRCPs_all$season <- factor(plotRCPs_all$season, levels = c("summer", "autumn",
                                                               "winter", "spring"))
+
+plotRCPs_all$final_RCP <- factor(plotRCPs_all$final_RCP, 
+                                 levels = c("RCP1", "RCP2"))
 
 ## Map it -------------------------------------------------------------------- #
 map_point_pred <-
@@ -677,6 +694,9 @@ ggsave(plot = map_point_pred,
        filename = 
          paste0("./results/NegBin/NegBin_07_RCP-point-predictions.png"),
        height = 10, width = 20, units = "cm", dpi = 300)
+
+save("map_point_pred",
+     file = "./results/NegBin/NegBin_07_RCP-point-predictions-ggplot.rda")
 
 rm("map_point_pred")
 
@@ -748,6 +768,10 @@ for (season_vec in seasons_vec) {
          filename = 
            paste0("./results/NegBin/NegBin_", as.character(season_vec), "_07_RCP-prob-predictions.png"),
          width = 11, height = 11, units = "cm", dpi = 300)
+  
+  save(plotRCPs,
+       filename = 
+         paste0("./results/NegBin/NegBin_", as.character(season_vec), "_07_RCP-prob-predictions-ggplot.rda"))
   
   ## Clean environment
   rm("best_model_files", "best_model",

### 2023-06-05 scripts/09_figs-patchwork.R

@@ -1,0 +1,65 @@
+##
+## ASG -- Figures
+## 
+## Nicholas W Daudt
+## ****************************************************************************#
+
+## This code put together some Figs for the manuscript.
+
+## Libraries ####
+
+library(ggplot2)
+library(patchwork)
+library(magick)
+
+## SVG
+
+# https://cran.r-project.org/web/packages/grImport/vignettes/import.pdf
+# grImport 
+
+# https://github.com/m-jahn/fluctuator
+
+## Figures #### 
+
+## FIGURE 1 ------------------------------------------------------------------ #
+
+# Based on:
+# https://stackoverflow.com/questions/76079316/including-vector-graphics-file-svg-in-a-figure-generated-with-the-patchwork-pa
+
+oceanogr <- magick::image_read_svg("./EDA/EDA_map-oceanogr1.svg")
+gg_oceanogr <- magick::image_ggplot(oceanogr)
+
+load("./EDA/EDA_map-effort-raw.rda")
+
+fig1_study_area <-
+  (gg_oceanogr + effort_raw_seasons) +
+  patchwork::plot_layout(ncol = 2) + #, widths = c(2,1), heights = c(8,1)
+  patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
+
+ggsave(fig1_study_area, 
+       filename = "./results/Fig1_study-area-w-oceanogr.png",
+       height = 10, width = 20, units = "cm", dpi = 300)
+
+## FIGURE 2 ------------------------------------------------------------------ #
+
+load("./results/Bernoulli/Bernoulli_07_RCP-point-predictions-ggplot.rda")
+# As both files were originally names the same ('map_point_pred'), 
+# create a new object for Bernoulli fig (_Ber)
+map_point_pred_Ber <- map_point_pred
+
+load("./results/NegBin/NegBin_07_RCP-point-predictions-ggplot.rda")
+
+fig2_point_pred <-
+  (map_point_pred_Ber + theme(panel.spacing.x = unit(0.8, "lines"))) / 
+  (map_point_pred + theme(panel.spacing.x = unit(0.8, "lines"))) +
+  patchwork::plot_layout(nrow = 2) +
+  patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
+
+ggsave(fig2_point_pred, 
+       filename = "./results/Fig2-point-predictions.png",
+       height = 13, width = 15, units = "cm", dpi = 300)
+
+## FIGURE 3 ------------------------------------------------------------------ #
+
+
+

### 2023-06-06 scripts/05_data_EDA.R
added tableS1

@@ -56,6 +56,23 @@ length(unique(data_raw$order))    # > 5 orders
 
 sum(data_raw$total_ct)            # > In total, 142,646 individuals recorded
 
+## Effort summary, by voyage (Table S1) ####
+
+tableS1 <-
+  data_raw %>% 
+  dplyr::group_by(voyage) %>% 
+  dplyr::summarise(date_start = min(date),
+                   date_end = max(date),
+                   lat_range = paste(round(min(latitude)), "–", round(max(latitude))),
+                   lon_range = paste(round(min(longitude)), "–", round(max(longitude))),
+                   no_records = n(),
+                   no_birds = sum(total_ct),
+                   no_spp = n_distinct(species)) %>% 
+  dplyr::ungroup() %>% 
+  dplyr::arrange(date_start)
+
+write.csv(tableS1, "./results/tableS1-effort-summary.csv")
+
 ## Number of birds (total) by grid/season ----------------------------------####
 df <- 
   data_gridded %>% 

### 2023-06-09 docs/ms-draft.Rmd
found a bug and needed to re-run results, including some new files as well

@@
 ---
-title: "Probabilistic bioregion models suggest seabird assemblages are linked to the main, fast-changing oceanic current off eastern Australia"
+title: "Probabilistic bioregion models suggest seabird assemblages are linked to the boundary current off eastern Australia"
 # "Probabilistic bioregion models suggest seabird assemblages are linked to large scale oceanography off eastern Australia"
@@
   - Eric J. Woehler:
-      institute: [BirdLife, IMAS]
+      institute: [ASG]
       correspondence: false
@@
   - MarSciUO: Department of Marine Science, University of Otago, Aotearoa New Zealand
-  - BirdLife: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
-  - IMAS: Institute Marine and Antarctic Studies, University of Tasmania, Hobart, Australia
+  - ASG: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
   - MathStatUO: Department of Mathematics and Statistics, University of Otago, Aotearoa New Zealand
@@
 Seabird data were gathered aboard RV Investigator and RV Falkor during 15 voyages, between 2016--2021, over \~37° of latitude (**Fig.
-1b**; **Table SX**).
+1b**; **Table S1**).
 One observer continuously recorded seabirds within a 300-m forward quadrant (eye height **18-m, Eric?** a.s.l) throughout daylight hours while the ship was underway at a constant speed.
@@
 
-knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
+knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
 ```
@@
 The best fitting models generally had **1--3** covariates selected, although the abundance-autumn model had five covariates (**Table 2 - full-model/selected model**).
-For the P/A-autumn model, we have deliberately chosen the 'second' best model based on parsimony---the top model had six covariates but was \<1 BIC unit lower than the second model with only two covariates.
 All the best models selected SST or SSS, variables essential for characterising water masses.
@@
 
-<!-- Although we did not predict the models onto a future projection, a note on possible climate change effects on seabird off eastern Australia must be done.  -->
-
 ## Fitting RCP models to high mobile animals
@@
 Our choice of 1° latitude $\times$ 1° longitude was mainly influenced by data availability and the trade-off between number of species per grid/season (see also \@ref(limitations)); nonetheless, the chosen spatial scale is similar with seabird distribution analyses elsewhere [@harris2007; @santora2013; @sojitra2022].
-Therefore, for describing seabird assemblages at a large scale, we believe the chosen spatial scale was a good fit, and models described **expected?** patterns of species composition.
+Therefore, for describing seabird assemblages at a large scale, we believe the chosen spatial scale was a good fit, and models described expected patterns of species composition.
 

### 2023-06-09 scripts/05_data_EDA.R

@@ -127,7 +127,7 @@ boxplot_spp <-
 #        filename = "./EDA/EDA_boxplot_spp-seasons.png",
 #        height = 13, width = 20, units = "cm", dpi = 300)
 
-## Number of occurrences, Frequency of Occurrence, Numeric Frequency by grid/season -------- ####
+## Number of occurrences, Frequency of Occurrence, Numeric Frequency by grid/season (Figure S1) ####
 
 funs <- list(
   # Frequency of Occurrence
@@ -351,7 +351,7 @@ ggsave(effort_sum.all,
        filename = "./EDA/EDA_map-effort-total-birds-recorded-seasons.png",
        height = 10, width = 20, units = "cm", dpi = 300)
 
-## Supp Mat Fig: # richness & total # birds ####
+## patchwork #richness & #birds (Figure S2) ####
 
 SuppMat_Rich.Number <-
   effort_richness / effort_sum.all +

### 2023-06-09 scripts/08_1_modelling_fit-RCPs-Bernoulli.R

@@ -371,6 +371,8 @@ for (season_vec in seasons_vec) {
     rm("i")
   }
   
+  fit_BICs <- fit_BICs %>% dplyr::filter(!is.infinite(bic))
+  
   ## Min BIC
   BIC_min <- min(fit_BICs$bic)
   
@@ -378,7 +380,7 @@ for (season_vec in seasons_vec) {
   fit_BICs <- 
     fit_BICs %>%
     dplyr::mutate(is_min_BIC = as.factor(ifelse(fit_BICs$bic == BIC_min, 'yes', 'no'))) %>%
-    arrange(bic) %>%
+    dplyr::arrange(bic) %>%
     dplyr::mutate(delta_bic = bic - lag(bic)) %>%
     dplyr::mutate(delta_bic = (tidyr::replace_na(delta_bic, 0)))
   
@@ -416,24 +418,37 @@ for (season_vec in seasons_vec) {
   ## Save 'best_model' for diagnostics
   
   # Get the best model name
-  if(season_vec == "autumn") {
-    # Given model parsimony, I deliberately chose this one (delta-BIC < 1)
-    best_model_name <- "bat_mean_sst_mean" 
-  } else {
-    best_model_name <- fit_BICs[1,]$model
-  }
+  best_model_name <- fit_BICs[1,]$model
   
   # Get it from the 'fit' list
   best_model <- fit[[best_model_name]]
   
-  # Save it
+  # Save the best model object
   save("best_model", 
        file = paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_04_best-model.rda"))
   
+  ### Get the 'full model name & full/best model infos
+  full_model_name <- paste(env_cols_models_season[[season_vec]], collapse = "_")
+  
+  full_and_best_models <- data.frame(
+    data_type = rep("bernoulli", times = 2),
+    season = rep(season_vec, times = 2),
+    model = c("full",
+              "best"),
+    model_specification = c(full_model_name,
+                            best_model_name),
+    bic = c(fit_BICs[fit_BICs$model==full_model_name,]$bic,
+            fit_BICs[fit_BICs$model==best_model_name,]$bic)
+  )
+  
+  write.csv(full_and_best_models, 
+            file = paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_04_full_and_best_models.csv"))
+  
   ## Clean environment
-  rm("season_vec", "fit_files", "fit_BICs", "BIC_min", 
[…]

### 2023-06-09 scripts/08_2_modelling_fit-RCPs-NegBin.R

@@ -344,6 +344,8 @@ for (season_vec in seasons_vec) {
     rm("i")
   }
   
+  fit_BICs <- fit_BICs %>% dplyr::filter(!is.infinite(bic))
+  
   ## Min BIC
   BIC_min <- min(fit_BICs$bic)
   
@@ -394,14 +396,32 @@ for (season_vec in seasons_vec) {
   # Get it from the 'fit' list
   best_model <- fit[[best_model_name]]
 
-  # Save it
+  # Save the best model object
   save("best_model",
        file = paste0("./results/NegBin/NegBin_", as.character(season_vec), "_04_best-model.rda"))
-
+  
+  ### Get the 'full model name & full/best model infos
+  full_model_name <- paste(env_cols_models_season[[season_vec]], collapse = "_")
+  
+  full_and_best_models <- data.frame(
+    data_type = rep("neg_bin", times = 2),
+    season = rep(season_vec, times = 2),
+    model = c("full",
+              "best"),
+    model_specification = c(full_model_name,
+                            best_model_name),
+    bic = c(fit_BICs[fit_BICs$model==full_model_name,]$bic,
+            fit_BICs[fit_BICs$model==best_model_name,]$bic)
+  )
+  
+  write.csv(full_and_best_models, 
+            file = paste0("./results/NegBin/NegBin_", as.character(season_vec), "_04_full_and_best_models.csv"))
+  
   ## Clean environment
   rm("season_vec", "fit_files", "fit_BICs", "BIC_min",
      "plot_model_selection_topBIC", "fit",
-     "best_model_name", "best_model")
+     "best_model_name", "best_model",
+     "full_model_name", "full_and_best_models")
   gc()
 }
 
@@ -721,12 +741,6 @@ rm("map_point_pred")
 
 for (season_vec in seasons_vec) {
   
-  # ## Read RCP data ----------------------------------------------------------- #
-  # load("./data_out/inputs/rcp-data-season.rda")
-  # rcp_data <- rcp_data_seasons[grepl(pattern = as.character(season_vec), 
-  #                                    x = names(rcp_data_seasons))]
-  # rcp_data <- rcp_data[[1]]
-  
   ## Read best_model --------------------------------------------------------- #
   best_model_files <- list.files(path = "./results/NegBin", pattern = "_04_best-model.rda$",
                                  full.names = TRUE, recursive = TRUE)
[…]

### 2023-06-09 scripts/09_figs-patchwork.R

@@ -1,10 +1,10 @@
 ##
-## ASG -- Figures
+## ASG -- Figures & Tables
 ## 
 ## Nicholas W Daudt
 ## ****************************************************************************#
 
-## This code put together some Figs for the manuscript.
+## This code put together some Figs & tables for the manuscript.
 
 ## Libraries ####
 
@@ -61,5 +61,9 @@ ggsave(fig2_point_pred,
 
 ## FIGURE 3 ------------------------------------------------------------------ #
 
+load("./results/Bernoulli/NegBin_07_.rda")
+load("./results/NegBin/NegBin_07_.rda")
 
-
+ggsave(fig3_prob_maps_spring, 
+       filename = "./results/Fig3-probability-maps-spring.png",
+       height = 13, width = 16, units = "cm", dpi = 300)

### 2023-06-11 scripts/08_2_modelling_fit-RCPs-NegBin.R
fixed colour bug in negbin summer/spring prob-prediction maps

@@ -776,6 +776,23 @@ for (season_vec in seasons_vec) {
   
   rcp_map = c("RCP_1", "RCP_2")
   
+  if(season_vec == "spring" | season_vec == "summer") {
+    
+    # As RCP group colours are depicted from 'south' (RCP1) to 'north' (RCP2),
+    # we had to do a hack here to the probability maps visually agree with 
+    # point-predictions in a meaningful way (basically, swap "RCP1" to "RCP2", and vice-versa)
+    
+    new_col_names <- c("IDgrid", "lon", "lat", 
+                       "RCP_2", "RCP_1", 
+                       "RCP_2.lower", "RCP_1.lower", 
+                       "RCP_2.upper", "RCP_1.upper", 
+                       "geometry")
+    
+    colnames(pred_data) <- new_col_names
+    
+    rm("new_col_names")
+  }
+  
   plotRCPs <- FUN_prob_maps(column_names = rcp_map)
   
   ggsave(plot = plotRCPs, 

### 2023-06-12 scripts/08_1_modelling_fit-RCPs-Bernoulli.R
fixed a few bugs (iNEXT) and colouring issues (prob-maps) and ran the whole RCP-Bernoulli script again

@@ -600,11 +600,11 @@ for (season_vec in seasons_vec) {
           lowerCI = spp_profile_boot_lower$lowerCI,
           upperCI = spp_profile_boot_upper$upperCI)
   
+  rm("spp_profile_boot_mean", "spp_profile_boot_lower", "spp_profile_boot_upper")
+  
   spp_profile_boot$rcp <- factor(spp_profile_boot$rcp, levels = c("RCP1",
                                                                   "RCP2",
                                                                   "RCP3"))
-  
-  rm("spp_profile_boot_mean", "spp_profile_boot_lower", "spp_profile_boot_upper")
   
   ## Plot -------------------------------------------------------------------- #
   plot_spp_profiles_CI <- 
@@ -811,6 +811,23 @@ for (season_vec in seasons_vec) {
     rcp_map = c("RCP_1", "RCP_2", "RCP_3")
   } else {
     rcp_map = c("RCP_1", "RCP_2")
+  }
+  
+  if(season_vec == "summer") {
+    
+    # As RCP group colours are being depicted from 'south' (RCP1) to 'north' (RCP2),
+    # we had to do a hack here for the probability maps visually agree with 
+    # point-predictions in a meaningful way (basically, swap "RCP1" to "RCP2", and vice-versa)
+    
+    new_col_names <- c("IDgrid", "lon", "lat", 
+                       "RCP_2", "RCP_1", 
+                       "RCP_2.lower", "RCP_1.lower", 
+                       "RCP_2.upper", "RCP_1.upper", 
+                       "geometry")
+    
+    colnames(pred_data) <- new_col_names
+    
+    rm("new_col_names")
   }
   
   plotRCPs <- FUN_prob_maps(column_names = rcp_map)
@@ -1045,12 +1062,12 @@ for (season_vec in seasons_vec) {
                       names(.)[which.max(dplyr::c_across(cols = everything()))]))
   
   finalRCP$final_RCP <- gsub(pattern = "_", replacement = "", x = finalRCP$final_RCP)
+  
   if(season_vec == "autumn"){
     finalRCP$final_RCP <- factor(finalRCP$final_RCP, levels = c("RCP1", "RCP2", "RCP3"))
   } else {
     finalRCP$final_RCP <- factor(finalRCP$final_RCP, levels = c("RCP1", "RCP2"))
   }
-  
   
   ## Bind with 'rcp_data' to get 'IDgrid'
   finalRCP <- 
@@ -1079,7 +1096,7 @@ for (season_vec in seasons_vec) {
[…]

### 2023-06-13 docs/ms-draft.Rmd
runing figs again and sorting plots for ms; few adjustments on the model fitting/results scripts to export ggplot objs as well as constantly updating figs-patchwork script; in source script I noticed it was missing one require() inside the prob-maps plot function so I added it

@@
 ---
-title: "Probabilistic bioregion models suggest seabird assemblages are linked to the boundary current off eastern Australia"
+title: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
 # "Probabilistic bioregion models suggest seabird assemblages are linked to large scale oceanography off eastern Australia"
@@
 
-```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions for seabird RCP models, by season, based on presence/absence data (top row) and abundance data (bottom row) off eastern Australia.
-Each cell was assigned to the RCP group with highest probability (i.e.
-hard-classes).", fig.align = "center", echo = FALSE}
+```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird Region of Common Profile (RCP) models, by season, based on presence-absence (a) and abundance (b) models off eastern Australia.
+Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}
 
@@
 
-```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probabilities of RCP membership in in spring presence/absence (a) and abundance (b) models.", fig.align = "center", echo = FALSE}
+```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probabilities of Region of Common Profiles membership for each grid, based on selected covariates (see Table 2) of spring presence-absence (a) and abundance (b) models for seabird assemblages off eastern Australia.
+Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
 
@@
 However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross *Thalassarche cauta* and Australasian gannet *Morus serrator* standing out with high probabilities of occurrence.
+
+```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges off eastern Australia given by spring presence-absence (a) and abundance (b) Region of Common Profile models.
+Note that values in (b) are log10 transformed.
+Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
+
+knitr::include_graphics("../results/Fig5-spp-profiles-spring.png")
+```
 

### 2023-06-13 scripts/08_1_modelling_fit-RCPs-Bernoulli.R

@@ -185,6 +185,9 @@ for (season_vec in seasons_vec) {
            paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_01_multifit-plot.png"),
          width = 13, height = 6, units = "cm", dpi = 300)
   
+  save("gg_multifitBIC",
+       file = paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_01_multifit-plot-ggplot.rda"))
+  
   ## Save results 'multifit' ------------------------------------------------- #
   save("nRCPs_samp",
        file = paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_01_multifit.rda"))

### 2023-06-13 scripts/08_2_modelling_fit-RCPs-NegBin.R

@@ -175,6 +175,9 @@ for (season_vec in seasons_vec) {
            paste0("./results/NegBin/NegBin_", as.character(season_vec), "_01_multifit-plot.png"),
          width = 13, height = 6, units = "cm", dpi = 300)
   
+  save("gg_multifitBIC",
+       file = paste0("./results/NegBin/NegBin_", as.character(season_vec), "_01_multifit-plot-ggplot.rda"))
+  
   ## Save results 'multifit' ------------------------------------------------- #
   save("nRCPs_samp",
        file = paste0("./results/NegBin/NegBin_", as.character(season_vec), "_01_multifit.rda"))
@@ -572,10 +575,22 @@ for (season_vec in seasons_vec) {
           lowerCI = spp_profile_boot_lower$lowerCI,
           upperCI = spp_profile_boot_upper$upperCI)
   
+  rm("spp_profile_boot_mean", "spp_profile_boot_lower", "spp_profile_boot_upper")  
+  
+  if(season_vec == "spring"){
+    # As RCP group colours are depicted from 'south' (RCP1) to 'north' (RCP2),
+    # we had to do a hack here for the species profiles visually agree with 
+    # point-predictions in a meaningful way (basically, swap "RCP1" to "RCP2", and vice-versa)
+    
+    spp_profile_boot <- 
+      spp_profile_boot %>% 
+      dplyr::mutate(rcp = ifelse(rcp == "RCP1", "RCPx", rcp)) %>% 
+      dplyr::mutate(rcp = ifelse(rcp == "RCP2", "RCP1", rcp)) %>% 
+      dplyr::mutate(rcp = ifelse(rcp == "RCPx", "RCP2", rcp))
+  }
+  
   spp_profile_boot$rcp <- factor(spp_profile_boot$rcp, levels = c("RCP1",
                                                                   "RCP2"))
-  
-  rm("spp_profile_boot_mean", "spp_profile_boot_lower", "spp_profile_boot_upper")
   
   ## Plot -------------------------------------------------------------------- #
   plot_spp_profiles_CI <- 
@@ -596,6 +611,10 @@ for (season_vec in seasons_vec) {
          filename = 
            paste0("./results/NegBin/NegBin_", as.character(season_vec), "_07_species-profiles.png"),
          width = 16, height = 8, units = "cm", dpi = 300)
+  
+  # Save ggplot objects
+  save("plot_spp_profiles_CI",
[…]

### 2023-06-13 scripts/09_figs-patchwork.R

@@ -19,7 +19,7 @@ library(magick)
 
 # https://github.com/m-jahn/fluctuator
 
-## Figures #### 
+## Figures main text #### 
 
 ## FIGURE 1 ------------------------------------------------------------------ #
 
@@ -50,8 +50,12 @@ map_point_pred_Ber <- map_point_pred
 load("./results/NegBin/NegBin_07_RCP-point-predictions-ggplot.rda")
 
 fig2_point_pred <-
-  (map_point_pred_Ber + theme(panel.spacing.x = unit(0.8, "lines"))) / 
-  (map_point_pred + theme(panel.spacing.x = unit(0.8, "lines"))) +
+  (map_point_pred_Ber + 
+     theme(# panel.spacing.x = unit(0.8, "lines"),
+           axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))) / 
+  (map_point_pred + 
+     theme(# panel.spacing.x = unit(0.8, "lines"),
+           axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))) +
   patchwork::plot_layout(nrow = 2) +
   patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
 
@@ -59,11 +63,197 @@ ggsave(fig2_point_pred,
        filename = "./results/Fig2-point-predictions.png",
        height = 13, width = 15, units = "cm", dpi = 300)
 
+rm("map_point_pred_Ber", "map_point_pred", "fig2_point_pred")
+
 ## FIGURE 3 ------------------------------------------------------------------ #
 
-load("./results/Bernoulli/NegBin_07_.rda")
-load("./results/NegBin/NegBin_07_.rda")
+load("./results/Bernoulli/Bernoulli_spring_07_RCP-prob-predictions-ggplot.rda")
+# As both files were originally names the same ('plotRCPs'), 
+# create a new object for Bernoulli fig (_Ber)
+plotRCPs_Ber <- plotRCPs
+
+load("./results/NegBin/NegBin_spring_07_RCP-prob-predictions-ggplot.rda")
+
+fig3_prob_maps_spring <- 
+  (plotRCPs_Ber + theme(legend.position = "none")) / 
+  plotRCPs +
+  patchwork::plot_layout(ncol = 2) +
+  patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
 
 ggsave(fig3_prob_maps_spring, 
        filename = "./results/Fig3-probability-maps-spring.png",
[…]

### 2023-06-13 scripts/source.R

@@ -30,6 +30,7 @@ FUN_prob_maps <- function(df = pred_data,
   # Required libraries
   library(dplyr)
   library(ggplot2)
+  library(tidyr)
   library(sf)
   library(classInt)
   library(colorspace)

### 2023-06-13 scripts/09_figs-patchwork.R
finished plots for ms for now -- still waiting for partial plots

@@ -257,3 +257,227 @@ ggsave(
 
 rm("sum", "aut", "win", "spr")
 
+## FIGURE S5 ----------------------------------------------------------------- #
+
+## --------------------- S5.1 Bernoulli ------------------------------------- ##
+
+load("./results/Bernoulli/Bernoulli_summer_07_RCP-prob-predictions-ggplot.rda")
+plotRCPs_sum <- plotRCPs
+
+load("./results/Bernoulli/Bernoulli_autumn_07_RCP-prob-predictions-ggplot.rda")
+plotRCPs_aut <- plotRCPs
+
+load("./results/Bernoulli/Bernoulli_winter_07_RCP-prob-predictions-ggplot.rda")
+plotRCPs_win <- plotRCPs
+
+load("./results/Bernoulli/Bernoulli_spring_07_RCP-prob-predictions-ggplot.rda")
+
+figS5.1_prob_pred <-
+  (plotRCPs_sum + ggtitle("Summer") + 
+     theme(plot.title = element_text(size = 8, face = "italic"),
+           legend.position = "none",
+           axis.text = element_text(size = 3.5),
+           strip.text = element_text(size = 4))) +
+  (plotRCPs_aut + ggtitle("Autumn") + 
+     theme(plot.title = element_text(size = 8, face = "italic"),
+           legend.position = "none",
+           axis.text = element_text(size = 3.5),
+           strip.text = element_text(size = 4))) + 
+  (plotRCPs_win + ggtitle("Winter") + 
+     theme(plot.title = element_text(size = 8, face = "italic"),
+           legend.position = "none",
+           axis.text = element_text(size = 3.5),
+           strip.text = element_text(size = 4))) + 
+  (plotRCPs + ggtitle("Spring") + 
+     theme(plot.title = element_text(size = 8, face = "italic"),
+           axis.text = element_text(size = 3.5),
+           strip.text = element_text(size = 4),
+           legend.key.size = unit(0.3, 'cm'),
+           legend.text = element_text(size = 4),
+           legend.title = element_text(size = 5))) + 
+  patchwork::plot_layout(ncol = 4)
+
+ggsave(figS5.1_prob_pred, 
+       filename = "./results/FigS5_1_prob-pred-Bernoulli.png",
+       height = 9, width = 20, units = "cm", dpi = 300)
+
+## ---------------------- S5.2 NegBin --------------------------------------- ##
+##      Note: I'm overwriting previous objects by using the same names
+
[…]

### 2023-06-27 docs/ms-draft.Rmd
fixed tables 1 (cfs) and 2 (-Inf)

@@
 title: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
-# "Probabilistic bioregion models suggest seabird assemblages are linked to large scale oceanography off eastern Australia"
-# author: Nicholas W Daudt
-# date: '`r format(Sys.Date())`'
 author:
@@
 output: 
-  # html_document: default
-  # word_document:
-  #   reference_docx: "../docs/docx-template.docx"
-  # officedown::rdocx_document:
-  #   reference_docx: "../docs/docx-template.docx"
   # rticles::sim_article:
@@
 library(gt)
-# library(officedown)
-# library(officer)
 ```
@@
 Seabird assemblages have been described and related to environmental characteristics worldwide.
-Considering the hierarchical nature of oceanographic processes and its resulting features (**Haury78**), it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987].
+Considering the hierarchical nature of oceanographic processes and its resulting features [@haury1978patterns], it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987].
 In the marine environment, large scale oceanographic features (water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e.
@@
 
-```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records by season on top of the 1° latitude × 1° longitude grid cells (b).
+```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
 In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue color.
@@
 
-knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
+# knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
+
+knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
 ```
@@
   Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
-  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "cf", "cf", "cf", "0.25° × 0.25°", "cf", "cf", "NA"),
+  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
   `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
@@
 
-<!-- summer \~ bat+slope+eke+sst+sss+mld+sst_grad -->
-
-<!-- autumn \~ bat+slope+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad -->
-
-<!-- winter \~ bat+slope+eke+sst+sss+mld+sst_grad+clim_sst_grad -->
-
-<!-- spring \~ bat+slope+dist_coast+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad+clim_eke_mean -->
-
 # Results
@@
 The five most abundant species summed 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*), whereas the five most frequent species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig.
-S1 - FO/nOCC/FN**).
-An increase in species richness towards the south was clear in all but the winter season (**Fig S2.
-- spp richness/total abund**).
+S1**).
+An increase in species richness towards the south was clear in all but the winter season (**Fig S2**).
 Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig.
@@
 
-All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3.
-- multifit plots**).
-The best fitting models generally had **1--3** covariates selected, although the abundance-autumn model had five covariates (**Table 2 - full-model/selected model**).
+All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3**).
+The best fitting models generally had 1--2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**).
 All the best models selected SST or SSS, variables essential for characterising water masses.
 The best models show a good fit with no apparent deviances in assumptions (**Fig.
-S4 - residuals**).
+S4**).
+
+```{r table2, echo = FALSE, message = FALSE}
+
+df <- 
+  rbind(
+    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
+    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
+  purrr::map_df(~readr::read_csv(.)) %>% 
+  dplyr::select(- `...1`) %>% 
+  dplyr::mutate(data_type = factor(data_type, levels = c("bernoulli", "neg_bin")),
+                season = factor(season, levels = c("summer", "autumn", "winter", "spring")),
+                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
+                bic = round(bic, digits = 2)) %>%
+  dplyr::group_by(data_type) %>% 
+  dplyr::arrange(data_type, season) %>% 
+  as.data.frame(.) %>% 
+  dplyr::ungroup(.)
+
+# Correct a value in a cell
+# Full abundance-autumn model did not fit properly, returning a value of -Inf
+df[df$data_type=='neg_bin' & df$season=='autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"
+
+# A few tweaks to make the table look better
+df$data_type <- c("Bernoulli", rep("", 7), "Negative Binomial", rep("", 7))
+df$season <- rep(c("Summer", "", "Autumn", "", "Winter", "", "Spring", ""), 2)
+
+gt::gt(df)|>
+  ## Improve cols names
+  gt::cols_label(
+    data_type = gt::md("**Data type**"),
+    season = gt::md("**Season**"),
+    model = gt::md(" "),          ## To be excluded in the Word doc
+    model_specification = gt::md("**Model specification**"), 
+    bic = gt::md("**BIC**")) |>
+  ## Table title
+  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli error distribution; abundance, based on a Negative Binomial error distribution)"))
+
+rm("df")
+## Can't figure out why this code below doesn't work...
+# kableExtra::kbl(df, booktabs = T, align = "l", linesep = "") %>% 
+#   kableExtra::column_spec(1, bold=T) %>%
+#   kableExtra::collapse_rows(columns = c(1:3), 
+#                             latex_hline = 'custom', 
+#                             custom_latex_hline = 1:3, 
+#                             row_group_label_position = 'stack')
+```
 
@@
 
-```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird Region of Common Profile (RCP) models, by season, based on presence-absence (a) and abundance (b) models off eastern Australia.
+```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
 Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}
@@
 
-```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probabilities of Region of Common Profiles membership for each grid, based on selected covariates (see Table 2) of spring presence-absence (a) and abundance (b) models for seabird assemblages off eastern Australia.
+```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.
 Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
@@
 *Relationship with environmental covariates (highlight SST) --**waiting for 'partial plots' (Fig. 4, Fig. S6)***
+
+```{r fig4-partial-plots, fig.cap = "**FIGURE 4.**  Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
+
+knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
+```
 
@@
 
-```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges off eastern Australia given by spring presence-absence (a) and abundance (b) Region of Common Profile models.
-Note that values in (b) are log10 transformed.
+```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges (Region of Common Profile; RCP) off eastern Australia for spring presence-absence (a) and abundance (b) models.
+Values are the average and confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
+Values in (b) were log10-transformed to accomodate the high variation between species.
 Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
@@
 In addition, ideally, we could have had more information within each grid cell by season (our sampling units).
-Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8 - rarefaction curves**).
+Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8**).
 

### 2023-07-10 docs/ms-draft.Rmd
review from Will, plus partial plots sentences

@@
 abstract: |
-  **Aim**: To identify assemblages of seabird species at large scale, probabilistically, and related them to environmental conditions off eastern Australia, a world seabird hotspot.
-  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and counts).
-RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within it but different from other regions.
-These groups also vary according to covariates, on which we included oceanographic static and climatic, as well as physiographic layers.
-  **Results**: All models suggested two large scale assemblages ('northern' and 'southern' assemblages), except for the abundance-autumn model which indicated three groups.
-Model results seem to agree in a biogeographic barrier at ~32°S, where the East Australian Current mostly dissipates to the east and loses force.
-Furthermore, sea surface temperature and, in one case, sea surface salinity, were selected in all final models, further indicating close relationship between seabird assemblages and water masses.
-Results from both data types, presence-absence and counts resulted in similar spatial and species profile patterns.
-  **Conclusion**: Two seabird assemblages were clearly identified from RCP models, suggesting the persistence of these groups at temporal and spatial scales.
-Given the rapid change that the East Australian Current is experiencing, and its importance on driving seabirds' distributions, the methods applied in our study could be replicated in future analyses to assess possible spatial/temporal changes in seabird assemblages, including on its structure.
+  **Aim**: To identify assemblages of seabird species, probabilistically, and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different among them.
+These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic layers.
+  **Results**: All models suggested two large scale assemblages ('northern' and 'southern'), except for the autumn presence-absence model which indicated three groups.
+Model results seem to agree in a biogeographic barrier at ~34°S, where the East Australian Current dissipates as it flows eastward and loses force.
+Moreover, sea surface temperature or sea surface salinity, were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
+Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
+  **Conclusion**: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and large spatial scales.
+Given the rapid change that the East Australian Current is experiencing, and its importance in driving distributions of seabirds, the methods applied in our study could be replicated in future analyses to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 output: 
@@
 
-<!-- 'rticles::sim_article' example? -->
-
-<!-- https://github.com/emitanaka/combine2019/tree/master/output/articles/article_template_sim_article -->
-
 ```{r setup, include = FALSE, echo = FALSE}
+library(readr)
+library(dplyr)
+library(purrr)
 library(gt)
@@
 Bioregionalization is the discipline that identifies, delimits, characterises, and names biogeographical regions [@vilhena2015; @morrone2018].
-Although there is no agreement on the naming systems between proposed regionalizations [but see @ebach2008], the hierarchical nature of bioregions is well recognized [@vilhena2015; @kreft2010].
+Although there is often no agreement on the naming systems between proposed regionalizations [but see @ebach2008], the hierarchical nature of bioregions is well recognized [@vilhena2015; @kreft2010].
 Thus, bioregions are usually described within geographically larger classifications, such as 'provinces' or 'realms' [@spalding2007; @costello2017].
-The delimitation of bioregions is made through identifying sets of species that are endemic or with high levels of co-occurrence, distinct enough from other regions, which is interpreted as a natural representation of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010].
-In particular, the identification of bioregions allows the representation of the distribution of multiple species at a given time and space as well as to relate these regions to environmental characteristics.
+The delimitation of bioregions is made through identifying sets of species that are endemic or have high levels of co-occurrence, and are distinct enough from other regions.
+In this sense, bioregions are interpreted as a natural representations of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010].
+In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space as well as relating these regions to environmental characteristics.
 By doing so, bioregions can be interpreted as species assemblages [@lyons2017; @hill2017].
-Given these features, bioregions have a fundamental role in conservation biogeography [@whittaker2005], by being used as management units and even shedding light into conservation gaps [@spalding2007].
+Given these features, bioregions can be used as surrogates for biodiversity monitoring in data-poor regions and in regional conservation planning [@spalding2007; @ferrier2002a], playing a fundamental role in conservation biogeography [@whittaker2005].
 
-Seabird assemblages have been described and related to environmental characteristics worldwide.
-Considering the hierarchical nature of oceanographic processes and its resulting features [@haury1978patterns], it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987].
-In the marine environment, large scale oceanographic features (water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e.
-the occurrence of species and its relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014].
-On the other hand, at smaller scales, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014].
-Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013].
-Noticeably, sea surface temperature is a key driver of distribution of many seabird species, at individual- and assemblage-levels, as well as at large (water masses) and smaller (fronts) scales [@dunlop1988; @ribic1997; @quillfeldt2015].
-Nonetheless, many regions still lack information regarding multiple seabird species distribution and how they assemble in space and time.
+Seabird assemblages have been widely described and related to environmental characteristics.
+Considering the hierarchical nature of oceanographic processes and their resulting features [@haury1978patterns], it is not surprising that the distribution of highly mobile seabirds are linked differently across scales [@hunt1987].
+In the marine environment, large scale oceanographic features (e.g.
+water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e.
+the occurrence of species and their relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014].
+In contrast, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014].
+Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces, and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013].
+Noticeably, sea surface temperature is a key driver of seabird distribution, at individual- and assemblage-levels, operating over a range of scales from large (e.g.
+water masses) to small (e.g.
+fronts) [@dunlop1988; @ribic1997; @quillfeldt2015].
+Nonetheless, many regions still lack information regarding distribution of seabird species and how they assemble in space and time.
 
-One such area that could benefit from better understanding of seabird distributions is eastern Australia, considered a world seabird hotspot [@karpouzi2007; @ramírez2017].
-Surprisingly, however, there are not many studies on seabird distribution relative to the region's relevance for this group [@mott2018; @bernard2021].
-Off northwest Australia, at-sea surveys, multivariate analyses [@dunlop1988] and ensemble species distribution models [@lavers2014] of species assemblages have shown that preferential habitat use of tropical species are related to sea surface temperature and salinity, as well as indicated the importance of the continental shelf for some species.
-In particular, @dunlop1988 suggests that salinity is the main variable influencing species distribution in oceanic waters whereas temperature would be the key factor on the continental shelf.
-On the other hand, off eastern Australia, there is a scarcity of studies at larger scales.
-@reid2000 and @evans2021a analysed seabird assemblages and their relationship with oceanography and prey data, however at coarse scale (tens of km), off southeast Tasmania.
-At a multi-species, assemblage-level, the latter studies seem to be the only ones on eastern Australia; other studies were mostly descriptive species accounts at local scale [@norman1992distribution; @blaber1994distribution], or focused on a single species [@priddel2014].
+One area that could benefit from better understanding of seabird distributions is Australia, considered a world seabird hotspot [@karpouzi2007; @ramírez2017].
+Given the region's importance for seabird diversity, there are surprisingly few studies on seabird distribution [@mott2018; @bernard2021].
+Off northwest Australia, at-sea surveys, multivariate analyses [@dunlop1988] and ensemble species distribution models of species assemblages [@lavers2014] have shown that habitat use by tropical species is related to sea surface temperature and salinity.
+In particular, @dunlop1988 suggested that salinity is the main variable influencing species distribution in oceanic waters whereas temperature is the key factor on the continental shelf.
+In contrast, there are few studies at large scales off eastern Australia.
+@reid2000 and @evans2021a analysed seabird assemblages and their relationships with oceanography and prey data at broad scales (tens of km), off southeast Tasmania.
+The remaining studies are mostly descriptive accounts at local scales [e.g.
+@norman1992distribution; @blaber1994distribution], or focused on a single species [e.g.
+@priddel2014].
 
-The East Australian Current (EAC) is the dominant current in eastern Australia, the major western boundary current in the South Pacific [@ganachaud2014].
-The EAC has been extensively studied by physical oceanographers, although its link between oceanography and biological processes is still in its infancy.
-The dynamic features provided by EAC, such as eddies and upwellings, can likely influence marine biota [@suthers2011; @scales2014].
-While most of the biological studies done within this system were focused on plankton entrapment within the cores of eddies [@garcia2022; @firme2023; @hassler2011], there is a growing body of work focusing on fisheries [@hobday2010; @young2011; @brieva2015].
-These studies emphasize how the assemblage structure and probability of occurrence of a given taxa is driven by EAC.
-In particular, the potential biological consequence triggered by climate changes is the tropicalization of the biota, given the EAC is strengthening and getting warmer [@cai2005; @poloczanska2007; @hobday2010; @phillips2022], as shown elsewhere [@hastings2020].
-Notwithstanding, the relationship between oceanography and marine megafauna along the EAC is poorly known [@suthers2011], albeit they are considered environmental sentinels [@hazen2019].
-Some progress has been made using shark and ray movement data [@jaine2014; @niella2020; @niella2022], and satellite tracking have also demonstrated that increased sea surface temperature in EAC decreases penguin foraging success [@carroll2016].
+The East Australian Current (EAC) is the major western boundary current in the South Pacific, and the dominant current in eastern Australia [@ganachaud2014].
+The EAC has been extensively studied, although its link between oceanography and biological processes is still in its infancy.
+The dynamic features caused by the EAC, such as eddies and upwellings, are likely to influence marine biota [@suthers2011; @scales2014].
+While most of the biological studies done in this system were focused on plankton entrapment within the cores of eddies [@garcia2022; @firme2023; @hassler2011], there is a growing body of work focusing on fisheries [e.g.
+@hobday2010; @young2011; @brieva2015].
+These studies emphasize how the probability of occurrence of a given taxon, and the structure of species assemblages, is driven by the EAC.
+Notwithstanding, the relationship between oceanography and marine megafauna along the EAC is poorly known [@suthers2011].
+Some progress has been made using shark and ray movement data [@jaine2014; @niella2020; @niella2022], and satellite tracking has also demonstrated that increased sea surface temperature in EAC decreases penguin foraging success [@carroll2016].
 However, information at the assemblage-level is still scarce [but see @evans2021a].
 
-Given east Australia's importance for seabirds and the recent attention on how climate changes in the EAC can affect marine biota [@poloczanska2007], an assessment of seabird spatial distribution at large scale is timely.
-Seabirds can serve as sentinels for tracking EAC ecological changes in many facets [e.g.
-@sydeman2012], including by their distribution.
-Surprisingly, though, seabird distribution is under-represented in climate change studies, despite being pointed as a priority topic [@grémillet2009; @sydeman2012].
-In this study, we aim to identify assemblages of seabird species probabilistically and related them to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
-We used the mixture-model Region of Common Profiles [RCP; @foster2013], which is under the best current available tools for describing bioregions [@hill2020; @woolley2020].
-RCP is model-based approach, which include a series of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
-East Australia supports a high diversity of seabirds and offers wide environmental gradients, from the tropics to the subantarctics.
+Marine species range shifts are more likely to occur faster than terrestrial species due to their higher sensitivity to warming [@pinsky2019; @lenoir2020].
+As a result, climate change is rearranging the structure of species assemblages around the globe [@poloczanska2013; @antão2020; @dornelas2014].
+Top predators, for instance, may change their core habitats [@hazen2013a; @reisinger2022], with unprecedented population consequences.
+For seabirds, @péron2010 showed that sea surface temperature anomalies affected the occurrence and abundance of albatrosses and petrels species along a latitudinal range in the Indian Ocean, thus affecting assemblage structure.
+In eastern Australia, changing EAC conditions have affected a range of species.
+Changes in zooplankton assemblage structure and fish distribution patterns have been recorded, and shark distribution shifts are expected to occur [@johnson2011; @niella2022].
+One of the strongest marine heatwave ever recorded, in the Tasman Sea, showed drastic environmental impacts [@oliver2017].
+Therefore, given that the EAC is strengthening and getting warmer [@cai2005; @phillips2022], biological changes are expected to occur [@poloczanska2007; @hobday2010].
+
+Given east Australia's importance for seabirds and the recent attention on how climate change in the EAC can affect marine biota [@poloczanska2007], an assessment of the spatial distribution of seabirds at large scales is timely.
+Seabirds can serve as sentinels for tracking ecological changes in the EAC in many ways [e.g.
+@sydeman2012].
+Surprisingly, though, seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic [@grémillet2009; @sydeman2012].
+In this study, we aim to identify assemblages of seabird species probabilistically and relate them to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
+We used the mixture-model Region of Common Profiles [RCP; @foster2013], regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
+RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
+East Australia supports a high diversity of seabirds and offers wide environmental gradients, from the tropics to the subantarctic.
 Therefore, we expect to find different assemblages relating to different environmental conditions.
-In addition, we provide a repeatable framework on which future studies can build up to assess how EAC changes will influence seabird assemblage distributions.
+In addition, we run RCP models for presence-absence and abundance as input data, to test if the different numeric resolution would result in similar assemblage patterns.
+Furthermore, we provide a repeatable framework on which future studies can build to assess how changes in the EAC will influence the distribution of seabird assemblages.
 
@@
 
-The EAC originates from the South Equatorial Current (SEC), a tropical, nutrient-poor water mass.
+The EAC originates from the South Equatorial Current (SEC), a tropical, nutrient-poor water.
 When the SEC's North Caledonian Jet hits the Queensland Plateau (\~18°S), it splits into a north branch running towards the Gulf of Papua and a south branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
@@
 1a**; @kessler2013; @ganachaud2014).
-The EAC interacts with topographical features on the seafloor, from which oceanographic features and processes arise, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
-The EAC is mainly attached to the shelf break, but at \~32°S, it spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; typically mentioned as the 'Tasman Front').
+**SST/SSS characteristics?
+ROB?** The EAC interacts with the topography of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
+The EAC is mainly attached to the shelf break, but at \~32°S it spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
 Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.
@@
 One observer continuously recorded seabirds within a 300-m forward quadrant (eye height **18-m, Eric?** a.s.l) throughout daylight hours while the ship was underway at a constant speed.
-Species, the number of birds and behaviour were logged directly onto a data entry system through a ship-based web server.
+The species, the number of birds per group, and behaviour were logged directly into a data entry system through a ship-based web server.
 Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species) **(Eric: is there any ref for the protocol?)**.
+Birds following the boat were previously excluded from analyses.
 
 ```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
-In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue color.
-The EAC system originates from the South Equatorial Current (SEC) and its branches, North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), drawn in red.
-The EAC 'eastern extension' is an eddy field (blue shaded area) often mentioned as the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
+In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
+The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
+The EAC 'eastern extension' is an eddy field (blue shaded area) often called the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
 Figure (a) is adapted after Oke, Roughan, et al.
@@
 To explain and predict species distribution and biogeography, a thoughtful choice of covariates is important [@barbet-massin2014].
-We used seven oceanographic variables and calculated three climatic layers from them, and three physiographical variables (**Table 1**).
+We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table 1**).
 These variables are the most used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of scales from large-scale water-mass properties (e.g.
-sea surface temperature) to increased predictability of prey availability in mesoscale oceanographic features (e.g.
+sea surface temperature) to prey availability in mesoscale oceanographic features (e.g.
 fronts and eddies; @weimerskirch2007; @scales2014).
-Seabird records were matched with the best open-access, remote-sensed or interpolated data available, and distance from the coast was calculated in R software (see details below).
+Seabird records were matched with the best open-access, remote-sensed or interpolated data available, and distance from the coast was calculated in R software, version 4.2.0 [@r2022] (see details below).
 
@@
   Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
-  Abbreviation = c("sst", "sstgrad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke", "clim_ekesd", "clim_sstgrad", "bat", "slope", "dist_coast"),
+  Abbreviation = c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"),
   Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
@@
   ## Table title
-  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources"))
+  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources."))
 
@@
 
-Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022], using packages **X, Y, ..., Z**. Code and (gridded) data are available in **Daudt et al. 2023 (Zenodo repo)**.
+Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022], using packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021], 'iNEXT' 3.0.0 [@chao2014; @hsieh2016], and 'ecomix' 1.0.0 [@ecomix].
+Code and the modelling dataset (see \@ref(modellingdatasets)) are available in **Daudt et al.
+2023 (Zenodo repo)**.
 
-Before data extraction, a few environmental layers needed to be calculated.
-Spatial and temporal persistence of features such as eddy kinetic energy (EKE) and thermal fronts (SSTgrad) may influence seabird movement and distribution by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
-Therefore, for EKE and SSTgrad, we calculated climatic, seasonal average layers based on 10-year data (2011--2021), as well as EKE standard deviation (EKEsd) over the same period.
-EKE was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ (**ROB: Ref here?**).
-Sea surface temperature gradient (SSTgrad) was calculated as the spatial gradient by identifying the maximum value for each neighbour cell and averaging them (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
-EKE and SSTgrad calculations were done layer-by-layer to match seabird data on the finest scale possible.
-The slope was calculated using `raster::terrain()`, and distance from the coast was calculated through `sf::st_distance()` using Australia and Tasmania as mainland (ignoring smaller islands).
-The most proximate value for each seabird record was extracted from each environmental layer using built-in functions from `raster`, `terra`, and `sf` packages.
+### Environmental layers
+
+Spatial and temporal persistence of features such as eddy kinetic energy (eke) and thermal fronts (sea surface temperature gradient; sst_grad) may influence the movement and distribution of seabirds by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
+Therefore, before extracting environmental data for each seabird record, 'eke' and 'sst_grad' environmental layers were calculated.
+For both, 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table 1**) and calculations were done layer-by-layer to match seabird data on the finest scale possible.
+Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ (**ROB: Ref here?**).
+The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
+In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on 10-years of data (2011--2021) as a proxy for persistent oceanographic features, as well as 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table 1**).
+Seabed slope was calculated using 'raster::terrain()', and distance from the coast was calculated through 'sf::st_distance()' using Australia and Tasmania as the mainland (i.e.
+ignoring smaller islands).
+Then, for each seabird record, the most proximate value for each environmental layer (original or calculated) was extracted using built-in functions from 'raster', 'terra', and 'sf' packages.
+
+### Modelling datasets {#modellingdatasets}
 
@@
 Therefore, seasonal differences in species composition were expected [e.g.
-@reid2023], as well as in their probability of occurrence and total numbers.
-Given that, we aggregated data by 1° latitude $\times$ 1° longitude grid cells (c.
+@reid2023], as well as in probability of occurrence and total numbers.
+Thus, we build separate modelling dataset for each season.
+We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (c.
 110 $\times$ 110 km) and season (**Fig.
 1b**), by summing seabirds from the same species.
-Birds following the boat were previously excluded from analyses.
-Considering our aim to test if presence-absence (P/A) and abundance (count) data would result in similar assemblage patterns, we built two species matrices, where the P/A matrix transformed values \>1 to 1.
-In addition, to avoid spurious correlations with assemblage/environment and numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12; autumn = 21; winter = 20; spring = 33).
-During data aggregation, the environmental data were averaged.
-Before modelling, environmental data were scaled and centred.
+Considering our aim to test if presence-absence and abundance (count) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
+Furthermore, to avoid numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12 species; autumn = 21; winter = 20; spring = 33).
+The associated environmental data of each seabird record were averaged within the same grid cells and seasons.
+Environmental data were scaled and centred before modelling.
 
@@
 We used the mixture model Region of Common Profiles (RCP) to identify and describe seabird assemblages.
-RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates biotic (multi-species) and abiotic (environmental) data [@woolley2020; @hill2020].
+RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates response (multi-species data) and predictor (environmental data) [@woolley2020; @hill2020].
 The model defines 'regions' of similar environmental spaces in which the probability of observing a set of species 'profiles' is constant but different from other regions [@foster2013; @foster2017].
-As such, RCPs can be directly interpreted as assemblages, and its species profiles as the assemblage structure.
+As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
 RCP memberships are latent (unobserved) and vary as a function of environmental data, assuming sites within each RCP have constant species profiles.
-Given a set of environmental variables, the model also enables predicting assemblage probabilities onto sites where no biotic data exist.
-Additionally, as a feature of model-based approaches, RCPs allow the opportunity to compare models through information criteria, present uncertainty around the results, and check for model assumptions through checking distribution of residuals [@warton2015; @hill2020].
+Given a set of environmental variables, the model also enables prediction of assemblage probabilities onto sites where no biotic data exist.
+Additionally, as a feature of model-based approaches, RCPs allow the opportunity to compare models through information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals [@warton2015; @hill2020].
 
-Before fitting models, covariates were checked for collinearity within each season using the modelling dataset.
-We removed variables $\left|x\right| >0.7$ [@dormann2013], favouring direct measures over the 'climatic' layers we calculated (e.g.
-if SSTgrad and 'climatic' SSTgrad were correlated, we chose SSTgrad).
-Thus, 'full models' differed between seasons as the covariates kept vary for each season, although most covariates were common.
+Before fitting models, covariates were checked for collinearity within each seasonal modelling dataset.
+If a pairwise collinearity of $\left|x\right| >0.7$ was found we removed one of the variables [@dormann2013], favouring direct measures over the 'climatic' layers we calculated (e.g.
+if 'sst_grad' and 'clim_sst_grad' were correlated, we chose 'sst_grad').
+Thus, 'full models' differed between seasons as the retained covariates varied, although most covariates were common for all seasons.
 
-Due to seabird migration, seasons could act as 'sampling artifacts' [@foster2017], biasing results [@ponti2023].
-Given the difference in species between seasons that met our occurrence threshold, we decided to run RCP models separately by season and data type (P/A, count).
-RCPs were fitted using the `ecomix` package.
-All models follow the workflow: (i) finding the best number of groups that describes the data, (ii) re-fit the model by fixing the best number of groups and going through covariates selection, (iii) checking residuals, (iv) bootstrapping the best model to get uncertainty around parameter estimates, and (v) producing results.
-For steps (i) and (ii), we used Bayesian Information Criterion (BIC) as the metric for model selection, and diagnostics (iii) were assessed visually.
+Due to seabird migration, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
+Given the difference among seasons in the number of species that met our occurrence threshold (see \@ref(modellingdatasets)), we decided to run RCP models separately by season and data type (presence-absence and abundance).
+RCP models were used to relate the response variables (i.e.
+presence-absence or abundance of each seabird species in each grid cell) to the independent predictor covariates (i.e.
+averaged values of environmental variables in each cell).
+Presence-absence data were modelled based on a Bernoulli error distribution, whereas abundance data based on a Negative Binomial error distribution.
+RCPs were fitted using the 'ecomix' package [@ecomix], with the following workflow: (i) find the best number of groups that describes the data, (ii) re-fit the model by fixing the best number of groups and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
 
-To identify the best number of groups that describe the data, we ran full models (all covariates) with 100 random start numbers [see @foster2013], between 2--7 groups, and chose the number of RCPs that minimized BIC.
-Then, to assess the best set of environmental covariates, we fixed the number of groups and fitted models for every possible linear, additive combination of covariates specified in the full model.
+To identify the best number of groups that described the data, we ran full models (all covariates) with 100 random start numbers [see @foster2013] between 2--7 groups, and chose the number of RCPs that minimized the Bayesian Information Criterion (BIC).
+Then, we fixed the number of groups at the optimal number and fitted models for every possible linear, additive combination of covariates.
 We chose the best model as the one with the lowest BIC.
-Randomized quantile residual plots modified for mixture models were used to verify model assumptions [@dunstan2013].
-To obtain estimates of uncertainty for model parameters (confidence intervals), we used the default settings of `ecomix::regional_mix.bootstrap()`, which applies Bayesian bootstrap 1000 times.
+Visual inspection of randomized quantile residual plots modified for mixture models was used to verify model assumptions [@dunstan2013].
+To obtain estimates of uncertainty for model parameters (confidence intervals), we used the default settings of 'ecomix::regional_mix.bootstrap()', which applies a Bayesian bootstrap 1000 times.
 A custom code was written to generate partial plots for exploring the influence of covariates on RCP membership.
@@
 These variables were then averaged within 1° latitude $\times$ 1° longitude grid cells, by season, between 9--47°S and 141--161°E.
-Worth noticing that climatic EKEmean and SSTgrad were processed beforehand as they may provide persistent oceanographic features (described above) and thus represent the same value as EKE and SSTgrad.
+Note that 'clim_eke_mean' and 'clim_sst_grad' were processed beforehand as they may characterise persistent oceanographic features (described above) and thus represent the same value as 'eke' and 'sst_grad'.
 Data for prediction were scaled and centred using the same coefficients from the modelling dataset.
 
-Additionally, to investigate if the amount of information used to run the RCP models was representative, we ran rarefaction curves based on the P/A matrix and the assigned RCP P/A group for each sample (grid).
-Rarefaction curves were fit using the `iNEXT` package [@chao2014; @hsieh2016], and results are presented based on alpha diversity (i.e.
+Additionally, to investigate if the amount of information used to run the RCP models was representative, we ran rarefaction curves based on the presence-absence matrix and the assigned RCP group for each sample (grid cell).
+Rarefaction curves were fit using the 'iNEXT' package, and results are presented based on alpha diversity (i.e.
 species richness).
@@
 
-A total of 10,261 seabird records were obtained.
-They summed to 142,646 individuals from 80 species, belonging to 11 families and five orders.
-The five most abundant species summed 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*), whereas the five most frequent species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig.
+A total of 10,261 seabird records were obtained during the 15 voyages, comprising 142,646 individuals from 80 species, belonging to 11 families and five orders.
+The five most abundant species made up 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*).
+The five most frequently encountered species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig.
 S1**).
-An increase in species richness towards the south was clear in all but the winter season (**Fig S2**).
+An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig S2**).
 Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig.
 S2**).
-However, one cell in the Coral Sea summed to numbers of the same order of magnitude as off Tasmania in summer.
+However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
 
@@
 
-All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3**).
-The best fitting models generally had 1--2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**).
-All the best models selected SST or SSS, variables essential for characterising water masses.
-The best models show a good fit with no apparent deviances in assumptions (**Fig.
+For all models two RCPs was the optimal number of groups, except for presence-absence-autumn, which indicated three RCPs as the best fit (**Fig S3**).
+The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**).
+The best models showed a good fit with no apparent deviations from the assumptions (**Fig.
 S4**).
+All the best models selected 'sst' or 'sss', variables essential for characterising water masses.
+The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21°C, or 'sss' 35.1 PSU for summer (**Fig.
+4**, **Fig.
+S6**), similar to EAC's temperature and salinity thresholds.
+Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa (**Fig.
+S6**).
+Interestingly, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) were selected only in autumn and spring abundance models.
 
@@
   ## Table title
-  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli error distribution; abundance, based on a Negative Binomial error distribution)"))
+  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli error distribution; abundance, based on a Negative Binomial error distribution).
+Refer to Table 1 for the environmental data acronyms."))
 
@@
 
+```{r fig4-partial-plots, fig.cap = "**FIGURE 4.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S6).
+Refer to Table 1 for the environmental data acronyms.", fig.align = "center", echo = FALSE}
+
+knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
+```
+
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic change around \~34°S, with both P/A and abundance models pointing to similar patterns (**Fig.
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic change around \~34°S, with both presence-absence and abundance models pointing to similar patterns (**Fig.
 2**).
@@
 S5**), showing that around \~34°S the 'northern' RCP tends to have higher probabilities in coastal regions, whereas the 'southern' RCP spreads northward offshore.
-The P/A-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig.
+The presence-absence-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig.
 2**, **Fig.
@@
 The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.
-Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S5).", fig.align = "center", echo = FALSE}
 
@@
 
-*Relationship with environmental covariates (highlight SST) --**waiting for 'partial plots' (Fig. 4, Fig. S6)***
-
-```{r fig4-partial-plots, fig.cap = "**FIGURE 4.**  Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
-
-knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
-```
-
-The broad pattern of two large-scale assemblages is reflected in their species composition.
-Although there are overlaps between groups, both P/A and abundance models captured typical tropical species in the northern RCP (e.g.
+The broad pattern of two large-scale assemblages is reflected in their species compositions.
+Although there are overlaps between groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.
 boobies *Sula* spp., lesser frigatebird *Fregata ariel*) and typical Southern Ocean species in the southern RCP (e.g.
-mollymawks *Thalassarche* spp., fairy prion *Pachyptila turtur*) with higher occurrence/number probabilities in their usual habitat (**Fig.
+mollymawks *Thalassarche* spp., fairy prion *Pachyptila turtur*) (**Fig.
 5**, **Fig.
 S7**).
-The P/A-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
+The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
 However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross *Thalassarche cauta* and Australasian gannet *Morus serrator* standing out with high probabilities of occurrence.
 
-```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges (Region of Common Profile; RCP) off eastern Australia for spring presence-absence (a) and abundance (b) models.
-Values are the average and confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
-Values in (b) were log10-transformed to accomodate the high variation between species.
+```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
+Values are the average and 95% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
+Values in (b) were log10-transformed to accommodate the high variation between species.
 Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
@@
 
-This study shows that seabird assemblages off eastern Australia relate to oceanography and how the East Australian Current (EAC) may be playing a key role in structuring seabird distribution at a large scale.
+This study shows that seabird assemblages off eastern Australia are correlated with oceanographic conditions, and that the East Australian Current (EAC) may be playing a key role in structuring seabird distribution at a large scale.
 Our results are based on a cutting-edge multivariate statistical model [@hill2020], which brings advantages over classical methods for describing species assemblages.
-For instance, it includes uncertainty around model parameters that allow representing the fuzziness between biogeographical boundaries in a probabilistic framework, which can be monitored given the noticeable anomalies in EAC due to climate change [@phillips2022; @morrone2023biogeographical].
-Our results draw an important baseline for seabird assemblages on a world hotspot for this group.
-**Furthermore, data and code are available to allow reproducibility and future applications to verify spatial and temporal persistence of the described assemblages.**
+For instance, RCP models allow estimating uncertainty around the point probability of a sample (grid cell) belonging to an assemblage, thus representing the fuzziness between bioregion boundaries.
+These transition zones between bioregions are essential for monitoring biological changes due to environmental changes [@morrone2023biogeographical].
+As such, the boundary between the two consistently identified bioregions in our study may be a crucial region to monitor, given noticeable anomalies in the EAC at this same region [@phillips2022].
+Our results draw an important baseline for seabird assemblages in a world hotspot for this group.
+
+## Seabird bioregions off eastern Australia
+
+The relationship between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997].
+Therefore, it is not surprising that these two assemblages emerged off eastern Australia.
+However, the spatial consistency of these patterns throughout seasons is remarkable.
+Resident species seem to characterise these assemblages.
+The shy albatross [@hedd2001; @mason2018] and the fairy prions [@fromant2022] in the southern assemblage, and boobies [@bunce2015; @miller2018] and the tropical wedge-tailed shearwater *Ardenna pacifica* [@weimerskirch2020; @mcduie2015] in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species.
+Given that changes in assemblages over time are tipically driven by a few species [@gotelli2022], and that those species are likely residents [@souza2023], the relationship between resident seabirds and the EAC system should be closely monitored.
+For instance, coastal seabirds such as the little penguin *Eudyptula minor* are signalling the effects climate change might have in changing seabird behaviour and decreasing foraging success [@carroll2016], which ultimately influence their breeding success and population dynamics [@johnson2021].
+Therefore, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring seabird responses to the fast-changing EAC.
+
+The seasonal models also captured the influence of migratory species, particularly in the southern assemblage.
+For instance, summer and autumn models reflected the high abundance of the short-tailed shearwater *Ardenna tenuirostris* before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
+Given the geographic location of the southern assemblage, however, we believe subantarctic taxa were not well represented in our models due to their low number of occurrences (**Fig.
+S1 - FO/nOCC/FN**; hence they were removed before modelling).
+More intensive surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.
+albatrosses *Diomedea* spp.
+and *Phoebetria* spp.), given that more effort translates as higher likelihood of recording these species.
+In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the presence-absence-autumn model) in all but the summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**.
+
+Our analysis clearly revealed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia.
+There is a noticeable link with EAC dynamics; the boundaries of RCPs occurred where the EAC turns into its 'eastern extension' and the eddy field spawns [@oke2019].
+However, the 'northern/southern' assemblage pattern does not agree with popular, large scale regionalization proposals, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007], or with the Large Marine Ecosystems [LME; @sherman1994].
+The MEOW scheme was designed using coastal and shelf biota and was meant to represent habitats up to the 200-m isobath, which clearly does not describe the entirety of seabird habitat.
+The LME scheme is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management areas and thus a 'political' component to it, which does not necessarily align with distribution of seabirds.
+@longhurst2007 designated only one biogeochemical province for east Australia based on productivity patterns, which may work well at larger scales but seems oversimplified specifically for this region.
+
+The described pattern does however agree with the marine biogeographic units proposed by @costello2017, who identified a boundary at \~34°S between their 'Coral Sea' and 'South Australia' realms.
+This approach used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus encompassing more complexity of the biogeographic boundaries in the marine environment.
+Other biological data also suggest 'northern/southern' assemblages off eastern Australia, including tuna and billfish [also highly mobile top predators; @reygondeau2012], mesopelagic fauna [which includes seabird prey; @sutton2017a] and deep-sea invertebrates [@ohara2020] and fish [@ohara2020a].
+All these studies focused mainly on describing biological assemblages, as opposed to classifying environment.
+It is worth noting that the Longhurst Provinces approach was updated to include SST, SSS and bathymetry (the same variables retained by our best RCP models) and a temporal component to the classification [@reygondeau2013].
+The updated provinces are more similar to the seabird assemblages described by us, although still not a perfect match.
+Australia's bioregionalization has a long story of disagreements [@ebach2012history], however, the inherent link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
+
+## Fitting RCP models to high mobile animals
+
+To the best of our knowledge, this is the first study applying RCPs to highly mobile animals such as seabirds.
+RCP models have been previously applied to plants [@lyons2017], and relatively sedentary animals, such as macrobenthic organisms [@pantiukhin2021] and demersal fish [@hill2017].
+Pelagic seabirds however can have daily movements of hundreds of kilometres [e.g.
+@mcduie2015], during which they traverse a variety of habitats and environmental gradients.
+As such, capturing the distribution of highly mobile species, and their relationships with environmental parameters is not a trivial task, and likely much harder than for sessile organisms.
+
+While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009], they fail to perceive how groups of species (assemblages) relate to the environment and each other on a shared seascape.
+Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
+Furthermore, the spatial scale on which assemblages relate to their environment is also a field of debate [@hunt1987; @ribic1997; @scales2017].
+Understanding the biology of the studied group is thus essential to set the appropriate scale of analysis.
+Our choice of 1° latitude $\times$ 1° longitude was mainly influenced by data availability and the trade-off between number of species per grid/season (see also \@ref(limitations)).
+Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere [@harris2007; @santora2013; @sojitra2022].
+Therefore, for describing seabird assemblages at a large scale, we believe the chosen resolution was a good fit, and the models described expected patterns of species composition.
+
+RCP models based on presence-absence and abundance data could complement each other.
+Previous studies using RCPs have opted to use only one type of response data.
+@lyons2017, @hill2017, and @pantiukhin2021 used presence-absence data, whereas @lee2019good and @receveur2020 used abundance data.
+The added information contained in abundance datasets make them inherently more powerful than presence-absence data.
+However, the use of different data inputs in multivariate, biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
+At larger scales, presence-absence datasets describe similar patterns to abundance datasets [@anderson2005], as we found in our results.
+In contrast, at smaller scales, abundance datasets may highlight local conditions and therefore reveal hidden patterns [@marchant1990; @anderson2005; @melo2005; @waters2008marine].
+If combined with ecological knowledge of a particular group, the information contained in both presence-absence and abundance datasets could be leverage to better interpret the resulting patterns.
+For example, seabirds will likely occur in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance is related to fine scale processes and prey availability [@hunt1987; @weimerskirch2007].
+Therefore, future studies assessing bioregions of highly mobile animals could describe the make-up of assemblages using smaller resolution grid cells and presence-absence models.
+Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore 'new bioregions') likely nested within the bioregions described at large scale.
 
@@
 Although the results suggest consistent patterns, we should point out some caveats of our analyses.
-Seabird data were gathered opportunistically while taking part on several projects, whose sampling design were not planned to account for seabird spatial ecology specifically.
-As a result, there are spatial discontinuities on the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors more robustly.
-Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution to date [@mott2018; @bernard2021].
-Dedicated seabird effort could thus benefit future studies advancing seabird research in this region, as well as improving patterns described here.
+Seabird data were gathered during multi-disciplinary cruises, whose sampling designs were not specifically planned to account for spatial ecology of seabirds.
+As a result, there are spatial discontinuities in the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors.
+Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution in the region [@mott2018; @bernard2021].
+Dedicated seabird surveys would help to advance seabird research in this region, as well as improving the patterns described here.
 By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at a finer scale.
-In addition, ideally, we could have had more information within each grid cell by season (our sampling units).
-Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8**).
-
-## Seabird bioregions off eastern Australia
-
-Spatial patterns clearly showed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia.
-There is a noticeable link with EAC dynamics, although the weak statistical relationship between RCPs and environmental covariates---RCPs boundaries are closely related with where the EAC turns into its 'eastern extension' and the eddy field spawns [@oke2019].
-However, the 'northern/southern' assemblage pattern does not agree with popular, large scale regionalization proposals, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007] province's and ecoregion's, and neither with the Large Marine Ecosystems [LME; @sherman1994].
-The MEOW scheme was designed using coastal and shelf biota and it was meant to represent habitats up to the 200-m isobath, which clearly is not representative enough of seabirds' environment.
-The LME scheme is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management areas and thus a 'political' component on it, which seabirds do not seem to agree with.
-@longhurst2007 designated only one biogeochemical province for east Australia based on productivity patterns, which may work well at larger scales but seems oversimplified specifically for this region.
-
-The described pattern, on the other hand, agree with the marine biogeographic units proposed by @costello2017, who points out a boundary at \~34°S between their 'Coral Sea' and 'South Australia' realms.
-The latter regionalization used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus encompassing more complexity of the biogeographic boundaries in the marine environment.
-Other biological data also suggests 'northern/southern' assemblages off eastern Australia, including tuna and billfish [also highly mobile top predators; @reygondeau2012], mesopelagic fauna [which includes seabird prey; @sutton2017a] and deep-sea invertebrates [@ohara2020] and fish [@ohara2020a].
-All these studies focused mainly on describing biological assemblages, whereas the ones mentioned previously focused on the environment.
-Worth mentioning, the Longhurst' provinces had gone through an update including SST, SSS, bathymetry (the same variables kept by our selected best RCP models) and a temporal component to its equation [@reygondeau2013].
-The updated provinces are easily comparable to the seabird assemblages described by us, although still not a perfect match.
-Australia's bioregionalization has a long story of disagreements [@ebach2012history], however, the inherently link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
-
-The relationship between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997].
-Therefore, it is not surprising that at the large scale these two assemblages emerged off eastern Australia.
-However, the spatial consistency of these patterns throughout seasons is remarkable.
-Resident species seem to characterise these assemblages.
-The shy albatross [@hedd2001; @mason2018] and the fairy prions [@fromant2022] in the southern assemblage, and boobies [@bunce2015; @miller2018] and the tropical wedge-tailed shearwater *Ardenna pacifica* [@weimerskirch2020; @mcduie2015] in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species.
-Given that assemblage changes over time are driven by few species [@gotelli2022] and that those species are likely residents [@souza2023], the relationship between resident seabirds and the EAC system should be closely monitored.
-For instance, coastal seabirds such as the little penguin *Eudyptula minor* are signalling the effects climate change might have in changing seabird behaviour and decreasing foraging success [@carroll2016], which ultimately influence their breeding success and population dynamics [@johnson2021].
-Therefore, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring seabird responses to the fast-changing EAC.
-
-The seasonal models also captured the influence of migratory species, particularly in the southern assemblage.
-For instance, summer and autumn models reflected the abundant presence of the short-tailed shearwater *Ardenna tenuirostris* before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
-Given the geographic location of the southern assemblage, however, we believe subantarctic taxa were not well represented in our models due to their low number of occurrences (**Fig.
-S1 - FO/nOCC/FN**; hence they were removed before modelling).
-If increased, dedicated seabird effort is carried out, the data could reflect the seasonal northward pulse of these taxa into the southern assemblage (e.g.
-albatrosses *Diomedea* spp.
-and *Phoebetria* spp.), given that more effort translates as higher likelihood of recording these species.
-In turn, including these taxa onto the models could point towards a more stable 'three assemblage pattern' (as revealed by the P/A-autumn model) in all but summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**, instead of the 'southern/northern' assemblages described here.
-
-## Fitting RCP models to high mobile animals
-
-To the best of our knowledge, this is the first study applying RCPs on high mobile animals such as seabirds.
-RCP models have been previously applied to plants [@lyons2017], and sessile and **avg?** mobile animals, such as macrobenthic organisms [@pantiukhin2021] and demersal fish [@hill2017].
-Differences in species energy and physiological requirements can led animals to using larger areas given their needs and, commonly, pelagic seabirds have daily movements of about hundreds of kilometres per day [e.g.
-@mcduie2015].
-During their movements, seabirds can wander through a variety of habitats and gradients including, for instance, shallow/deep water and poor/productive waters.
-As such, capturing high mobile species distribution and their relationship with environmental parameters is not a trivial task, and likely harder than for sessile organisms.
-
-While remote tracking studies are advancing our understanding on how seabirds use their environment at fine-, individual-scale level [@tremblay2009], they fail to perceive how groups of species (assemblages) relate to the environment and each other on a shared seascape.
-Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
-On the other hand, the spatial scale on which assemblages relate to their environment is also a field of debate [@hunt1987; @ribic1997; @scales2017].
-Understanding the biology of the studied group is thus essential to set the appropriate scale of analysis.
-Our choice of 1° latitude $\times$ 1° longitude was mainly influenced by data availability and the trade-off between number of species per grid/season (see also \@ref(limitations)); nonetheless, the chosen spatial scale is similar with seabird distribution analyses elsewhere [@harris2007; @santora2013; @sojitra2022].
-Therefore, for describing seabird assemblages at a large scale, we believe the chosen spatial scale was a good fit, and models described expected patterns of species composition.
-
-RCP models based on presence-absence and abundance (count) data could complement each other.
-Studies that previously applied RCPs opted by one or another type of data as input.
-@lyons2017, @hill2017, and @pantiukhin2021 used presence-absence data to run their models, whereas @lee2019good and @receveur2020 used abundance data.
-Information contained in abundance datasets are inherently better than simplified presence-absence ones.
-However, the use of different data input in multivariate, biodiversity assessments and assemblage descriptions have shown contrasting patterns at different spatial scales.
-At larger scales, presence-absence data to equate (relative) abundance in terms of patterns described, while at smaller scales abundance data may highlight local conditions and therefore reveal different patterns, within the broad picture [@marchant1990; @anderson2005; @melo2005; @waters2008marine].
-If tied up with biological/ecological knowledge of a particular group, the information contained in both presence-absence and abundance datasets could be leverage to better interpret the resulted patterns.
-For example, seabirds will likely occur (*presence-absence data*) in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance (*count/abundance data*) are related to fine scale processes and prey availability [@hunt1987; @weimerskirch2007].
-As such, the make-up of assemblages can be described through P/A models and its species probability of occurrence within RCPs at large scales, whereas the relationship with environmental variables at finer scales could improve our understanding on how seabirds relate to their environment and where feeding/aggregation hotspots might be, likely nested within the large scale bioregions.
+In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
+Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Fig S8**).
 
@@
 We applied a robust, repeatable method to assess seabird assemblages off eastern Australia for the first time.
-The models point toward a consistent pattern, throughout seasons, of a 'northern' and a 'southern' assemblage, where the spatial division clearly links with the EAC behaviour and SST seems to be the key variable discriminating these groups.
-As the EAC is predicted to strengthen [@poloczanska2007] as well as extreme weather events such as marine heatwaves to happen more often [@oliver2018], seabirds will likely respond to these changes.
+The models point toward a consistent pattern of a 'northern' and a 'southern' assemblage, where the spatial division is clearly linked with the behaviour of the EAC and SST seems to be the key variable discriminating these groups.
+As the EAC is predicted to strengthen [@poloczanska2007], and extreme weather events such as marine heatwaves to happen more often [@oliver2018], seabirds will likely respond to these changes.
 Southern ocean albatrosses off southeast Australia, for instance, have shown 'early stages of a southward range shift' due to environmental variability [@sojitra2022].
-Resident species such as the little penguins and wedge-tailed shearwater are changing their behaviour and decrease their foraging success when SST is higher [@carroll2016; @peck2004].
-Given that eastern Australia is a biodiversity and abundance hotspot for seabirds, it is imperative to set a monitoring program on which the assemblages and species profiles presented here could be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
-In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the EAC eastern extension/eddy-spawning area, may be fundamental to effectively monitoring EAC strengthening and its impact on the marine biota.
+Resident species such as the little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher [@carroll2016; @peck2004].
+Given that eastern Australia is a biodiversity and abundance hotspot for seabirds, it is imperative establish a monitoring program allowing the assemblages and species profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
+In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
 

### 2023-07-11 docs/ms-draft.Rmd
small adjustments on fig numbering and I've sent this version off to coauthors

@@
 In this study, we aim to identify assemblages of seabird species probabilistically and relate them to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
-We used the mixture-model Region of Common Profiles [RCP; @foster2013], regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
+We used the mixture model Region of Common Profiles [RCP; @foster2013], regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
 RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
@@
 
-Spatial and temporal persistence of features such as eddy kinetic energy (eke) and thermal fronts (sea surface temperature gradient; sst_grad) may influence the movement and distribution of seabirds by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
+Spatial and temporal persistence of features such as eddy kinetic energy (eke) and thermal fronts (sea surface temperature gradient; 'sst_grad') may influence the movement and distribution of seabirds by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
 Therefore, before extracting environmental data for each seabird record, 'eke' and 'sst_grad' environmental layers were calculated.
@@
 The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21°C, or 'sss' 35.1 PSU for summer (**Fig.
-4**, **Fig.
-S6**), similar to EAC's temperature and salinity thresholds.
+2**, **Fig.
+S5**), similar to EAC's temperature and salinity thresholds.
 Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa (**Fig.
-S6**).
+S5**).
 Interestingly, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) were selected only in autumn and spring abundance models.
@@
 
-```{r fig4-partial-plots, fig.cap = "**FIGURE 4.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+```{r fig4-partial-plots, fig.cap = "**FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
 Results for other seasons and models can be found in the Supplementary Material (Fig.
-S6).
+S5).
 Refer to Table 1 for the environmental data acronyms.", fig.align = "center", echo = FALSE}
@@
 When predicting RCP onto space, patterns of point-predictions suggest a biogeographic change around \~34°S, with both presence-absence and abundance models pointing to similar patterns (**Fig.
-2**).
+3**).
 These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.
+4**, **Fig.
+S6**), showing that around \~34°S the 'northern' RCP tends to have higher probabilities in coastal regions, whereas the 'southern' RCP spreads northward offshore.
+The presence-absence-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig.
 3**, **Fig.
-S5**), showing that around \~34°S the 'northern' RCP tends to have higher probabilities in coastal regions, whereas the 'southern' RCP spreads northward offshore.
-The presence-absence-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig.
-2**, **Fig.
-S5**).
+S6**).
 
-```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+```{r fig2-point-pred, fig.cap = "**FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
 Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}
@@
 
-```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+```{r fig3-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
 The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.
 Results for other seasons and models can be found in the Supplementary Material (Fig.
-S5).", fig.align = "center", echo = FALSE}
+S6).", fig.align = "center", echo = FALSE}
 
@@
 Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S7).", fig.align = "center", echo = FALSE}
 
@@
 and *Phoebetria* spp.), given that more effort translates as higher likelihood of recording these species.
-In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the presence-absence-autumn model) in all but the summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**.
+In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the autumn presence-absence model) in all but the summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**.
 

### 2023-07-26 docs/ms-draft.Rmd
coauthors ms review

@@
 ---
-title: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
+title: "Seabird assemblages are linked to the major western boundary current off eastern Australia"
 author:
@@
 abstract: |
-  **Aim**: To identify assemblages of seabird species, probabilistically, and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
-  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
-RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different among them.
-These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic layers.
-  **Results**: All models suggested two large scale assemblages ('northern' and 'southern'), except for the autumn presence-absence model which indicated three groups.
-Model results seem to agree in a biogeographic barrier at ~34°S, where the East Australian Current dissipates as it flows eastward and loses force.
-Moreover, sea surface temperature or sea surface salinity, were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
+  **Aim**: To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+  **Methods**: We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
+These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
+  **Results**: Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
 Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
-  **Conclusion**: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and large spatial scales.
-Given the rapid change that the East Australian Current is experiencing, and its importance in driving distributions of seabirds, the methods applied in our study could be replicated in future analyses to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+  **Conclusion**: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 output: 
@@
 Bioregionalization is the discipline that identifies, delimits, characterises, and names biogeographical regions [@vilhena2015; @morrone2018].
-Although there is often no agreement on the naming systems between proposed regionalizations [but see @ebach2008], the hierarchical nature of bioregions is well recognized [@vilhena2015; @kreft2010].
+Although there is often no agreement on the naming systems among proposed regionalization schema [but see @ebach2008], the hierarchical nature of bioregions is well recognized [@vilhena2015; @kreft2010].
 Thus, bioregions are usually described within geographically larger classifications, such as 'provinces' or 'realms' [@spalding2007; @costello2017].
@@
 In this sense, bioregions are interpreted as a natural representations of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010].
-In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space as well as relating these regions to environmental characteristics.
+In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space, and relating these regions to environmental characteristics.
 By doing so, bioregions can be interpreted as species assemblages [@lyons2017; @hill2017].
@@
 
-Seabird assemblages have been widely described and related to environmental characteristics.
-Considering the hierarchical nature of oceanographic processes and their resulting features [@haury1978patterns], it is not surprising that the distribution of highly mobile seabirds are linked differently across scales [@hunt1987].
-In the marine environment, large scale oceanographic features (e.g.
-water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e.
+Seabird assemblages have been widely described and related to environmental characteristics [@ribic1997; @woehler2003; @hyrenbach2007].
+Considering the hierarchical nature of oceanographic processes and their resulting features [@haury1978patterns], it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales [@haury1978patterns; @hunt1987].
+In the marine environment, large scale (\>100 km length scale) oceanographic features (e.g.
+boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.
 the occurrence of species and their relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014].
-In contrast, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014].
-Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces, and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013].
-Noticeably, sea surface temperature is a key driver of seabird distribution, at individual- and assemblage-levels, operating over a range of scales from large (e.g.
-water masses) to small (e.g.
+In contrast, local abundance is related to meso- (10--100 km) and sub-mesoscale (1--10 km) processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014].
+Such processes, although relatively ephemeral, are primarily driven by physical and atmospheric forces, and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013].
+Sea surface temperatures are a key driver of seabird distributions, at individual- and assemblage-levels, operating over a range of spatial and temporal scales from macro (e.g.
+water masses) to fine (e.g.
 fronts) [@dunlop1988; @ribic1997; @quillfeldt2015].
-Nonetheless, many regions still lack information regarding distribution of seabird species and how they assemble in space and time.
+Despite extensive surveys [e.g.
+@ballance1997] and tracking studies [e.g.
+@carneiro2020], many oceanic regions still lack information on the distributions and abundances of seabird species, and how they assemble in space and time.
 
-One area that could benefit from better understanding of seabird distributions is Australia, considered a world seabird hotspot [@karpouzi2007; @ramírez2017].
-Given the region's importance for seabird diversity, there are surprisingly few studies on seabird distribution [@mott2018; @bernard2021].
-Off northwest Australia, at-sea surveys, multivariate analyses [@dunlop1988] and ensemble species distribution models of species assemblages [@lavers2014] have shown that habitat use by tropical species is related to sea surface temperature and salinity.
-In particular, @dunlop1988 suggested that salinity is the main variable influencing species distribution in oceanic waters whereas temperature is the key factor on the continental shelf.
-In contrast, there are few studies at large scales off eastern Australia.
-@reid2000 and @evans2021a analysed seabird assemblages and their relationships with oceanography and prey data at broad scales (tens of km), off southeast Tasmania.
-The remaining studies are mostly descriptive accounts at local scales [e.g.
-@norman1992distribution; @blaber1994distribution], or focused on a single species [e.g.
-@priddel2014].
+One area in need of better understanding of seabird distributions is Australia, considered a global seabird hotspot [@karpouzi2007; @ramírez2017].
+Given the region's importance for seabird diversity, there are surprisingly few studies on seabird distributions at sea [@mott2018; @bernard2021].
+Off northwest Australia, at-sea surveys, multivariate analyses [@dunlop1988] and ensemble species distribution models of species assemblages [@lavers2014] showed that habitat use by tropical seabird species was related to sea surface temperatures and salinity.
+@dunlop1988 suggested that salinity was the main variable influencing species' distributions in oceanic waters while temperature was the key factor for species on the continental shelf.
+In contrast, there are few studies at large spatial and temporal scales off eastern Australia.
+@reid2000 and @evans2021a analysed seabird assemblages and their relationships with oceanographic and prey data at coarse scales (tens of km) off southeast Tasmania.
+The remaining studies are mostly descriptive accounts locally [@norman1992distribution; @blaber1994distribution], or focused on a single species [@priddel2014].
 
-The East Australian Current (EAC) is the major western boundary current in the South Pacific, and the dominant current in eastern Australia [@ganachaud2014].
-The EAC has been extensively studied, although its link between oceanography and biological processes is still in its infancy.
-The dynamic features caused by the EAC, such as eddies and upwellings, are likely to influence marine biota [@suthers2011; @scales2014].
-While most of the biological studies done in this system were focused on plankton entrapment within the cores of eddies [@garcia2022; @firme2023; @hassler2011], there is a growing body of work focusing on fisheries [e.g.
-@hobday2010; @young2011; @brieva2015].
-These studies emphasize how the probability of occurrence of a given taxon, and the structure of species assemblages, is driven by the EAC.
-Notwithstanding, the relationship between oceanography and marine megafauna along the EAC is poorly known [@suthers2011].
-Some progress has been made using shark and ray movement data [@jaine2014; @niella2020; @niella2022], and satellite tracking has also demonstrated that increased sea surface temperature in EAC decreases penguin foraging success [@carroll2016].
-However, information at the assemblage-level is still scarce [but see @evans2021a].
+The East Australian Current (EAC) is the major western boundary current in the South Pacific Ocean, and the dominant boundary current off eastern Australia [@ganachaud2014].
+The EAC has been extensively studied, although research into the links between the EAC's oceanography and marine predators are not well known.
+The dynamic features associated with the EAC, such as mesoscale eddies and upwelling, are likely to influence marine biota [@suthers2011; @scales2014].
+While most of the biological studies in this system were focused on plankton entrapment within the cores of eddies [@garcia2022; @firme2023; @hassler2011], there is a growing body of work focusing on fisheries [@hobday2010; @young2011; @brieva2015].
+These studies emphasized how the probability of occurrence of a given taxon, and the structure of species assemblages, were driven by the EAC.
+However, the relationships between oceanography and marine megafauna within the EAC are poorly known [@suthers2011].
+Some progress has been made using shark and ray movement data [@jaine2014; @niella2020; @niella2022], and satellite tracking has also demonstrated that increased sea surface temperatures in the EAC decreases little penguin *Eudyptula minor* foraging success [@carroll2016], but information at the assemblage-level for seabirds in the EAC is absent.
 
@@
 As a result, climate change is rearranging the structure of species assemblages around the globe [@poloczanska2013; @antão2020; @dornelas2014].
-Top predators, for instance, may change their core habitats [@hazen2013a; @reisinger2022], with unprecedented population consequences.
-For seabirds, @péron2010 showed that sea surface temperature anomalies affected the occurrence and abundance of albatrosses and petrels species along a latitudinal range in the Indian Ocean, thus affecting assemblage structure.
-In eastern Australia, changing EAC conditions have affected a range of species.
-Changes in zooplankton assemblage structure and fish distribution patterns have been recorded, and shark distribution shifts are expected to occur [@johnson2011; @niella2022].
-One of the strongest marine heatwave ever recorded, in the Tasman Sea, showed drastic environmental impacts [@oliver2017].
-Therefore, given that the EAC is strengthening and getting warmer [@cai2005; @phillips2022], biological changes are expected to occur [@poloczanska2007; @hobday2010].
+Top predators, for instance, may change their core habitats [@hazen2013a; @reisinger2022], with unprecedented population consequences [@woehler2023].
+For seabirds, @péron2010 and @sojitra2022 showed that sea surface temperature anomalies affected the occurrence and abundances of albatrosses and petrels species with latitude, thus affecting assemblage structure.
+In eastern Australia, changing EAC conditions have affected a range of marine species.
+For example, changes in zooplankton assemblage structure and fish distribution patterns have been recorded and attributed to warming water temperatures, and shifts in shark distributions are expected [@johnson2011; @niella2022].
+A severe marine heatwave driven by a poleward strengthening of the EAC, showed drastic environmental impacts including disease outbreaks in aquaculture farms and wildlife mortalities [@oliver2017].
+Given that the EAC is strengthening and extending farther southward and getting warmer [@cai2005; @phillips2022], biological changes are expected to occur [@poloczanska2007; @hobday2010].
 
-Given east Australia's importance for seabirds and the recent attention on how climate change in the EAC can affect marine biota [@poloczanska2007], an assessment of the spatial distribution of seabirds at large scales is timely.
+In light of eastern Australia's importance for seabirds, and the recent attention on how predicted climate change will influence the EAC's marine biota [@poloczanska2007], an assessment of the contemporary spatial distributions of seabirds within the EAC at large temporal and spatial scales is timely.
 Seabirds can serve as sentinels for tracking ecological changes in the EAC in many ways [e.g.
 @sydeman2012].
-Surprisingly, though, seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic [@grémillet2009; @sydeman2012].
-In this study, we aim to identify assemblages of seabird species probabilistically and relate them to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
-We used the mixture model Region of Common Profiles [RCP; @foster2013], regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
+Surprisingly though, seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic [@grémillet2009; @sydeman2012].
+
+In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
+We use the Region of Common Profiles [RCP; @foster2013], a mixture model regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
 RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
-East Australia supports a high diversity of seabirds and offers wide environmental gradients, from the tropics to the subantarctic.
-Therefore, we expect to find different assemblages relating to different environmental conditions.
-In addition, we run RCP models for presence-absence and abundance as input data, to test if the different numeric resolution would result in similar assemblage patterns.
-Furthermore, we provide a repeatable framework on which future studies can build to assess how changes in the EAC will influence the distribution of seabird assemblages.
+Eastern Australia supports a high diversity of seabirds and offers wide environmental gradients over a near 40° latitudinal range, from the tropics to the cool temperate (**ref?**) We hypothesised to find different seabird assemblages relating to different environmental conditions within the EAC.
+In addition, presence-absence and abundance data were used to test if the different numerical resolutions would result in the identification of similar assemblages.
+This study provides a repeatable framework on which future studies can compare to assess how future changes in the EAC may influence the distribution of seabird assemblages.
 
@@
 
-The EAC originates from the South Equatorial Current (SEC), a tropical, nutrient-poor water.
-When the SEC's North Caledonian Jet hits the Queensland Plateau (\~18°S), it splits into a north branch running towards the Gulf of Papua and a south branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
+The EAC originates from the South Equatorial Current (SEC), a tropical nutrient-poor current.
+When the SEC's North Caledonian Jet meets the Queensland Plateau (\~18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
 The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.
@@
 **SST/SSS characteristics?
-ROB?** The EAC interacts with the topography of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
-The EAC is mainly attached to the shelf break, but at \~32°S it spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
+ROB?** The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
+The EAC flow polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
 Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.
-1a**), possibly carrying eddy-like features as a transient current [@oke2019].
+1a**), carrying eddy-like features as a transient current [@oke2019].
 
@@
 
-Seabird data were gathered aboard RV Investigator and RV Falkor during 15 voyages, between 2016--2021, over \~37° of latitude (**Fig.
+Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.
 1b**; **Table S1**).
-One observer continuously recorded seabirds within a 300-m forward quadrant (eye height **18-m, Eric?** a.s.l) throughout daylight hours while the ship was underway at a constant speed.
-The species, the number of birds per group, and behaviour were logged directly into a data entry system through a ship-based web server.
-Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species) **(Eric: is there any ref for the protocol?)**.
-Birds following the boat were previously excluded from analyses.
+One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
+The species, the numbers of birds, and their behaviours were logged in real time into a ship-based web server.
+**Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.** Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species).
+Birds following the vessels were excluded from analyses.
+Seabird observation protocols followed @raymond2010, in use around Australia since 1980.
 
@@
 
-To explain and predict species distribution and biogeography, a thoughtful choice of covariates is important [@barbet-massin2014].
+To investigate and predict species' distributions and relationships with environmental variables, a considered choice of **broad-scale climatic and oceanographic** covariates is important [@barbet-massin2014].
 We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table 1**).
-These variables are the most used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of scales from large-scale water-mass properties (e.g.
-sea surface temperature) to prey availability in mesoscale oceanographic features (e.g.
-fronts and eddies; @weimerskirch2007; @scales2014).
-Seabird records were matched with the best open-access, remote-sensed or interpolated data available, and distance from the coast was calculated in R software, version 4.2.0 [@r2022] (see details below).
+These variables are the most frequently used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of spatial and temporal scales from macro-scale water-mass properties to prey availability in meso- to submesoscale oceanographic features [@haury1978patterns; @weimerskirch2007; @scales2014].
 
@@
 
-Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022], using packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021], 'iNEXT' 3.0.0 [@chao2014; @hsieh2016], and 'ecomix' 1.0.0 [@ecomix].
+Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material **XX**.
 Code and the modelling dataset (see \@ref(modellingdatasets)) are available in **Daudt et al.
@@
 
+<!-- using packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021].
+-->
+
 ### Environmental layers
 
-Spatial and temporal persistence of features such as eddy kinetic energy (eke) and thermal fronts (sea surface temperature gradient; 'sst_grad') may influence the movement and distribution of seabirds by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
-Therefore, before extracting environmental data for each seabird record, 'eke' and 'sst_grad' environmental layers were calculated.
-For both, 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table 1**) and calculations were done layer-by-layer to match seabird data on the finest scale possible.
-Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ (**ROB: Ref here?**).
+The spatial and temporal persistence of oceanographic features such as eddy kinetic energy ('eke') and thermal fronts (sea surface temperature gradient: 'sst_grad') may influence the movements and distributions of seabirds by offering a higher likelihood of finding food [@weimerskirch2007; @scales2014a].
+Thus, 'eke' and 'sst_grad' environmental layers were calculated before extracting environmental data for each seabird record.
+For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table 1**) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
+Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ [@sun2022].
 The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
-In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on 10-years of data (2011--2021) as a proxy for persistent oceanographic features, as well as 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table 1**).
-Seabed slope was calculated using 'raster::terrain()', and distance from the coast was calculated through 'sf::st_distance()' using Australia and Tasmania as the mainland (i.e.
+In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table 1**).
+Seabed slope was calculated using 'raster::terrain()' [@raster], and distance from the coast was calculated through 'sf::st_distance()' [@sf] using Australia and Tasmania as the mainland (i.e.
 ignoring smaller islands).
-Then, for each seabird record, the most proximate value for each environmental layer (original or calculated) was extracted using built-in functions from 'raster', 'terra', and 'sf' packages.
+Then, for each seabird record, the most proximate value for each environmental layer (original or calculated) was extracted using built-in functions from 'raster' 3.5-21, 'terra' 1.6-7, and 'sf' 1.0-8 packages [@sf; @raster; @terra].
 
-### Modelling datasets {#modellingdatasets}
+### Data {#modellingdatasets}
 
-Seabirds are highly mobile animals, and many species migrate during non-breeding periods [@biology2002].
+Seabirds are highly mobile animals, and many species migrate and/or disperse during their non-breeding periods [@biology2002].
 Therefore, seasonal differences in species composition were expected [e.g.
-@reid2023], as well as in probability of occurrence and total numbers.
-Thus, we build separate modelling dataset for each season.
+@reid2023], and in the probabilities of occurrence and in total numbers.
+Thus, we built separate modelling dataset for each season.
 We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (c.
@@
 1b**), by summing seabirds from the same species.
-Considering our aim to test if presence-absence and abundance (count) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
-Furthermore, to avoid numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12 species; autumn = 21; winter = 20; spring = 33).
-The associated environmental data of each seabird record were averaged within the same grid cells and seasons.
-Environmental data were scaled and centred before modelling.
+To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
+To avoid numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12 species; autumn = 21; winter = 20; spring = 33).
+The associated environmental data of each seabird record were averaged within the same grid cells and seasons, and environmental data were **standardised** before modelling.
 
@@
 
-We used the mixture model Region of Common Profiles (RCP) to identify and describe seabird assemblages.
+We used the mixture model Region of Common Profiles (RCP) to identify and describe seabird species assemblages.
 RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates response (multi-species data) and predictor (environmental data) [@woolley2020; @hill2020].
-The model defines 'regions' of similar environmental spaces in which the probability of observing a set of species 'profiles' is constant but different from other regions [@foster2013; @foster2017].
+Given a set of environmental variables, the model can predict assemblage probabilities of sites where no biotic (observational) data exist.
+As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals [@warton2015; @hill2020].
+
+The model is specified in terms of assemblages and site-specific probabilities.
+There are $K$ assemblages (i.e.
+RCPs) that are common across the study area, where $K = (1, ..., k)$.
+Depending on whether the model is specified in terms of abundance or presence-absence, the assemblages are defined in terms of their structure at each $i$th site (a vector $y_{i}$), i.e.
+frequency or relative frequency of the species, respectively.
+The assemblage structure is denoted by $y_{i} = (y_{i1}, ...
+y_{iS})$, where $S$ is the number of species.
+The model estimates the probability of belonging to the $K$ assemblage given the site's assemblage structure.
+Thus, each site within the study area has a $k$-dimensional probability vector that gives the probability of the corresponding assemblage for that site.
+RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) [@foster2013].
+However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the $K$ assemblage [@foster2013].
 As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
-RCP memberships are latent (unobserved) and vary as a function of environmental data, assuming sites within each RCP have constant species profiles.
-Given a set of environmental variables, the model also enables prediction of assemblage probabilities onto sites where no biotic data exist.
-Additionally, as a feature of model-based approaches, RCPs allow the opportunity to compare models through information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals [@warton2015; @hill2020].
 
-Before fitting models, covariates were checked for collinearity within each seasonal modelling dataset.
+Before fitting models, covariates were checked for collinearity within each seasonal dataset.
 If a pairwise collinearity of $\left|x\right| >0.7$ was found we removed one of the variables [@dormann2013], favouring direct measures over the 'climatic' layers we calculated (e.g.
@@
 
-Due to seabird migration, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
-Given the difference among seasons in the number of species that met our occurrence threshold (see \@ref(modellingdatasets)), we decided to run RCP models separately by season and data type (presence-absence and abundance).
+Due to seabird migration and dispersion, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
+In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \@ref(modellingdatasets)), we decided to use RCP models by season and data type (presence-absence and abundance).
 RCP models were used to relate the response variables (i.e.
@@
 averaged values of environmental variables in each cell).
-Presence-absence data were modelled based on a Bernoulli error distribution, whereas abundance data based on a Negative Binomial error distribution.
-RCPs were fitted using the 'ecomix' package [@ecomix], with the following workflow: (i) find the best number of groups that describes the data, (ii) re-fit the model by fixing the best number of groups and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data based on a Negative Binomial model.
+RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($k$) that describes the data, (ii) re-fit the model by fixing $k$ and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
 
-To identify the best number of groups that described the data, we ran full models (all covariates) with 100 random start numbers [see @foster2013] between 2--7 groups, and chose the number of RCPs that minimized the Bayesian Information Criterion (BIC).
-Then, we fixed the number of groups at the optimal number and fitted models for every possible linear, additive combination of covariates.
+To identify the best number of RCP groups ($K$) that described the data, we used full models (all covariates) with 100 random start numbers [see @foster2013] between $k = (2, ..., 7)$ groups, and chose the number of RCPs that minimized the Bayesian Information Criterion (BIC).
+Then, we fixed $K$ at the optimal number and fitted models for every possible linear, additive combination of covariates.
 We chose the best model as the one with the lowest BIC.
@@
 
-We built a seasonal 10-year (2011--2021) average layer for predicting model results for each environmental covariate.
+We built a seasonal 11-year (2011--2021) average layer for predicting model results for each environmental covariate.
 These variables were then averaged within 1° latitude $\times$ 1° longitude grid cells, by season, between 9--47°S and 141--161°E.
 Note that 'clim_eke_mean' and 'clim_sst_grad' were processed beforehand as they may characterise persistent oceanographic features (described above) and thus represent the same value as 'eke' and 'sst_grad'.
-Data for prediction were scaled and centred using the same coefficients from the modelling dataset.
+Data for prediction were **standardised** using the same mean and standard deviation from the modelling dataset.
 
-Additionally, to investigate if the amount of information used to run the RCP models was representative, we ran rarefaction curves based on the presence-absence matrix and the assigned RCP group for each sample (grid cell).
-Rarefaction curves were fit using the 'iNEXT' package, and results are presented based on alpha diversity (i.e.
+Finally, we ran rarefaction curves based on the presence-absence matrix and the assigned RCP group for each sample (grid cell) to assess if the amount of information in the RCP model was representative.
+Rarefaction curves were fit using the 'iNEXT' 3.0.0 package [@chao2014; @hsieh2016], and results are presented based on alpha diversity (i.e.
 species richness).
@@
 
-A total of 10,261 seabird records were obtained during the 15 voyages, comprising 142,646 individuals from 80 species, belonging to 11 families and five orders.
-The five most abundant species made up 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*).
-The five most frequently encountered species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig.
+A total of 10,261 seabird records was obtained during the 15 voyages, comprising 142,646 individuals from 80 species, belonging to 11 families and five orders.
+The five most abundant species made up 83% of total birds counted (43.0% short-tailed shearwater *Ardenna tenuirostris*, 17.2% fairy prion *Pachyptila turtur*, 14.3% wedge-tailed shearwater *Ardenna pacifica*, 5.2% sooty tern *Onychoprion fuscatus*, 3.2% shy albatross *Thalassarche cauta*).
+The five most frequently encountered species were *A.
+pacifica* (frequency of occurrence = 35.3%), *T.
+cauta* (33.2%), brown booby *Sula leucogaster* (25.4%), red-footed booby *S.
+sula* (25.4%) and black-browed albatross *T.
+melanophris* (25%) (**Fig.
 S1**).
 An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig S2**).
-Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig.
+Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Fig.
 S2**).
@@
 
-For all models two RCPs was the optimal number of groups, except for presence-absence-autumn, which indicated three RCPs as the best fit (**Fig S3**).
+For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Fig.
+S3**).
 The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**).
@@
 S4**).
-All the best models selected 'sst' or 'sss', variables essential for characterising water masses.
-The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21°C, or 'sss' 35.1 PSU for summer (**Fig.
+All the best models selected 'sst' or 'sss', as essential for characterising water masses.
+The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21 °C, or 'sss' 35.1 PSU for summer (**Fig.
 2**, **Fig.
-S5**), similar to EAC's temperature and salinity thresholds.
-Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa (**Fig.
+S5**), similar to the EAC's temperature and salinity thresholds.
+Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Fig.
 S5**).
-Interestingly, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) were selected only in autumn and spring abundance models.
+Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
 
@@
   ## Table title
-  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli error distribution; abundance, based on a Negative Binomial error distribution).
+  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli model; abundance, based on a Negative Binomial model).
 Refer to Table 1 for the environmental data acronyms."))
@@
 rm("df")
+
 ## Can't figure out why this code below doesn't work...
@@
 ```{r fig4-partial-plots, fig.cap = "**FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
-The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S5).
+The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
+All other predictors are held at their mean values.
+Results for other seasons and models can be found in the Supplementary Material (**Fig.
+S5**).
 Refer to Table 1 for the environmental data acronyms.", fig.align = "center", echo = FALSE}
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic change around \~34°S, with both presence-absence and abundance models pointing to similar patterns (**Fig.
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.
 3**).
@@
 4**, **Fig.
-S6**), showing that around \~34°S the 'northern' RCP tends to have higher probabilities in coastal regions, whereas the 'southern' RCP spreads northward offshore.
-The presence-absence-autumn model (three RCPs) has a similar broad-scale pattern, although the southern RCP is further south, and a strong 'Tasmania/Bass Straight' assemblage seems to appear (**Fig.
+S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
+The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Straight' assemblage appears (**Fig.
 3**, **Fig.
@@
 
-```{r fig3-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
-The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S6).", fig.align = "center", echo = FALSE}
+```{r fig3-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
+Results for other seasons and models can be found in the Supplementary Material (**Fig.
+S6**).", fig.align = "center", echo = FALSE}
 
@@
 The broad pattern of two large-scale assemblages is reflected in their species compositions.
-Although there are overlaps between groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.
+Although there are overlaps among the seabird groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.
 boobies *Sula* spp., lesser frigatebird *Fregata ariel*) and typical Southern Ocean species in the southern RCP (e.g.
-mollymawks *Thalassarche* spp., fairy prion *Pachyptila turtur*) (**Fig.
+small albatrosses *Thalassarche* spp., fairy prion *P.
+turtur*) (**Fig.
 5**, **Fig.
@@
 The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
-However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross *Thalassarche cauta* and Australasian gannet *Morus serrator* standing out with high probabilities of occurrence.
+However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross and Australasian gannet *Morus serrator* prominent with high probabilities of occurrence.
 
@@
 Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S7).", fig.align = "center", echo = FALSE}
+Results for other seasons and models can be found in the Supplementary Material (**Fig.
+S7**).", fig.align = "center", echo = FALSE}
 
@@
 
-This study shows that seabird assemblages off eastern Australia are correlated with oceanographic conditions, and that the East Australian Current (EAC) may be playing a key role in structuring seabird distribution at a large scale.
-Our results are based on a cutting-edge multivariate statistical model [@hill2020], which brings advantages over classical methods for describing species assemblages.
-For instance, RCP models allow estimating uncertainty around the point probability of a sample (grid cell) belonging to an assemblage, thus representing the fuzziness between bioregion boundaries.
+This study showed that seabird assemblages off eastern Australia were correlated with oceanographic conditions, and that the East Australian Current appeared to be key in structuring seabird distributions at meso- to macro scales [*sensu* @hunt1987].
+Our results are based on a multivariate statistical model [@hill2020], which brings advantages over classical methods for describing species assemblages.
+RCP models allow estimating uncertainty around the point probability of a sample (grid cell) belonging to an assemblage, thus representing the fuzziness between bioregion boundaries.
 These transition zones between bioregions are essential for monitoring biological changes due to environmental changes [@morrone2023biogeographical].
-As such, the boundary between the two consistently identified bioregions in our study may be a crucial region to monitor, given noticeable anomalies in the EAC at this same region [@phillips2022].
-Our results draw an important baseline for seabird assemblages in a world hotspot for this group.
+As such, the boundary between the two consistently identified bioregions in this study may be a crucial region to monitor in future, given noticeable anomalies in the EAC at this same region [@phillips2022].
+Our results draw an important baseline for seabird assemblages in a world hotspot for seabirds.
 
@@
 
-The relationship between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997].
-Therefore, it is not surprising that these two assemblages emerged off eastern Australia.
-However, the spatial consistency of these patterns throughout seasons is remarkable.
+The relationships between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997].
+Thus, it is not surprising that two seabird assemblages were identified off eastern Australia.
+However, the spatial consistency of these patterns throughout seasons is surprising and remarkable.
 Resident species seem to characterise these assemblages.
-The shy albatross [@hedd2001; @mason2018] and the fairy prions [@fromant2022] in the southern assemblage, and boobies [@bunce2015; @miller2018] and the tropical wedge-tailed shearwater *Ardenna pacifica* [@weimerskirch2020; @mcduie2015] in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species.
-Given that changes in assemblages over time are tipically driven by a few species [@gotelli2022], and that those species are likely residents [@souza2023], the relationship between resident seabirds and the EAC system should be closely monitored.
-For instance, coastal seabirds such as the little penguin *Eudyptula minor* are signalling the effects climate change might have in changing seabird behaviour and decreasing foraging success [@carroll2016], which ultimately influence their breeding success and population dynamics [@johnson2021].
-Therefore, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring seabird responses to the fast-changing EAC.
+Shy albatross [@hedd2001; @mason2018] and fairy prions [@fromant2022] in the southern assemblage, and boobies [@bunce2015; @miller2018] and wedge-tailed shearwaters [@weimerskirch2020; @mcduie2015] in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species in the same groups.
+Given that changes in assemblages over time are typically driven by a few species [@gotelli2022], and that those species are likely residents [@souza2023], the relationships between these resident seabirds and the EAC system over the broad latitudinal range detailed here should be closely monitored.
+Elsewhere, coastal seabirds such as little penguin are signalling the adverse effects that climate change might have in changing seabird behaviours and decreasing foraging success [@carroll2016; @woehler2023], which ultimately influence their breeding success and population dynamics [@johnson2021].
+Thus, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring future seabird responses to the fast-changing EAC.
 
 The seasonal models also captured the influence of migratory species, particularly in the southern assemblage.
-For instance, summer and autumn models reflected the high abundance of the short-tailed shearwater *Ardenna tenuirostris* before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
-Given the geographic location of the southern assemblage, however, we believe subantarctic taxa were not well represented in our models due to their low number of occurrences (**Fig.
-S1 - FO/nOCC/FN**; hence they were removed before modelling).
-More intensive surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.
+The summer and autumn models reflected the high abundance of short-tailed shearwater before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by the autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
+Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (**Fig.
+S1**; hence they were removed before modelling).
+Additional surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.
 albatrosses *Diomedea* spp.
 and *Phoebetria* spp.), given that more effort translates as higher likelihood of recording these species.
-In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the autumn presence-absence model) in all but the summer season, when the majority of Antarctic species are breeding **[Harrison 1983 - is there a better ref here, seabirders?]**.
+In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the autumn presence-absence model) in all but the summer season, when the majority of Antarctic species are breeding [@marchant1990handbook].
 
-Our analysis clearly revealed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia.
-There is a noticeable link with EAC dynamics; the boundaries of RCPs occurred where the EAC turns into its 'eastern extension' and the eddy field spawns [@oke2019].
-However, the 'northern/southern' assemblage pattern does not agree with popular, large scale regionalization proposals, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007], or with the Large Marine Ecosystems [LME; @sherman1994].
-The MEOW scheme was designed using coastal and shelf biota and was meant to represent habitats up to the 200-m isobath, which clearly does not describe the entirety of seabird habitat.
-The LME scheme is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management areas and thus a 'political' component to it, which does not necessarily align with distribution of seabirds.
-@longhurst2007 designated only one biogeochemical province for east Australia based on productivity patterns, which may work well at larger scales but seems oversimplified specifically for this region.
+Our analyses clearly revealed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia within the EAC.
+There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \~34°S, near the average separation latitude (\~32°S) of the EAC and its eastward extension into the Tasman Sea [@oke2019].
+However, the 'northern/southern' assemblage pattern identified herein does not agree with previously-proposed large-scale regionalizations, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007], or the Large Marine Ecosystems [LME; @sherman1994].
+The MEOW schema was based on coastal and shelf biota and was meant to represent habitats to the 200 m isobath, which clearly does not encompass the entirety of seabird habitats.
+The LME schema is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management of areas and thus there is a political component embedded within it, which does not align with seabird distributions.
+@longhurst2007 designated only one biogeochemical province for eastern Australia based on productivity patterns, which may work well at larger (macro) scales, but appears to be oversimplified for the EAC.
 
-The described pattern does however agree with the marine biogeographic units proposed by @costello2017, who identified a boundary at \~34°S between their 'Coral Sea' and 'South Australia' realms.
-This approach used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus encompassing more complexity of the biogeographic boundaries in the marine environment.
-Other biological data also suggest 'northern/southern' assemblages off eastern Australia, including tuna and billfish [also highly mobile top predators; @reygondeau2012], mesopelagic fauna [which includes seabird prey; @sutton2017a] and deep-sea invertebrates [@ohara2020] and fish [@ohara2020a].
-All these studies focused mainly on describing biological assemblages, as opposed to classifying environment.
+The results do agree with the marine biogeographic units proposed by @costello2017, who identified a boundary at \~34°S between their 'Coral Sea' and 'South Australia' realms.
+Their approach used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus integrating more complexity of the biogeographic boundaries in the marine environment.
+Other biological data also suggest 'northern/southern' assemblages off eastern Australia, including tuna and billfish [also highly mobile top predators; @reygondeau2012], mesopelagic fauna [which includes seabird prey; @sutton2017a] and deep-sea invertebrates and fish [@ohara2020; @ohara2020a].
+All these studies focused mainly on describing biological assemblages, as opposed to classifying environments.
 It is worth noting that the Longhurst Provinces approach was updated to include SST, SSS and bathymetry (the same variables retained by our best RCP models) and a temporal component to the classification [@reygondeau2013].
-The updated provinces are more similar to the seabird assemblages described by us, although still not a perfect match.
+The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage at EAC region includes several long-distance migratory species.
 Australia's bioregionalization has a long story of disagreements [@ebach2012history], however, the inherent link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
@@
 RCP models have been previously applied to plants [@lyons2017], and relatively sedentary animals, such as macrobenthic organisms [@pantiukhin2021] and demersal fish [@hill2017].
-Pelagic seabirds however can have daily movements of hundreds of kilometres [e.g.
+Oceanic seabirds can have daily movements of hundreds of kilometres [e.g.
 @mcduie2015], during which they traverse a variety of habitats and environmental gradients.
-As such, capturing the distribution of highly mobile species, and their relationships with environmental parameters is not a trivial task, and likely much harder than for sessile organisms.
+As such, capturing the distribution of highly mobile species, and their relationships with environmental parameters is not a trivial task, and much harder than for sessile organisms.
 
-While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009], they fail to perceive how groups of species (assemblages) relate to the environment and each other on a shared seascape.
+While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009], they fail to perceive how species assemblages interact with the environment and each other.
 Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
-Furthermore, the spatial scale on which assemblages relate to their environment is also a field of debate [@hunt1987; @ribic1997; @scales2017].
-Understanding the biology of the studied group is thus essential to set the appropriate scale of analysis.
-Our choice of 1° latitude $\times$ 1° longitude was mainly influenced by data availability and the trade-off between number of species per grid/season (see also \@ref(limitations)).
+Further, the spatial scales on which assemblages relate to their environment is variable and species-specific [@hunt1987; @ribic1997; @scales2017].
+Understanding the biology of the studied group is thus essential to set the appropriate scales of analyses.
+Our choice of 1° latitude $\times$ 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \@ref(limitations)).
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere [@harris2007; @santora2013; @sojitra2022].
-Therefore, for describing seabird assemblages at a large scale, we believe the chosen resolution was a good fit, and the models described expected patterns of species composition.
+Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to had been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
 
-RCP models based on presence-absence and abundance data could complement each other.
+RCP models based on presence-absence and abundance data are able to complement each other.
 Previous studies using RCPs have opted to use only one type of response data.
-@lyons2017, @hill2017, and @pantiukhin2021 used presence-absence data, whereas @lee2019good and @receveur2020 used abundance data.
-The added information contained in abundance datasets make them inherently more powerful than presence-absence data.
+@lyons2017, @hill2017, and @pantiukhin2021 all used presence-absence data, whereas @lee2019good and @receveur2020 used abundance (count) data.
+The added information contained in abundance data make them inherently more powerful than binary presence-absence data.
 However, the use of different data inputs in multivariate, biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
-At larger scales, presence-absence datasets describe similar patterns to abundance datasets [@anderson2005], as we found in our results.
-In contrast, at smaller scales, abundance datasets may highlight local conditions and therefore reveal hidden patterns [@marchant1990; @anderson2005; @melo2005; @waters2008marine].
-If combined with ecological knowledge of a particular group, the information contained in both presence-absence and abundance datasets could be leverage to better interpret the resulting patterns.
-For example, seabirds will likely occur in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance is related to fine scale processes and prey availability [@hunt1987; @weimerskirch2007].
-Therefore, future studies assessing bioregions of highly mobile animals could describe the make-up of assemblages using smaller resolution grid cells and presence-absence models.
-Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore 'new bioregions') likely nested within the bioregions described at large scale.
+At larger spatial and temporal scales, presence-absence data describe similar patterns to abundance data [@anderson2005], as we found.
+In contrast, at smaller scales, abundance data may highlight local conditions and therefore reveal hidden patterns at coarse to fine scales [@marchant1990; @anderson2005; @melo2005; @waters2008marine].
+If combined with ecological knowledge of a particular group, the information contained in both presence-absence and abundance data could be leveraged to better interpret the resulting patterns.
+For example, seabirds will likely occur in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance is related to coarse- to fine-scale processes that enhance prey availability [@hunt1987; @weimerskirch2007].
+Future studies assessing bioregions of highly mobile animals could describe the compositions of assemblages using smaller resolution grid cells and presence-absence models.
+Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore 'new bioregions') likely nested within the bioregions described at large scales.
+Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect not exist at present.
 
@@
 
-Although the results suggest consistent patterns, we should point out some caveats of our analyses.
+Although the results suggest consistent patterns, there are a number of caveats to our analyses.
 Seabird data were gathered during multi-disciplinary cruises, whose sampling designs were not specifically planned to account for spatial ecology of seabirds.
-As a result, there are spatial discontinuities in the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors.
-Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution in the region [@mott2018; @bernard2021].
-Dedicated seabird surveys would help to advance seabird research in this region, as well as improving the patterns described here.
-By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at a finer scale.
+As a result, there are spatial discontinuities in the data, which might have limited the ability of our models to detect the relationships between the seabird assemblages and the full spectrum of environmental predictors.
+Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is a paucity of information on seabird distributions in the region [@mott2018; @bernard2021].
+Dedicated seabird surveys would help to advance seabird research in this region, and improve our understanding of the patterns described here.
+By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at finer spatial and temporal scales.
 In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
@@
 
-We applied a robust, repeatable method to assess seabird assemblages off eastern Australia for the first time.
-The models point toward a consistent pattern of a 'northern' and a 'southern' assemblage, where the spatial division is clearly linked with the behaviour of the EAC and SST seems to be the key variable discriminating these groups.
-As the EAC is predicted to strengthen [@poloczanska2007], and extreme weather events such as marine heatwaves to happen more often [@oliver2018], seabirds will likely respond to these changes.
-Southern ocean albatrosses off southeast Australia, for instance, have shown 'early stages of a southward range shift' due to environmental variability [@sojitra2022].
+We applied a robust, repeatable method to assess seabird assemblages off eastern Australia within the EAC for the first time.
+The models point toward a consistent pattern of a 'northern' and a 'southern' seabird assemblage, where the spatial division is attributable to the eastward separation of the EAC from the coast, and that SST seems to be the key variable discriminating these groups.
+The EAC has penetrated poleward over the past half-century [@li2022; @hill2008], a trend which is anticipated to continue into the future [@yang2016], and extreme weather events such as marine heatwaves to happen more often [@oliver2018].
+As such, seabirds will have to respond to these changes [@woehler2023].
+Southern Ocean albatrosses off southeast Australia, for instance, have shown 'early stages of a southward range shift' due to environmental variability [@sojitra2022].
 Resident species such as the little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher [@carroll2016; @peck2004].
-Given that eastern Australia is a biodiversity and abundance hotspot for seabirds, it is imperative establish a monitoring program allowing the assemblages and species profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
+Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species' profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
 In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.

### 2023-07-28 scripts/00_environment-set-up.R
I've done a few tweaks in the code to make better partial plots and fixed spring-abund partial plot

@@ -21,7 +21,7 @@ install.packages("pacman")
 
 needed_libraries <- c("plyr", "dplyr", "tidyr", "readr", "tibble",
                       "lubridate", "stringr", "purrr", 
-                      "ggplot2", "ggspatial", "patchwork", "RColorBrewer", "rnaturalearth", 
+                      "ggplot2", "ggspatial", "patchwork", "RColorBrewer", "colorspace", "rnaturalearth", 
                       "sp", "sf", "mapview", "raster", "terra", "rerddap", "rerddapXtracto", "tmap", 
                       "corrplot", "iNEXT")
 

### 2023-07-28 scripts/08_1_modelling_fit-RCPs-Bernoulli.R

@@ -969,7 +969,9 @@ for (season_vec in seasons_vec) {
                      xlb = stringr::str_sub(name, end = -6), 
                      ylb = "Probability", 
                      location = "right", 
-                     coluse = coluse_pal)
+                     coluse = coluse_pal,
+                     model = "pres_abs",
+                     season = season_vec)
     dev.off()
     
     rm("name", "i")

### 2023-07-28 scripts/08_2_modelling_fit-RCPs-NegBin.R

@@ -7,6 +7,9 @@
 ## This code implements RCP models, 
 ## using count data based on 'Negative Binomial' error distribution.
 
+## Note: To keep colour congruence between all visualizations, 
+## for the *spring* results, I had to swap 'RCP1' / 'RCP2' labels.
+
 ## Libraries ####
 
 library(plyr)
@@ -821,8 +824,8 @@ for (season_vec in seasons_vec) {
   if(season_vec == "spring" | season_vec == "summer") {
     
     # As RCP group colours are depicted from 'south' (RCP1) to 'north' (RCP2),
-    # we had to do a hack here for the probability maps visually agree with 
-    # point-predictions in a meaningful way (basically, swap "RCP1" to "RCP2", and vice-versa)
+    # I had to do a hack here for the probability maps visually agree with 
+    # point-predictions in a meaningful way (basically, swap "RCP1" / "RCP2")
     
     new_col_names <- c("IDgrid", "lon", "lat", 
                        "RCP_2", "RCP_1", 
@@ -950,7 +953,9 @@ for (season_vec in seasons_vec) {
                      xlb = stringr::str_sub(name, end = -6), 
                      ylb = "Probability", 
                      location = "right", 
-                     coluse = coluse_pal)
+                     coluse = coluse_pal,
+                     model = "abund",
+                     season = season_vec)
     dev.off()
     
     rm("name", "i")

### 2023-07-28 scripts/source.R

@@ -124,7 +124,9 @@ FUN_partial_plot = function(beta.mat,
                             xlb = "", 
                             ylb = "Probability", 
                             location = "topright", 
-                            coluse = palette()[1:(nrow(beta.mat)+1)]){
+                            coluse = palette()[1:(nrow(beta.mat)+1)],
+                            model = c("pres_abs", "abund"), 
+                            season = c("summer", "autumn", "winter", "spring")){
   ## beta.mat is a matrix of beta values.  This is obtained from coef(model)
   ## predplot is a vector (dimension = # of predictors).  The predictor with NA is that plotted.  The value for all other predictors gives the fixed value of that predictor.
   ## predplot = c(NA,0) plots the 1st predictor, setting the value of the 2nd predictor to 0.
@@ -135,6 +137,8 @@ FUN_partial_plot = function(beta.mat,
   ## ylb is the y label
   ## location gives the location of the legend (?legend for options)
   ## coluse is a vector giving the colours of the profiles (defaults to first 3 colours in the palette)
+  ## model is the type of the specified model -- only needed because for a specific model, we need to swap colours/labels
+  ## season is respective season -- only needed because for a specific model, we need to swap colours/labels
   
   npar = ncol(beta.mat)  # number of parameters (includes intercept)
   nprof = nrow(beta.mat)+1 # number of profiles to plot
@@ -161,12 +165,17 @@ FUN_partial_plot = function(beta.mat,
   sum.mat = matrix(rowSums(eta.mat),n,nprof)
   prob.mat = eta.mat/sum.mat
   
-  ## plot the output
-  plot(NA,NA,xlim = xlm.real, ylim = c(0,1), xlab = xlb, ylab = ylb)
-  for(j in 1:nprof){
-    lines(xuse.real, prob.mat[,j], col = coluse[j], lwd = 3)
+  if(model == "abund" & season == "spring"){
+    prob.mat <- prob.mat[nrow(prob.mat):1, ]
   }
-  legend(location, legend = paste("RCP",1:nprof), col = coluse, lty = 1, bty = "n")
+  
+  ## plot the output
+  plot(NA,NA,xlim = xlm.real, ylim = c(0,1), xlab = toupper(xlb), ylab = ylb,
+       cex.lab = 1.5, cex.axis = 1)
+  for(j in 1:nprof){
+    lines(xuse.real, prob.mat[,j], col = coluse[j], lwd = 4)
+  }
[…]

### 2023-10-13 ms_preprint/ms_preprint.Rmd
{rticle} with the ms content updated

--- 2023-07-26 docs/ms-draft.Rmd
@@
 ---
-title: "Seabird assemblages are linked to the major western boundary current off eastern Australia"
-author:
-  - Nicholas W. Daudt:
+title: Seabird assemblages are linked to the major western boundary current off eastern Australia
+authors:
+  - name: Nicholas W. Daudt
+    department: Department of Marine Science
+    affiliation: University of Otago
+    location: Dunedin, Aotearoa New Zealand
     email: nicholaswdaudt@gmail.com
-      institute: [MarSciUO]
-      correspondence: true
-  - Eric J. Woehler:
-      institute: [ASG]
-      correspondence: false
-  - Matthew R. Schofield:
-      institute: [MathStatUO]
-      correspondence: false
-  - Robert O. Smith:
-      institute: [MarSciUO]
-      correspondence: false
-  - Leandro Bugoni:
-      institute: [FURG]
-      correspondence: false
-  - William J. Rayment:
-      institute: [MarSciUO]
-      correspondence: false
-institute:
-  - MarSciUO: Department of Marine Science, University of Otago, Aotearoa New Zealand
-  - ASG: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
-  - MathStatUO: Department of Mathematics and Statistics, University of Otago, Aotearoa New Zealand
-  - FURG: Institute of Biological Sciences, Universidade Federal do Rio Grande--FURG, Rio Grande, RS, Brazil
+  - name: Eric J. Woehler
+    department: Australasian Seabird Group
+    affiliation: BirdLife Australia
+    location: Hobart, Australia
+  - name: Matthew R. Schofield
+    department: Department of Mathematics and Statistics
+    affiliation: University of Otago
+    location: Dunedin, Aotearoa New Zealand
+  - name: Robert O. Smith
+    department: Department of Marine Science
+    affiliation: University of Otago
+    location: Dunedin, Aotearoa New Zealand
+  - name: Leandro Bugoni
+    department: Institute of Biological Sciences
+    affiliation: Universidade Federal do Rio Grande--FURG
+    location: Rio Grande, RS, Brazil
+  - name: William J. Rayment
+    department: Department of Marine Science
+    affiliation: University of Otago
+    location: Dunedin, Aotearoa New Zealand
 abstract: |
@@
 Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
-output: 
-  # rticles::sim_article:
-  #   keep_tex: TRUE
-  bookdown::word_document2:
-    pandoc_args:
-      - --lua-filter=../docs/scholarly-metadata.lua
-      - --lua-filter=../docs/author-info-blocks.lua
-      - --reference-doc=../docs/docx-template.docx 
+keywords:
+  - Assemblages
+  - Community ecology
+  - East Australian Current
+  - Mixture models
+  - Region of Common Profiles
 bibliography: ../references.bib
+# biblio-style: 
 csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
+output: rticles::arxiv_article
+header-includes:
+- \usepackage{setspace}\doublespacing
+- \usepackage{pdflscape}
+- \DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}}
+- \newcommand{\blandscape}{\begin{landscape}}
+- \newcommand{\elandscape}{\end{landscape}}
 ---
 
-<!-- YAML for bookdown+lua edited following: -->
-
-<!-- https://www.r-bloggers.com/2019/08/rmarkdown-template-that-manages-academic-affiliations-docx-or-pdf-output/ -->
+<!-- Running title: Seabird assemblages off East Australia -->
+<!-- Note: I've changed the 'arxiv.sty' source file to include the running title in the PDF -->
 
@@
 library(purrr)
-library(gt)
+library(stringr)
+library(kableExtra)
+
+### If using {gt}:
+# library(gt)
+# options(gt.html_tag_check = FALSE)
 ```
-
-\pagebreak
 
@@
 RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
-Eastern Australia supports a high diversity of seabirds and offers wide environmental gradients over a near 40° latitudinal range, from the tropics to the cool temperate (**ref?**) We hypothesised to find different seabird assemblages relating to different environmental conditions within the EAC.
+Eastern Australia supports a high diversity of seabirds and offers wide environmental gradients over a near 40° latitudinal range, from the tropics to the cool temperate.
+We hypothesised to find different seabird assemblages relating to different environmental conditions within the EAC.
 In addition, presence-absence and abundance data were used to test if the different numerical resolutions would result in the identification of similar assemblages.
@@
 
-```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
-In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
-The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
-The EAC 'eastern extension' is an eddy field (blue shaded area) often called the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
-Figure (a) is adapted after Oke, Roughan, et al.
-(2019); for colour references see the online version.", fig.align = "center", echo = FALSE}
-
-# knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
-
-knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
-```
-
 To investigate and predict species' distributions and relationships with environmental variables, a considered choice of **broad-scale climatic and oceanographic** covariates is important [@barbet-massin2014].
@@
 
-```{r table1, echo = FALSE}
-
-data.frame(
-  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
-  Abbreviation = c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"),
-  Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
-  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
-  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
-  Source = c("Huang et al.
-(2021)", "Derived from Huang et al.
-(2021)", "E.U.
-Copernicus Marine Service Information", "E.U.
-Copernicus Marine Service Information", "ERDDAP/NOAA", "Derived from ERDDAP/NOAA", "IMOS/AODN", "Derived from IMOS/AODN", "Derived from IMOS/AODN", "Derived from Huang et al.
-(2021)", "GEBCO", "Derived from GEBCO", "Calculated in R (see main text)")
-) |>
-  gt::gt() |>
-  ## SST & SSTgrad
-  gt::tab_footnote(footnote = 
-                     gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
-                   locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
-  ## SSS & MLD
-  gt::tab_footnote(footnote = 
-                     gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
-                   locations = gt::cells_body(columns = Source, rows = 3:4)) |>
-  ## CHL8 & log10-CHL8
-  gt::tab_footnote(footnote = 
-                     gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
-                   locations = gt::cells_body(columns = Source, rows = 5:6)) |>
-  ## EKE and derivations
-  gt::tab_footnote(footnote = 
-                     gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
-                   locations = gt::cells_body(columns = Source, rows = 7:9)) |>
-  ## GEBCO & slope
-  gt::tab_footnote(footnote = 
-                     gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
-                   locations = gt::cells_body(columns = Source, rows = 11:12)) |>
-  ## Improve cols names
-  gt::cols_label(
-    Variable = gt::md("**Variable**"),
-    Abbreviation = gt::md("**Abbreviation**"),
-    Units = gt::md("**Units**"),
-    `Spatial.resolution` = gt::md("**Spatial resolution**"), 
-    `Temporal.resolution` = gt::md("**Temporal resolution**"),
-    Source = gt::md("**Source**")) |>
-  ## Table title
-  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources."))
-
-```
-
 ## Data treatment
@@
 Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material **XX**.
-Code and the modelling dataset (see \@ref(modellingdatasets)) are available in **Daudt et al.
+Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in **Daudt et al.
 2023 (Zenodo repo)**.
-
-<!-- using packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021].
--->
 
@@
 ### Data {#modellingdatasets}
+\label{sec:modellingdatasets}
 
@@
 To avoid numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12 species; autumn = 21; winter = 20; spring = 33).
-The associated environmental data of each seabird record were averaged within the same grid cells and seasons, and environmental data were **standardised** before modelling.
+The associated environmental data of each seabird record were averaged within the same grid cells and seasons, and environmental data were standardised before modelling.
 
@@
 There are $K$ assemblages (i.e.
-RCPs) that are common across the study area, where $K = (1, ..., k)$.
-Depending on whether the model is specified in terms of abundance or presence-absence, the assemblages are defined in terms of their structure at each $i$th site (a vector $y_{i}$), i.e.
-frequency or relative frequency of the species, respectively.
-The assemblage structure is denoted by $y_{i} = (y_{i1}, ...
-y_{iS})$, where $S$ is the number of species.
-The model estimates the probability of belonging to the $K$ assemblage given the site's assemblage structure.
-Thus, each site within the study area has a $k$-dimensional probability vector that gives the probability of the corresponding assemblage for that site.
+RCPs) that are common across the study area, $K = (1, ..., k)$.
+The outcome variable is a vector $y_{i} = (y_{i1}, ...
+y_{iS})$, that give the count or presence-absence of each of the $S$ species at site $i$.
+Depending on whether we model the counts or presence-absence data, the assemblages are defined in terms of the expected frequency or relative frequency of the species, respectively.
+An output of the model is the estimated probability that site $i$ has each of the $K$ possible assemblages.
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) [@foster2013].
@@
 Due to seabird migration and dispersion, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
-In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \@ref(modellingdatasets)), we decided to use RCP models by season and data type (presence-absence and abundance).
+In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
 RCP models were used to relate the response variables (i.e.
@@
 Presence-absence data were modelled based on a Bernoulli model, whereas abundance data based on a Negative Binomial model.
-RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($k$) that describes the data, (ii) re-fit the model by fixing $k$ and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($K$) that describes the data, (ii) re-fit the model by fixing $K$ and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
 
@@
 Note that 'clim_eke_mean' and 'clim_sst_grad' were processed beforehand as they may characterise persistent oceanographic features (described above) and thus represent the same value as 'eke' and 'sst_grad'.
-Data for prediction were **standardised** using the same mean and standard deviation from the modelling dataset.
+Data for prediction were standardised using the same mean and standard deviation from the modelling dataset.
 
@@
 
-```{r table2, echo = FALSE, message = FALSE}
-
-df <- 
-  rbind(
-    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
-    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
-  purrr::map_df(~readr::read_csv(.)) %>% 
-  dplyr::select(- `...1`) %>% 
-  dplyr::mutate(data_type = factor(data_type, levels = c("bernoulli", "neg_bin")),
-                season = factor(season, levels = c("summer", "autumn", "winter", "spring")),
-                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
-                bic = round(bic, digits = 2)) %>%
-  dplyr::group_by(data_type) %>% 
-  dplyr::arrange(data_type, season) %>% 
-  as.data.frame(.) %>% 
-  dplyr::ungroup(.)
-
-# Correct a value in a cell
-# Full abundance-autumn model did not fit properly, returning a value of -Inf
-df[df$data_type=='neg_bin' & df$season=='autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"
-
-# A few tweaks to make the table look better
-df$data_type <- c("Bernoulli", rep("", 7), "Negative Binomial", rep("", 7))
-df$season <- rep(c("Summer", "", "Autumn", "", "Winter", "", "Spring", ""), 2)
-
-gt::gt(df)|>
-  ## Improve cols names
-  gt::cols_label(
-    data_type = gt::md("**Data type**"),
-    season = gt::md("**Season**"),
-    model = gt::md(" "),          ## To be excluded in the Word doc
-    model_specification = gt::md("**Model specification**"), 
-    bic = gt::md("**BIC**")) |>
-  ## Table title
-  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli model; abundance, based on a Negative Binomial model).
-Refer to Table 1 for the environmental data acronyms."))
-
-rm("df")
-
-## Can't figure out why this code below doesn't work...
-# kableExtra::kbl(df, booktabs = T, align = "l", linesep = "") %>% 
-#   kableExtra::column_spec(1, bold=T) %>%
-#   kableExtra::collapse_rows(columns = c(1:3), 
-#                             latex_hline = 'custom', 
-#                             custom_latex_hline = 1:3, 
-#                             row_group_label_position = 'stack')
-```
-
-```{r fig4-partial-plots, fig.cap = "**FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
-The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
-All other predictors are held at their mean values.
-Results for other seasons and models can be found in the Supplementary Material (**Fig.
-S5**).
-Refer to Table 1 for the environmental data acronyms.", fig.align = "center", echo = FALSE}
-
-knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
-```
-
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
@@
 
-```{r fig2-point-pred, fig.cap = "**FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
-Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}
-
-knitr::include_graphics("../results/Fig2-point-predictions.png")
-```
-
-```{r fig3-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
-The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
-Results for other seasons and models can be found in the Supplementary Material (**Fig.
-S6**).", fig.align = "center", echo = FALSE}
-
-knitr::include_graphics("../results/Fig3-probability-maps-spring.png")
-```
-
 The broad pattern of two large-scale assemblages is reflected in their species compositions.
@@
 However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross and Australasian gannet *Morus serrator* prominent with high probabilities of occurrence.
-
-```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
-Values are the average and 95% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
-Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material (**Fig.
-S7**).", fig.align = "center", echo = FALSE}
-
-knitr::include_graphics("../results/Fig5-spp-profiles-spring.png")
-```
 
@@
 Understanding the biology of the studied group is thus essential to set the appropriate scales of analyses.
-Our choice of 1° latitude $\times$ 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \@ref(limitations)).
+Our choice of 1° latitude $\times$ 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \ref{sec:limitations}).
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere [@harris2007; @santora2013; @sojitra2022].
@@
 
-## Limitations {#limitations}
+## Limitations
+\label{sec:limitations}
 
@@
 
+# Acknowledgements
+
+We thank all observers who assisted in data collection, the crews of RV *Investigator* and RV *Falkor*, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
+The authors would like to thank Dr.
+Skipton N.C.
+Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
+NWD is supported by a Doctoral scholarship from the University of Otago.
+
 # References
+
+<div id="refs"></div>
+
+\newpage
+
+# Tables and Figures {-}
+
+- **TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources
+- **TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli model; abundance, based on a Negative Binomial model).
+Refer to Table 1 for the environmental data acronyms
+- **FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
+In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
+The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
+The EAC 'eastern extension' is an eddy field (blue shaded area) often called the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
+Figure (a) is adapted after Oke, Roughan, et al.
+(2019); for colour references see the online version
+- **FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
+All other predictors are held at their mean values.
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S5).
+Refer to Table 1 for the environmental data acronyms
+- **FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+Each grid was assigned to the RCP group with highest membership probability
+- **FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
+Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S6)
+- **FIGURE 5.** Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
+Values are the average and 95% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
+Values in (b) were log10-transformed to accommodate the high variation between species.
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S7)
+
+\blandscape
+
+```{r table1, echo = FALSE}
+
+data.frame(
+  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
+  # Abbreviation needs the `str_replace_all` because I'm using `escape=F` in `kbl()`
+  Abbreviation = stringr::str_replace_all(c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"), "_", "\\\\_"),
+  Units = c("°Celsius", paste0(parse(text = "\u0394"), "°Celsius"), "PSU", "m", "mg/m${^3}$", "$log_{10}$(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", paste0("mean(", paste0(parse(text = "\u0394"), "°Celsius"), ")/season/10 years"), "m", "degrees (°)", "km"),
+  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
+  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
+  Source = c("Huang et al.
+(2021)${^a}$", "Derived from Huang et al.
+(2021)${^a}$", "E.U.
+Copernicus Marine Service Information${^b}{^c}$", "E.U.
+Copernicus Marine Service Information${^b}{^c}$", "ERDDAP/NOAA${^d}$", "Derived from ERDDAP/NOAA${^d}$", "IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from Huang et al.
+(2021)${^a}$", "GEBCO${^f}$", "Derived from GEBCO${^f}$", "Calculated in R (see main text)")
+) |>
+  kableExtra::kbl(caption = " ", booktabs = TRUE, escape = FALSE)  %>% 
+  kableExtra::row_spec(0, bold = TRUE) %>% 
+  kableExtra::column_spec(1, width = "4cm") %>% 
+  kableExtra::column_spec(6, width = "3cm") %>% 
+  kableExtra::footnote(
+    alphabet = c(
+    "NOAA OISST: (https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html); ", 
+    "2011--2019: GLOBAL_MULTIYEAR_PHY_001_030 (https://doi.org/10.48670/moi-00021); ", 
+    "2020--2021: GLOBAL_ANALYSIS_FORECAST_PHY_001_024 (https://doi.org/10.48670/moi-00016); ",
+    "ERDDAP/NOAA (https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html); ",
+    "IMOS/AODN (https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set; ",
+    "GEBCO (https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"))
+
+### Using {gt} -- just pipe this onto the data.frame
+  # gt::gt() |>
+  # ## SST & SSTgrad
+  # gt::tab_footnote(footnote = 
+  #                    gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
+  #                  locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
+  # ## SSS & MLD
+  # gt::tab_footnote(footnote = 
+  #                    gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
+  #                  locations = gt::cells_body(columns = Source, rows = 3:4)) |>
+  # ## CHL8 & log10-CHL8
+  # gt::tab_footnote(footnote = 
+  #                    gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
+  #                  locations = gt::cells_body(columns = Source, rows = 5:6)) |>
+  # ## EKE and derivations
+  # gt::tab_footnote(footnote = 
+  #                    gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
+  #                  locations = gt::cells_body(columns = Source, rows = 7:9)) |>
+  # ## GEBCO & slope
+  # gt::tab_footnote(footnote = 
+  #                    gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
+  #                  locations = gt::cells_body(columns = Source, rows = 11:12)) |>
+  # ## Improve cols names
+  # gt::cols_label(
+  #   Variable = gt::md("**Variable**"),
+  #   Abbreviation = gt::md("**Abbreviation**"),
+  #   Units = gt::md("**Units**"),
+  #   `Spatial.resolution` = gt::md("**Spatial resolution**"), 
+  #   `Temporal.resolution` = gt::md("**Temporal resolution**"),
+  #   Source = gt::md("**Source**")) |>
+  # ## Table title
+  # gt::tab_header(title = gt::md("Table 1.")) |>
+  # gt::fmt_markdown(columns = gt::starts_with("Var")) #|> #columns = c("Variable", "Source")
+  # # gt::tab_options(table.width = gt::pct(100))
+  # # gt::tab_options(container.width = 300, container.height = 300)
+```
+
+\elandscape
+
+\newpage
+
+```{r table2, echo = FALSE, message = FALSE}
+
+df <- 
+  rbind(
+    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
+    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
+  purrr::map_df(~readr::read_csv(.)) %>% 
+  dplyr::select(- `...1`) %>% 
+  dplyr::mutate(data_type = factor(data_type, 
+                                   levels = c("bernoulli", "neg_bin"),
+                                   labels = c("Bernoulli", "Negative Binomial")),
+                season = factor(season, 
+                                levels = c("summer", "autumn", "winter", "spring"),
+                                labels = c("Summer", "Autumn", "Winter", "Spring")),
+                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
+                bic = round(bic, digits = 2)) %>%
+  dplyr::group_by(data_type) %>% 
+  dplyr::arrange(data_type, season) %>% 
+  as.data.frame(.) %>% 
+  dplyr::ungroup(.)
+
+# Correct a value in a cell
+# Full abundance-autumn model did not fit properly, returning a value of -Inf
+df[df$data_type=='Negative Binomial' & df$season=='Autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"
+
+# A few adjustments for nicer output
+# df$model_specification <- stringr::str_replace_all(df$model_specification, " ", "")
+colnames(df)[3] <- 'Model'
+colnames(df)[4] <- 'Model specification'
+colnames(df)[5] <- 'BIC'
+
+kableExtra::kbl(df, caption = " ", booktabs = TRUE) %>% 
+  kableExtra::column_spec(0, bold = TRUE) %>%
+  kableExtra::column_spec(1, bold = TRUE) %>%
+  kableExtra::column_spec(4, width = "11cm") %>% 
+  kableExtra::collapse_rows(columns = c(1:3),
+                            latex_hline = 'custom',
+                            custom_latex_hline = 1:3,
+                            row_group_label_position = 'stack')
+
+### Using {gt}:
+# A few tweaks to make the table look better
+# df$data_type <- c("Bernoulli", rep("", 7), "Negative Binomial", rep("", 7))
+# df$season <- rep(c("Summer", "", "Autumn", "", "Winter", "", "Spring", ""), 2)
+
+# gt::gt(df)|>
+#   ## Improve cols names
+#   gt::cols_label(
+#     data_type = gt::md("**Data type**"),
+#     season = gt::md("**Season**"),
+#     model = gt::md("**Model**"),
+#     model_specification = gt::md("**Model specification**"), 
+#     bic = gt::md("**BIC**")) |>
+#   ## Table title
+#   gt::tab_header(title = gt::md("Table 2."))
+
+rm("df")
+```
+
+\newpage
+
+\blandscape
+
+```{r fig1-study-area, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
+
+# knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
+
+knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
+```
+
+\elandscape
+
+\newpage
+
+```{r fig2-partial-plots, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
+```
+
+\newpage
+
+```{r fig3-point-pred, fig.cap = " ", fig.align = "center", out.width = "110%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig2_point-predictions.png")
+```
+
+\newpage
+
+\blandscape
+
+```{r fig4-prob-maps, fig.cap = " ", fig.align = "center", out.width = "110%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig3_probability-maps-spring.png")
+```
+
+\newpage
+
+```{r fig5-spp-profiles, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig5_spp-profiles-spring.png")
+```
+
+\elandscape

### 2023-10-13 ms_preprint/ms_preprint.tex
{rticle} with the ms content updated

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Seabird assemblages are linked to the major western boundary
current off eastern Australia}

\author{
    Nicholas W. Daudt
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{\href{mailto:nicholaswdaudt@gmail.com}{\nolinkurl{nicholaswdaudt@gmail.com}}} \\
   \And
    Eric J. Woehler
   \\
    Australasian Seabird Group \\
    BirdLife Australia \\
  Hobart, Australia \\
  \texttt{} \\
   \And
    Matthew R. Schofield
   \\
    Department of Mathematics and Statistics \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
   \And
    Robert O. Smith
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
   \And
    Leandro Bugoni
   \\
    Institute of Biological Sciences \\
    Universidade Federal do Rio Grande--FURG \\
  Rio Grande, RS, Brazil \\
  \texttt{} \\
   \And
    William J. Rayment
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{setspace}\doublespacing
\usepackage{pdflscape}
\DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\textbf{Aim}: To identify assemblages of seabird species and relate them
to environmental conditions off eastern Australia, a world seabird
hotspot. \textbf{Methods}: We used 15 seabird at-sea ship-based survey
data collected over 37° of latitude, from 2016 to 2021, and fitted
Region of Common Profile (RCP) mixture models for each season and type
of data (presence-absence and abundance). RCP groups are defined as a
set of regions where the probability of encountering a specific species
profile is constant within regions, but different amongst them. These
groups also vary according to covariates, which in our case included
oceanographic, climatic, and physiographic parameters. \textbf{Results}:
Based on 142,646 seabirds recorded, from 80 species, all models
suggested two macro-scale assemblages (`northern' and `southern'),
except for the autumn presence-absence model that identified three
groups. The model results consistently show a biogeographic transition
at \textasciitilde34°S, near the latitude at which the East Australian
Current separates from the Australian continental slope. Sea surface
temperatures or sea surface salinities were selected in all final
models, further indicating a close relationship between seabird
assemblages and water masses. Results from both data types,
presence-absence and abundance, resulted in similar spatial and species
profile patterns. \textbf{Conclusion}: RCP models clearly identified two
seabird assemblages off the east coast of Australia, suggesting the
persistence of these groups at seasonal and macro spatial scales. Given
the rapid climatic changes that the East Australian Current is
experiencing and its importance in influencing distributions of
seabirds, the methods applied in our study could be replicated to assess
possible changes in seabird assemblages, and how they are affected by
changing environmental conditions.
\end{abstract}

\keywords{
    Assemblages
   \and
    Community ecology
   \and
    East Australian Current
   \and
    Mixture models
   \and
    Region of Common Profiles
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Bioregionalization is the discipline that identifies, delimits,
characterises, and names biogeographical regions (Morrone, 2018; Vilhena
\& Antonelli, 2015). Although there is often no agreement on the naming
systems among proposed regionalization schema (but see Ebach et al.,
2008), the hierarchical nature of bioregions is well recognized (Kreft
\& Jetz, 2010; Vilhena \& Antonelli, 2015). Thus, bioregions are usually
described within geographically larger classifications, such as
`provinces' or `realms' (Costello et al., 2017; Spalding et al., 2007).
The delimitation of bioregions is made through identifying sets of
species that are endemic or have high levels of co-occurrence, and are
distinct enough from other regions. In this sense, bioregions are
interpreted as a natural representations of common historical and/or
ecological processes shaped by physical and biological forces (Kreft \&
Jetz, 2010). In particular, the identification of bioregions allows
representation of the distribution of multiple species at a given time
and space, and relating these regions to environmental characteristics.
By doing so, bioregions can be interpreted as species assemblages (N. A.
Hill et al., 2017; Lyons et al., 2017). Given these features, bioregions
can be used as surrogates for biodiversity monitoring in data-poor
regions and in regional conservation planning (Ferrier, 2002; Spalding
et al., 2007), playing a fundamental role in conservation biogeography
(Whittaker et al., 2005).

Seabird assemblages have been widely described and related to
environmental characteristics (Hyrenbach et al., 2007; Ribic et al.,
1997; Woehler et al., 2003). Considering the hierarchical nature of
oceanographic processes and their resulting features (Haury et al.,
1978), it is not surprising that the distribution of highly mobile
seabirds are linked differently across temporal and spatial scales
(Haury et al., 1978; Hunt \& Schneider, 1987). In the marine
environment, large scale (\textgreater100 km length scale) oceanographic
features (e.g.~boundary currents and water masses) act as terrestrial
biomes (based on vegetation), and influence seabird assemblage
structure--i.e.~the occurrence of species and their relative numerical
composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al.,
1997). In contrast, local abundance is related to meso- (10--100 km) and
sub-mesoscale (1--10 km) processes that aggregate prey and therefore
predators, such as fronts and eddies (Scales, Miller, Hawkes, et al.,
2014; Weimerskirch, 2007). Such processes, although relatively
ephemeral, are primarily driven by physical and atmospheric forces, and
are thus somewhat predictable in space and time, providing consistent
feeding areas for marine predators when they occur (Davoren, 2013;
Hazen, Suryan, et al., 2013). Sea surface temperatures are a key driver
of seabird distributions, at individual- and assemblage-levels,
operating over a range of spatial and temporal scales from macro
(e.g.~water masses) to fine (e.g.~fronts) (Dunlop et al., 1988;
Quillfeldt et al., 2015; Ribic et al., 1997). Despite extensive surveys
(e.g. Ballance et al., 1997) and tracking studies (e.g. Carneiro et al.,
2020), many oceanic regions still lack information on the distributions
and abundances of seabird species, and how they assemble in space and
time.

One area in need of better understanding of seabird distributions is
Australia, considered a global seabird hotspot (Karpouzi et al., 2007;
Ramírez et al., 2017). Given the region's importance for seabird
diversity, there are surprisingly few studies on seabird distributions
at sea (Bernard et al., 2021; Mott \& Clarke, 2018). Off northwest
Australia, at-sea surveys, multivariate analyses (Dunlop et al., 1988)
and ensemble species distribution models of species assemblages (Lavers
et al., 2014) showed that habitat use by tropical seabird species was
related to sea surface temperatures and salinity. Dunlop et al. (1988)
suggested that salinity was the main variable influencing species'
distributions in oceanic waters while temperature was the key factor for
species on the continental shelf. In contrast, there are few studies at
large spatial and temporal scales off eastern Australia. T. A. Reid \&
Hindell (2000) and Evans et al. (2021) analysed seabird assemblages and
their relationships with oceanographic and prey data at coarse scales
(tens of km) off southeast Tasmania. The remaining studies are mostly
descriptive accounts locally (Blaber \& Milton, 1994; Norman, 1992), or
focused on a single species (Priddel et al., 2014).

The East Australian Current (EAC) is the major western boundary current
in the South Pacific Ocean, and the dominant boundary current off
eastern Australia (Ganachaud et al., 2014). The EAC has been extensively
studied, although research into the links between the EAC's oceanography
and marine predators are not well known. The dynamic features associated
with the EAC, such as mesoscale eddies and upwelling, are likely to
influence marine biota (Scales, Miller, Hawkes, et al., 2014; Suthers et
al., 2011). While most of the biological studies in this system were
focused on plankton entrapment within the cores of eddies (Firme et al.,
2023; Garcia et al., 2022; Hassler et al., 2011), there is a growing
body of work focusing on fisheries (Brieva et al., 2015; Hobday, 2010;
Young et al., 2011). These studies emphasized how the probability of
occurrence of a given taxon, and the structure of species assemblages,
were driven by the EAC. However, the relationships between oceanography
and marine megafauna within the EAC are poorly known (Suthers et al.,
2011). Some progress has been made using shark and ray movement data
(Jaine et al., 2014; Niella et al., 2020; Niella et al., 2022), and
satellite tracking has also demonstrated that increased sea surface
temperatures in the EAC decreases little penguin \emph{Eudyptula minor}
foraging success (Carroll et al., 2016), but information at the
assemblage-level for seabirds in the EAC is absent.

Marine species range shifts are more likely to occur faster than
terrestrial species due to their higher sensitivity to warming (Lenoir
et al., 2020; Pinsky et al., 2019). As a result, climate change is
rearranging the structure of species assemblages around the globe (Antão
et al., 2020; Dornelas et al., 2014; Poloczanska et al., 2013). Top
predators, for instance, may change their core habitats (Hazen,
Jorgensen, et al., 2013; Reisinger et al., 2022), with unprecedented
population consequences (Woehler \& Hobday, 2023). For seabirds, Péron
et al. (2010) and Sojitra et al. (2022) showed that sea surface
temperature anomalies affected the occurrence and abundances of
albatrosses and petrels species with latitude, thus affecting assemblage
structure. In eastern Australia, changing EAC conditions have affected a
range of marine species. For example, changes in zooplankton assemblage
structure and fish distribution patterns have been recorded and
attributed to warming water temperatures, and shifts in shark
distributions are expected (C. R. Johnson et al., 2011; Niella et al.,
2022). A severe marine heatwave driven by a poleward strengthening of
the EAC, showed drastic environmental impacts including disease
outbreaks in aquaculture farms and wildlife mortalities (Oliver et al.,
2017). Given that the EAC is strengthening and extending farther
southward and getting warmer (Cai et al., 2005; Phillips et al., 2022),
biological changes are expected to occur (Hobday, 2010; Poloczanska et
al., 2007).

In light of eastern Australia's importance for seabirds, and the recent
attention on how predicted climate change will influence the EAC's
marine biota (Poloczanska et al., 2007), an assessment of the
contemporary spatial distributions of seabirds within the EAC at large
temporal and spatial scales is timely. Seabirds can serve as sentinels
for tracking ecological changes in the EAC in many ways (e.g. Sydeman et
al., 2012). Surprisingly though, seabird distribution is
under-represented in climate change studies, despite being highlighted
as a priority topic (Grémillet \& Boulinier, 2009; Sydeman et al.,
2012).

In this study, we aim to identify assemblages of seabird species within
the EAC and adjacent areas, and relate these assemblages to
environmental conditions, by using at-sea surveys data and a
state-of-the-art modelling technique. We use the Region of Common
Profiles {[}RCP; Foster et al. (2013){]}, a mixture model regarded as
one of the best available tools for describing bioregions (N. A. Hill et
al., 2020; S. N. C. Woolley et al., 2020). RCP is a model-based
approach, and has a range of advantages over traditional algorithm-based
statistics (N. A. Hill et al., 2020; Warton et al., 2015). Eastern
Australia supports a high diversity of seabirds and offers wide
environmental gradients over a near 40° latitudinal range, from the
tropics to the cool temperate. We hypothesised to find different seabird
assemblages relating to different environmental conditions within the
EAC. In addition, presence-absence and abundance data were used to test
if the different numerical resolutions would result in the
identification of similar assemblages. This study provides a repeatable
framework on which future studies can compare to assess how future
changes in the EAC may influence the distribution of seabird
assemblages.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{study-area}{%
\subsection{Study area}\label{study-area}}

The EAC originates from the South Equatorial Current (SEC), a tropical
nutrient-poor current. When the SEC's North Caledonian Jet meets the
Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing
branch towards the Gulf of Papua and a south-flowing branch, which marks
the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte,
2013). The South Caledonian Jet completes the EAC system at
\textasciitilde22°S after encountering the shelf break in the southern
portion of the Great Barrier Reef (\textbf{Fig. 1a}; Kessler \& Cravatte
(2013); Ganachaud et al. (2014)). \textbf{SST/SSS characteristics? ROB?}
The EAC interacts with the bathymetry of the seafloor, giving rise to
oceanographic features and processes, such as smaller currents,
upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019;
Oke \& Middleton, 2000; Ridgway \& Dunn, 2003). The EAC flow polewards
and is primarily associated with the shelf break, but at
\textasciitilde32°S it detaches from the shelf and spawns an area of
eddy generation that dissipates throughout the western Tasman Sea (Mata
et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension'
(Oke, Pilo, et al. (2019); often termed the `Tasman Front'). Although
most EAC energy is converted into eddies, the EAC still runs southwards
following the shelf break towards Tasmania (the `EAC southern
extension'; \textbf{Fig. 1a}), carrying eddy-like features as a
transient current (Oke, Roughan, et al., 2019).

\hypertarget{seabird-and-environmental-data-sources}{%
\subsection{Seabird and environmental data
sources}\label{seabird-and-environmental-data-sources}}

Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV
Falkor} on 15 voyages, between 2016/17--2020/21, spanning over
\textasciitilde37° of latitude (\textbf{Fig. 1b}; \textbf{Table S1}).
One or more observers continuously recorded seabirds within a 300 m
forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours
while the ship was underway. The species, the numbers of birds, and
their behaviours were logged in real time into a ship-based web server.
\textbf{Environmental parameters for atmospheric and oceanographic
conditions were logged continuously by automated-ship based sensors,
with seabird observations supplementing relevant environmental
conditions such as visibility or associations with marine mammals.}
Individual records received a date/time/geographic coordinate label
(multi-species records have the same label but are logged as separate
entries for each species). Birds following the vessels were excluded
from analyses. Seabird observation protocols followed Raymond et al.
(2010), in use around Australia since 1980.

To investigate and predict species' distributions and relationships with
environmental variables, a considered choice of \textbf{broad-scale
climatic and oceanographic} covariates is important (Barbet-Massin \&
Jetz, 2014). We used three physiographic variables and seven
oceanographic variables, from which a further three climatic layers were
calculated (\textbf{Table 1}). These variables are the most frequently
used in seabird-environment relationship studies (Tremblay et al.,
2009), capturing processes at a range of spatial and temporal scales
from macro-scale water-mass properties to prey availability in meso- to
submesoscale oceanographic features (Haury et al., 1978; Scales, Miller,
Hawkes, et al., 2014; Weimerskirch, 2007).

\hypertarget{data-treatment}{%
\subsection{Data treatment}\label{data-treatment}}

Data wrangling, modelling and visualization were done in R 4.2.0 (R Core
Team, 2022); the full list of packages can be found in the Supplementary
Material \textbf{XX}. Code and the modelling dataset (see
\ref{sec:modellingdatasets}) are available in \textbf{Daudt et al.~2023
(Zenodo repo)}.

\hypertarget{environmental-layers}{%
\subsubsection{Environmental layers}\label{environmental-layers}}

The spatial and temporal persistence of oceanographic features such as
eddy kinetic energy (`eke') and thermal fronts (sea surface temperature
gradient: `sst\_grad') may influence the movements and distributions of
seabirds by offering a higher likelihood of finding food (Scales,
Miller, Embling, et al., 2014; Weimerskirch, 2007). Thus, `eke' and
`sst\_grad' environmental layers were calculated before extracting
environmental data for each seabird record. For both `eke' and
`sst\_grad', we used their original cell spatial resolution
(\textbf{Table 1}) and calculations were done layer-by-layer to match
seabird data on the finest spatial scale possible. Eddy kinetic energy
was calculated as \(0.5 (u'² + v'²)\), where \(u' = u - \bar{u}\) and
\(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian
(longitudinal) geostrophic currents, respectively, using a rolling mean
of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022). The
spatial `sst\_grad' for each cell was calculated by averaging the
maximum value for each neighbouring cell (average neighbourhood
maximum), from north-to-south (NS) and west-to-east (WE); then, the
magnitude of the gradient was defined as \(NS² + WE²\) (Burrows et al.,
2011). In addition, we calculated climatic seasonal average layers for
`eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a
proxy for persistent oceanographic features, and the `eke' standard
deviation over the same period (named with the prefix `clim\_' in
\textbf{Table 1}). Seabed slope was calculated using `raster::terrain()'
(Hijmans, 2022a), and distance from the coast was calculated through
`sf::st\_distance()' (Pebesma, 2018) using Australia and Tasmania as the
mainland (i.e.~ignoring smaller islands). Then, for each seabird record,
the most proximate value for each environmental layer (original or
calculated) was extracted using built-in functions from `raster' 3.5-21,
`terra' 1.6-7, and `sf' 1.0-8 packages (Hijmans, 2022a, 2022b; Pebesma,
2018).

\hypertarget{modellingdatasets}{%
\subsubsection{Data}\label{modellingdatasets}}

\label{sec:modellingdatasets}

Seabirds are highly mobile animals, and many species migrate and/or
disperse during their non-breeding periods (Schreiber \& Burger, 2002).
Therefore, seasonal differences in species composition were expected
(e.g. K. Reid et al., 2023), and in the probabilities of occurrence and
in total numbers. Thus, we built separate modelling dataset for each
season. We aggregated seabird data by 1° latitude \(\times\) 1°
longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Fig.
1b}), by summing seabirds from the same species. To test if the
presence-absence and abundance (raw counts) data would result in similar
assemblage patterns, we built two species matrices, where the
presence-absence matrix transformed values \textgreater1 to 1. To avoid
numerical instabilities in the models prompted by unusual or rare
species, we kept only species occurring in at least six grids within
each season (summer = 12 species; autumn = 21; winter = 20; spring =
33). The associated environmental data of each seabird record were
averaged within the same grid cells and seasons, and environmental data
were standardised before modelling.

\hypertarget{statistical-modelling}{%
\subsection{Statistical modelling}\label{statistical-modelling}}

We used the mixture model Region of Common Profiles (RCP) to identify
and describe seabird species assemblages. RCP is a one-stage
bioregionalization approach, as it identifies bioregions by specifying a
statistical model that simultaneously relates response (multi-species
data) and predictor (environmental data) (N. A. Hill et al., 2020; S. N.
C. Woolley et al., 2020). Given a set of environmental variables, the
model can predict assemblage probabilities of sites where no biotic
(observational) data exist. As a model-based approach, RCPs can be
compared by information criteria, present uncertainty around the
results, and check for model assumptions through assessing distribution
of residuals (N. A. Hill et al., 2020; Warton et al., 2015).

The model is specified in terms of assemblages and site-specific
probabilities. There are \(K\) assemblages (i.e.~RCPs) that are common
across the study area, \(K = (1, ..., k)\). The outcome variable is a
vector \(y_{i} = (y_{i1}, ... y_{iS})\), that give the count or
presence-absence of each of the \(S\) species at site \(i\). Depending
on whether we model the counts or presence-absence data, the assemblages
are defined in terms of the expected frequency or relative frequency of
the species, respectively. An output of the model is the estimated
probability that site \(i\) has each of the \(K\) possible assemblages.
RCPs are a multivariate adaptation of a mixture-of-experts model; as
such, the mixing proportions (assemblage structure) vary according to
covariates (predictors) (Foster et al., 2013). However, the final
estimated assemblage structure is assumed to be constant for all sites
belonging to the \(K\) assemblage (Foster et al., 2013). As such, RCPs
can be directly interpreted as assemblages, and their species profiles
as the assemblage structure.

Before fitting models, covariates were checked for collinearity within
each seasonal dataset. If a pairwise collinearity of
\(\left|x\right| >0.7\) was found we removed one of the variables
(Dormann et al., 2013), favouring direct measures over the `climatic'
layers we calculated (e.g.~if `sst\_grad' and `clim\_sst\_grad' were
correlated, we chose `sst\_grad'). Thus, `full models' differed between
seasons as the retained covariates varied, although most covariates were
common for all seasons.

Due to seabird migration and dispersion, seasons could act as `sampling
artifacts' (Foster et al., 2017), thus biasing results (Ponti \&
Sannolo, 2023). In light of the difference among seasons in the numbers
of species that met our occurrence threshold (see
\ref{sec:modellingdatasets}), we decided to use RCP models by season and
data type (presence-absence and abundance). RCP models were used to
relate the response variables (i.e.~presence-absence or abundance of
each seabird species in each grid cell) to the independent predictor
covariates (i.e.~averaged values of environmental variables in each
cell). Presence-absence data were modelled based on a Bernoulli model,
whereas abundance data based on a Negative Binomial model. RCPs were
fitted using the `ecomix' 1.0.0 package (S. Woolley et al., 2022), with
the following workflow: (i) find the best number of groups (\(K\)) that
describes the data, (ii) re-fit the model by fixing \(K\) and selecting
optimal combination of covariates, (iii) check residuals, (iv) bootstrap
the best model to estimate uncertainty around parameter estimates, and
(v) produce results.

To identify the best number of RCP groups (\(K\)) that described the
data, we used full models (all covariates) with 100 random start numbers
(see Foster et al., 2013) between \(k = (2, ..., 7)\) groups, and chose
the number of RCPs that minimized the Bayesian Information Criterion
(BIC). Then, we fixed \(K\) at the optimal number and fitted models for
every possible linear, additive combination of covariates. We chose the
best model as the one with the lowest BIC. Visual inspection of
randomized quantile residual plots modified for mixture models was used
to verify model assumptions (Dunstan et al., 2013). To obtain estimates
of uncertainty for model parameters (confidence intervals), we used the
default settings of `ecomix::regional\_mix.bootstrap()', which applies a
Bayesian bootstrap 1000 times. A custom code was written to generate
partial plots for exploring the influence of covariates on RCP
membership.

We built a seasonal 11-year (2011--2021) average layer for predicting
model results for each environmental covariate. These variables were
then averaged within 1° latitude \(\times\) 1° longitude grid cells, by
season, between 9--47°S and 141--161°E. Note that `clim\_eke\_mean' and
`clim\_sst\_grad' were processed beforehand as they may characterise
persistent oceanographic features (described above) and thus represent
the same value as `eke' and `sst\_grad'. Data for prediction were
standardised using the same mean and standard deviation from the
modelling dataset.

Finally, we ran rarefaction curves based on the presence-absence matrix
and the assigned RCP group for each sample (grid cell) to assess if the
amount of information in the RCP model was representative. Rarefaction
curves were fit using the `iNEXT' 3.0.0 package (Chao et al., 2014;
Hsieh et al., 2016), and results are presented based on alpha diversity
(i.e.~species richness).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{summary-of-species-richness-and-numbers}{%
\subsection{Summary of species richness and
numbers}\label{summary-of-species-richness-and-numbers}}

A total of 10,261 seabird records was obtained during the 15 voyages,
comprising 142,646 individuals from 80 species, belonging to 11 families
and five orders. The five most abundant species made up 83\% of total
birds counted (43.0\% short-tailed shearwater \emph{Ardenna
tenuirostris}, 17.2\% fairy prion \emph{Pachyptila turtur}, 14.3\%
wedge-tailed shearwater \emph{Ardenna pacifica}, 5.2\% sooty tern
\emph{Onychoprion fuscatus}, 3.2\% shy albatross \emph{Thalassarche
cauta}). The five most frequently encountered species were \emph{A.
pacifica} (frequency of occurrence = 35.3\%), \emph{T. cauta} (33.2\%),
brown booby \emph{Sula leucogaster} (25.4\%), red-footed booby \emph{S.
sula} (25.4\%) and black-browed albatross \emph{T. melanophris} (25\%)
(\textbf{Fig. S1}). An increase in species richness towards the south of
the study region was clear in all but the winter season (\textbf{Fig
S2}). Noticeably higher counts were made around Tasmania, possibly due
to the `colony effect' and/or the presence of Southern Ocean species
(\textbf{Fig. S2}). However, one cell in the Coral Sea had counts of the
same order of magnitude as off Tasmania in summer.

\hypertarget{rcp-models}{%
\subsection{RCP models}\label{rcp-models}}

For all models, two RCPs was the optimal number of seabird groups,
except for the presence-absence-autumn model, which indicated three RCPs
as the best fit (\textbf{Fig. S3}). The best fitting models generally
had 1 or 2 covariates selected, although the abundance-autumn model had
five covariates (\textbf{Table 2}). The best models showed a good fit
with no apparent deviations from the assumptions (\textbf{Fig. S4}). All
the best models selected `sst' or `sss', as essential for characterising
water masses. The probability of belonging to one or another RCP varied
its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer
(\textbf{Fig. 2}, \textbf{Fig. S5}), similar to the EAC's temperature
and salinity thresholds. Furthermore, most of the best models selected
`bat', suggesting a gradient between `coastal' and `oceanic' taxa based
on bathymetry (\textbf{Fig. S5}). Of particular note, proxies for
persistent oceanographic features that may aggregate food resources (the
`climatic' variables) for seabirds were selected only in the autumn and
spring abundance models.

Grid point-predictions are the assignment of the RCP with the highest
probability value for a given cell. When predicting RCP onto space,
patterns of point-predictions suggest a biogeographic transition in
seabird species around \textasciitilde34°S, with both presence-absence
and abundance models pointing to similar results (\textbf{Fig. 3}).
These patterns are further reinforced when looking at the mean
probability along with the lower and upper confidence intervals of
predictions (\textbf{Fig. 4}, \textbf{Fig. S6}), showing that around
\textasciitilde34°S, the `northern' RCP tended to have higher
probabilities in coastal regions, whereas the `southern' RCP spread
northward offshore. The presence-absence-autumn model (three RCPs) had a
similar broad-scale pattern, although the southern RCP is farther south,
and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Fig.
3}, \textbf{Fig. S6}).

The broad pattern of two large-scale assemblages is reflected in their
species compositions. Although there are overlaps among the seabird
groups, both presence-absence and abundance models captured typical
tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp.,
lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean
species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche}
spp., fairy prion \emph{P. turtur}) (\textbf{Fig. 5}, \textbf{Fig. S7}).
The presence-absence-autumn model, which had three RCPs, also showed the
characteristic taxa representing the `northern' and `southern' RCPs.
However, the intermediate `Tasmania/Bass Straight' assemblage showed a
mixture of both groups with shy albatross and Australasian gannet
\emph{Morus serrator} prominent with high probabilities of occurrence.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This study showed that seabird assemblages off eastern Australia were
correlated with oceanographic conditions, and that the East Australian
Current appeared to be key in structuring seabird distributions at meso-
to macro scales (\emph{sensu} Hunt \& Schneider, 1987). Our results are
based on a multivariate statistical model (N. A. Hill et al., 2020),
which brings advantages over classical methods for describing species
assemblages. RCP models allow estimating uncertainty around the point
probability of a sample (grid cell) belonging to an assemblage, thus
representing the fuzziness between bioregion boundaries. These
transition zones between bioregions are essential for monitoring
biological changes due to environmental changes (Morrone, 2023). As
such, the boundary between the two consistently identified bioregions in
this study may be a crucial region to monitor in future, given
noticeable anomalies in the EAC at this same region (Phillips et al.,
2022). Our results draw an important baseline for seabird assemblages in
a world hotspot for seabirds.

\hypertarget{seabird-bioregions-off-eastern-australia}{%
\subsection{Seabird bioregions off eastern
Australia}\label{seabird-bioregions-off-eastern-australia}}

The relationships between seabird assemblages and water masses has been
known for a long time (Elphick \& Hunt, 1993; Pocklington, 1979; Ribic
et al., 1997). Thus, it is not surprising that two seabird assemblages
were identified off eastern Australia. However, the spatial consistency
of these patterns throughout seasons is surprising and remarkable.
Resident species seem to characterise these assemblages. Shy albatross
(Hedd et al., 2001; Mason et al., 2018) and fairy prions (Fromant et
al., 2022) in the southern assemblage, and boobies (Bunce, 2015; Miller
et al., 2018) and wedge-tailed shearwaters (McDuie et al., 2015;
Weimerskirch et al., 2020) in the northern assemblage, were always
profiled with higher probabilities of occurrence and predicted
abundances compared to other species in the same groups. Given that
changes in assemblages over time are typically driven by a few species
(Gotelli et al., 2022), and that those species are likely residents
(Souza \& Santos, 2023), the relationships between these resident
seabirds and the EAC system over the broad latitudinal range detailed
here should be closely monitored. Elsewhere, coastal seabirds such as
little penguin are signalling the adverse effects that climate change
might have in changing seabird behaviours and decreasing foraging
success (Carroll et al., 2016; Woehler \& Hobday, 2023), which
ultimately influence their breeding success and population dynamics (B.
Johnson \& Colombelli-Négrel, 2021). Thus, the species profiles and the
spatial predictions of assemblages from RCP models could be used as
tools for monitoring future seabird responses to the fast-changing EAC.

The seasonal models also captured the influence of migratory species,
particularly in the southern assemblage. The summer and autumn models
reflected the high abundance of short-tailed shearwater before they
depart on their trans-equatorial migration (Carey et al., 2014); the
presence of the Australasian gannet was picked up by the autumn and
winter models (Ismar et al., 2011); and the white-headed petrel
\emph{Pterodroma lessonii} movements during chick-rearing and pre-laying
exodus (Taylor et al., 2021) were detected in autumn and spring models,
respectively. Given the geographic location of the southern assemblage,
however, we believe sub-Antarctic taxa were not well represented in our
models due to their low number of occurrences and relative paucity of
survey effort to date (\textbf{Fig. S1}; hence they were removed before
modelling). Additional surveys may reveal the seasonal northward pulse
of these taxa into the southern assemblage (e.g.~albatrosses
\emph{Diomedea} spp. and \emph{Phoebetria} spp.), given that more effort
translates as higher likelihood of recording these species. In turn,
including these taxa in the models could result in a more stable `three
assemblage pattern' (as revealed by the autumn presence-absence model)
in all but the summer season, when the majority of Antarctic species are
breeding (S. Marchant \& Higgins, 1990).

Our analyses clearly revealed a `northern' and a `southern' assemblage
of seabirds off eastern Australia within the EAC. There is a noticeable
link with the EAC dynamics: the boundaries of RCPs occurred at
\textasciitilde34°S, near the average separation latitude
(\textasciitilde32°S) of the EAC and its eastward extension into the
Tasman Sea (Oke, Roughan, et al., 2019). However, the
`northern/southern' assemblage pattern identified herein does not agree
with previously-proposed large-scale regionalizations, such as the
pelagic Longhurst Provinces (Longhurst, 2007), the Marine Ecoregions of
the World {[}MEOW; Spalding et al. (2007){]}, or the Large Marine
Ecosystems {[}LME; Sherman (1994){]}. The MEOW schema was based on
coastal and shelf biota and was meant to represent habitats to the 200 m
isobath, which clearly does not encompass the entirety of seabird
habitats. The LME schema is characterised by physiographical,
hydrological and productivity data but has a strong focus on practical
management of areas and thus there is a political component embedded
within it, which does not align with seabird distributions. Longhurst
(2007) designated only one biogeochemical province for eastern Australia
based on productivity patterns, which may work well at larger (macro)
scales, but appears to be oversimplified for the EAC.

The results do agree with the marine biogeographic units proposed by
Costello et al. (2017), who identified a boundary at \textasciitilde34°S
between their `Coral Sea' and `South Australia' realms. Their approach
used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats,
thus integrating more complexity of the biogeographic boundaries in the
marine environment. Other biological data also suggest
`northern/southern' assemblages off eastern Australia, including tuna
and billfish {[}also highly mobile top predators; Reygondeau et al.
(2012){]}, mesopelagic fauna {[}which includes seabird prey; Sutton et
al. (2017){]} and deep-sea invertebrates and fish (O'Hara, Williams,
Althaus, et al., 2020; O'Hara, Williams, Woolley, et al., 2020). All
these studies focused mainly on describing biological assemblages, as
opposed to classifying environments. It is worth noting that the
Longhurst Provinces approach was updated to include SST, SSS and
bathymetry (the same variables retained by our best RCP models) and a
temporal component to the classification (Reygondeau et al., 2013). The
updated provinces are more similar to the seabird assemblages described
here, although still not a perfect match, which could be partially
explained because the seabird assemblage at EAC region includes several
long-distance migratory species. Australia's bioregionalization has a
long story of disagreements (Ebach, 2012), however, the inherent link
that RCPs develop between environment and biological data can be a
strong tool to support marine bioregionalization.

\hypertarget{fitting-rcp-models-to-high-mobile-animals}{%
\subsection{Fitting RCP models to high mobile
animals}\label{fitting-rcp-models-to-high-mobile-animals}}

To the best of our knowledge, this is the first study applying RCPs to
highly mobile animals such as seabirds. RCP models have been previously
applied to plants (Lyons et al., 2017), and relatively sedentary
animals, such as macrobenthic organisms (Pantiukhin et al., 2021) and
demersal fish (N. A. Hill et al., 2017). Oceanic seabirds can have daily
movements of hundreds of kilometres (e.g. McDuie et al., 2015), during
which they traverse a variety of habitats and environmental gradients.
As such, capturing the distribution of highly mobile species, and their
relationships with environmental parameters is not a trivial task, and
much harder than for sessile organisms.

While remote tracking studies are advancing our understanding on how
individual seabirds use their environment at fine scales (Tremblay et
al., 2009), they fail to perceive how species assemblages interact with
the environment and each other. Therefore, using at-sea counts of
seabirds is a well-established method to capture these relationships
(Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
Further, the spatial scales on which assemblages relate to their
environment is variable and species-specific (Hunt \& Schneider, 1987;
Ribic et al., 1997; Scales et al., 2017). Understanding the biology of
the studied group is thus essential to set the appropriate scales of
analyses. Our choice of 1° latitude \(\times\) 1° longitude was mainly
determined by data availability reflecting survey effort, and the
trade-off between the number of species per grid/season (see also
\ref{sec:limitations}). Nonetheless, the chosen spatial scale is similar
to other seabird distribution analyses conducted elsewhere (Harris et
al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022). Therefore, for
describing seabird assemblages at a macro-scale, the chosen resolution
seems to had been a good fit, accords with studies elsewhere, and the
resultant models described patterns of seabird species' composition
consistent with our current understanding of species' ecologies and life
history strategies.

RCP models based on presence-absence and abundance data are able to
complement each other. Previous studies using RCPs have opted to use
only one type of response data. Lyons et al. (2017), N. A. Hill et al.
(2017), and Pantiukhin et al. (2021) all used presence-absence data,
whereas Lee et al. (2019) and Receveur et al. (2020) used abundance
(count) data. The added information contained in abundance data make
them inherently more powerful than binary presence-absence data.
However, the use of different data inputs in multivariate, biodiversity
assessments and assemblage descriptions has shown contrasting patterns
at different spatial scales. At larger spatial and temporal scales,
presence-absence data describe similar patterns to abundance data
(Anderson et al., 2005), as we found. In contrast, at smaller scales,
abundance data may highlight local conditions and therefore reveal
hidden patterns at coarse to fine scales (Anderson et al., 2005; R.
Marchant, 1990; Melo, 2005; Waters, 2008). If combined with ecological
knowledge of a particular group, the information contained in both
presence-absence and abundance data could be leveraged to better
interpret the resulting patterns. For example, seabirds will likely
occur in specific water masses or currents (Hunt \& Schneider, 1987;
Ribic et al., 1997), whereas their abundance is related to coarse- to
fine-scale processes that enhance prey availability (Hunt \& Schneider,
1987; Weimerskirch, 2007). Future studies assessing bioregions of highly
mobile animals could describe the compositions of assemblages using
smaller resolution grid cells and presence-absence models. Then,
abundance models using higher resolution grid cells could improve our
understanding of how these animals relate to their environment and
possibly reveal multi-species aggregation areas (therefore `new
bioregions') likely nested within the bioregions described at large
scales. Such a hierarchical approach would require substantial
concurrent observational and environmental data sets, which we suspect
not exist at present.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\label{sec:limitations}

Although the results suggest consistent patterns, there are a number of
caveats to our analyses. Seabird data were gathered during
multi-disciplinary cruises, whose sampling designs were not specifically
planned to account for spatial ecology of seabirds. As a result, there
are spatial discontinuities in the data, which might have limited the
ability of our models to detect the relationships between the seabird
assemblages and the full spectrum of environmental predictors. Even
though Australia is a hotspot for seabirds (Karpouzi et al., 2007;
Ramírez et al., 2017), there is a paucity of information on seabird
distributions in the region (Bernard et al., 2021; Mott \& Clarke,
2018). Dedicated seabird surveys would help to advance seabird research
in this region, and improve our understanding of the patterns described
here. By increasing the amount of data, future studies could set the
grid cells at higher resolution, and potentially reveal assemblages at
finer spatial and temporal scales. In addition, ideally, we would have
had more information within each grid cell by season (our sampling
units). Nonetheless, rarefaction curves show that the sample coverage
(see Chao \& Jost, 2012) was high at the RCP level, although there is
room for detecting more species in most of them (\textbf{Fig S8}).

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

We applied a robust, repeatable method to assess seabird assemblages off
eastern Australia within the EAC for the first time. The models point
toward a consistent pattern of a `northern' and a `southern' seabird
assemblage, where the spatial division is attributable to the eastward
separation of the EAC from the coast, and that SST seems to be the key
variable discriminating these groups. The EAC has penetrated poleward
over the past half-century (K. L. Hill et al., 2008; Li et al., 2022), a
trend which is anticipated to continue into the future (Yang et al.,
2016), and extreme weather events such as marine heatwaves to happen
more often (Oliver et al., 2018). As such, seabirds will have to respond
to these changes (Woehler \& Hobday, 2023). Southern Ocean albatrosses
off southeast Australia, for instance, have shown `early stages of a
southward range shift' due to environmental variability (Sojitra et al.,
2022). Resident species such as the little penguins and wedge-tailed
shearwater change their behaviour and have decreased foraging success
when SST is higher (Carroll et al., 2016; Peck et al., 2004). Given that
eastern Australia is a biodiversity and numerical hotspot for seabirds,
it is imperative to establish a monitoring program allowing the
assemblages and species' profiles presented here to be used as the
baseline for understanding possible effects of a warming ocean on
seabird distribution. In particular, the transition zone between
assemblages (Morrone, 2023), associated with the eastern extension of
the EAC, may be fundamental to effectively monitoring strengthening of
the EAC and the subsequent impact on marine biota.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank all observers who assisted in data collection, the crews of RV
\emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird
Group (BirdLife Australia) for logistical support. The authors would
like to thank Dr.~Skipton N.C. Woolley, \texttt{\{ecomix\}} developer,
who kindly helped NWD build preliminary models and gave advice in
several stages of model construction. NWD is supported by a Doctoral
scholarship from the University of Otago.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ainley2012}{}}%
Ainley, D., Ribic, C., \& Woehler, E. (2012). Adding the ocean to the
study of seabirds: a brief history of at-sea seabird research.
\emph{Marine Ecology Progress Series}, \emph{451}, 231--243.
\url{https://doi.org/10.3354/meps09524}

\leavevmode\vadjust pre{\hypertarget{ref-anderson2005}{}}%
Anderson, M. J., Connell, S. D., Gillanders, B. M., Diebel, C. E., Blom,
W. M., Saunders, J. E., \& Landers, T. J. (2005). Relationships between
taxonomic resolution and spatial scales of multivariate variation.
\emph{Journal of Animal Ecology}, \emph{74}(4), 636--646.
\url{https://doi.org/10.1111/j.1365-2656.2005.00959.x}

\leavevmode\vadjust pre{\hypertarget{ref-antuxe3o2020}{}}%
Antão, L. H., Bates, A. E., Blowes, S. A., Waldock, C., Supp, S. R.,
Magurran, A. E., Dornelas, M., \& Schipper, A. M. (2020).
Temperature-related biodiversity change across temperate marine and
terrestrial systems. \emph{Nature Ecology \& Evolution}, \emph{4}(7),
927--933. \url{https://doi.org/10.1038/s41559-020-1185-7}

\leavevmode\vadjust pre{\hypertarget{ref-ballance1997}{}}%
Ballance, L. T., Pitman, R. L., \& Reilly, S. B. (1997). Seabird
community structure along a productivity gradient: importance of
competition and energetic constraint. \emph{Ecology}, \emph{78}(5),
1502--1518.
\url{https://doi.org/10.1890/0012-9658(1997)078\%5B1502:SCSAAP\%5D2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-barbet-massin2014}{}}%
Barbet-Massin, M., \& Jetz, W. (2014). A 40-year, continent-wide,
multispecies assessment of relevant climate predictors for species
distribution modelling. \emph{Diversity and Distributions},
\emph{20}(11), 1285--1295. \url{https://doi.org/10.1111/ddi.12229}

\leavevmode\vadjust pre{\hypertarget{ref-bernard2021}{}}%
Bernard, A., Rodrigues, A. S. L., Cazalis, V., \& Grémillet, D. (2021).
Toward a global strategy for seabird tracking. \emph{Conservation
Letters}, \emph{14}(3), e12804. \url{https://doi.org/10.1111/conl.12804}

\leavevmode\vadjust pre{\hypertarget{ref-blaber1994distribution}{}}%
Blaber, S., \& Milton, D. (1994). Distribution of seabirds at sea in the
{G}ulf of {C}arpentaria, {A}ustralia. \emph{Marine and Freshwater
Research}, \emph{45}(3), 445--454.

\leavevmode\vadjust pre{\hypertarget{ref-brieva2015}{}}%
Brieva, D., Ribbe, J., \& Lemckert, C. (2015). Is the East Australian
Current causing a marine ecological hot-spot and an important fisheries
near Fraser Island, Australia? \emph{Estuarine, Coastal and Shelf
Science}, \emph{153}, 121--134.
\url{https://doi.org/10.1016/j.ecss.2014.12.012}

\leavevmode\vadjust pre{\hypertarget{ref-bunce2015}{}}%
Bunce, A. (2015). Foraging behaviour of a declining population of brown
boobies ({\emph{Sula leucogaster}}) breeding in the Swain Reefs, Great
Barrier Reef. \emph{Emu - Austral Ornithology}, \emph{115}(4), 368--372.
\url{https://doi.org/10.1071/MU14037}

\leavevmode\vadjust pre{\hypertarget{ref-burrows2011}{}}%
Burrows, M. T., Schoeman, D. S., Buckley, L. B., Moore, P., Poloczanska,
E. S., Brander, K. M., Brown, C., Bruno, J. F., Duarte, C. M., Halpern,
B. S., Holding, J., Kappel, C. V., Kiessling, W., O'Connor, M. I.,
Pandolfi, J. M., Parmesan, C., Schwing, F. B., Sydeman, W. J., \&
Richardson, A. J. (2011). The pace of shifting climate in marine and
terrestrial ecosystems. \emph{Science}, \emph{334}(6056), 652--655.
\url{https://doi.org/10.1126/science.1210288}

\leavevmode\vadjust pre{\hypertarget{ref-cai2005}{}}%
Cai, W., Shi, G., Cowan, T., Bi, D., \& Ribbe, J. (2005). The response
of the Southern Annular Mode, the East Australian Current, and the
southern mid-latitude ocean circulation to global warming.
\emph{Geophysical Research Letters}, \emph{32}(23), L23706.
\url{https://doi.org/10.1029/2005GL024701}

\leavevmode\vadjust pre{\hypertarget{ref-carey2014}{}}%
Carey, M. J., Phillips, R. A., Silk, J. R. D., \& Shaffer, S. A. (2014).
Trans-equatorial migration of short-tailed shearwaters revealed by
geolocators. \emph{Emu - Austral Ornithology}, \emph{114}(4), 352--359.
\url{https://doi.org/10.1071/MU13115}

\leavevmode\vadjust pre{\hypertarget{ref-carneiro2020}{}}%
Carneiro, A. P. B., Pearmain, E. J., Oppel, S., Clay, T. A., Phillips,
R. A., Bonnet-Lebrun, A.-S., Wanless, R. M., Abraham, E., Richard, Y.,
Rice, J., Handley, J., Davies, T. E., Dilley, B. J., Ryan, P. G., Small,
C., Arata, J., Arnould, J. P. Y., Bell, E., Bugoni, L., \ldots{} Dias,
M. P. (2020). A framework for mapping the distribution of seabirds by
integrating tracking, demography and phenology. \emph{Journal of Applied
Ecology}, \emph{57}(3), 514--525.
\url{https://doi.org/10.1111/1365-2664.13568}

\leavevmode\vadjust pre{\hypertarget{ref-carroll2016}{}}%
Carroll, G., Everett, J. D., Harcourt, R., Slip, D., \& Jonsen, I.
(2016). High sea surface temperatures driven by a strengthening current
reduce foraging success by penguins. \emph{Scientific Reports},
\emph{6}(1), 22236. \url{https://doi.org/10.1038/srep22236}

\leavevmode\vadjust pre{\hypertarget{ref-chao2014}{}}%
Chao, A., Chiu, C.-H., \& Jost, L. (2014). Unifying species diversity,
phylogenetic diversity, functional diversity, and related similarity and
differentiation measures through Hill numbers. \emph{Annual Review of
Ecology, Evolution, and Systematics}, \emph{45}(1), 297--324.
\url{https://doi.org/10.1146/annurev-ecolsys-120213-091540}

\leavevmode\vadjust pre{\hypertarget{ref-chao2012}{}}%
Chao, A., \& Jost, L. (2012). Coverage-based rarefaction and
extrapolation: standardizing samples by completeness rather than size.
\emph{Ecology}, \emph{93}(12), 2533--2547.
\url{https://doi.org/10.1890/11-1952.1}

\leavevmode\vadjust pre{\hypertarget{ref-commins2014}{}}%
Commins, M. L., Ansorge, I., \& Ryan, P. G. (2014). Multi-scale factors
influencing seabird assemblages in the African sector of the Southern
Ocean. \emph{Antarctic Science}, \emph{26}(1), 38--48.
\url{https://doi.org/10.1017/S0954102013000138}

\leavevmode\vadjust pre{\hypertarget{ref-costello2017}{}}%
Costello, M. J., Tsai, P., Wong, P. S., Cheung, A. K. L., Basher, Z., \&
Chaudhary, C. (2017). Marine biogeographic realms and species
endemicity. \emph{Nature Communications}, \emph{8}(1), 1057.
\url{https://doi.org/10.1038/s41467-017-01121-2}

\leavevmode\vadjust pre{\hypertarget{ref-davoren2013}{}}%
Davoren, G. K. (2013). Distribution of marine predator hotspots
explained by persistent areas of prey. \emph{Marine Biology},
\emph{160}(12), 3043--3058.
\url{https://doi.org/10.1007/s00227-013-2294-5}

\leavevmode\vadjust pre{\hypertarget{ref-dormann2013}{}}%
Dormann, C. F., Elith, J., Bacher, S., Buchmann, C., Carl, G., Carré,
G., Marquéz, J. R. G., Gruber, B., Lafourcade, B., Leitão, P. J.,
Münkemüller, T., McClean, C., Osborne, P. E., Reineking, B., Schröder,
B., Skidmore, A. K., Zurell, D., \& Lautenbach, S. (2013). Collinearity:
a review of methods to deal with it and a simulation study evaluating
their performance. \emph{Ecography}, \emph{36}(1), 27--46.
\url{https://doi.org/10.1111/j.1600-0587.2012.07348.x}

\leavevmode\vadjust pre{\hypertarget{ref-dornelas2014}{}}%
Dornelas, M., Gotelli, N. J., McGill, B., Shimadzu, H., Moyes, F.,
Sievers, C., \& Magurran, A. E. (2014). Assemblage time series reveal
biodiversity change but not systematic loss. \emph{Science},
\emph{344}(6181), 296--299.
\url{https://doi.org/10.1126/science.1248484}

\leavevmode\vadjust pre{\hypertarget{ref-dunlop1988}{}}%
Dunlop, J., Wooller, R., \& Cheshire, N. (1988). Distribution and
abundance of marine birds in the Eastern Indian Ocean. \emph{Marine and
Freshwater Research}, \emph{39}(5), 661.
\url{https://doi.org/10.1071/MF9880661}

\leavevmode\vadjust pre{\hypertarget{ref-dunstan2013}{}}%
Dunstan, P. K., Foster, S. D., Hui, F. K. C., \& Warton, D. I. (2013).
Finite mixture of regression modeling for high-dimensional count and
biomass data in ecology. \emph{Journal of Agricultural, Biological, and
Environmental Statistics}, \emph{18}(3), 357--375.
\url{https://doi.org/10.1007/s13253-013-0146-x}

\leavevmode\vadjust pre{\hypertarget{ref-ebach2012history}{}}%
Ebach, M. C. (2012). A history of bioregionalisation in {A}ustralia.
\emph{Zootaxa}, \emph{3392}, 1--34.
\url{https://doi.org/10.11646/ZOOTAXA.3392.1.1}

\leavevmode\vadjust pre{\hypertarget{ref-ebach2008}{}}%
Ebach, M. C., Morrone, J. J., Parenti, L. R., \& Viloria, Á. L. (2008).
International Code of Area Nomenclature. \emph{Journal of Biogeography},
\emph{35}(7), 1153--1157.
\url{https://doi.org/10.1111/j.1365-2699.2008.01920.x}

\leavevmode\vadjust pre{\hypertarget{ref-elphick1993}{}}%
Elphick, C. S., \& Hunt, G. L. (1993). Variations in the distributions
of marine birds with water mass in the northern Bering Sea.
\emph{Condor}, \emph{95}(1), 33--44.
\url{https://doi.org/10.2307/1369384}

\leavevmode\vadjust pre{\hypertarget{ref-evans2021a}{}}%
Evans, R., Lea, M.-A., \& Hindell, M. A. (2021). Predicting the
distribution of foraging seabirds during a period of heightened
environmental variability. \emph{Ecological Applications}, \emph{31}(5),
e02343. \url{https://doi.org/10.1002/eap.2343}

\leavevmode\vadjust pre{\hypertarget{ref-ferrier2002a}{}}%
Ferrier, S. (2002). Mapping spatial pattern in biodiversity for regional
conservation planning: where to from here? \emph{Systematic Biology},
\emph{51}(2), 331--363. \url{https://doi.org/10.1080/10635150252899806}

\leavevmode\vadjust pre{\hypertarget{ref-firme2023}{}}%
Firme, G. F., Hughes, D. J., Laiolo, L., Roughan, M., Suthers, I. M., \&
Doblin, M. A. (2023). Contrasting phytoplankton composition and primary
productivity in multiple mesoscale eddies along the East Australian
coast. \emph{Deep Sea Research Part I: Oceanographic Research Papers},
\emph{193}, 103952. \url{https://doi.org/10.1016/j.dsr.2022.103952}

\leavevmode\vadjust pre{\hypertarget{ref-foster2013}{}}%
Foster, S. D., Givens, G. H., Dornan, G. J., Dunstan, P. K., \& Darnell,
R. (2013). Modelling biological regions from multi-species and
environmental data. \emph{Environmetrics}, \emph{24}(7), 489--499.
\url{https://doi.org/10.1002/env.2245}

\leavevmode\vadjust pre{\hypertarget{ref-foster2017}{}}%
Foster, S. D., Hill, N. A., \& Lyons, M. (2017). Ecological grouping of
survey sites when sampling artefacts are present. \emph{Journal of the
Royal Statistical Society: Series C (Applied Statistics)}, \emph{66}(5),
1031--1047. \url{https://doi.org/10.1111/rssc.12211}

\leavevmode\vadjust pre{\hypertarget{ref-fromant2022}{}}%
Fromant, A., Eizenberg, Y. H., Poupart, T., Bustamante, P., \& Arnould,
J. P. Y. (2022). Year-round at-sea movements of fairy prions from
southeastern {A}ustralia. \emph{Royal Society Open Science},
\emph{9}(5), 220134. \url{https://doi.org/10.1098/rsos.220134}

\leavevmode\vadjust pre{\hypertarget{ref-ganachaud2014}{}}%
Ganachaud, A., Cravatte, S., Melet, A., Schiller, A., Holbrook, N. J.,
Sloyan, B. M., Widlansky, M. J., Bowen, M., Verron, J., Wiles, P.,
Ridgway, K., Sutton, P., Sprintall, J., Steinberg, C., Brassington, G.,
Cai, W., Davis, R., Gasparin, F., Gourdeau, L., \ldots{} Send, U.
(2014). The Southwest Pacific Ocean circulation and climate experiment
(SPICE). \emph{Journal of Geophysical Research: Oceans}, \emph{119}(11),
7660--7686. \url{https://doi.org/10.1002/2013JC009678}

\leavevmode\vadjust pre{\hypertarget{ref-garcia2022}{}}%
Garcia, V., Schilling, H., Cruz, D., Hawes, S., Everett, J., Roughan,
M., Miskiewicz, A., Pakhomov, E., Jeffs, A., \& Suthers, I. (2022).
Entrainment and development of larval fish assemblages in two
contrasting cold core eddies of the East Australian Current system.
\emph{Marine Ecology Progress Series}, \emph{685}, 1--18.
\url{https://doi.org/10.3354/meps13982}

\leavevmode\vadjust pre{\hypertarget{ref-gotelli2022}{}}%
Gotelli, N. J., Moyes, F., Antão, L. H., Blowes, S. A., Dornelas, M.,
McGill, B. J., Penny, A., Schipper, A. M., Shimadzu, H., Supp, S. R.,
Waldock, C. A., \& Magurran, A. E. (2022). Long-term changes in
temperate marine fish assemblages are driven by a small subset of
species. \emph{Global Change Biology}, \emph{28}(1), 46--53.
\url{https://doi.org/10.1111/gcb.15947}

\leavevmode\vadjust pre{\hypertarget{ref-gruxe9millet2009}{}}%
Grémillet, D., \& Boulinier, T. (2009). Spatial ecology and conservation
of seabirds facing global climate change: a review. \emph{Marine Ecology
Progress Series}, \emph{391}, 121--137.
\url{https://doi.org/10.3354/meps08212}

\leavevmode\vadjust pre{\hypertarget{ref-harris2007}{}}%
Harris, J., Haward, M., Jabour, J., \& Woehler, E. J. (2007). A new
approach to selecting Marine Protected Areas (MPAs) in the Southern
Ocean. \emph{Antarctic Science}, \emph{19}(2), 189--194.
\url{https://doi.org/10.1017/S0954102007000260}

\leavevmode\vadjust pre{\hypertarget{ref-hassler2011}{}}%
Hassler, C. S., Djajadikarta, J. R., Doblin, M. A., Everett, J. D., \&
Thompson, P. A. (2011). Characterisation of water masses and
phytoplankton nutrient limitation in the East Australian Current
separation zone during spring 2008. \emph{Deep Sea Research Part II:
Topical Studies in Oceanography}, \emph{58}(5), 664--677.
\url{https://doi.org/10.1016/j.dsr2.2010.06.008}

\leavevmode\vadjust pre{\hypertarget{ref-haury1978patterns}{}}%
Haury, L., McGowan, J., \& Wiebe, P. (1978). Patterns and processes in
the time-space scales of plankton distributions. In J. Steele (Ed.),
\emph{Spatial pattern in plankton communities} (pp. 277--327). Springer.
\url{https://doi.org/10.1007/978-1-4899-2195-6_12}

\leavevmode\vadjust pre{\hypertarget{ref-hazen2013a}{}}%
Hazen, E. L., Jorgensen, S., Rykaczewski, R. R., Bograd, S. J., Foley,
D. G., Jonsen, I. D., Shaffer, S. A., Dunne, J. P., Costa, D. P.,
Crowder, L. B., \& Block, B. A. (2013). Predicted habitat shifts of
Pacific top predators in a changing climate. \emph{Nature Climate
Change}, \emph{3}(3), 234--238.
\url{https://doi.org/10.1038/nclimate1686}

\leavevmode\vadjust pre{\hypertarget{ref-hazen2013}{}}%
Hazen, E. L., Suryan, R. M., Santora, J. A., Bograd, S., Watanuki, Y.,
\& Wilson, R. P. (2013). Scales and mechanisms of marine hotspot
formation. \emph{Marine Ecology Progress Series}, \emph{487}, 177--183.
\url{https://doi.org/10.3354/meps10477}

\leavevmode\vadjust pre{\hypertarget{ref-hedd2001}{}}%
Hedd, A., Gales, R., \& Brothers, N. (2001). Foraging strategies of shy
albatross {\emph{Thalassarche cauta}} breeding at Albatross Island,
Tasmania, Australia. \emph{Marine Ecology Progress Series}, \emph{224},
267--282. \url{https://doi.org/10.3354/meps224267}

\leavevmode\vadjust pre{\hypertarget{ref-raster}{}}%
Hijmans, R. J. (2022a). \emph{Raster: Geographic data analysis and
modeling}. \url{https://CRAN.R-project.org/package=raster}

\leavevmode\vadjust pre{\hypertarget{ref-terra}{}}%
Hijmans, R. J. (2022b). \emph{Terra: Spatial data analysis}.
\url{https://CRAN.R-project.org/package=terra}

\leavevmode\vadjust pre{\hypertarget{ref-hill2008}{}}%
Hill, K. L., Rintoul, S. R., Coleman, R., \& Ridgway, K. R. (2008). Wind
forced low frequency variability of the East Australia Current.
\emph{Geophysical Research Letters}, \emph{35}(8).
\url{https://doi.org/10.1029/2007GL032912}

\leavevmode\vadjust pre{\hypertarget{ref-hill2017}{}}%
Hill, N. A., Foster, S. D., Duhamel, G., Welsford, D., Koubbi, P., \&
Johnson, C. R. (2017). Model-based mapping of assemblages for ecology
and conservation management: a case study of demersal fish on the
Kerguelen Plateau. \emph{Diversity and Distributions}, \emph{23}(10),
1216--1230. https://doi.org/\url{https://doi.org/10.1111/ddi.12613}

\leavevmode\vadjust pre{\hypertarget{ref-hill2020}{}}%
Hill, N. A., Woolley, S. N. C., Foster, S., Dunstan, P. K., McKinlay,
J., Ovaskainen, O., \& Johnson, C. (2020). Determining marine
bioregions: a comparison of quantitative approaches. \emph{Methods in
Ecology and Evolution}, \emph{11}(10), 1258--1272.
\url{https://doi.org/10.1111/2041-210X.13447}

\leavevmode\vadjust pre{\hypertarget{ref-hobday2010}{}}%
Hobday, A. J. (2010). Ensemble analysis of the future distribution of
large pelagic fishes off Australia. \emph{Progress in Oceanography},
\emph{86}(1), 291--301.
\url{https://doi.org/10.1016/j.pocean.2010.04.023}

\leavevmode\vadjust pre{\hypertarget{ref-hsieh2016}{}}%
Hsieh, T. C., Ma, K. H., \& Chao, A. (2016). iNEXT: an R package for
rarefaction and extrapolation of species diversity (Hill numbers).
\emph{Methods in Ecology and Evolution}, \emph{7}(12), 1451--1456.
https://doi.org/\url{https://doi.org/10.1111/2041-210X.12613}

\leavevmode\vadjust pre{\hypertarget{ref-hunt1987}{}}%
Hunt, G. L., \& Schneider, D. C. (1987). Scale-dependent processes in
the physical and biological environment of marine birds. In J. P.
Croxall (Ed.), \emph{Seabirds: feeding ecology and roles in marine
ecosystem} (pp. 7--41). Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-hyrenbach2007}{}}%
Hyrenbach, K. D., Veit, R. R., Weimerskirch, H., Metzl, N., \& Hunt, G.
L. (2007). Community structure across a large-scale ocean productivity
gradient: marine bird assemblages of the Southern Indian Ocean.
\emph{Deep Sea Research Part I: Oceanographic Research Papers},
\emph{54}(7), 1129--1145.
\url{https://doi.org/10.1016/j.dsr.2007.05.002}

\leavevmode\vadjust pre{\hypertarget{ref-ismar2011}{}}%
Ismar, S. M. H., Phillips, R. A., Rayner, M. J., \& Hauber, M. E.
(2011). Geolocation tracking of the annual migration of adult
Australasian gannets ({\emph{Morus serrator}}) breeding in New Zealand.
\emph{Wilson Journal of Ornithology}, \emph{123}(1), 121--125.
\url{https://doi.org/10.1676/10-072.1}

\leavevmode\vadjust pre{\hypertarget{ref-jaine2014}{}}%
Jaine, F., Rohner, C., Weeks, S., Couturier, L., Bennett, M., Townsend,
K., \& Richardson, A. (2014). Movements and habitat use of reef manta
rays off eastern Australia: offshore excursions, deep diving and eddy
affinity revealed by satellite telemetry. \emph{Marine Ecology Progress
Series}, \emph{510}, 73--86. \url{https://doi.org/10.3354/meps10910}

\leavevmode\vadjust pre{\hypertarget{ref-johnson2021}{}}%
Johnson, B., \& Colombelli-Négrel, D. (2021). Breeding success in
Southern Australian little penguins is negatively correlated with high
wind speeds and sea surface temperatures. \emph{Ornithological
Applications}, \emph{123}(1), duaa062.
\url{https://doi.org/10.1093/ornithapp/duaa062}

\leavevmode\vadjust pre{\hypertarget{ref-johnson2011}{}}%
Johnson, C. R., Banks, S. C., Barrett, N. S., Cazassus, F., Dunstan, P.
K., Edgar, G. J., Frusher, S. D., Gardner, C., Haddon, M., Helidoniotis,
F., Hill, K. L., Holbrook, N. J., Hosie, G. W., Last, P. R., Ling, S.
D., Melbourne-Thomas, J., Miller, K., Pecl, G. T., Richardson, A. J.,
\ldots{} Taw, N. (2011). Climate change cascades: Shifts in
oceanography, species' ranges and subtidal marine community dynamics in
eastern Tasmania. \emph{Journal of Experimental Marine Biology and
Ecology}, \emph{400}(1), 17--32.
\url{https://doi.org/10.1016/j.jembe.2011.02.032}

\leavevmode\vadjust pre{\hypertarget{ref-karpouzi2007}{}}%
Karpouzi, V., Watson, R., \& Pauly, D. (2007). Modelling and mapping
resource overlap between seabirds and fisheries on a global scale: a
preliminary assessment. \emph{Marine Ecology Progress Series},
\emph{343}, 87--99. \url{https://doi.org/10.3354/meps06860}

\leavevmode\vadjust pre{\hypertarget{ref-kessler2013}{}}%
Kessler, W. S., \& Cravatte, S. (2013). Mean circulation of the Coral
Sea. \emph{Journal of Geophysical Research: Oceans}, \emph{118}(12),
6385--6410. \url{https://doi.org/10.1002/2013JC009117}

\leavevmode\vadjust pre{\hypertarget{ref-kreft2010}{}}%
Kreft, H., \& Jetz, W. (2010). A framework for delineating
biogeographical regions based on species distributions. \emph{Journal of
Biogeography}, \emph{37}(11), 2029--2053.
\url{https://doi.org/10.1111/j.1365-2699.2010.02375.x}

\leavevmode\vadjust pre{\hypertarget{ref-lavers2014}{}}%
Lavers, J. L., Miller, M. G. R., Carter, M. J., Swann, G., \& Clarke, R.
H. (2014). Predicting the spatial distribution of a seabird community to
identify priority conservation areas in the Timor Sea.
\emph{Conservation Biology}, \emph{28}(6), 1699--1709.
\url{https://doi.org/10.1111/cobi.12324}

\leavevmode\vadjust pre{\hypertarget{ref-lee2019good}{}}%
Lee, M. R., Powell, J. R., Oberle, B., Cornwell, W. K., Lyons, M., Rigg,
J. L., \& Zanne, A. E. (2019). Good neighbors aplenty: Fungal endophytes
rarely exhibit competitive exclusion patterns across a span of woody
habitats. \emph{Ecology}, \emph{100}(9), e02790.
\url{https://doi.org/10.1002/ecy.2790}

\leavevmode\vadjust pre{\hypertarget{ref-lenoir2020}{}}%
Lenoir, J., Bertrand, R., Comte, L., Bourgeaud, L., Hattab, T.,
Murienne, J., \& Grenouillet, G. (2020). Species better track climate
warming in the oceans than on land. \emph{Nature Ecology \& Evolution},
\emph{4}(8), 1044--1059. \url{https://doi.org/10.1038/s41559-020-1198-2}

\leavevmode\vadjust pre{\hypertarget{ref-li2022}{}}%
Li, J., Roughan, M., \& Kerry, C. (2022). Drivers of ocean warming in
the western boundary currents of the Southern Hemisphere. \emph{Nature
Climate Change}, \emph{12}(10), 901--909.
\url{https://doi.org/10.1038/s41558-022-01473-8}

\leavevmode\vadjust pre{\hypertarget{ref-longhurst2007}{}}%
Longhurst, A. R. (2007). \emph{Ecological geography of the sea} (2nd
ed). Academic Press.

\leavevmode\vadjust pre{\hypertarget{ref-lyons2017}{}}%
Lyons, M. B., Foster, S. D., \& Keith, D. A. (2017). Simultaneous
vegetation classification and mapping at large spatial scales.
\emph{Journal of Biogeography}, \emph{44}(12), 2891--2902.
\url{https://doi.org/10.1111/jbi.13088}

\leavevmode\vadjust pre{\hypertarget{ref-marchant1990}{}}%
Marchant, R. (1990). Robustness of classification and ordination
techniques applied to macroinvertebrate communities from the La Trobe
River, Victoria. \emph{Marine and Freshwater Research}, \emph{41}(4),
493--504. \url{https://doi.org/10.1071/MF9900493}

\leavevmode\vadjust pre{\hypertarget{ref-marchant1990handbook}{}}%
Marchant, S., \& Higgins, P. J. (1990). \emph{Handbook of australian,
new zealand \& antarctic birds. Vol. 1, ratites to ducks, p. AB}. Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mason2018}{}}%
Mason, C., Alderman, R., McGowan, J., Possingham, H. P., Hobday, A. J.,
Sumner, M., \& Shaw, J. (2018). Telemetry reveals existing marine
protected areas are worse than random for protecting the foraging
habitat of threatened shy albatross ({\emph{Thalassarche cauta}}).
\emph{Diversity and Distributions}, \emph{24}(12), 1744--1755.
\url{https://doi.org/10.1111/ddi.12830}

\leavevmode\vadjust pre{\hypertarget{ref-mata2006}{}}%
Mata, M. M., Wijffels, S. E., Church, J. A., \& Tomczak, M. (2006). Eddy
shedding and energy conversions in the East Australian Current.
\emph{Journal of Geophysical Research: Oceans}, \emph{111}, C09034.
\url{https://doi.org/10.1029/2006JC003592}

\leavevmode\vadjust pre{\hypertarget{ref-mcduie2015}{}}%
McDuie, F., Weeks, S. J., Miller, M. G. R., \& Congdon, B. C. (2015).
Breeding tropical shearwaters use distant foraging sites when
self-provisioning. \emph{Marine Ornithology}, \emph{43}, 123--129.

\leavevmode\vadjust pre{\hypertarget{ref-melo2005}{}}%
Melo, A. S. (2005). Effects of taxonomic and numeric resolution on the
ability to detect ecological patterns at a local scale using stream
macroinvertebrates. \emph{Archiv Für Hydrobiologie}, \emph{164}(3),
309--323. \url{https://doi.org/10.1127/0003-9136/2005/0164-0309}

\leavevmode\vadjust pre{\hypertarget{ref-miller2018}{}}%
Miller, M. G. R., Silva, F. R. O., Machovsky-Capuska, G. E., \& Congdon,
B. C. (2018). Sexual segregation in tropical seabirds: drivers of
sex-specific foraging in the brown booby {\emph{Sula leucogaster}}.
\emph{Journal of Ornithology}, \emph{159}(2), 425--437.
\url{https://doi.org/10.1007/s10336-017-1512-1}

\leavevmode\vadjust pre{\hypertarget{ref-morrone2018}{}}%
Morrone, J. J. (2018). The spectre of biogeographical regionalization.
\emph{Journal of Biogeography}, \emph{45}(2), 282--288.
\url{https://doi.org/10.1111/jbi.13135}

\leavevmode\vadjust pre{\hypertarget{ref-morrone2023biogeographical}{}}%
Morrone, J. J. (2023). Why biogeographical transition zones matter.
\emph{Journal of Biogeography}. \url{https://doi.org/10.1111/jbi.14632}

\leavevmode\vadjust pre{\hypertarget{ref-mott2018}{}}%
Mott, R., \& Clarke, R. H. (2018). Systematic review of geographic
biases in the collection of at-sea distribution data for seabirds.
\emph{Emu - Austral Ornithology}, \emph{118}(3), 235--246.
\url{https://doi.org/10.1080/01584197.2017.1416957}

\leavevmode\vadjust pre{\hypertarget{ref-niella2022}{}}%
Niella, Y., Butcher, P., Holmes, B., Barnett, A., \& Harcourt, R.
(2022). Forecasting intraspecific changes in distribution of a
wide-ranging marine predator under climate change. \emph{Oecologia},
\emph{198}(1), 111--124.
\url{https://doi.org/10.1007/s00442-021-05075-7}

\leavevmode\vadjust pre{\hypertarget{ref-niella2020}{}}%
Niella, Y., Smoothey, A., Peddemors, V., \& Harcourt, R. (2020).
Predicting changes in distribution of a large coastal shark in the face
of the strengthening East Australian Current. \emph{Marine Ecology
Progress Series}, \emph{642}, 163--177.
\url{https://doi.org/10.3354/meps13322}

\leavevmode\vadjust pre{\hypertarget{ref-norman1992distribution}{}}%
Norman, F. (1992). Distribution and abundance of seabirds off {P}hillip
{I}sland and within {P}ort {P}hillip {B}ay, {V}ictoria, 1986--1988.
\emph{Emu - Austral Ornithology}, \emph{91}(5), 377--394.

\leavevmode\vadjust pre{\hypertarget{ref-ohara2020}{}}%
O'Hara, T. D., Williams, A., Althaus, F., Ross, A. S., \& Bax, N. J.
(2020). Regional-scale patterns of deep seafloor biodiversity for
conservation assessment. \emph{Diversity and Distributions},
\emph{26}(4), 479--494. \url{https://doi.org/10.1111/ddi.13034}

\leavevmode\vadjust pre{\hypertarget{ref-ohara2020a}{}}%
O'Hara, T. D., Williams, A., Woolley, S. N. C., Nau, A. W., \& Bax, N.
J. (2020). Deep-sea temperate-tropical faunal transition across uniform
environmental gradients. \emph{Deep Sea Research Part I: Oceanographic
Research Papers}, \emph{161}, 103283.
\url{https://doi.org/10.1016/j.dsr.2020.103283}

\leavevmode\vadjust pre{\hypertarget{ref-oke2000}{}}%
Oke, P. R., \& Middleton, J. H. (2000). Topographically induced
upwelling off eastern Australia. \emph{Journal of Physical
Oceanography}, \emph{30}(3), 512--531.
\url{https://doi.org/10.1175/1520-0485(2000)030\%3C0512:TIUOEA\%3E2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-oke2019a}{}}%
Oke, P. R., Pilo, G. S., Ridgway, K., Kiss, A., \& Rykova, T. (2019). A
search for the Tasman Front. \emph{Journal of Marine Systems},
\emph{199}, 103217. \url{https://doi.org/10.1016/j.jmarsys.2019.103217}

\leavevmode\vadjust pre{\hypertarget{ref-oke2019}{}}%
Oke, P. R., Roughan, M., Cetina-Heredia, P., Pilo, G. S., Ridgway, K.
R., Rykova, T., Archer, M. R., Coleman, R. C., Kerry, C. G., Rocha, C.,
Schaeffer, A., \& Vitarelli, E. (2019). Revisiting the circulation of
the East Australian Current: its path, separation, and eddy field.
\emph{Progress in Oceanography}, \emph{176}, 102139.
\url{https://doi.org/10.1016/j.pocean.2019.102139}

\leavevmode\vadjust pre{\hypertarget{ref-oliver2017}{}}%
Oliver, E. C. J., Benthuysen, J. A., Bindoff, N. L., Hobday, A. J.,
Holbrook, N. J., Mundy, C. N., \& Perkins-Kirkpatrick, S. E. (2017). The
unprecedented 2015/16 Tasman Sea marine heatwave. \emph{Nature
Communications}, \emph{8}(1), 16101.
\url{https://doi.org/10.1038/ncomms16101}

\leavevmode\vadjust pre{\hypertarget{ref-oliver2018}{}}%
Oliver, E. C. J., Lago, V., Hobday, A. J., Holbrook, N. J., Ling, S. D.,
\& Mundy, C. N. (2018). Marine heatwaves off eastern Tasmania: trends,
interannual variability, and predictability. \emph{Progress in
Oceanography}, \emph{161}, 116--130.
\url{https://doi.org/10.1016/j.pocean.2018.02.007}

\leavevmode\vadjust pre{\hypertarget{ref-pantiukhin2021}{}}%
Pantiukhin, D., Piepenburg, D., Hansen, M. L. S., \& Kraan, C. (2021).
Data-driven bioregionalization: a seascape-scale study of macrobenthic
communities in the Eurasian Arctic. \emph{Journal of Biogeography},
\emph{48}(11), 2877--2890. \url{https://doi.org/10.1111/jbi.14247}

\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
Pebesma, E. (2018). {Simple features for R: standardized support for
spatial vector data}. \emph{{The R Journal}}, \emph{10}(1), 439--446.
\url{https://doi.org/10.32614/RJ-2018-009}

\leavevmode\vadjust pre{\hypertarget{ref-peck2004}{}}%
Peck, D., Smithers, B., Krockenberger, A., \& Congdon, B. (2004). Sea
surface temperature constrains wedge-tailed shearwater foraging success
within breeding seasons. \emph{Marine Ecology Progress Series},
\emph{281}, 259--266. \url{https://doi.org/10.3354/meps281259}

\leavevmode\vadjust pre{\hypertarget{ref-puxe9ron2010}{}}%
Péron, C., Authier, M., Barbraud, C., Delord, K., Besson, D., \&
Weimerskirch, H. (2010). Interdecadal changes in at-sea distribution and
abundance of subantarctic seabirds along a latitudinal gradient in the
Southern Indian Ocean. \emph{Global Change Biology}, \emph{16}(7),
1895--1909. \url{https://doi.org/10.1111/j.1365-2486.2010.02169.x}

\leavevmode\vadjust pre{\hypertarget{ref-phillips2022}{}}%
Phillips, L. R., Malan, N., Roughan, M., Harcourt, R., Jonsen, I., Cox,
M., Brierley, A. S., Slip, D., Wilkins, A., \& Carroll, G. (2022).
Coastal seascape variability in the intensifying {E}ast {A}ustralian
{C}urrent southern extension. \emph{Frontiers in Marine Science},
\emph{9}, 925123.
\url{https://www.frontiersin.org/articles/10.3389/fmars.2022.925123}

\leavevmode\vadjust pre{\hypertarget{ref-pinsky2019}{}}%
Pinsky, M. L., Eikeset, A. M., McCauley, D. J., Payne, J. L., \& Sunday,
J. M. (2019). Greater vulnerability to warming of marine versus
terrestrial ectotherms. \emph{Nature}, \emph{569}(7754), 108--111.
\url{https://doi.org/10.1038/s41586-019-1132-4}

\leavevmode\vadjust pre{\hypertarget{ref-pocklington1979}{}}%
Pocklington, R. (1979). An oceanographic interpretation of seabird
distributions in the Indian Ocean. \emph{Marine Biology}, \emph{51}(1),
9--21. \url{https://doi.org/10.1007/BF00389026}

\leavevmode\vadjust pre{\hypertarget{ref-poloczanska2007}{}}%
Poloczanska, E. S., Babcock, R. C., Butler, A., Hobday, A. J.,
Hoegh-Guldberg, O., Kunz, T. J., Matear, R., Milton, D., Okey, T. A., \&
Richardson, A. J. (2007). Climate change and Australian marine life. In
R. N. Gibson, R. J. A. Atkinson, \& J. D. M. Gordon (Eds.),
\emph{Oceanography and marine biology: an annual review} (Vol. 45, pp.
407--478). CRC Press, Taylor \& Francis Group.

\leavevmode\vadjust pre{\hypertarget{ref-poloczanska2013}{}}%
Poloczanska, E. S., Brown, C. J., Sydeman, W. J., Kiessling, W.,
Schoeman, D. S., Moore, P. J., Brander, K., Bruno, J. F., Buckley, L.
B., Burrows, M. T., Duarte, C. M., Halpern, B. S., Holding, J., Kappel,
C. V., O'Connor, M. I., Pandolfi, J. M., Parmesan, C., Schwing, F.,
Thompson, S. A., \& Richardson, A. J. (2013). Global imprint of climate
change on marine life. \emph{Nature Climate Change}, \emph{3}(10),
919--925. \url{https://doi.org/10.1038/nclimate1958}

\leavevmode\vadjust pre{\hypertarget{ref-ponti2023}{}}%
Ponti, R., \& Sannolo, M. (2023). The importance of including phenology
when modelling species ecological niche. \emph{Ecography},
\emph{2023}(4), e06143. \url{https://doi.org/10.1111/ecog.06143}

\leavevmode\vadjust pre{\hypertarget{ref-priddel2014}{}}%
Priddel, D., Carlile, N., Portelli, D., Kim, Y., O'Neill, L.,
Bretagnolle, V., Ballance, L. T., Phillips, R. A., Pitman, R. L., \&
Raynei, M. J. (2014). Pelagic distribution of Gould's petrel
({\emph{Pterodroma leucoptera}}): linking shipboard and onshore
observations with remote-tracking data. \emph{Emu - Austral
Ornithology}, \emph{114}(4), 360--370.
\url{https://doi.org/10.1071/MU14021}

\leavevmode\vadjust pre{\hypertarget{ref-quillfeldt2015}{}}%
Quillfeldt, P., Cherel, Y., Delord, K., \& Weimerkirch, H. (2015). Cool,
cold or colder? Spatial segregation of prions and blue petrels is
explained by differences in preferred sea surface temperatures.
\emph{Biology Letters}, \emph{11}(4), 20141090.
\url{https://doi.org/10.1098/rsbl.2014.1090}

\leavevmode\vadjust pre{\hypertarget{ref-r2022}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-ramuxedrez2017}{}}%
Ramírez, F., Afán, I., Davis, L. S., \& Chiaradia, A. (2017). Climate
impacts on global hot spots of marine biodiversity. \emph{Science
Advances}, \emph{3}(2), e1601198.
\url{https://doi.org/10.1126/sciadv.1601198}

\leavevmode\vadjust pre{\hypertarget{ref-raymond2010}{}}%
Raymond, B., Shaffer, S. A., Sokolov, S., Woehler, E. J., Costa, D. P.,
Einoder, L., Hindell, M., Hosie, G., Pinkerton, M., Sagar, P. M., Scott,
D., Smith, A., Thompson, D. R., Vertigan, C., \& Weimerskirch, H.
(2010). Shearwater foraging in the Southern Ocean: the roles of prey
availability and winds. \emph{PLoS ONE}, \emph{5}(6), e10960.
\url{https://doi.org/10.1371/journal.pone.0010960}

\leavevmode\vadjust pre{\hypertarget{ref-receveur2020}{}}%
Receveur, A., Vourey, E., Lebourges-Dhaussy, A., Menkes, C., Ménard, F.,
\& Allain, V. (2020). Biogeography of micronekton assemblages in the
Natural Park of the Coral Sea. \emph{Frontiers in Marine Science},
\emph{7}, 449. \url{https://doi.org/10.3389/fmars.2020.00449}

\leavevmode\vadjust pre{\hypertarget{ref-reid2023}{}}%
Reid, K., Baker, G. B., \& Woehler, E. J. (2023). An ecological risk
assessment for the impacts of offshore wind farms on birds in Australia.
\emph{Austral Ecology}, \emph{48}(2), 418--439.
\url{https://doi.org/10.1111/aec.13278}

\leavevmode\vadjust pre{\hypertarget{ref-reid2000}{}}%
Reid, T. A., \& Hindell, M. A. (2000). Coarse-scale relationships
between seabirds and zooplankton off south-eastern Tasmania.
\emph{Marine and Freshwater Research}, \emph{51}(8), 789--798.
\url{https://doi.org/10.1071/MF98160}

\leavevmode\vadjust pre{\hypertarget{ref-reisinger2022}{}}%
Reisinger, R. R., Corney, S., Raymond, B., Lombard, A. T., Bester, M.
N., Crawford, R. J. M., Davies, D., de Bruyn, P. J. N., Dilley, B. J.,
Kirkman, S. P., Makhado, A. B., Ryan, P. G., Schoombie, S., Stevens, K.
L., Tosh, C. A., Wege, M., Whitehead, T. O., Sumner, M. D., Wotherspoon,
S., \ldots{} Pistorius, P. A. (2022). Habitat model forecasts suggest
potential redistribution of marine predators in the southern Indian
Ocean. \emph{Diversity and Distributions}, \emph{28}(1), 142--159.
\url{https://doi.org/10.1111/ddi.13447}

\leavevmode\vadjust pre{\hypertarget{ref-reygondeau2013}{}}%
Reygondeau, G., Longhurst, A., Martinez, E., Beaugrand, G., Antoine, D.,
\& Maury, O. (2013). Dynamic biogeochemical provinces in the global
ocean. \emph{Global Biogeochemical Cycles}, \emph{27}(4), 1046--1058.
\url{https://doi.org/10.1002/gbc.20089}

\leavevmode\vadjust pre{\hypertarget{ref-reygondeau2012}{}}%
Reygondeau, G., Maury, O., Beaugrand, G., Fromentin, J. M., Fonteneau,
A., \& Cury, P. (2012). Biogeography of tuna and billfish communities.
\emph{Journal of Biogeography}, \emph{39}(1), 114--129.
\url{https://doi.org/10.1111/j.1365-2699.2011.02582.x}

\leavevmode\vadjust pre{\hypertarget{ref-ribic1997}{}}%
Ribic, C., Ainley, D., \& Spear, L. (1997). Scale-related
seabird-environmental relationships in Pacific equatorial waters, with
reference to El Niño-Southern Oscillation events. \emph{Marine Ecology
Progress Series}, \emph{156}, 183--203.
\url{https://doi.org/10.3354/meps156183}

\leavevmode\vadjust pre{\hypertarget{ref-ridgway2003a}{}}%
Ridgway, K. R., \& Dunn, J. R. (2003). Mesoscale structure of the mean
East Australian Current System and its relationship with topography.
\emph{Progress in Oceanography}, \emph{56}(2), 189--222.
\url{https://doi.org/10.1016/S0079-6611(03)00004-1}

\leavevmode\vadjust pre{\hypertarget{ref-santora2013}{}}%
Santora, J., \& Veit, R. (2013). Spatio-temporal persistence of top
predator hotspots near the Antarctic Peninsula. \emph{Marine Ecology
Progress Series}, \emph{487}, 287--304.
\url{https://doi.org/10.3354/meps10350}

\leavevmode\vadjust pre{\hypertarget{ref-scales2017}{}}%
Scales, K. L., Hazen, E. L., Jacox, M. G., Edwards, C. A., Boustany, A.
M., Oliver, M. J., \& Bograd, S. J. (2017). Scale of inference: on the
sensitivity of habitat models for wide-ranging marine predators to the
resolution of environmental data. \emph{Ecography}, \emph{40}(1),
210--220. \url{https://doi.org/10.1111/ecog.02272}

\leavevmode\vadjust pre{\hypertarget{ref-scales2014a}{}}%
Scales, K. L., Miller, P. I., Embling, C. B., Ingram, S. N., Pirotta,
E., \& Votier, S. C. (2014). Mesoscale fronts as foraging habitats:
composite front mapping reveals oceanographic drivers of habitat use for
a pelagic seabird. \emph{Journal of The Royal Society Interface},
\emph{11}(100), 20140679. \url{https://doi.org/10.1098/rsif.2014.0679}

\leavevmode\vadjust pre{\hypertarget{ref-scales2014}{}}%
Scales, K. L., Miller, P. I., Hawkes, L. A., Ingram, S. N., Sims, D. W.,
\& Votier, S. C. (2014). On the front line: frontal zones as priority
at-sea conservation areas for mobile marine vertebrates. \emph{Journal
of Applied Ecology}, \emph{51}(6), 1575--1583.
\url{https://doi.org/10.1111/1365-2664.12330}

\leavevmode\vadjust pre{\hypertarget{ref-biology2002}{}}%
Schreiber, E. A., \& Burger, J. (Eds.). (2002). \emph{Biology of marine
birds}. CRC Press.

\leavevmode\vadjust pre{\hypertarget{ref-sherman1994}{}}%
Sherman, K. (1994). Sustainability, biomass yields, and health of
coastal ecosystems: an ecological perspective. \emph{Marine Ecology
Progress Series}, \emph{112}, 277--301.
\url{https://doi.org/10.3354/meps112277}

\leavevmode\vadjust pre{\hypertarget{ref-sojitra2022}{}}%
Sojitra, M., Woehler, E. J., Lea, M.-A., \& Wotherspoon, S. (2022).
Multi-decadal changes in the at-sea distribution and abundance of
black-browed and light-mantled sooty albatrosses in the {S}outhwest
{P}acific {O}cean. \emph{ICES Journal of Marine Science}, fsac197.
\url{https://doi.org/10.1093/icesjms/fsac197}

\leavevmode\vadjust pre{\hypertarget{ref-souza2023}{}}%
Souza, J. S., \& Santos, L. N. (2023). Resident species, not immigrants,
drive reorganization of estuarine fish assemblages in response to
warming. \emph{Ecology}, \emph{104}(5), e3987.
\url{https://doi.org/10.1002/ecy.3987}

\leavevmode\vadjust pre{\hypertarget{ref-spalding2007}{}}%
Spalding, M. D., Fox, H. E., Allen, G. R., Davidson, N., Ferdaña, Z. A.,
Finlayson, M., Halpern, B. S., Jorge, M. A., Lombana, A., Lourie, S. A.,
Martin, K. D., McManus, E., Molnar, J., Recchia, C. A., \& Robertson, J.
(2007). Marine {E}coregions of the {W}orld: A bioregionalization of
coastal and shelf areas. \emph{BioScience}, \emph{57}(7), 573--583.
\url{https://doi.org/10.1641/B570707}

\leavevmode\vadjust pre{\hypertarget{ref-sun2022}{}}%
Sun, C., Zhang, A., Jin, B., Wang, X., Zhang, X., \& Zhang, L. (2022).
Seasonal variability of eddy kinetic energy in the north indian ocean.
\emph{Frontiers in Marine Science}, \emph{9}, 1032699.
\url{https://doi.org/10.3389/fmars.2022.1032699}

\leavevmode\vadjust pre{\hypertarget{ref-suthers2011}{}}%
Suthers, I. M., Young, J. W., Baird, M. E., Roughan, M., Everett, J. D.,
Brassington, G. B., Byrne, M., Condie, S. A., Hartog, J. R., Hassler, C.
S., Hobday, A. J., Holbrook, N. J., Malcolm, H. A., Oke, P. R.,
Thompson, P. A., \& Ridgway, K. (2011). The strengthening East
Australian Current, its eddies and biological effects {\textemdash} an
introduction and overview. \emph{Deep Sea Research Part II: Topical
Studies in Oceanography}, \emph{58}(5), 538--546.
\url{https://doi.org/10.1016/j.dsr2.2010.09.029}

\leavevmode\vadjust pre{\hypertarget{ref-sutton2017a}{}}%
Sutton, T. T., Clark, M. R., Dunn, D. C., Halpin, P. N., Rogers, A. D.,
Guinotte, J., Bograd, S. J., Angel, M. V., Perez, J. A. A., Wishner, K.,
Haedrich, R. L., Lindsay, D. J., Drazen, J. C., Vereshchaka, A.,
Piatkowski, U., Morato, T., Błachowiak-Samołyk, K., Robison, B. H.,
Gjerde, K. M., \ldots{} Heino, M. (2017). A global biogeographic
classification of the mesopelagic zone. \emph{Deep Sea Research Part I:
Oceanographic Research Papers}, \emph{126}, 85--102.
\url{https://doi.org/10.1016/j.dsr.2017.05.006}

\leavevmode\vadjust pre{\hypertarget{ref-sydeman2012}{}}%
Sydeman, W., Thompson, S., \& Kitaysky, A. (2012). Seabirds and climate
change: roadmap for the future. \emph{Marine Ecology Progress Series},
\emph{454}, 107--117. \url{https://doi.org/10.3354/meps09806}

\leavevmode\vadjust pre{\hypertarget{ref-taylor2021}{}}%
Taylor, G., Elliott, G., Walker, K., \& Bose, S. (2021). Year-round
distribution, breeding cycle, and activity of white-headed petrels
({\emph{Pterodroma lessonii}}) nesting on adams island, auckland
islands. \emph{Notornis}, \emph{67}, 369--386.

\leavevmode\vadjust pre{\hypertarget{ref-tremblay2009}{}}%
Tremblay, Y., Bertrand, S., Henry, R., Kappes, M., Costa, D., \&
Shaffer, S. (2009). Analytical approaches to investigating
seabird--environment interactions: a review. \emph{Marine Ecology
Progress Series}, \emph{391}, 153--163.
\url{https://doi.org/10.3354/meps08146}

\leavevmode\vadjust pre{\hypertarget{ref-vilhena2015}{}}%
Vilhena, D. A., \& Antonelli, A. (2015). A network approach for
identifying and delimiting biogeographical regions. \emph{Nature
Communications}, \emph{6}(1), 6848.
\url{https://doi.org/10.1038/ncomms7848}

\leavevmode\vadjust pre{\hypertarget{ref-warton2015}{}}%
Warton, D. I., Foster, S. D., De'ath, G., Stoklosa, J., \& Dunstan, P.
K. (2015). Model-based thinking for community ecology. \emph{Plant
Ecology}, \emph{216}(5), 669--682.
\url{https://doi.org/10.1007/s11258-014-0366-3}

\leavevmode\vadjust pre{\hypertarget{ref-waters2008marine}{}}%
Waters, J. M. (2008). Marine biogeographical disjunction in temperate
{A}ustralia: Historical landbridge, contemporary currents, or both?
\emph{Diversity and Distributions}, \emph{14}(4), 692--700.
\url{https://doi.org/10.1111/j.1472-4642.2008.00481.x}

\leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2007}{}}%
Weimerskirch, H. (2007). Are seabirds foraging for unpredictable
resources? \emph{Deep Sea Research Part II: Topical Studies in
Oceanography}, \emph{54}(3-4), 211--223.
\url{https://doi.org/10.1016/j.dsr2.2006.11.013}

\leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2020}{}}%
Weimerskirch, H., De Grissac, S., Ravache, A., Prudor, A., Corbeau, A.,
Congdon, B., McDuie, F., Bourgeois, K., Dromzée, S., Butscher, J.,
Menkes, C., Allain, V., Vidal, E., Jaeger, A., \& Borsa, P. (2020).
At-sea movements of wedge-tailed shearwaters during and outside the
breeding season from four colonies in New Caledonia. \emph{Marine
Ecology Progress Series}, \emph{633}, 225--238.
\url{https://doi.org/10.3354/meps13171}

\leavevmode\vadjust pre{\hypertarget{ref-whittaker2005}{}}%
Whittaker, R. J., Araújo, M. B., Jepson, P., Ladle, R. J., Watson, J. E.
M., \& Willis, K. J. (2005). Conservation Biogeography: assessment and
prospect. \emph{Diversity and Distributions}, \emph{11}(1), 3--23.
\url{https://doi.org/10.1111/j.1366-9516.2005.00143.x}

\leavevmode\vadjust pre{\hypertarget{ref-woehler2023}{}}%
Woehler, E., \& Hobday, A. (2023). Impacts of marine heatwaves may be
mediated by seabird life history strategies. \emph{Marine Ecology
Progress Series}, \emph{HEAT}. \url{https://doi.org/10.3354/meps14333}

\leavevmode\vadjust pre{\hypertarget{ref-woehler2003}{}}%
Woehler, E., Raymond, B., \& Watts, D. (2003). Decadal-scale seabird
assemblages in Prydz Bay, East Antarctica. \emph{Marine Ecology Progress
Series}, \emph{251}, 299--310. \url{https://doi.org/10.3354/meps251299}

\leavevmode\vadjust pre{\hypertarget{ref-woolley2020}{}}%
Woolley, S. N. C., Foster, S. D., Bax, N. J., Currie, J. C., Dunn, D.
C., Hansen, C., Hill, N. A., O'Hara, T. D., Ovaskainen, O., Sayre, R.,
Vanhatalo, J. P., \& Dunstan, P. K. (2020). Bioregions in marine
environments: Combining biological and environmental data for management
and scientific understanding. \emph{BioScience}, \emph{70}(1), 48--59.
\url{https://doi.org/10.1093/biosci/biz133}

\leavevmode\vadjust pre{\hypertarget{ref-ecomix}{}}%
Woolley, S., Foster, S., \& Dunstan, P. (2022). \emph{Ecomix: Fitting
finite mixture models to ecological data}.
\url{https://github.com/skiptoniam/ecomix}

\leavevmode\vadjust pre{\hypertarget{ref-yang2016}{}}%
Yang, H., Lohmann, G., Wei, W., Dima, M., Ionita, M., \& Liu, J. (2016).
Intensification and poleward shift of subtropical western boundary
currents in a warming climate. \emph{Journal of Geophysical Research:
Oceans}, \emph{121}(7), 4928--4945.
\url{https://doi.org/10.1002/2015JC011513}

\leavevmode\vadjust pre{\hypertarget{ref-young2011}{}}%
Young, J. W., Hobday, A. J., Campbell, R. A., Kloser, R. J., Bonham, P.
I., Clementson, L. A., \& Lansdell, M. J. (2011). The biological
oceanography of the East Australian Current and surrounding waters in
relation to tuna and billfish catches off eastern Australia. \emph{Deep
Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5),
720--733. \url{https://doi.org/10.1016/j.dsr2.2010.10.005}

\end{CSLReferences}

\newpage

\hypertarget{tables-and-figures}{%
\section*{Tables and Figures}\label{tables-and-figures}}
\addcontentsline{toc}{section}{Tables and Figures}

\begin{itemize}
\tightlist
\item
  \textbf{TABLE 1.} Environmental data used in the Region of Common
  Profile models applied to seabird data off eastern Australia, their
  spatial/temporal resolutions, units, and sources
\item
  \textbf{TABLE 2.} Full (a) and best (b) Region of Common Profile
  Models applied to seabird data off eastern Australia, for both data
  types used (presence-absence, based on a Bernoulli model; abundance,
  based on a Negative Binomial model). Refer to Table 1 for the
  environmental data acronyms
\item
  \textbf{FIGURE 1.} Study area showing the main ocean currents (a), and
  seabird records made by season on top of the 1° latitude × 1°
  longitude grid cells (b). In (a), the East Australian Current (EAC)
  system is highlighted with its acronyms in bold and drawn in blue. The
  South Equatorial Current (SEC) and its branches, the North Caledonian
  Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red. The EAC
  `eastern extension' is an eddy field (blue shaded area) often called
  the `Tasman Front', from which the East Auckland Current originates
  (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et
  al.~(2019); for colour references see the online version
\item
  \textbf{FIGURE 2.} Partial plots for the retained covariates in the
  spring (a) presence-absence and (b) abundance best models. The plot
  shows the fitted probability of belonging to a Region of Common
  Profiles (RCP) against the environmental value. All other predictors
  are held at their mean values. Results for other seasons and models
  can be found in the Supplementary Material (Fig. S5). Refer to Table 1
  for the environmental data acronyms
\item
  \textbf{FIGURE 3.} Spatialized point predictions (hard-classes) for
  seabird assemblages (Region of Common Profile; RCP), by season, based
  on presence-absence (a) and abundance (b) models, off eastern
  Australia. Each grid was assigned to the RCP group with highest
  membership probability
\item
  \textbf{FIGURE 4.} Predicted probability membership of each seabird
  assemblage (Region of Common Profiles; RCP) and grid, off eastern
  Australia, from spring presence-absence (a) and abundance (b) models.
  The central column, `mean', corresponds to the point prediction and
  Bayesian boot-strapped, lower and upper confidence intervals (CI), on
  its sides. Note that, in case of two RCPs, RCP2 probablility equals to
  (1 - RCP1 probability). Results for other seasons and models can be
  found in the Supplementary Material (Fig. S6)
\item
  \textbf{FIGURE 5.} Species profiles of seabird assemblages (Region of
  Common Profile; RCP) off eastern Australia in spring. Values are the
  average and 95\% confidence intervals of probability of occurrence (a)
  and predicted mean abundance (b) for each species, based on 1000
  Bayesian bootstraps. Values in (b) were log10-transformed to
  accommodate the high variation between species. Results for other
  seasons and models can be found in the Supplementary Material (Fig.
  S7)
\end{itemize}

\begin{landscape}

\begin{table}

\caption{\label{tab:table1} }
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}llll>{\raggedright\arraybackslash}p{3cm}}
\toprule
\textbf{Variable} & \textbf{Abbreviation} & \textbf{Units} & \textbf{Spatial.resolution} & \textbf{Temporal.resolution} & \textbf{Source}\\
\midrule
Sea Surface Temperature & sst & °Celsius & 0.25° × 0.25° & daily & Huang et al. (2021)${^a}$\\
Sea Surface Temperature gradient & sst\_grad & Δ°Celsius & 0.25° × 0.25° & daily & Derived from Huang et al. (2021)${^a}$\\
Sea Surface Salinity & sss & PSU & 0.083° × 0.083° & daily & E.U. Copernicus Marine Service Information${^b}{^c}$\\
Mixed Layer Depth & mld & m & 0.083° × 0.083° & daily & E.U. Copernicus Marine Service Information${^b}{^c}$\\
8-day average Chlorophyll-a & chl8 & mg/m${^3}$ & 0.036° × 0.036° & 8-days composite & ERDDAP/NOAA${^d}$\\
\addlinespace
Log10 8-day average Chlorophyll-a & log10\_chl8 & $log_{10}$(chl8) & 0.036° × 0.036° & 8-days composite & Derived from ERDDAP/NOAA${^d}$\\
Eddy Kinetic Energy & eke & m/s & 0.2° × 0.2° & daily & IMOS/AODN${^e}$\\
Climatic mean EKE & clim\_eke\_mean & mean(m/s)/season/10 years & 0.2° × 0.2° & seasonal climatology & Derived from IMOS/AODN${^e}$\\
Climatic standard deviation EKE & clim\_eke\_sd & sd(m/s)/season/10 years & 0.2° × 0.2° & seasonal climatology & Derived from IMOS/AODN${^e}$\\
Climatic mean SSTgrad & clim\_sst\_grad & mean(Δ°Celsius)/season/10 years & 0.25° × 0.25° & seasonal climatology & Derived from Huang et al. (2021)${^a}$\\
\addlinespace
Bathymetry & bat & m & 0.004° × 0.004° & NA & GEBCO${^f}$\\
Slope & slope & degrees (°) & 0.004° × 0.004° & NA & Derived from GEBCO${^f}$\\
Distance from coast & dist\_coast & km & NA & NA & Calculated in R (see main text)\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{a} NOAA OISST: (https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{b} 2011--2019: GLOBAL\_MULTIYEAR\_PHY\_001\_030 (https://doi.org/10.48670/moi-00021); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{c} 2020--2021: GLOBAL\_ANALYSIS\_FORECAST\_PHY\_001\_024 (https://doi.org/10.48670/moi-00016); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{d} ERDDAP/NOAA (https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{e} IMOS/AODN (https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set; }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{f} GEBCO (https://www.gebco.net/data\_and\_products/gridded\_bathymetry\_data/)}\\
\end{tabular}
\end{table}

\end{landscape}

\newpage

\begin{table}

\caption{\label{tab:table2} }
\centering
\begin{tabular}[t]{>{}lll>{\raggedright\arraybackslash}p{11cm}l}
\toprule
 &  & Model & Model specification & BIC\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{\textbf{Bernoulli}}}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Summer}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad & 332.69\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sss & 319.34\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Autumn}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad & 1184.29\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sst & 1147.94\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Winter}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad + clim\_sst\_grad & 1089.43\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst & 1062.98\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Spring}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + dist\_coast + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad + clim\_eke\_mean & 2345.6\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sst & 2306.7\\
\cmidrule{1-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{\textbf{Negative Binomial}}}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Summer}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad & 1518.96\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sss & 1505.13\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Autumn}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad & -Inf (n.f.)\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + chl + sst + mld + clim\_sst\_grad & 3082.14\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Winter}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad + clim\_sst\_grad & 3262.45\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst & 3234.21\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Spring}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + dist\_coast + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad + clim\_eke\_mean & 5854.32\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst + clim\_eke\_mean & 5824.38\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{landscape}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../EDA/Fig1-study-area-raw-counts}

}

\caption{ }\label{fig:fig1-study-area}
\end{figure}

\end{landscape}

\newpage

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../results/Fig4_partial-plots-spring}

}

\caption{ }\label{fig:fig2-partial-plots}
\end{figure}

\newpage

\begin{figure}

{\centering \includegraphics[width=1.1\linewidth]{../results/Fig2_point-predictions}

}

\caption{ }\label{fig:fig3-point-pred}
\end{figure}

\newpage

\begin{landscape}

\begin{figure}

{\centering \includegraphics[width=1.1\linewidth]{../results/Fig3_probability-maps-spring}

}

\caption{ }\label{fig:fig4-prob-maps}
\end{figure}

\newpage

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../results/Fig5_spp-profiles-spring}

}

\caption{ }\label{fig:fig5-spp-profiles}
\end{figure}

\end{landscape}

\bibliographystyle{unsrt}
\bibliography{../references.bib}

\end{document}

### 2023-10-13 docs/ms-draft.Rmd
updated figs 2 and 5 dimensions, and README (docs/* file uptaded too but I may drop this one off as now the ms lives in ms_preprint/)

@@
 
-```{r fig4-partial-plots, fig.cap = "**FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+```{r fig2-partial-plots, fig.cap = "**FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
 The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 
-```{r fig2-point-pred, fig.cap = "**FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+```{r fig3-point-pred, fig.cap = "**FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
 Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}
@@
 
-```{r fig3-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+```{r fig4-prob-maps, fig.cap = "**FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
 The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.

### 2023-10-13 scripts/08_1_modelling_fit-RCPs-Bernoulli.R

@@ -635,14 +635,15 @@ for (season_vec in seasons_vec) {
     ylab("") + xlab("Probability of Occurrence + Confidence Interval") +
     theme_bw() +
     theme(legend.position = "none",
-          axis.title = element_text(size = 8),
-          axis.text = element_text(size = 7),
+          axis.title = element_text(size = 7),
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7.2),
           strip.text = element_text(size = 8))
   
   ggsave(plot = plot_spp_profiles_CI, 
          filename = 
            paste0("./results/Bernoulli/Bernoulli_", as.character(season_vec), "_07_species-profiles.png"),
-         width = 16, height = 8, units = "cm", dpi = 300)
+         width = 16, height = 10, units = "cm", dpi = 300)
   
   # Save ggplot objects
   save("plot_spp_profiles_CI",
@@ -748,11 +749,11 @@ map_point_pred <-
   geom_sf(data = plotRCPs_all, aes(fill = final_RCP)) + 
   coord_sf(xlim = c(140, 161), ylim = c(-9, -48)) +
   scale_x_continuous(breaks = c(140, 150, 160)) +
-  scale_fill_brewer(palette = "Dark2", name = "Hard-class RCP") + 
+  scale_fill_brewer(palette = "Dark2", name = "Hard-class RCP:") + 
   facet_wrap(~season, ncol = 4) + 
-  theme(legend.position = "right",
-        legend.title = element_text(size = 9),
-        legend.text = element_text(size = 7.5),
+  theme(legend.position = "bottom",
+        legend.title = element_text(size = 6),
+        legend.text = element_text(size = 6),
         legend.key.size = unit(0.5, "cm"),
         axis.text = element_text(size = 6))
 

### 2023-10-13 scripts/08_2_modelling_fit-RCPs-NegBin.R

@@ -629,14 +629,15 @@ for (season_vec in seasons_vec) {
     ylab("") + xlab("Log(predicted mean abundance) + Confidence Interval") +
     theme_bw() +
     theme(legend.position = "none",
-          axis.title = element_text(size = 8),
-          axis.text = element_text(size = 7),
+          axis.title = element_text(size = 7),
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7.2),
           strip.text = element_text(size = 8))
   
   ggsave(plot = plot_spp_profiles_CI, 
          filename = 
            paste0("./results/NegBin/NegBin_", as.character(season_vec), "_07_species-profiles.png"),
-         width = 16, height = 8, units = "cm", dpi = 300)
+         width = 16, height = 10, units = "cm", dpi = 300)
   
   # Save ggplot objects
   save("plot_spp_profiles_CI",
@@ -749,10 +750,10 @@ map_point_pred <-
   scale_x_continuous(breaks = c(140, 150, 160)) +
   scale_fill_brewer(palette = "Dark2", name = "Hard-class RCP") + 
   facet_wrap(~season, ncol = 4) + 
-  theme(legend.position = "right",
-        legend.title = element_text(size = 9),
-        legend.text = element_text(size = 7.5),
-        legend.key.size = unit(0.5, "cm"),
+  theme(legend.position = "none",
+        # legend.title = element_text(size = 7),
+        # legend.text = element_text(size = 7),
+        # legend.key.size = unit(0.5, "cm"),
         axis.text = element_text(size = 6))
 
 ggsave(plot = map_point_pred, 

### 2023-10-13 scripts/09_figs-patchwork.R

@@ -33,6 +33,9 @@ ggsave(fig1_study_area,
        filename = "./results/Fig1_study-area-w-oceanogr.png",
        height = 10, width = 20, units = "cm", dpi = 300)
 
+## The quality of this plot was not great. So I ended up using a .svg version
+## that I put both maps together by hand in Inkscape software
+
 ## FIGURE 2 ------------------------------------------------------------------ #
 
 load("./results/Bernoulli/Bernoulli_07_RCP-point-predictions-ggplot.rda")
@@ -53,8 +56,8 @@ fig2_point_pred <-
   patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
 
 ggsave(fig2_point_pred, 
-       filename = "./results/Fig2-point-predictions.png",
-       height = 13, width = 15, units = "cm", dpi = 300)
+       filename = "./results/Fig2_point-predictions.png",
+       height = 15, width = 11.5, units = "cm", dpi = 300)
 
 rm("map_point_pred_Ber", "map_point_pred", "fig2_point_pred")
 
@@ -74,7 +77,7 @@ fig3_prob_maps_spring <-
   patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")")
 
 ggsave(fig3_prob_maps_spring, 
-       filename = "./results/Fig3-probability-maps-spring.png",
+       filename = "./results/Fig3_probability-maps-spring.png",
        height = 9, width = 16, units = "cm", dpi = 300)
 
 rm("plotRCPs_Ber", "plotRCPs", "fig3_prob_maps_spring")
@@ -120,14 +123,16 @@ plot_spp_profiles_CI_Ber <- plot_spp_profiles_CI
 load("./results/NegBin/NegBin_spring_07_species-profiles-ggplot.rda")
 
 fig5_spp_profiles <- 
-  (plot_spp_profiles_CI_Ber + xlab("Probability of Occurrence\n + Confidence Interval") + theme(axis.text.y = element_text(size = 5))) / 
-  (plot_spp_profiles_CI + xlab("log(Predicted mean abundance)\n + Confidence Interval") + theme(axis.text.y = element_blank())) +
+  (plot_spp_profiles_CI_Ber + xlab("Probability of Occurrence\n + Confidence Interval") + theme(axis.text.y = element_text(size = 8),
+                                                                                                axis.title.x = element_text(size = 7.5))) / 
[…]

### 2023-10-17 ms_preprint/ms_preprint.Rmd
ms tidy-up and supplementary material in rmd

@@
   - Assemblages
+  - Biogeography
   - Community ecology
@@
 bibliography: ../references.bib
-# biblio-style: 
 csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
-output: rticles::arxiv_article
+# output: rticles::arxiv_article
+output:
+  bookdown::pdf_book:
+    base_format: rticles::arxiv_article
 header-includes:
@@
 - \usepackage{pdflscape}
-- \DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}}
+- \DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}} # needed to understand unicode for greek Delta in Table 1
 - \newcommand{\blandscape}{\begin{landscape}}
@@
 
-<!-- Running title: Seabird assemblages off East Australia -->
+<!-- Running title: Seabird assemblages off Eastern Australia -->
+
 <!-- Note: I've changed the 'arxiv.sty' source file to include the running title in the PDF -->
@@
 The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.
-1a**; @kessler2013; @ganachaud2014).
+\@ref(fig:fig1-study-area)a**; @kessler2013; @ganachaud2014).
 **SST/SSS characteristics?
@@
 Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.
-1a**), carrying eddy-like features as a transient current [@oke2019].
+\@ref(fig:fig1-study-area)a**), carrying eddy-like features as a transient current [@oke2019].
 
@@
 Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.
-1b**; **Table S1**).
+\@ref(fig:fig1-study-area)b**; **Table S1**).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
 The species, the numbers of birds, and their behaviours were logged in real time into a ship-based web server.
-**Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.** Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species).
+Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.
+Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species).
 Birds following the vessels were excluded from analyses.
@@
 
-To investigate and predict species' distributions and relationships with environmental variables, a considered choice of **broad-scale climatic and oceanographic** covariates is important [@barbet-massin2014].
-We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table 1**).
+To investigate and predict species' distributions and relationships with environmental variables, a considered choice of broad-scale climatic and oceanographic covariates is important [@barbet-massin2014].
+We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table \@ref(tab:table1)**).
 These variables are the most frequently used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of spatial and temporal scales from macro-scale water-mass properties to prey availability in meso- to submesoscale oceanographic features [@haury1978patterns; @weimerskirch2007; @scales2014].
@@
 
-Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material **XX**.
-Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in **Daudt et al.
-2023 (Zenodo repo)**.
+Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material.
+Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in **(Zenodo repo)**.
 
@@
 Thus, 'eke' and 'sst_grad' environmental layers were calculated before extracting environmental data for each seabird record.
-For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table 1**) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
+For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table \@ref(tab:table1)**) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
 Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ [@sun2022].
 The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
-In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table 1**).
+In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table \@ref(tab:table1)**).
 Seabed slope was calculated using 'raster::terrain()' [@raster], and distance from the coast was calculated through 'sf::st_distance()' [@sf] using Australia and Tasmania as the mainland (i.e.
@@
 ### Data {#modellingdatasets}
+
 \label{sec:modellingdatasets}
@@
 110 $\times$ 110 km) and season (**Fig.
-1b**), by summing seabirds from the same species.
+\@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 S1**).
-An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig S2**).
+An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig.
+S2**).
 Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Fig.
@@
 S3**).
-The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**).
+The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table \@ref(tab:table2)**).
 The best models showed a good fit with no apparent deviations from the assumptions (**Fig.
@@
 The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21 °C, or 'sss' 35.1 PSU for summer (**Fig.
-2**, **Fig.
+\@ref(fig:fig2-partial-plots)**, **Fig.
 S5**), similar to the EAC's temperature and salinity thresholds.
@@
 When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.
-3**).
+\@ref(fig:fig3-point-pred)**).
 These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.
-4**, **Fig.
+\@ref(fig:fig4-prob-maps)**, **Fig.
 S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
 The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Straight' assemblage appears (**Fig.
-3**, **Fig.
+\@ref(fig:fig3-point-pred)**, **Fig.
 S6**).
@@
 turtur*) (**Fig.
-5**, **Fig.
+\@ref(fig:fig5-spp-profiles)**, **Fig.
 S7**).
@@
 ## Limitations
+
 \label{sec:limitations}
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-# Tables and Figures {-}
+# Tables and Figures {.unnumbered}
 
@@
 (2021)${^a}$", "GEBCO${^f}$", "Derived from GEBCO${^f}$", "Calculated in R (see main text)")
-) |>
+) %>% 
   kableExtra::kbl(caption = " ", booktabs = TRUE, escape = FALSE)  %>% 

### 2023-10-17 ms_preprint/ms_preprint.tex
ms tidy-up and supplementary material in rmd

@@
 
-\title{Seabird assemblages are linked to the major western boundary
-current off eastern Australia}
+\title{Seabird assemblages are linked to the major western boundary current off eastern Australia}
 
@@
 
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 % Pandoc citation processing
@@
 \begin{abstract}
-\textbf{Aim}: To identify assemblages of seabird species and relate them
-to environmental conditions off eastern Australia, a world seabird
-hotspot. \textbf{Methods}: We used 15 seabird at-sea ship-based survey
-data collected over 37° of latitude, from 2016 to 2021, and fitted
-Region of Common Profile (RCP) mixture models for each season and type
-of data (presence-absence and abundance). RCP groups are defined as a
-set of regions where the probability of encountering a specific species
-profile is constant within regions, but different amongst them. These
-groups also vary according to covariates, which in our case included
-oceanographic, climatic, and physiographic parameters. \textbf{Results}:
-Based on 142,646 seabirds recorded, from 80 species, all models
-suggested two macro-scale assemblages (`northern' and `southern'),
-except for the autumn presence-absence model that identified three
-groups. The model results consistently show a biogeographic transition
-at \textasciitilde34°S, near the latitude at which the East Australian
-Current separates from the Australian continental slope. Sea surface
-temperatures or sea surface salinities were selected in all final
-models, further indicating a close relationship between seabird
-assemblages and water masses. Results from both data types,
-presence-absence and abundance, resulted in similar spatial and species
-profile patterns. \textbf{Conclusion}: RCP models clearly identified two
-seabird assemblages off the east coast of Australia, suggesting the
-persistence of these groups at seasonal and macro spatial scales. Given
-the rapid climatic changes that the East Australian Current is
-experiencing and its importance in influencing distributions of
-seabirds, the methods applied in our study could be replicated to assess
-possible changes in seabird assemblages, and how they are affected by
-changing environmental conditions.
+\textbf{Aim}: To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+\textbf{Methods}: We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
+These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
+\textbf{Results}: Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
+Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
+\textbf{Conclusion}: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 \end{abstract}
@@
     Assemblages
+   \and
+    Biogeography
    \and
@@
 
-Bioregionalization is the discipline that identifies, delimits,
-characterises, and names biogeographical regions (Morrone, 2018; Vilhena
-\& Antonelli, 2015). Although there is often no agreement on the naming
-systems among proposed regionalization schema (but see Ebach et al.,
-2008), the hierarchical nature of bioregions is well recognized (Kreft
-\& Jetz, 2010; Vilhena \& Antonelli, 2015). Thus, bioregions are usually
-described within geographically larger classifications, such as
-`provinces' or `realms' (Costello et al., 2017; Spalding et al., 2007).
-The delimitation of bioregions is made through identifying sets of
-species that are endemic or have high levels of co-occurrence, and are
-distinct enough from other regions. In this sense, bioregions are
-interpreted as a natural representations of common historical and/or
-ecological processes shaped by physical and biological forces (Kreft \&
-Jetz, 2010). In particular, the identification of bioregions allows
-representation of the distribution of multiple species at a given time
-and space, and relating these regions to environmental characteristics.
-By doing so, bioregions can be interpreted as species assemblages (N. A.
-Hill et al., 2017; Lyons et al., 2017). Given these features, bioregions
-can be used as surrogates for biodiversity monitoring in data-poor
-regions and in regional conservation planning (Ferrier, 2002; Spalding
-et al., 2007), playing a fundamental role in conservation biogeography
-(Whittaker et al., 2005).
-
-Seabird assemblages have been widely described and related to
-environmental characteristics (Hyrenbach et al., 2007; Ribic et al.,
-1997; Woehler et al., 2003). Considering the hierarchical nature of
-oceanographic processes and their resulting features (Haury et al.,
-1978), it is not surprising that the distribution of highly mobile
-seabirds are linked differently across temporal and spatial scales
-(Haury et al., 1978; Hunt \& Schneider, 1987). In the marine
-environment, large scale (\textgreater100 km length scale) oceanographic
-features (e.g.~boundary currents and water masses) act as terrestrial
-biomes (based on vegetation), and influence seabird assemblage
-structure--i.e.~the occurrence of species and their relative numerical
-composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al.,
-1997). In contrast, local abundance is related to meso- (10--100 km) and
-sub-mesoscale (1--10 km) processes that aggregate prey and therefore
-predators, such as fronts and eddies (Scales, Miller, Hawkes, et al.,
-2014; Weimerskirch, 2007). Such processes, although relatively
-ephemeral, are primarily driven by physical and atmospheric forces, and
-are thus somewhat predictable in space and time, providing consistent
-feeding areas for marine predators when they occur (Davoren, 2013;
-Hazen, Suryan, et al., 2013). Sea surface temperatures are a key driver
-of seabird distributions, at individual- and assemblage-levels,
-operating over a range of spatial and temporal scales from macro
-(e.g.~water masses) to fine (e.g.~fronts) (Dunlop et al., 1988;
-Quillfeldt et al., 2015; Ribic et al., 1997). Despite extensive surveys
-(e.g. Ballance et al., 1997) and tracking studies (e.g. Carneiro et al.,
-2020), many oceanic regions still lack information on the distributions
-and abundances of seabird species, and how they assemble in space and
-time.
-
-One area in need of better understanding of seabird distributions is
-Australia, considered a global seabird hotspot (Karpouzi et al., 2007;
-Ramírez et al., 2017). Given the region's importance for seabird
-diversity, there are surprisingly few studies on seabird distributions
-at sea (Bernard et al., 2021; Mott \& Clarke, 2018). Off northwest
-Australia, at-sea surveys, multivariate analyses (Dunlop et al., 1988)
-and ensemble species distribution models of species assemblages (Lavers
-et al., 2014) showed that habitat use by tropical seabird species was
-related to sea surface temperatures and salinity. Dunlop et al. (1988)
-suggested that salinity was the main variable influencing species'
-distributions in oceanic waters while temperature was the key factor for
-species on the continental shelf. In contrast, there are few studies at
-large spatial and temporal scales off eastern Australia. T. A. Reid \&
-Hindell (2000) and Evans et al. (2021) analysed seabird assemblages and
-their relationships with oceanographic and prey data at coarse scales
-(tens of km) off southeast Tasmania. The remaining studies are mostly
-descriptive accounts locally (Blaber \& Milton, 1994; Norman, 1992), or
-focused on a single species (Priddel et al., 2014).
-
-The East Australian Current (EAC) is the major western boundary current
-in the South Pacific Ocean, and the dominant boundary current off
-eastern Australia (Ganachaud et al., 2014). The EAC has been extensively
-studied, although research into the links between the EAC's oceanography
-and marine predators are not well known. The dynamic features associated
-with the EAC, such as mesoscale eddies and upwelling, are likely to
-influence marine biota (Scales, Miller, Hawkes, et al., 2014; Suthers et
-al., 2011). While most of the biological studies in this system were
-focused on plankton entrapment within the cores of eddies (Firme et al.,
-2023; Garcia et al., 2022; Hassler et al., 2011), there is a growing
-body of work focusing on fisheries (Brieva et al., 2015; Hobday, 2010;
-Young et al., 2011). These studies emphasized how the probability of
-occurrence of a given taxon, and the structure of species assemblages,
-were driven by the EAC. However, the relationships between oceanography
-and marine megafauna within the EAC are poorly known (Suthers et al.,
-2011). Some progress has been made using shark and ray movement data
-(Jaine et al., 2014; Niella et al., 2020; Niella et al., 2022), and
-satellite tracking has also demonstrated that increased sea surface
-temperatures in the EAC decreases little penguin \emph{Eudyptula minor}
-foraging success (Carroll et al., 2016), but information at the
-assemblage-level for seabirds in the EAC is absent.
-
-Marine species range shifts are more likely to occur faster than
-terrestrial species due to their higher sensitivity to warming (Lenoir
-et al., 2020; Pinsky et al., 2019). As a result, climate change is
-rearranging the structure of species assemblages around the globe (Antão
-et al., 2020; Dornelas et al., 2014; Poloczanska et al., 2013). Top
-predators, for instance, may change their core habitats (Hazen,
-Jorgensen, et al., 2013; Reisinger et al., 2022), with unprecedented
-population consequences (Woehler \& Hobday, 2023). For seabirds, Péron
-et al. (2010) and Sojitra et al. (2022) showed that sea surface
-temperature anomalies affected the occurrence and abundances of
-albatrosses and petrels species with latitude, thus affecting assemblage
-structure. In eastern Australia, changing EAC conditions have affected a
-range of marine species. For example, changes in zooplankton assemblage
-structure and fish distribution patterns have been recorded and
-attributed to warming water temperatures, and shifts in shark
-distributions are expected (C. R. Johnson et al., 2011; Niella et al.,
-2022). A severe marine heatwave driven by a poleward strengthening of
-the EAC, showed drastic environmental impacts including disease
-outbreaks in aquaculture farms and wildlife mortalities (Oliver et al.,
-2017). Given that the EAC is strengthening and extending farther
-southward and getting warmer (Cai et al., 2005; Phillips et al., 2022),
-biological changes are expected to occur (Hobday, 2010; Poloczanska et
-al., 2007).
-
-In light of eastern Australia's importance for seabirds, and the recent
-attention on how predicted climate change will influence the EAC's
-marine biota (Poloczanska et al., 2007), an assessment of the
-contemporary spatial distributions of seabirds within the EAC at large
-temporal and spatial scales is timely. Seabirds can serve as sentinels
-for tracking ecological changes in the EAC in many ways (e.g. Sydeman et
-al., 2012). Surprisingly though, seabird distribution is
-under-represented in climate change studies, despite being highlighted
-as a priority topic (Grémillet \& Boulinier, 2009; Sydeman et al.,
-2012).
-
-In this study, we aim to identify assemblages of seabird species within
-the EAC and adjacent areas, and relate these assemblages to
-environmental conditions, by using at-sea surveys data and a
-state-of-the-art modelling technique. We use the Region of Common
-Profiles {[}RCP; Foster et al. (2013){]}, a mixture model regarded as
-one of the best available tools for describing bioregions (N. A. Hill et
-al., 2020; S. N. C. Woolley et al., 2020). RCP is a model-based
-approach, and has a range of advantages over traditional algorithm-based
-statistics (N. A. Hill et al., 2020; Warton et al., 2015). Eastern
-Australia supports a high diversity of seabirds and offers wide
-environmental gradients over a near 40° latitudinal range, from the
-tropics to the cool temperate. We hypothesised to find different seabird
-assemblages relating to different environmental conditions within the
-EAC. In addition, presence-absence and abundance data were used to test
-if the different numerical resolutions would result in the
-identification of similar assemblages. This study provides a repeatable
-framework on which future studies can compare to assess how future
-changes in the EAC may influence the distribution of seabird
-assemblages.
+Bioregionalization is the discipline that identifies, delimits, characterises, and names biogeographical regions (Morrone, 2018; Vilhena \& Antonelli, 2015).
+Although there is often no agreement on the naming systems among proposed regionalization schema (but see Ebach et al., 2008), the hierarchical nature of bioregions is well recognized (Kreft \& Jetz, 2010; Vilhena \& Antonelli, 2015).
+Thus, bioregions are usually described within geographically larger classifications, such as `provinces' or `realms' (Costello et al., 2017; Spalding et al., 2007).
+The delimitation of bioregions is made through identifying sets of species that are endemic or have high levels of co-occurrence, and are distinct enough from other regions.
+In this sense, bioregions are interpreted as a natural representations of common historical and/or ecological processes shaped by physical and biological forces (Kreft \& Jetz, 2010).
+In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space, and relating these regions to environmental characteristics.
+By doing so, bioregions can be interpreted as species assemblages (N.
+A.
+Hill et al., 2017; Lyons et al., 2017).
+Given these features, bioregions can be used as surrogates for biodiversity monitoring in data-poor regions and in regional conservation planning (Ferrier, 2002; Spalding et al., 2007), playing a fundamental role in conservation biogeography (Whittaker et al., 2005).
+
+Seabird assemblages have been widely described and related to environmental characteristics (Hyrenbach et al., 2007; Ribic et al., 1997; Woehler et al., 2003).
+Considering the hierarchical nature of oceanographic processes and their resulting features (Haury et al., 1978), it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales (Haury et al., 1978; Hunt \& Schneider, 1987).
+In the marine environment, large scale (\textgreater100 km length scale) oceanographic features (e.g.~boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.~the occurrence of species and their relative numerical composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al., 1997).
+In contrast, local abundance is related to meso- (10--100 km) and sub-mesoscale (1--10 km) processes that aggregate prey and therefore predators, such as fronts and eddies (Scales, Miller, Hawkes, et al., 2014; Weimerskirch, 2007).
+Such processes, although relatively ephemeral, are primarily driven by physical and atmospheric forces, and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur (Davoren, 2013; Hazen, Suryan, et al., 2013).
+Sea surface temperatures are a key driver of seabird distributions, at individual- and assemblage-levels, operating over a range of spatial and temporal scales from macro (e.g.~water masses) to fine (e.g.~fronts) (Dunlop et al., 1988; Quillfeldt et al., 2015; Ribic et al., 1997).
+Despite extensive surveys (e.g.
+Ballance et al., 1997) and tracking studies (e.g.
+Carneiro et al., 2020), many oceanic regions still lack information on the distributions and abundances of seabird species, and how they assemble in space and time.
+
+One area in need of better understanding of seabird distributions is Australia, considered a global seabird hotspot (Karpouzi et al., 2007; Ramírez et al., 2017).
+Given the region's importance for seabird diversity, there are surprisingly few studies on seabird distributions at sea (Bernard et al., 2021; Mott \& Clarke, 2018).
+Off northwest Australia, at-sea surveys, multivariate analyses (Dunlop et al., 1988) and ensemble species distribution models of species assemblages (Lavers et al., 2014) showed that habitat use by tropical seabird species was related to sea surface temperatures and salinity.
+Dunlop et al.
+(1988) suggested that salinity was the main variable influencing species' distributions in oceanic waters while temperature was the key factor for species on the continental shelf.
+In contrast, there are few studies at large spatial and temporal scales off eastern Australia.
+T.
+A.
+Reid \& Hindell (2000) and Evans et al.
+(2021) analysed seabird assemblages and their relationships with oceanographic and prey data at coarse scales (tens of km) off southeast Tasmania.
+The remaining studies are mostly descriptive accounts locally (Blaber \& Milton, 1994; Norman, 1992), or focused on a single species (Priddel et al., 2014).
+
+The East Australian Current (EAC) is the major western boundary current in the South Pacific Ocean, and the dominant boundary current off eastern Australia (Ganachaud et al., 2014).
+The EAC has been extensively studied, although research into the links between the EAC's oceanography and marine predators are not well known.
+The dynamic features associated with the EAC, such as mesoscale eddies and upwelling, are likely to influence marine biota (Scales, Miller, Hawkes, et al., 2014; Suthers et al., 2011).
+While most of the biological studies in this system were focused on plankton entrapment within the cores of eddies (Firme et al., 2023; Garcia et al., 2022; Hassler et al., 2011), there is a growing body of work focusing on fisheries (Brieva et al., 2015; Hobday, 2010; Young et al., 2011).
+These studies emphasized how the probability of occurrence of a given taxon, and the structure of species assemblages, were driven by the EAC.
+However, the relationships between oceanography and marine megafauna within the EAC are poorly known (Suthers et al., 2011).
+Some progress has been made using shark and ray movement data (Jaine et al., 2014; Niella et al., 2020; Niella et al., 2022), and satellite tracking has also demonstrated that increased sea surface temperatures in the EAC decreases little penguin \emph{Eudyptula minor} foraging success (Carroll et al., 2016), but information at the assemblage-level for seabirds in the EAC is absent.
+
+Marine species range shifts are more likely to occur faster than terrestrial species due to their higher sensitivity to warming (Lenoir et al., 2020; Pinsky et al., 2019).
+As a result, climate change is rearranging the structure of species assemblages around the globe (Antão et al., 2020; Dornelas et al., 2014; Poloczanska et al., 2013).
+Top predators, for instance, may change their core habitats (Hazen, Jorgensen, et al., 2013; Reisinger et al., 2022), with unprecedented population consequences (Woehler \& Hobday, 2023).
+For seabirds, Péron et al.
+(2010) and Sojitra et al.
+(2022) showed that sea surface temperature anomalies affected the occurrence and abundances of albatrosses and petrels species with latitude, thus affecting assemblage structure.
+In eastern Australia, changing EAC conditions have affected a range of marine species.
+For example, changes in zooplankton assemblage structure and fish distribution patterns have been recorded and attributed to warming water temperatures, and shifts in shark distributions are expected (C.
+R.
+Johnson et al., 2011; Niella et al., 2022).
+A severe marine heatwave driven by a poleward strengthening of the EAC, showed drastic environmental impacts including disease outbreaks in aquaculture farms and wildlife mortalities (Oliver et al., 2017).
+Given that the EAC is strengthening and extending farther southward and getting warmer (Cai et al., 2005; Phillips et al., 2022), biological changes are expected to occur (Hobday, 2010; Poloczanska et al., 2007).
+
+In light of eastern Australia's importance for seabirds, and the recent attention on how predicted climate change will influence the EAC's marine biota (Poloczanska et al., 2007), an assessment of the contemporary spatial distributions of seabirds within the EAC at large temporal and spatial scales is timely.
+Seabirds can serve as sentinels for tracking ecological changes in the EAC in many ways (e.g.
+Sydeman et al., 2012).
+Surprisingly though, seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic (Grémillet \& Boulinier, 2009; Sydeman et al., 2012).
+
+In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
+We use the Region of Common Profiles {[}RCP; Foster et al.
+(2013){]}, a mixture model regarded as one of the best available tools for describing bioregions (N.
+A.
+Hill et al., 2020; S.
+N.
+C.
+Woolley et al., 2020).
+RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics (N.
+A.
+Hill et al., 2020; Warton et al., 2015).
+Eastern Australia supports a high diversity of seabirds and offers wide environmental gradients over a near 40° latitudinal range, from the tropics to the cool temperate.
+We hypothesised to find different seabird assemblages relating to different environmental conditions within the EAC.
+In addition, presence-absence and abundance data were used to test if the different numerical resolutions would result in the identification of similar assemblages.
+This study provides a repeatable framework on which future studies can compare to assess how future changes in the EAC may influence the distribution of seabird assemblages.
 
@@
 
-The EAC originates from the South Equatorial Current (SEC), a tropical
-nutrient-poor current. When the SEC's North Caledonian Jet meets the
-Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing
-branch towards the Gulf of Papua and a south-flowing branch, which marks
-the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte,
-2013). The South Caledonian Jet completes the EAC system at
-\textasciitilde22°S after encountering the shelf break in the southern
-portion of the Great Barrier Reef (\textbf{Fig. 1a}; Kessler \& Cravatte
-(2013); Ganachaud et al. (2014)). \textbf{SST/SSS characteristics? ROB?}
-The EAC interacts with the bathymetry of the seafloor, giving rise to
-oceanographic features and processes, such as smaller currents,
-upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019;
-Oke \& Middleton, 2000; Ridgway \& Dunn, 2003). The EAC flow polewards
-and is primarily associated with the shelf break, but at
-\textasciitilde32°S it detaches from the shelf and spawns an area of
-eddy generation that dissipates throughout the western Tasman Sea (Mata
-et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension'
-(Oke, Pilo, et al. (2019); often termed the `Tasman Front'). Although
-most EAC energy is converted into eddies, the EAC still runs southwards
-following the shelf break towards Tasmania (the `EAC southern
-extension'; \textbf{Fig. 1a}), carrying eddy-like features as a
-transient current (Oke, Roughan, et al., 2019).
+The EAC originates from the South Equatorial Current (SEC), a tropical nutrient-poor current.
+When the SEC's North Caledonian Jet meets the Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte, 2013).
+The South Caledonian Jet completes the EAC system at \textasciitilde22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (\textbf{Fig.
+\ref{fig:fig1-study-area}a}; Kessler \& Cravatte (2013); Ganachaud et al.
+(2014)).
+\textbf{SST/SSS characteristics?
+ROB?} The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019; Oke \& Middleton, 2000; Ridgway \& Dunn, 2003).
+The EAC flow polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension' (Oke, Pilo, et al.
+(2019); often termed the `Tasman Front').
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the `EAC southern extension'; \textbf{Fig.
+\ref{fig:fig1-study-area}a}), carrying eddy-like features as a transient current (Oke, Roughan, et al., 2019).
 
 \hypertarget{seabird-and-environmental-data-sources}{%
-\subsection{Seabird and environmental data
-sources}\label{seabird-and-environmental-data-sources}}
-
-Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV
-Falkor} on 15 voyages, between 2016/17--2020/21, spanning over
-\textasciitilde37° of latitude (\textbf{Fig. 1b}; \textbf{Table S1}).
-One or more observers continuously recorded seabirds within a 300 m
-forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours
-while the ship was underway. The species, the numbers of birds, and
-their behaviours were logged in real time into a ship-based web server.
-\textbf{Environmental parameters for atmospheric and oceanographic
-conditions were logged continuously by automated-ship based sensors,
-with seabird observations supplementing relevant environmental
-conditions such as visibility or associations with marine mammals.}
-Individual records received a date/time/geographic coordinate label
-(multi-species records have the same label but are logged as separate
-entries for each species). Birds following the vessels were excluded
-from analyses. Seabird observation protocols followed Raymond et al.
+\subsection{Seabird and environmental data sources}\label{seabird-and-environmental-data-sources}}
+
+Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV Falkor} on 15 voyages, between 2016/17--2020/21, spanning over \textasciitilde37° of latitude (\textbf{Fig.
+\ref{fig:fig1-study-area}b}; \textbf{Table S1}).
+One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
+The species, the numbers of birds, and their behaviours were logged in real time into a ship-based web server.
+Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.
+Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species).
+Birds following the vessels were excluded from analyses.
+Seabird observation protocols followed Raymond et al.
 (2010), in use around Australia since 1980.
 
-To investigate and predict species' distributions and relationships with
-environmental variables, a considered choice of \textbf{broad-scale
-climatic and oceanographic} covariates is important (Barbet-Massin \&
-Jetz, 2014). We used three physiographic variables and seven
-oceanographic variables, from which a further three climatic layers were
-calculated (\textbf{Table 1}). These variables are the most frequently
-used in seabird-environment relationship studies (Tremblay et al.,
-2009), capturing processes at a range of spatial and temporal scales
-from macro-scale water-mass properties to prey availability in meso- to
-submesoscale oceanographic features (Haury et al., 1978; Scales, Miller,
-Hawkes, et al., 2014; Weimerskirch, 2007).
+To investigate and predict species' distributions and relationships with environmental variables, a considered choice of broad-scale climatic and oceanographic covariates is important (Barbet-Massin \& Jetz, 2014).
+We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (\textbf{Table \ref{tab:table1}}).
+These variables are the most frequently used in seabird-environment relationship studies (Tremblay et al., 2009), capturing processes at a range of spatial and temporal scales from macro-scale water-mass properties to prey availability in meso- to submesoscale oceanographic features (Haury et al., 1978; Scales, Miller, Hawkes, et al., 2014; Weimerskirch, 2007).
 
@@
 
-Data wrangling, modelling and visualization were done in R 4.2.0 (R Core
-Team, 2022); the full list of packages can be found in the Supplementary
-Material \textbf{XX}. Code and the modelling dataset (see
-\ref{sec:modellingdatasets}) are available in \textbf{Daudt et al.~2023
-(Zenodo repo)}.
+Data wrangling, modelling and visualization were done in R 4.2.0 (R Core Team, 2022); the full list of packages can be found in the Supplementary Material.
+Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in \textbf{(Zenodo repo)}.
 
@@
 
-The spatial and temporal persistence of oceanographic features such as
-eddy kinetic energy (`eke') and thermal fronts (sea surface temperature
-gradient: `sst\_grad') may influence the movements and distributions of
-seabirds by offering a higher likelihood of finding food (Scales,
-Miller, Embling, et al., 2014; Weimerskirch, 2007). Thus, `eke' and
-`sst\_grad' environmental layers were calculated before extracting
-environmental data for each seabird record. For both `eke' and
-`sst\_grad', we used their original cell spatial resolution
-(\textbf{Table 1}) and calculations were done layer-by-layer to match
-seabird data on the finest spatial scale possible. Eddy kinetic energy
-was calculated as \(0.5 (u'² + v'²)\), where \(u' = u - \bar{u}\) and
-\(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian
-(longitudinal) geostrophic currents, respectively, using a rolling mean
-of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022). The
-spatial `sst\_grad' for each cell was calculated by averaging the
-maximum value for each neighbouring cell (average neighbourhood
-maximum), from north-to-south (NS) and west-to-east (WE); then, the
-magnitude of the gradient was defined as \(NS² + WE²\) (Burrows et al.,
-2011). In addition, we calculated climatic seasonal average layers for
-`eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a
-proxy for persistent oceanographic features, and the `eke' standard
-deviation over the same period (named with the prefix `clim\_' in
-\textbf{Table 1}). Seabed slope was calculated using `raster::terrain()'
-(Hijmans, 2022a), and distance from the coast was calculated through
-`sf::st\_distance()' (Pebesma, 2018) using Australia and Tasmania as the
-mainland (i.e.~ignoring smaller islands). Then, for each seabird record,
-the most proximate value for each environmental layer (original or
-calculated) was extracted using built-in functions from `raster' 3.5-21,
-`terra' 1.6-7, and `sf' 1.0-8 packages (Hijmans, 2022a, 2022b; Pebesma,
-2018).
+The spatial and temporal persistence of oceanographic features such as eddy kinetic energy (`eke') and thermal fronts (sea surface temperature gradient: `sst\_grad') may influence the movements and distributions of seabirds by offering a higher likelihood of finding food (Scales, Miller, Embling, et al., 2014; Weimerskirch, 2007).
+Thus, `eke' and `sst\_grad' environmental layers were calculated before extracting environmental data for each seabird record.
+For both `eke' and `sst\_grad', we used their original cell spatial resolution (\textbf{Table \ref{tab:table1}}) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
+Eddy kinetic energy was calculated as \(0.5 (u'² + v'²)\), where \(u' = u - \bar{u}\) and \(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022).
+The spatial `sst\_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as \(NS² + WE²\) (Burrows et al., 2011).
+In addition, we calculated climatic seasonal average layers for `eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the `eke' standard deviation over the same period (named with the prefix `clim\_' in \textbf{Table \ref{tab:table1}}).
+Seabed slope was calculated using `raster::terrain()' (Hijmans, 2022a), and distance from the coast was calculated through `sf::st\_distance()' (Pebesma, 2018) using Australia and Tasmania as the mainland (i.e.~ignoring smaller islands).
+Then, for each seabird record, the most proximate value for each environmental layer (original or calculated) was extracted using built-in functions from `raster' 3.5-21, `terra' 1.6-7, and `sf' 1.0-8 packages (Hijmans, 2022a, 2022b; Pebesma, 2018).
 
@@
 
-Seabirds are highly mobile animals, and many species migrate and/or
-disperse during their non-breeding periods (Schreiber \& Burger, 2002).
-Therefore, seasonal differences in species composition were expected
-(e.g. K. Reid et al., 2023), and in the probabilities of occurrence and
-in total numbers. Thus, we built separate modelling dataset for each
-season. We aggregated seabird data by 1° latitude \(\times\) 1°
-longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Fig.
-1b}), by summing seabirds from the same species. To test if the
-presence-absence and abundance (raw counts) data would result in similar
-assemblage patterns, we built two species matrices, where the
-presence-absence matrix transformed values \textgreater1 to 1. To avoid
-numerical instabilities in the models prompted by unusual or rare
-species, we kept only species occurring in at least six grids within
-each season (summer = 12 species; autumn = 21; winter = 20; spring =
-33). The associated environmental data of each seabird record were
-averaged within the same grid cells and seasons, and environmental data
-were standardised before modelling.
+Seabirds are highly mobile animals, and many species migrate and/or disperse during their non-breeding periods (Schreiber \& Burger, 2002).
+Therefore, seasonal differences in species composition were expected (e.g.
+K.
+Reid et al., 2023), and in the probabilities of occurrence and in total numbers.
+Thus, we built separate modelling dataset for each season.
+We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Fig.
+\ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
+To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \textgreater1 to 1.
+To avoid numerical instabilities in the models prompted by unusual or rare species, we kept only species occurring in at least six grids within each season (summer = 12 species; autumn = 21; winter = 20; spring = 33).
+The associated environmental data of each seabird record were averaged within the same grid cells and seasons, and environmental data were standardised before modelling.
 
@@
 
-We used the mixture model Region of Common Profiles (RCP) to identify
-and describe seabird species assemblages. RCP is a one-stage
-bioregionalization approach, as it identifies bioregions by specifying a
-statistical model that simultaneously relates response (multi-species
-data) and predictor (environmental data) (N. A. Hill et al., 2020; S. N.
-C. Woolley et al., 2020). Given a set of environmental variables, the
-model can predict assemblage probabilities of sites where no biotic
-(observational) data exist. As a model-based approach, RCPs can be
-compared by information criteria, present uncertainty around the
-results, and check for model assumptions through assessing distribution
-of residuals (N. A. Hill et al., 2020; Warton et al., 2015).
-
-The model is specified in terms of assemblages and site-specific
-probabilities. There are \(K\) assemblages (i.e.~RCPs) that are common
-across the study area, \(K = (1, ..., k)\). The outcome variable is a
-vector \(y_{i} = (y_{i1}, ... y_{iS})\), that give the count or
-presence-absence of each of the \(S\) species at site \(i\). Depending
-on whether we model the counts or presence-absence data, the assemblages
-are defined in terms of the expected frequency or relative frequency of
-the species, respectively. An output of the model is the estimated
-probability that site \(i\) has each of the \(K\) possible assemblages.
-RCPs are a multivariate adaptation of a mixture-of-experts model; as
-such, the mixing proportions (assemblage structure) vary according to
-covariates (predictors) (Foster et al., 2013). However, the final
-estimated assemblage structure is assumed to be constant for all sites
-belonging to the \(K\) assemblage (Foster et al., 2013). As such, RCPs
-can be directly interpreted as assemblages, and their species profiles
-as the assemblage structure.
-
-Before fitting models, covariates were checked for collinearity within
-each seasonal dataset. If a pairwise collinearity of
-\(\left|x\right| >0.7\) was found we removed one of the variables
-(Dormann et al., 2013), favouring direct measures over the `climatic'
-layers we calculated (e.g.~if `sst\_grad' and `clim\_sst\_grad' were
-correlated, we chose `sst\_grad'). Thus, `full models' differed between
-seasons as the retained covariates varied, although most covariates were
-common for all seasons.
-
-Due to seabird migration and dispersion, seasons could act as `sampling
-artifacts' (Foster et al., 2017), thus biasing results (Ponti \&
-Sannolo, 2023). In light of the difference among seasons in the numbers
-of species that met our occurrence threshold (see
-\ref{sec:modellingdatasets}), we decided to use RCP models by season and
-data type (presence-absence and abundance). RCP models were used to
-relate the response variables (i.e.~presence-absence or abundance of
-each seabird species in each grid cell) to the independent predictor
-covariates (i.e.~averaged values of environmental variables in each
-cell). Presence-absence data were modelled based on a Bernoulli model,
-whereas abundance data based on a Negative Binomial model. RCPs were
-fitted using the `ecomix' 1.0.0 package (S. Woolley et al., 2022), with
-the following workflow: (i) find the best number of groups (\(K\)) that
-describes the data, (ii) re-fit the model by fixing \(K\) and selecting
-optimal combination of covariates, (iii) check residuals, (iv) bootstrap
-the best model to estimate uncertainty around parameter estimates, and
-(v) produce results.
-
-To identify the best number of RCP groups (\(K\)) that described the
-data, we used full models (all covariates) with 100 random start numbers
-(see Foster et al., 2013) between \(k = (2, ..., 7)\) groups, and chose
-the number of RCPs that minimized the Bayesian Information Criterion
-(BIC). Then, we fixed \(K\) at the optimal number and fitted models for
-every possible linear, additive combination of covariates. We chose the
-best model as the one with the lowest BIC. Visual inspection of
-randomized quantile residual plots modified for mixture models was used
-to verify model assumptions (Dunstan et al., 2013). To obtain estimates
-of uncertainty for model parameters (confidence intervals), we used the
-default settings of `ecomix::regional\_mix.bootstrap()', which applies a
-Bayesian bootstrap 1000 times. A custom code was written to generate
-partial plots for exploring the influence of covariates on RCP
-membership.
-
-We built a seasonal 11-year (2011--2021) average layer for predicting
-model results for each environmental covariate. These variables were
-then averaged within 1° latitude \(\times\) 1° longitude grid cells, by
-season, between 9--47°S and 141--161°E. Note that `clim\_eke\_mean' and
-`clim\_sst\_grad' were processed beforehand as they may characterise
-persistent oceanographic features (described above) and thus represent
-the same value as `eke' and `sst\_grad'. Data for prediction were
-standardised using the same mean and standard deviation from the
-modelling dataset.
-
-Finally, we ran rarefaction curves based on the presence-absence matrix
-and the assigned RCP group for each sample (grid cell) to assess if the
-amount of information in the RCP model was representative. Rarefaction
-curves were fit using the `iNEXT' 3.0.0 package (Chao et al., 2014;
-Hsieh et al., 2016), and results are presented based on alpha diversity
-(i.e.~species richness).
+We used the mixture model Region of Common Profiles (RCP) to identify and describe seabird species assemblages.
+RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates response (multi-species data) and predictor (environmental data) (N.
+A.
+Hill et al., 2020; S.
+N.
+C.
+Woolley et al., 2020).
+Given a set of environmental variables, the model can predict assemblage probabilities of sites where no biotic (observational) data exist.
+As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals (N.
+A.
+Hill et al., 2020; Warton et al., 2015).
+
+The model is specified in terms of assemblages and site-specific probabilities.
+There are \(K\) assemblages (i.e.~RCPs) that are common across the study area, \(K = (1, ..., k)\).
+The outcome variable is a vector \(y_{i} = (y_{i1}, ...
+y_{iS})\), that give the count or presence-absence of each of the \(S\) species at site \(i\).
+Depending on whether we model the counts or presence-absence data, the assemblages are defined in terms of the expected frequency or relative frequency of the species, respectively.
+An output of the model is the estimated probability that site \(i\) has each of the \(K\) possible assemblages.
+RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) (Foster et al., 2013).
+However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the \(K\) assemblage (Foster et al., 2013).
+As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
+
+Before fitting models, covariates were checked for collinearity within each seasonal dataset.
+If a pairwise collinearity of \(\left|x\right| >0.7\) was found we removed one of the variables (Dormann et al., 2013), favouring direct measures over the `climatic' layers we calculated (e.g.~if `sst\_grad' and `clim\_sst\_grad' were correlated, we chose `sst\_grad').
+Thus, `full models' differed between seasons as the retained covariates varied, although most covariates were common for all seasons.
+
+Due to seabird migration and dispersion, seasons could act as `sampling artifacts' (Foster et al., 2017), thus biasing results (Ponti \& Sannolo, 2023).
+In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
+RCP models were used to relate the response variables (i.e.~presence-absence or abundance of each seabird species in each grid cell) to the independent predictor covariates (i.e.~averaged values of environmental variables in each cell).
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data based on a Negative Binomial model.
+RCPs were fitted using the `ecomix' 1.0.0 package (S.
+Woolley et al., 2022), with the following workflow: (i) find the best number of groups (\(K\)) that describes the data, (ii) re-fit the model by fixing \(K\) and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+
+To identify the best number of RCP groups (\(K\)) that described the data, we used full models (all covariates) with 100 random start numbers (see Foster et al., 2013) between \(k = (2, ..., 7)\) groups, and chose the number of RCPs that minimized the Bayesian Information Criterion (BIC).
+Then, we fixed \(K\) at the optimal number and fitted models for every possible linear, additive combination of covariates.
+We chose the best model as the one with the lowest BIC.
+Visual inspection of randomized quantile residual plots modified for mixture models was used to verify model assumptions (Dunstan et al., 2013).
+To obtain estimates of uncertainty for model parameters (confidence intervals), we used the default settings of `ecomix::regional\_mix.bootstrap()', which applies a Bayesian bootstrap 1000 times.
+A custom code was written to generate partial plots for exploring the influence of covariates on RCP membership.
+
+We built a seasonal 11-year (2011--2021) average layer for predicting model results for each environmental covariate.
+These variables were then averaged within 1° latitude \(\times\) 1° longitude grid cells, by season, between 9--47°S and 141--161°E.
+Note that `clim\_eke\_mean' and `clim\_sst\_grad' were processed beforehand as they may characterise persistent oceanographic features (described above) and thus represent the same value as `eke' and `sst\_grad'.
+Data for prediction were standardised using the same mean and standard deviation from the modelling dataset.
+
+Finally, we ran rarefaction curves based on the presence-absence matrix and the assigned RCP group for each sample (grid cell) to assess if the amount of information in the RCP model was representative.
+Rarefaction curves were fit using the `iNEXT' 3.0.0 package (Chao et al., 2014; Hsieh et al., 2016), and results are presented based on alpha diversity (i.e.~species richness).
 
@@
 \hypertarget{summary-of-species-richness-and-numbers}{%
-\subsection{Summary of species richness and
-numbers}\label{summary-of-species-richness-and-numbers}}
-
-A total of 10,261 seabird records was obtained during the 15 voyages,
-comprising 142,646 individuals from 80 species, belonging to 11 families
-and five orders. The five most abundant species made up 83\% of total
-birds counted (43.0\% short-tailed shearwater \emph{Ardenna
-tenuirostris}, 17.2\% fairy prion \emph{Pachyptila turtur}, 14.3\%
-wedge-tailed shearwater \emph{Ardenna pacifica}, 5.2\% sooty tern
-\emph{Onychoprion fuscatus}, 3.2\% shy albatross \emph{Thalassarche
-cauta}). The five most frequently encountered species were \emph{A.
-pacifica} (frequency of occurrence = 35.3\%), \emph{T. cauta} (33.2\%),
-brown booby \emph{Sula leucogaster} (25.4\%), red-footed booby \emph{S.
-sula} (25.4\%) and black-browed albatross \emph{T. melanophris} (25\%)
-(\textbf{Fig. S1}). An increase in species richness towards the south of
-the study region was clear in all but the winter season (\textbf{Fig
-S2}). Noticeably higher counts were made around Tasmania, possibly due
-to the `colony effect' and/or the presence of Southern Ocean species
-(\textbf{Fig. S2}). However, one cell in the Coral Sea had counts of the
-same order of magnitude as off Tasmania in summer.
+\subsection{Summary of species richness and numbers}\label{summary-of-species-richness-and-numbers}}
+
+A total of 10,261 seabird records was obtained during the 15 voyages, comprising 142,646 individuals from 80 species, belonging to 11 families and five orders.
+The five most abundant species made up 83\% of total birds counted (43.0\% short-tailed shearwater \emph{Ardenna tenuirostris}, 17.2\% fairy prion \emph{Pachyptila turtur}, 14.3\% wedge-tailed shearwater \emph{Ardenna pacifica}, 5.2\% sooty tern \emph{Onychoprion fuscatus}, 3.2\% shy albatross \emph{Thalassarche cauta}).
+The five most frequently encountered species were \emph{A.
+pacifica} (frequency of occurrence = 35.3\%), \emph{T.
+cauta} (33.2\%), brown booby \emph{Sula leucogaster} (25.4\%), red-footed booby \emph{S.
+sula} (25.4\%) and black-browed albatross \emph{T.
+melanophris} (25\%) (\textbf{Fig.
+S1}).
+An increase in species richness towards the south of the study region was clear in all but the winter season (\textbf{Fig.
+S2}).
+Noticeably higher counts were made around Tasmania, possibly due to the `colony effect' and/or the presence of Southern Ocean species (\textbf{Fig.
+S2}).
+However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
 
@@
 
-For all models, two RCPs was the optimal number of seabird groups,
-except for the presence-absence-autumn model, which indicated three RCPs
-as the best fit (\textbf{Fig. S3}). The best fitting models generally
-had 1 or 2 covariates selected, although the abundance-autumn model had
-five covariates (\textbf{Table 2}). The best models showed a good fit
-with no apparent deviations from the assumptions (\textbf{Fig. S4}). All
-the best models selected `sst' or `sss', as essential for characterising
-water masses. The probability of belonging to one or another RCP varied
-its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer
-(\textbf{Fig. 2}, \textbf{Fig. S5}), similar to the EAC's temperature
-and salinity thresholds. Furthermore, most of the best models selected
-`bat', suggesting a gradient between `coastal' and `oceanic' taxa based
-on bathymetry (\textbf{Fig. S5}). Of particular note, proxies for
-persistent oceanographic features that may aggregate food resources (the
-`climatic' variables) for seabirds were selected only in the autumn and
-spring abundance models.
-
-Grid point-predictions are the assignment of the RCP with the highest
-probability value for a given cell. When predicting RCP onto space,
-patterns of point-predictions suggest a biogeographic transition in
-seabird species around \textasciitilde34°S, with both presence-absence
-and abundance models pointing to similar results (\textbf{Fig. 3}).
-These patterns are further reinforced when looking at the mean
-probability along with the lower and upper confidence intervals of
-predictions (\textbf{Fig. 4}, \textbf{Fig. S6}), showing that around
-\textasciitilde34°S, the `northern' RCP tended to have higher
-probabilities in coastal regions, whereas the `southern' RCP spread
-northward offshore. The presence-absence-autumn model (three RCPs) had a
-similar broad-scale pattern, although the southern RCP is farther south,
-and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Fig.
-3}, \textbf{Fig. S6}).
-
-The broad pattern of two large-scale assemblages is reflected in their
-species compositions. Although there are overlaps among the seabird
-groups, both presence-absence and abundance models captured typical
-tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp.,
-lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean
-species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche}
-spp., fairy prion \emph{P. turtur}) (\textbf{Fig. 5}, \textbf{Fig. S7}).
-The presence-absence-autumn model, which had three RCPs, also showed the
-characteristic taxa representing the `northern' and `southern' RCPs.
-However, the intermediate `Tasmania/Bass Straight' assemblage showed a
-mixture of both groups with shy albatross and Australasian gannet
-\emph{Morus serrator} prominent with high probabilities of occurrence.
+For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (\textbf{Fig.
+S3}).
+The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (\textbf{Table \ref{tab:table2}}).
+The best models showed a good fit with no apparent deviations from the assumptions (\textbf{Fig.
+S4}).
+All the best models selected `sst' or `sss', as essential for characterising water masses.
+The probability of belonging to one or another RCP varied its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer (\textbf{Fig.
+\ref{fig:fig2-partial-plots}}, \textbf{Fig.
+S5}), similar to the EAC's temperature and salinity thresholds.
+Furthermore, most of the best models selected `bat', suggesting a gradient between `coastal' and `oceanic' taxa based on bathymetry (\textbf{Fig.
+S5}).
+Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the `climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
+
+Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \textasciitilde34°S, with both presence-absence and abundance models pointing to similar results (\textbf{Fig.
+\ref{fig:fig3-point-pred}}).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Fig.
+\ref{fig:fig4-prob-maps}}, \textbf{Fig.
+S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
+The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Fig.
+\ref{fig:fig3-point-pred}}, \textbf{Fig.
+S6}).
+
+The broad pattern of two large-scale assemblages is reflected in their species compositions.
+Although there are overlaps among the seabird groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp., lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche} spp., fairy prion \emph{P.
+turtur}) (\textbf{Fig.
+\ref{fig:fig5-spp-profiles}}, \textbf{Fig.
+S7}).
+The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the `northern' and `southern' RCPs.
+However, the intermediate `Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross and Australasian gannet \emph{Morus serrator} prominent with high probabilities of occurrence.
 
@@
 
-This study showed that seabird assemblages off eastern Australia were
-correlated with oceanographic conditions, and that the East Australian
-Current appeared to be key in structuring seabird distributions at meso-
-to macro scales (\emph{sensu} Hunt \& Schneider, 1987). Our results are
-based on a multivariate statistical model (N. A. Hill et al., 2020),
-which brings advantages over classical methods for describing species
-assemblages. RCP models allow estimating uncertainty around the point
-probability of a sample (grid cell) belonging to an assemblage, thus
-representing the fuzziness between bioregion boundaries. These
-transition zones between bioregions are essential for monitoring
-biological changes due to environmental changes (Morrone, 2023). As
-such, the boundary between the two consistently identified bioregions in
-this study may be a crucial region to monitor in future, given
-noticeable anomalies in the EAC at this same region (Phillips et al.,
-2022). Our results draw an important baseline for seabird assemblages in
-a world hotspot for seabirds.
+This study showed that seabird assemblages off eastern Australia were correlated with oceanographic conditions, and that the East Australian Current appeared to be key in structuring seabird distributions at meso- to macro scales (\emph{sensu} Hunt \& Schneider, 1987).
+Our results are based on a multivariate statistical model (N.
+A.
+Hill et al., 2020), which brings advantages over classical methods for describing species assemblages.
+RCP models allow estimating uncertainty around the point probability of a sample (grid cell) belonging to an assemblage, thus representing the fuzziness between bioregion boundaries.
+These transition zones between bioregions are essential for monitoring biological changes due to environmental changes (Morrone, 2023).
+As such, the boundary between the two consistently identified bioregions in this study may be a crucial region to monitor in future, given noticeable anomalies in the EAC at this same region (Phillips et al., 2022).
+Our results draw an important baseline for seabird assemblages in a world hotspot for seabirds.
 
 \hypertarget{seabird-bioregions-off-eastern-australia}{%
-\subsection{Seabird bioregions off eastern
-Australia}\label{seabird-bioregions-off-eastern-australia}}
-
-The relationships between seabird assemblages and water masses has been
-known for a long time (Elphick \& Hunt, 1993; Pocklington, 1979; Ribic
-et al., 1997). Thus, it is not surprising that two seabird assemblages
-were identified off eastern Australia. However, the spatial consistency
-of these patterns throughout seasons is surprising and remarkable.
-Resident species seem to characterise these assemblages. Shy albatross
-(Hedd et al., 2001; Mason et al., 2018) and fairy prions (Fromant et
-al., 2022) in the southern assemblage, and boobies (Bunce, 2015; Miller
-et al., 2018) and wedge-tailed shearwaters (McDuie et al., 2015;
-Weimerskirch et al., 2020) in the northern assemblage, were always
-profiled with higher probabilities of occurrence and predicted
-abundances compared to other species in the same groups. Given that
-changes in assemblages over time are typically driven by a few species
-(Gotelli et al., 2022), and that those species are likely residents
-(Souza \& Santos, 2023), the relationships between these resident
-seabirds and the EAC system over the broad latitudinal range detailed
-here should be closely monitored. Elsewhere, coastal seabirds such as
-little penguin are signalling the adverse effects that climate change
-might have in changing seabird behaviours and decreasing foraging
-success (Carroll et al., 2016; Woehler \& Hobday, 2023), which
-ultimately influence their breeding success and population dynamics (B.
-Johnson \& Colombelli-Négrel, 2021). Thus, the species profiles and the
-spatial predictions of assemblages from RCP models could be used as
-tools for monitoring future seabird responses to the fast-changing EAC.
-
-The seasonal models also captured the influence of migratory species,
-particularly in the southern assemblage. The summer and autumn models
-reflected the high abundance of short-tailed shearwater before they
-depart on their trans-equatorial migration (Carey et al., 2014); the
-presence of the Australasian gannet was picked up by the autumn and
-winter models (Ismar et al., 2011); and the white-headed petrel
-\emph{Pterodroma lessonii} movements during chick-rearing and pre-laying
-exodus (Taylor et al., 2021) were detected in autumn and spring models,
-respectively. Given the geographic location of the southern assemblage,
-however, we believe sub-Antarctic taxa were not well represented in our
-models due to their low number of occurrences and relative paucity of
-survey effort to date (\textbf{Fig. S1}; hence they were removed before
-modelling). Additional surveys may reveal the seasonal northward pulse
-of these taxa into the southern assemblage (e.g.~albatrosses
-\emph{Diomedea} spp. and \emph{Phoebetria} spp.), given that more effort
-translates as higher likelihood of recording these species. In turn,
-including these taxa in the models could result in a more stable `three
-assemblage pattern' (as revealed by the autumn presence-absence model)
-in all but the summer season, when the majority of Antarctic species are
-breeding (S. Marchant \& Higgins, 1990).
-
-Our analyses clearly revealed a `northern' and a `southern' assemblage
-of seabirds off eastern Australia within the EAC. There is a noticeable
-link with the EAC dynamics: the boundaries of RCPs occurred at
-\textasciitilde34°S, near the average separation latitude
-(\textasciitilde32°S) of the EAC and its eastward extension into the
-Tasman Sea (Oke, Roughan, et al., 2019). However, the
-`northern/southern' assemblage pattern identified herein does not agree
-with previously-proposed large-scale regionalizations, such as the
-pelagic Longhurst Provinces (Longhurst, 2007), the Marine Ecoregions of
-the World {[}MEOW; Spalding et al. (2007){]}, or the Large Marine
-Ecosystems {[}LME; Sherman (1994){]}. The MEOW schema was based on
-coastal and shelf biota and was meant to represent habitats to the 200 m
-isobath, which clearly does not encompass the entirety of seabird
-habitats. The LME schema is characterised by physiographical,
-hydrological and productivity data but has a strong focus on practical
-management of areas and thus there is a political component embedded
-within it, which does not align with seabird distributions. Longhurst
-(2007) designated only one biogeochemical province for eastern Australia
-based on productivity patterns, which may work well at larger (macro)
-scales, but appears to be oversimplified for the EAC.
-
-The results do agree with the marine biogeographic units proposed by
-Costello et al. (2017), who identified a boundary at \textasciitilde34°S
-between their `Coral Sea' and `South Australia' realms. Their approach
-used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats,
-thus integrating more complexity of the biogeographic boundaries in the
-marine environment. Other biological data also suggest
-`northern/southern' assemblages off eastern Australia, including tuna
-and billfish {[}also highly mobile top predators; Reygondeau et al.
-(2012){]}, mesopelagic fauna {[}which includes seabird prey; Sutton et
-al. (2017){]} and deep-sea invertebrates and fish (O'Hara, Williams,
-Althaus, et al., 2020; O'Hara, Williams, Woolley, et al., 2020). All
-these studies focused mainly on describing biological assemblages, as
-opposed to classifying environments. It is worth noting that the
-Longhurst Provinces approach was updated to include SST, SSS and
-bathymetry (the same variables retained by our best RCP models) and a
-temporal component to the classification (Reygondeau et al., 2013). The
-updated provinces are more similar to the seabird assemblages described
-here, although still not a perfect match, which could be partially
-explained because the seabird assemblage at EAC region includes several
-long-distance migratory species. Australia's bioregionalization has a
-long story of disagreements (Ebach, 2012), however, the inherent link
-that RCPs develop between environment and biological data can be a
-strong tool to support marine bioregionalization.
+\subsection{Seabird bioregions off eastern Australia}\label{seabird-bioregions-off-eastern-australia}}
+
+The relationships between seabird assemblages and water masses has been known for a long time (Elphick \& Hunt, 1993; Pocklington, 1979; Ribic et al., 1997).
+Thus, it is not surprising that two seabird assemblages were identified off eastern Australia.
+However, the spatial consistency of these patterns throughout seasons is surprising and remarkable.
+Resident species seem to characterise these assemblages.
+Shy albatross (Hedd et al., 2001; Mason et al., 2018) and fairy prions (Fromant et al., 2022) in the southern assemblage, and boobies (Bunce, 2015; Miller et al., 2018) and wedge-tailed shearwaters (McDuie et al., 2015; Weimerskirch et al., 2020) in the northern assemblage, were always profiled with higher probabilities of occurrence and predicted abundances compared to other species in the same groups.
+Given that changes in assemblages over time are typically driven by a few species (Gotelli et al., 2022), and that those species are likely residents (Souza \& Santos, 2023), the relationships between these resident seabirds and the EAC system over the broad latitudinal range detailed here should be closely monitored.
+Elsewhere, coastal seabirds such as little penguin are signalling the adverse effects that climate change might have in changing seabird behaviours and decreasing foraging success (Carroll et al., 2016; Woehler \& Hobday, 2023), which ultimately influence their breeding success and population dynamics (B.
+Johnson \& Colombelli-Négrel, 2021).
+Thus, the species profiles and the spatial predictions of assemblages from RCP models could be used as tools for monitoring future seabird responses to the fast-changing EAC.
+
+The seasonal models also captured the influence of migratory species, particularly in the southern assemblage.
+The summer and autumn models reflected the high abundance of short-tailed shearwater before they depart on their trans-equatorial migration (Carey et al., 2014); the presence of the Australasian gannet was picked up by the autumn and winter models (Ismar et al., 2011); and the white-headed petrel \emph{Pterodroma lessonii} movements during chick-rearing and pre-laying exodus (Taylor et al., 2021) were detected in autumn and spring models, respectively.
+Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (\textbf{Fig.
+S1}; hence they were removed before modelling).
+Additional surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.~albatrosses \emph{Diomedea} spp.
+and \emph{Phoebetria} spp.), given that more effort translates as higher likelihood of recording these species.
+In turn, including these taxa in the models could result in a more stable `three assemblage pattern' (as revealed by the autumn presence-absence model) in all but the summer season, when the majority of Antarctic species are breeding (S.
+Marchant \& Higgins, 1990).
+
+Our analyses clearly revealed a `northern' and a `southern' assemblage of seabirds off eastern Australia within the EAC.
+There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \textasciitilde34°S, near the average separation latitude (\textasciitilde32°S) of the EAC and its eastward extension into the Tasman Sea (Oke, Roughan, et al., 2019).
+However, the `northern/southern' assemblage pattern identified herein does not agree with previously-proposed large-scale regionalizations, such as the pelagic Longhurst Provinces (Longhurst, 2007), the Marine Ecoregions of the World {[}MEOW; Spalding et al.
+(2007){]}, or the Large Marine Ecosystems {[}LME; Sherman (1994){]}.
+The MEOW schema was based on coastal and shelf biota and was meant to represent habitats to the 200 m isobath, which clearly does not encompass the entirety of seabird habitats.
+The LME schema is characterised by physiographical, hydrological and productivity data but has a strong focus on practical management of areas and thus there is a political component embedded within it, which does not align with seabird distributions.
+Longhurst (2007) designated only one biogeochemical province for eastern Australia based on productivity patterns, which may work well at larger (macro) scales, but appears to be oversimplified for the EAC.
+
+The results do agree with the marine biogeographic units proposed by Costello et al.
+(2017), who identified a boundary at \textasciitilde34°S between their `Coral Sea' and `South Australia' realms.
+Their approach used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats, thus integrating more complexity of the biogeographic boundaries in the marine environment.
+Other biological data also suggest `northern/southern' assemblages off eastern Australia, including tuna and billfish {[}also highly mobile top predators; Reygondeau et al.
+(2012){]}, mesopelagic fauna {[}which includes seabird prey; Sutton et al.
+(2017){]} and deep-sea invertebrates and fish (O'Hara, Williams, Althaus, et al., 2020; O'Hara, Williams, Woolley, et al., 2020).
+All these studies focused mainly on describing biological assemblages, as opposed to classifying environments.
+It is worth noting that the Longhurst Provinces approach was updated to include SST, SSS and bathymetry (the same variables retained by our best RCP models) and a temporal component to the classification (Reygondeau et al., 2013).
+The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage at EAC region includes several long-distance migratory species.
+Australia's bioregionalization has a long story of disagreements (Ebach, 2012), however, the inherent link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
 
 \hypertarget{fitting-rcp-models-to-high-mobile-animals}{%
-\subsection{Fitting RCP models to high mobile
-animals}\label{fitting-rcp-models-to-high-mobile-animals}}
-
-To the best of our knowledge, this is the first study applying RCPs to
-highly mobile animals such as seabirds. RCP models have been previously
-applied to plants (Lyons et al., 2017), and relatively sedentary
-animals, such as macrobenthic organisms (Pantiukhin et al., 2021) and
-demersal fish (N. A. Hill et al., 2017). Oceanic seabirds can have daily
-movements of hundreds of kilometres (e.g. McDuie et al., 2015), during
-which they traverse a variety of habitats and environmental gradients.
-As such, capturing the distribution of highly mobile species, and their
-relationships with environmental parameters is not a trivial task, and
-much harder than for sessile organisms.
-
-While remote tracking studies are advancing our understanding on how
-individual seabirds use their environment at fine scales (Tremblay et
-al., 2009), they fail to perceive how species assemblages interact with
-the environment and each other. Therefore, using at-sea counts of
-seabirds is a well-established method to capture these relationships
-(Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
-Further, the spatial scales on which assemblages relate to their
-environment is variable and species-specific (Hunt \& Schneider, 1987;
-Ribic et al., 1997; Scales et al., 2017). Understanding the biology of
-the studied group is thus essential to set the appropriate scales of
-analyses. Our choice of 1° latitude \(\times\) 1° longitude was mainly
-determined by data availability reflecting survey effort, and the
-trade-off between the number of species per grid/season (see also
-\ref{sec:limitations}). Nonetheless, the chosen spatial scale is similar
-to other seabird distribution analyses conducted elsewhere (Harris et
-al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022). Therefore, for
-describing seabird assemblages at a macro-scale, the chosen resolution
-seems to had been a good fit, accords with studies elsewhere, and the
-resultant models described patterns of seabird species' composition
-consistent with our current understanding of species' ecologies and life
-history strategies.
-
-RCP models based on presence-absence and abundance data are able to
-complement each other. Previous studies using RCPs have opted to use
-only one type of response data. Lyons et al. (2017), N. A. Hill et al.
-(2017), and Pantiukhin et al. (2021) all used presence-absence data,
-whereas Lee et al. (2019) and Receveur et al. (2020) used abundance
-(count) data. The added information contained in abundance data make
-them inherently more powerful than binary presence-absence data.
-However, the use of different data inputs in multivariate, biodiversity
-assessments and assemblage descriptions has shown contrasting patterns
-at different spatial scales. At larger spatial and temporal scales,
-presence-absence data describe similar patterns to abundance data
-(Anderson et al., 2005), as we found. In contrast, at smaller scales,
-abundance data may highlight local conditions and therefore reveal
-hidden patterns at coarse to fine scales (Anderson et al., 2005; R.
-Marchant, 1990; Melo, 2005; Waters, 2008). If combined with ecological
-knowledge of a particular group, the information contained in both
-presence-absence and abundance data could be leveraged to better
-interpret the resulting patterns. For example, seabirds will likely
-occur in specific water masses or currents (Hunt \& Schneider, 1987;
-Ribic et al., 1997), whereas their abundance is related to coarse- to
-fine-scale processes that enhance prey availability (Hunt \& Schneider,
-1987; Weimerskirch, 2007). Future studies assessing bioregions of highly
-mobile animals could describe the compositions of assemblages using
-smaller resolution grid cells and presence-absence models. Then,
-abundance models using higher resolution grid cells could improve our
-understanding of how these animals relate to their environment and
-possibly reveal multi-species aggregation areas (therefore `new
-bioregions') likely nested within the bioregions described at large
-scales. Such a hierarchical approach would require substantial
-concurrent observational and environmental data sets, which we suspect
-not exist at present.
+\subsection{Fitting RCP models to high mobile animals}\label{fitting-rcp-models-to-high-mobile-animals}}
+
+To the best of our knowledge, this is the first study applying RCPs to highly mobile animals such as seabirds.
+RCP models have been previously applied to plants (Lyons et al., 2017), and relatively sedentary animals, such as macrobenthic organisms (Pantiukhin et al., 2021) and demersal fish (N.
+A.
+Hill et al., 2017).
+Oceanic seabirds can have daily movements of hundreds of kilometres (e.g.
+McDuie et al., 2015), during which they traverse a variety of habitats and environmental gradients.
+As such, capturing the distribution of highly mobile species, and their relationships with environmental parameters is not a trivial task, and much harder than for sessile organisms.
+
+While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales (Tremblay et al., 2009), they fail to perceive how species assemblages interact with the environment and each other.
+Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships (Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
+Further, the spatial scales on which assemblages relate to their environment is variable and species-specific (Hunt \& Schneider, 1987; Ribic et al., 1997; Scales et al., 2017).
+Understanding the biology of the studied group is thus essential to set the appropriate scales of analyses.
+Our choice of 1° latitude \(\times\) 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \ref{sec:limitations}).
+Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere (Harris et al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022).
+Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to had been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
+
+RCP models based on presence-absence and abundance data are able to complement each other.
+Previous studies using RCPs have opted to use only one type of response data.
+Lyons et al.
+(2017), N.
+A.
+Hill et al.
+(2017), and Pantiukhin et al.
+(2021) all used presence-absence data, whereas Lee et al.
+(2019) and Receveur et al.
+(2020) used abundance (count) data.
+The added information contained in abundance data make them inherently more powerful than binary presence-absence data.
+However, the use of different data inputs in multivariate, biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
+At larger spatial and temporal scales, presence-absence data describe similar patterns to abundance data (Anderson et al., 2005), as we found.
+In contrast, at smaller scales, abundance data may highlight local conditions and therefore reveal hidden patterns at coarse to fine scales (Anderson et al., 2005; R.
+Marchant, 1990; Melo, 2005; Waters, 2008).
+If combined with ecological knowledge of a particular group, the information contained in both presence-absence and abundance data could be leveraged to better interpret the resulting patterns.
+For example, seabirds will likely occur in specific water masses or currents (Hunt \& Schneider, 1987; Ribic et al., 1997), whereas their abundance is related to coarse- to fine-scale processes that enhance prey availability (Hunt \& Schneider, 1987; Weimerskirch, 2007).
+Future studies assessing bioregions of highly mobile animals could describe the compositions of assemblages using smaller resolution grid cells and presence-absence models.
+Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore `new bioregions') likely nested within the bioregions described at large scales.
+Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect not exist at present.
 
@@
 
-Although the results suggest consistent patterns, there are a number of
-caveats to our analyses. Seabird data were gathered during
-multi-disciplinary cruises, whose sampling designs were not specifically
-planned to account for spatial ecology of seabirds. As a result, there
-are spatial discontinuities in the data, which might have limited the
-ability of our models to detect the relationships between the seabird
-assemblages and the full spectrum of environmental predictors. Even
-though Australia is a hotspot for seabirds (Karpouzi et al., 2007;
-Ramírez et al., 2017), there is a paucity of information on seabird
-distributions in the region (Bernard et al., 2021; Mott \& Clarke,
-2018). Dedicated seabird surveys would help to advance seabird research
-in this region, and improve our understanding of the patterns described
-here. By increasing the amount of data, future studies could set the
-grid cells at higher resolution, and potentially reveal assemblages at
-finer spatial and temporal scales. In addition, ideally, we would have
-had more information within each grid cell by season (our sampling
-units). Nonetheless, rarefaction curves show that the sample coverage
-(see Chao \& Jost, 2012) was high at the RCP level, although there is
-room for detecting more species in most of them (\textbf{Fig S8}).
+Although the results suggest consistent patterns, there are a number of caveats to our analyses.
+Seabird data were gathered during multi-disciplinary cruises, whose sampling designs were not specifically planned to account for spatial ecology of seabirds.
+As a result, there are spatial discontinuities in the data, which might have limited the ability of our models to detect the relationships between the seabird assemblages and the full spectrum of environmental predictors.
+Even though Australia is a hotspot for seabirds (Karpouzi et al., 2007; Ramírez et al., 2017), there is a paucity of information on seabird distributions in the region (Bernard et al., 2021; Mott \& Clarke, 2018).
+Dedicated seabird surveys would help to advance seabird research in this region, and improve our understanding of the patterns described here.
+By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at finer spatial and temporal scales.
+In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
+Nonetheless, rarefaction curves show that the sample coverage (see Chao \& Jost, 2012) was high at the RCP level, although there is room for detecting more species in most of them (\textbf{Fig S8}).
 
@@
 
-We applied a robust, repeatable method to assess seabird assemblages off
-eastern Australia within the EAC for the first time. The models point
-toward a consistent pattern of a `northern' and a `southern' seabird
-assemblage, where the spatial division is attributable to the eastward
-separation of the EAC from the coast, and that SST seems to be the key
-variable discriminating these groups. The EAC has penetrated poleward
-over the past half-century (K. L. Hill et al., 2008; Li et al., 2022), a
-trend which is anticipated to continue into the future (Yang et al.,
-2016), and extreme weather events such as marine heatwaves to happen
-more often (Oliver et al., 2018). As such, seabirds will have to respond
-to these changes (Woehler \& Hobday, 2023). Southern Ocean albatrosses
-off southeast Australia, for instance, have shown `early stages of a
-southward range shift' due to environmental variability (Sojitra et al.,
-2022). Resident species such as the little penguins and wedge-tailed
-shearwater change their behaviour and have decreased foraging success
-when SST is higher (Carroll et al., 2016; Peck et al., 2004). Given that
-eastern Australia is a biodiversity and numerical hotspot for seabirds,
-it is imperative to establish a monitoring program allowing the
-assemblages and species' profiles presented here to be used as the
-baseline for understanding possible effects of a warming ocean on
-seabird distribution. In particular, the transition zone between
-assemblages (Morrone, 2023), associated with the eastern extension of
-the EAC, may be fundamental to effectively monitoring strengthening of
-the EAC and the subsequent impact on marine biota.
+We applied a robust, repeatable method to assess seabird assemblages off eastern Australia within the EAC for the first time.
+The models point toward a consistent pattern of a `northern' and a `southern' seabird assemblage, where the spatial division is attributable to the eastward separation of the EAC from the coast, and that SST seems to be the key variable discriminating these groups.
+The EAC has penetrated poleward over the past half-century (K.
+L.
+Hill et al., 2008; Li et al., 2022), a trend which is anticipated to continue into the future (Yang et al., 2016), and extreme weather events such as marine heatwaves to happen more often (Oliver et al., 2018).
+As such, seabirds will have to respond to these changes (Woehler \& Hobday, 2023).
+Southern Ocean albatrosses off southeast Australia, for instance, have shown `early stages of a southward range shift' due to environmental variability (Sojitra et al., 2022).
+Resident species such as the little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher (Carroll et al., 2016; Peck et al., 2004).
+Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species' profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
+In particular, the transition zone between assemblages (Morrone, 2023), associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
 
@@
 
-We thank all observers who assisted in data collection, the crews of RV
-\emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird
-Group (BirdLife Australia) for logistical support. The authors would
-like to thank Dr.~Skipton N.C. Woolley, \texttt{\{ecomix\}} developer,
-who kindly helped NWD build preliminary models and gave advice in
-several stages of model construction. NWD is supported by a Doctoral
-scholarship from the University of Otago.
+We thank all observers who assisted in data collection, the crews of RV \emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
+The authors would like to thank Dr.~Skipton N.C.
+Woolley, \texttt{\{ecomix\}} developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
+NWD is supported by a Doctoral scholarship from the University of Otago.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ainley2012}{}}%
-Ainley, D., Ribic, C., \& Woehler, E. (2012). Adding the ocean to the
-study of seabirds: a brief history of at-sea seabird research.
+Ainley, D., Ribic, C., \& Woehler, E.
+(2012).
+Adding the ocean to the study of seabirds: a brief history of at-sea seabird research.
 \emph{Marine Ecology Progress Series}, \emph{451}, 231--243.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anderson2005}{}}%
-Anderson, M. J., Connell, S. D., Gillanders, B. M., Diebel, C. E., Blom,
-W. M., Saunders, J. E., \& Landers, T. J. (2005). Relationships between
-taxonomic resolution and spatial scales of multivariate variation.
+Anderson, M.
+J., Connell, S.
+D., Gillanders, B.
+M., Diebel, C.
+E., Blom, W.
+M., Saunders, J.
+E., \& Landers, T.
+J.
+(2005).
+Relationships between taxonomic resolution and spatial scales of multivariate variation.
 \emph{Journal of Animal Ecology}, \emph{74}(4), 636--646.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-antuxe3o2020}{}}%
-Antão, L. H., Bates, A. E., Blowes, S. A., Waldock, C., Supp, S. R.,
-Magurran, A. E., Dornelas, M., \& Schipper, A. M. (2020).
-Temperature-related biodiversity change across temperate marine and
-terrestrial systems. \emph{Nature Ecology \& Evolution}, \emph{4}(7),
-927--933. \url{https://doi.org/10.1038/s41559-020-1185-7}
+Antão, L.
+H., Bates, A.
+E., Blowes, S.
+A., Waldock, C., Supp, S.
+R., Magurran, A.
+E., Dornelas, M., \& Schipper, A.
+M.
+(2020).
+Temperature-related biodiversity change across temperate marine and terrestrial systems.
+\emph{Nature Ecology \& Evolution}, \emph{4}(7), 927--933.
+\url{https://doi.org/10.1038/s41559-020-1185-7}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ballance1997}{}}%
-Ballance, L. T., Pitman, R. L., \& Reilly, S. B. (1997). Seabird
-community structure along a productivity gradient: importance of
-competition and energetic constraint. \emph{Ecology}, \emph{78}(5),
-1502--1518.
+Ballance, L.
+T., Pitman, R.
+L., \& Reilly, S.
+B.
+(1997).
+Seabird community structure along a productivity gradient: importance of competition and energetic constraint.
+\emph{Ecology}, \emph{78}(5), 1502--1518.
 \url{https://doi.org/10.1890/0012-9658(1997)078\%5B1502:SCSAAP\%5D2.0.CO;2}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-barbet-massin2014}{}}%
-Barbet-Massin, M., \& Jetz, W. (2014). A 40-year, continent-wide,
-multispecies assessment of relevant climate predictors for species
-distribution modelling. \emph{Diversity and Distributions},
-\emph{20}(11), 1285--1295. \url{https://doi.org/10.1111/ddi.12229}
+Barbet-Massin, M., \& Jetz, W.
+(2014).
+A 40-year, continent-wide, multispecies assessment of relevant climate predictors for species distribution modelling.
+\emph{Diversity and Distributions}, \emph{20}(11), 1285--1295.
+\url{https://doi.org/10.1111/ddi.12229}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bernard2021}{}}%
-Bernard, A., Rodrigues, A. S. L., Cazalis, V., \& Grémillet, D. (2021).
-Toward a global strategy for seabird tracking. \emph{Conservation
-Letters}, \emph{14}(3), e12804. \url{https://doi.org/10.1111/conl.12804}
+Bernard, A., Rodrigues, A.
+S.
+L., Cazalis, V., \& Grémillet, D.
+(2021).
+Toward a global strategy for seabird tracking.
+\emph{Conservation Letters}, \emph{14}(3), e12804.
+\url{https://doi.org/10.1111/conl.12804}
 
 \leavevmode\vadjust pre{\hypertarget{ref-blaber1994distribution}{}}%
-Blaber, S., \& Milton, D. (1994). Distribution of seabirds at sea in the
-{G}ulf of {C}arpentaria, {A}ustralia. \emph{Marine and Freshwater
-Research}, \emph{45}(3), 445--454.
+Blaber, S., \& Milton, D. (1994). Distribution of seabirds at sea in the {G}ulf of {C}arpentaria, {A}ustralia. \emph{Marine and Freshwater Research}, \emph{45}(3), 445--454.
 
 \leavevmode\vadjust pre{\hypertarget{ref-brieva2015}{}}%
-Brieva, D., Ribbe, J., \& Lemckert, C. (2015). Is the East Australian
-Current causing a marine ecological hot-spot and an important fisheries
-near Fraser Island, Australia? \emph{Estuarine, Coastal and Shelf
-Science}, \emph{153}, 121--134.
+Brieva, D., Ribbe, J., \& Lemckert, C.
+(2015).
+Is the East Australian Current causing a marine ecological hot-spot and an important fisheries near Fraser Island, Australia?
+\emph{Estuarine, Coastal and Shelf Science}, \emph{153}, 121--134.
 \url{https://doi.org/10.1016/j.ecss.2014.12.012}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bunce2015}{}}%
-Bunce, A. (2015). Foraging behaviour of a declining population of brown
-boobies ({\emph{Sula leucogaster}}) breeding in the Swain Reefs, Great
-Barrier Reef. \emph{Emu - Austral Ornithology}, \emph{115}(4), 368--372.
+Bunce, A.
+(2015).
+Foraging behaviour of a declining population of brown boobies ({\emph{Sula leucogaster}}) breeding in the Swain Reefs, Great Barrier Reef.
+\emph{Emu - Austral Ornithology}, \emph{115}(4), 368--372.
 \url{https://doi.org/10.1071/MU14037}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-burrows2011}{}}%
-Burrows, M. T., Schoeman, D. S., Buckley, L. B., Moore, P., Poloczanska,
-E. S., Brander, K. M., Brown, C., Bruno, J. F., Duarte, C. M., Halpern,
-B. S., Holding, J., Kappel, C. V., Kiessling, W., O'Connor, M. I.,
-Pandolfi, J. M., Parmesan, C., Schwing, F. B., Sydeman, W. J., \&
-Richardson, A. J. (2011). The pace of shifting climate in marine and
-terrestrial ecosystems. \emph{Science}, \emph{334}(6056), 652--655.
+Burrows, M.
+T., Schoeman, D.
+S., Buckley, L.
+B., Moore, P., Poloczanska, E.
+S., Brander, K.
+M., Brown, C., Bruno, J.
+F., Duarte, C.
+M., Halpern, B.
+S., Holding, J., Kappel, C.
+V., Kiessling, W., O'Connor, M.
+I., Pandolfi, J.
+M., Parmesan, C., Schwing, F.
+B., Sydeman, W.
+J., \& Richardson, A.
+J.
+(2011).
+The pace of shifting climate in marine and terrestrial ecosystems.
+\emph{Science}, \emph{334}(6056), 652--655.
 \url{https://doi.org/10.1126/science.1210288}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cai2005}{}}%
-Cai, W., Shi, G., Cowan, T., Bi, D., \& Ribbe, J. (2005). The response
-of the Southern Annular Mode, the East Australian Current, and the
-southern mid-latitude ocean circulation to global warming.
+Cai, W., Shi, G., Cowan, T., Bi, D., \& Ribbe, J.
+(2005).
+The response of the Southern Annular Mode, the East Australian Current, and the southern mid-latitude ocean circulation to global warming.
 \emph{Geophysical Research Letters}, \emph{32}(23), L23706.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carey2014}{}}%
-Carey, M. J., Phillips, R. A., Silk, J. R. D., \& Shaffer, S. A. (2014).
-Trans-equatorial migration of short-tailed shearwaters revealed by
-geolocators. \emph{Emu - Austral Ornithology}, \emph{114}(4), 352--359.
+Carey, M.
+J., Phillips, R.
+A., Silk, J.
+R.
+D., \& Shaffer, S.
+A.
+(2014).
+Trans-equatorial migration of short-tailed shearwaters revealed by geolocators.
+\emph{Emu - Austral Ornithology}, \emph{114}(4), 352--359.
 \url{https://doi.org/10.1071/MU13115}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carneiro2020}{}}%
-Carneiro, A. P. B., Pearmain, E. J., Oppel, S., Clay, T. A., Phillips,
-R. A., Bonnet-Lebrun, A.-S., Wanless, R. M., Abraham, E., Richard, Y.,
-Rice, J., Handley, J., Davies, T. E., Dilley, B. J., Ryan, P. G., Small,
-C., Arata, J., Arnould, J. P. Y., Bell, E., Bugoni, L., \ldots{} Dias,
-M. P. (2020). A framework for mapping the distribution of seabirds by
-integrating tracking, demography and phenology. \emph{Journal of Applied
-Ecology}, \emph{57}(3), 514--525.
+Carneiro, A.
+P.
+B., Pearmain, E.
+J., Oppel, S., Clay, T.
+A., Phillips, R.
+A., Bonnet-Lebrun, A.-S., Wanless, R.
+M., Abraham, E., Richard, Y., Rice, J., Handley, J., Davies, T.
+E., Dilley, B.
+J., Ryan, P.
+G., Small, C., Arata, J., Arnould, J.
+P.
+Y., Bell, E., Bugoni, L., \ldots{} Dias, M.
+P.
+(2020).
+A framework for mapping the distribution of seabirds by integrating tracking, demography and phenology.
+\emph{Journal of Applied Ecology}, \emph{57}(3), 514--525.
 \url{https://doi.org/10.1111/1365-2664.13568}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carroll2016}{}}%
-Carroll, G., Everett, J. D., Harcourt, R., Slip, D., \& Jonsen, I.
-(2016). High sea surface temperatures driven by a strengthening current
-reduce foraging success by penguins. \emph{Scientific Reports},
-\emph{6}(1), 22236. \url{https://doi.org/10.1038/srep22236}
+Carroll, G., Everett, J.
+D., Harcourt, R., Slip, D., \& Jonsen, I.
+(2016).
+High sea surface temperatures driven by a strengthening current reduce foraging success by penguins.
+\emph{Scientific Reports}, \emph{6}(1), 22236.
+\url{https://doi.org/10.1038/srep22236}
 
 \leavevmode\vadjust pre{\hypertarget{ref-chao2014}{}}%
-Chao, A., Chiu, C.-H., \& Jost, L. (2014). Unifying species diversity,
-phylogenetic diversity, functional diversity, and related similarity and
-differentiation measures through Hill numbers. \emph{Annual Review of
-Ecology, Evolution, and Systematics}, \emph{45}(1), 297--324.
+Chao, A., Chiu, C.-H., \& Jost, L.
+(2014).
+Unifying species diversity, phylogenetic diversity, functional diversity, and related similarity and differentiation measures through Hill numbers.
+\emph{Annual Review of Ecology, Evolution, and Systematics}, \emph{45}(1), 297--324.
 \url{https://doi.org/10.1146/annurev-ecolsys-120213-091540}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-chao2012}{}}%
-Chao, A., \& Jost, L. (2012). Coverage-based rarefaction and
-extrapolation: standardizing samples by completeness rather than size.
+Chao, A., \& Jost, L.
+(2012).
+Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size.
 \emph{Ecology}, \emph{93}(12), 2533--2547.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-commins2014}{}}%
-Commins, M. L., Ansorge, I., \& Ryan, P. G. (2014). Multi-scale factors
-influencing seabird assemblages in the African sector of the Southern
-Ocean. \emph{Antarctic Science}, \emph{26}(1), 38--48.
+Commins, M.
+L., Ansorge, I., \& Ryan, P.
+G.
+(2014).
+Multi-scale factors influencing seabird assemblages in the African sector of the Southern Ocean.
+\emph{Antarctic Science}, \emph{26}(1), 38--48.
 \url{https://doi.org/10.1017/S0954102013000138}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-costello2017}{}}%
-Costello, M. J., Tsai, P., Wong, P. S., Cheung, A. K. L., Basher, Z., \&
-Chaudhary, C. (2017). Marine biogeographic realms and species
-endemicity. \emph{Nature Communications}, \emph{8}(1), 1057.
+Costello, M.
+J., Tsai, P., Wong, P.
+S., Cheung, A.
+K.
+L., Basher, Z., \& Chaudhary, C.
+(2017).
+Marine biogeographic realms and species endemicity.
+\emph{Nature Communications}, \emph{8}(1), 1057.
 \url{https://doi.org/10.1038/s41467-017-01121-2}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-davoren2013}{}}%
-Davoren, G. K. (2013). Distribution of marine predator hotspots
-explained by persistent areas of prey. \emph{Marine Biology},
-\emph{160}(12), 3043--3058.
+Davoren, G.
+K.
+(2013).
+Distribution of marine predator hotspots explained by persistent areas of prey.
+\emph{Marine Biology}, \emph{160}(12), 3043--3058.
 \url{https://doi.org/10.1007/s00227-013-2294-5}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dormann2013}{}}%
-Dormann, C. F., Elith, J., Bacher, S., Buchmann, C., Carl, G., Carré,
-G., Marquéz, J. R. G., Gruber, B., Lafourcade, B., Leitão, P. J.,
-Münkemüller, T., McClean, C., Osborne, P. E., Reineking, B., Schröder,
-B., Skidmore, A. K., Zurell, D., \& Lautenbach, S. (2013). Collinearity:
-a review of methods to deal with it and a simulation study evaluating
-their performance. \emph{Ecography}, \emph{36}(1), 27--46.
+Dormann, C.
+F., Elith, J., Bacher, S., Buchmann, C., Carl, G., Carré, G., Marquéz, J.
+R.
+G., Gruber, B., Lafourcade, B., Leitão, P.
+J., Münkemüller, T., McClean, C., Osborne, P.
+E., Reineking, B., Schröder, B., Skidmore, A.
+K., Zurell, D., \& Lautenbach, S.
+(2013).
+Collinearity: a review of methods to deal with it and a simulation study evaluating their performance.
+\emph{Ecography}, \emph{36}(1), 27--46.
 \url{https://doi.org/10.1111/j.1600-0587.2012.07348.x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dornelas2014}{}}%
-Dornelas, M., Gotelli, N. J., McGill, B., Shimadzu, H., Moyes, F.,
-Sievers, C., \& Magurran, A. E. (2014). Assemblage time series reveal
-biodiversity change but not systematic loss. \emph{Science},
-\emph{344}(6181), 296--299.
+Dornelas, M., Gotelli, N.
+J., McGill, B., Shimadzu, H., Moyes, F., Sievers, C., \& Magurran, A.
+E.
+(2014).
+Assemblage time series reveal biodiversity change but not systematic loss.
+\emph{Science}, \emph{344}(6181), 296--299.
 \url{https://doi.org/10.1126/science.1248484}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dunlop1988}{}}%
-Dunlop, J., Wooller, R., \& Cheshire, N. (1988). Distribution and
-abundance of marine birds in the Eastern Indian Ocean. \emph{Marine and
-Freshwater Research}, \emph{39}(5), 661.
+Dunlop, J., Wooller, R., \& Cheshire, N.
+(1988).
+Distribution and abundance of marine birds in the Eastern Indian Ocean.
+\emph{Marine and Freshwater Research}, \emph{39}(5), 661.
 \url{https://doi.org/10.1071/MF9880661}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dunstan2013}{}}%
-Dunstan, P. K., Foster, S. D., Hui, F. K. C., \& Warton, D. I. (2013).
-Finite mixture of regression modeling for high-dimensional count and
-biomass data in ecology. \emph{Journal of Agricultural, Biological, and
-Environmental Statistics}, \emph{18}(3), 357--375.
+Dunstan, P.
+K., Foster, S.
+D., Hui, F.
+K.
+C., \& Warton, D.
+I.
+(2013).
+Finite mixture of regression modeling for high-dimensional count and biomass data in ecology.
+\emph{Journal of Agricultural, Biological, and Environmental Statistics}, \emph{18}(3), 357--375.
 \url{https://doi.org/10.1007/s13253-013-0146-x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ebach2012history}{}}%
-Ebach, M. C. (2012). A history of bioregionalisation in {A}ustralia.
-\emph{Zootaxa}, \emph{3392}, 1--34.
-\url{https://doi.org/10.11646/ZOOTAXA.3392.1.1}
+Ebach, M. C. (2012). A history of bioregionalisation in {A}ustralia. \emph{Zootaxa}, \emph{3392}, 1--34. \url{https://doi.org/10.11646/ZOOTAXA.3392.1.1}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ebach2008}{}}%
-Ebach, M. C., Morrone, J. J., Parenti, L. R., \& Viloria, Á. L. (2008).
-International Code of Area Nomenclature. \emph{Journal of Biogeography},
-\emph{35}(7), 1153--1157.
+Ebach, M.
+C., Morrone, J.
+J., Parenti, L.
+R., \& Viloria, Á.
+L.
+(2008).
+International Code of Area Nomenclature.
+\emph{Journal of Biogeography}, \emph{35}(7), 1153--1157.
 \url{https://doi.org/10.1111/j.1365-2699.2008.01920.x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-elphick1993}{}}%
-Elphick, C. S., \& Hunt, G. L. (1993). Variations in the distributions
-of marine birds with water mass in the northern Bering Sea.
+Elphick, C.
+S., \& Hunt, G.
+L.
+(1993).
+Variations in the distributions of marine birds with water mass in the northern Bering Sea.
 \emph{Condor}, \emph{95}(1), 33--44.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-evans2021a}{}}%
-Evans, R., Lea, M.-A., \& Hindell, M. A. (2021). Predicting the
-distribution of foraging seabirds during a period of heightened
-environmental variability. \emph{Ecological Applications}, \emph{31}(5),
-e02343. \url{https://doi.org/10.1002/eap.2343}
+Evans, R., Lea, M.-A., \& Hindell, M.
+A.
+(2021).
+Predicting the distribution of foraging seabirds during a period of heightened environmental variability.
+\emph{Ecological Applications}, \emph{31}(5), e02343.
+\url{https://doi.org/10.1002/eap.2343}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ferrier2002a}{}}%
-Ferrier, S. (2002). Mapping spatial pattern in biodiversity for regional
-conservation planning: where to from here? \emph{Systematic Biology},
-\emph{51}(2), 331--363. \url{https://doi.org/10.1080/10635150252899806}
+Ferrier, S.
+(2002).
+Mapping spatial pattern in biodiversity for regional conservation planning: where to from here?
+\emph{Systematic Biology}, \emph{51}(2), 331--363.
+\url{https://doi.org/10.1080/10635150252899806}
 
 \leavevmode\vadjust pre{\hypertarget{ref-firme2023}{}}%
-Firme, G. F., Hughes, D. J., Laiolo, L., Roughan, M., Suthers, I. M., \&
-Doblin, M. A. (2023). Contrasting phytoplankton composition and primary
-productivity in multiple mesoscale eddies along the East Australian
-coast. \emph{Deep Sea Research Part I: Oceanographic Research Papers},
-\emph{193}, 103952. \url{https://doi.org/10.1016/j.dsr.2022.103952}
+Firme, G.
+F., Hughes, D.
+J., Laiolo, L., Roughan, M., Suthers, I.
+M., \& Doblin, M.
+A.
+(2023).
+Contrasting phytoplankton composition and primary productivity in multiple mesoscale eddies along the East Australian coast.
+\emph{Deep Sea Research Part I: Oceanographic Research Papers}, \emph{193}, 103952.
+\url{https://doi.org/10.1016/j.dsr.2022.103952}
 
 \leavevmode\vadjust pre{\hypertarget{ref-foster2013}{}}%
-Foster, S. D., Givens, G. H., Dornan, G. J., Dunstan, P. K., \& Darnell,
-R. (2013). Modelling biological regions from multi-species and
-environmental data. \emph{Environmetrics}, \emph{24}(7), 489--499.
+Foster, S.
+D., Givens, G.
+H., Dornan, G.
+J., Dunstan, P.
+K., \& Darnell, R.
+(2013).
+Modelling biological regions from multi-species and environmental data.
+\emph{Environmetrics}, \emph{24}(7), 489--499.
 \url{https://doi.org/10.1002/env.2245}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-foster2017}{}}%
-Foster, S. D., Hill, N. A., \& Lyons, M. (2017). Ecological grouping of
-survey sites when sampling artefacts are present. \emph{Journal of the
-Royal Statistical Society: Series C (Applied Statistics)}, \emph{66}(5),
-1031--1047. \url{https://doi.org/10.1111/rssc.12211}
+Foster, S.
+D., Hill, N.
+A., \& Lyons, M.
+(2017).
+Ecological grouping of survey sites when sampling artefacts are present.
+\emph{Journal of the Royal Statistical Society: Series C (Applied Statistics)}, \emph{66}(5), 1031--1047.
+\url{https://doi.org/10.1111/rssc.12211}
 
 \leavevmode\vadjust pre{\hypertarget{ref-fromant2022}{}}%
-Fromant, A., Eizenberg, Y. H., Poupart, T., Bustamante, P., \& Arnould,
-J. P. Y. (2022). Year-round at-sea movements of fairy prions from
-southeastern {A}ustralia. \emph{Royal Society Open Science},
-\emph{9}(5), 220134. \url{https://doi.org/10.1098/rsos.220134}
+Fromant, A., Eizenberg, Y.
+H., Poupart, T., Bustamante, P., \& Arnould, J.
+P.
+Y.
+(2022).
+Year-round at-sea movements of fairy prions from southeastern {A}ustralia.
+\emph{Royal Society Open Science}, \emph{9}(5), 220134.
+\url{https://doi.org/10.1098/rsos.220134}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ganachaud2014}{}}%
-Ganachaud, A., Cravatte, S., Melet, A., Schiller, A., Holbrook, N. J.,
-Sloyan, B. M., Widlansky, M. J., Bowen, M., Verron, J., Wiles, P.,
-Ridgway, K., Sutton, P., Sprintall, J., Steinberg, C., Brassington, G.,
-Cai, W., Davis, R., Gasparin, F., Gourdeau, L., \ldots{} Send, U.
-(2014). The Southwest Pacific Ocean circulation and climate experiment
-(SPICE). \emph{Journal of Geophysical Research: Oceans}, \emph{119}(11),
-7660--7686. \url{https://doi.org/10.1002/2013JC009678}
+Ganachaud, A., Cravatte, S., Melet, A., Schiller, A., Holbrook, N.
+J., Sloyan, B.
+M., Widlansky, M.
+J., Bowen, M., Verron, J., Wiles, P., Ridgway, K., Sutton, P., Sprintall, J., Steinberg, C., Brassington, G., Cai, W., Davis, R., Gasparin, F., Gourdeau, L., \ldots{} Send, U.
+(2014).
+The Southwest Pacific Ocean circulation and climate experiment (SPICE).
+\emph{Journal of Geophysical Research: Oceans}, \emph{119}(11), 7660--7686.
+\url{https://doi.org/10.1002/2013JC009678}
 
 \leavevmode\vadjust pre{\hypertarget{ref-garcia2022}{}}%
-Garcia, V., Schilling, H., Cruz, D., Hawes, S., Everett, J., Roughan,
-M., Miskiewicz, A., Pakhomov, E., Jeffs, A., \& Suthers, I. (2022).
-Entrainment and development of larval fish assemblages in two
-contrasting cold core eddies of the East Australian Current system.
+Garcia, V., Schilling, H., Cruz, D., Hawes, S., Everett, J., Roughan, M., Miskiewicz, A., Pakhomov, E., Jeffs, A., \& Suthers, I.
+(2022).
+Entrainment and development of larval fish assemblages in two contrasting cold core eddies of the East Australian Current system.
 \emph{Marine Ecology Progress Series}, \emph{685}, 1--18.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gotelli2022}{}}%
-Gotelli, N. J., Moyes, F., Antão, L. H., Blowes, S. A., Dornelas, M.,
-McGill, B. J., Penny, A., Schipper, A. M., Shimadzu, H., Supp, S. R.,
-Waldock, C. A., \& Magurran, A. E. (2022). Long-term changes in
-temperate marine fish assemblages are driven by a small subset of
-species. \emph{Global Change Biology}, \emph{28}(1), 46--53.
+Gotelli, N.
+J., Moyes, F., Antão, L.
+H., Blowes, S.
+A., Dornelas, M., McGill, B.
+J., Penny, A., Schipper, A.
+M., Shimadzu, H., Supp, S.
+R., Waldock, C.
+A., \& Magurran, A.
+E.
+(2022).
+Long-term changes in temperate marine fish assemblages are driven by a small subset of species.
+\emph{Global Change Biology}, \emph{28}(1), 46--53.
 \url{https://doi.org/10.1111/gcb.15947}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gruxe9millet2009}{}}%
-Grémillet, D., \& Boulinier, T. (2009). Spatial ecology and conservation
-of seabirds facing global climate change: a review. \emph{Marine Ecology
-Progress Series}, \emph{391}, 121--137.
+Grémillet, D., \& Boulinier, T.
+(2009).
+Spatial ecology and conservation of seabirds facing global climate change: a review.
+\emph{Marine Ecology Progress Series}, \emph{391}, 121--137.
 \url{https://doi.org/10.3354/meps08212}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-harris2007}{}}%
-Harris, J., Haward, M., Jabour, J., \& Woehler, E. J. (2007). A new
-approach to selecting Marine Protected Areas (MPAs) in the Southern
-Ocean. \emph{Antarctic Science}, \emph{19}(2), 189--194.
+Harris, J., Haward, M., Jabour, J., \& Woehler, E.
+J.
+(2007).
+A new approach to selecting Marine Protected Areas (MPAs) in the Southern Ocean.
+\emph{Antarctic Science}, \emph{19}(2), 189--194.
 \url{https://doi.org/10.1017/S0954102007000260}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hassler2011}{}}%
-Hassler, C. S., Djajadikarta, J. R., Doblin, M. A., Everett, J. D., \&
-Thompson, P. A. (2011). Characterisation of water masses and
-phytoplankton nutrient limitation in the East Australian Current
-separation zone during spring 2008. \emph{Deep Sea Research Part II:
-Topical Studies in Oceanography}, \emph{58}(5), 664--677.
+Hassler, C.
+S., Djajadikarta, J.
+R., Doblin, M.
+A., Everett, J.
+D., \& Thompson, P.
+A.
+(2011).
+Characterisation of water masses and phytoplankton nutrient limitation in the East Australian Current separation zone during spring 2008.
+\emph{Deep Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5), 664--677.
 \url{https://doi.org/10.1016/j.dsr2.2010.06.008}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-haury1978patterns}{}}%
-Haury, L., McGowan, J., \& Wiebe, P. (1978). Patterns and processes in
-the time-space scales of plankton distributions. In J. Steele (Ed.),
-\emph{Spatial pattern in plankton communities} (pp. 277--327). Springer.
+Haury, L., McGowan, J., \& Wiebe, P.
+(1978).
+Patterns and processes in the time-space scales of plankton distributions.
+In J.
+Steele (Ed.), \emph{Spatial pattern in plankton communities} (pp.
+277--327).
+Springer.
 \url{https://doi.org/10.1007/978-1-4899-2195-6_12}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hazen2013a}{}}%
-Hazen, E. L., Jorgensen, S., Rykaczewski, R. R., Bograd, S. J., Foley,
-D. G., Jonsen, I. D., Shaffer, S. A., Dunne, J. P., Costa, D. P.,
-Crowder, L. B., \& Block, B. A. (2013). Predicted habitat shifts of
-Pacific top predators in a changing climate. \emph{Nature Climate
-Change}, \emph{3}(3), 234--238.
+Hazen, E.
+L., Jorgensen, S., Rykaczewski, R.
+R., Bograd, S.
+J., Foley, D.
+G., Jonsen, I.
+D., Shaffer, S.
+A., Dunne, J.
+P., Costa, D.
+P., Crowder, L.
+B., \& Block, B.
+A.
+(2013).
+Predicted habitat shifts of Pacific top predators in a changing climate.
+\emph{Nature Climate Change}, \emph{3}(3), 234--238.
 \url{https://doi.org/10.1038/nclimate1686}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hazen2013}{}}%
-Hazen, E. L., Suryan, R. M., Santora, J. A., Bograd, S., Watanuki, Y.,
-\& Wilson, R. P. (2013). Scales and mechanisms of marine hotspot
-formation. \emph{Marine Ecology Progress Series}, \emph{487}, 177--183.
+Hazen, E.
+L., Suryan, R.
+M., Santora, J.
+A., Bograd, S., Watanuki, Y., \& Wilson, R.
+P.
+(2013).
+Scales and mechanisms of marine hotspot formation.
+\emph{Marine Ecology Progress Series}, \emph{487}, 177--183.
 \url{https://doi.org/10.3354/meps10477}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hedd2001}{}}%
-Hedd, A., Gales, R., \& Brothers, N. (2001). Foraging strategies of shy
-albatross {\emph{Thalassarche cauta}} breeding at Albatross Island,
-Tasmania, Australia. \emph{Marine Ecology Progress Series}, \emph{224},
-267--282. \url{https://doi.org/10.3354/meps224267}
+Hedd, A., Gales, R., \& Brothers, N.
+(2001).
+Foraging strategies of shy albatross {\emph{Thalassarche cauta}} breeding at Albatross Island, Tasmania, Australia.
+\emph{Marine Ecology Progress Series}, \emph{224}, 267--282.
+\url{https://doi.org/10.3354/meps224267}
 
 \leavevmode\vadjust pre{\hypertarget{ref-raster}{}}%
-Hijmans, R. J. (2022a). \emph{Raster: Geographic data analysis and
-modeling}. \url{https://CRAN.R-project.org/package=raster}
+Hijmans, R. J. (2022a). \emph{Raster: Geographic data analysis and modeling}. \url{https://CRAN.R-project.org/package=raster}
 
 \leavevmode\vadjust pre{\hypertarget{ref-terra}{}}%
-Hijmans, R. J. (2022b). \emph{Terra: Spatial data analysis}.
-\url{https://CRAN.R-project.org/package=terra}
+Hijmans, R. J. (2022b). \emph{Terra: Spatial data analysis}. \url{https://CRAN.R-project.org/package=terra}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hill2008}{}}%
-Hill, K. L., Rintoul, S. R., Coleman, R., \& Ridgway, K. R. (2008). Wind
-forced low frequency variability of the East Australia Current.
+Hill, K.
+L., Rintoul, S.
+R., Coleman, R., \& Ridgway, K.
+R.
+(2008).
+Wind forced low frequency variability of the East Australia Current.
 \emph{Geophysical Research Letters}, \emph{35}(8).
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hill2017}{}}%
-Hill, N. A., Foster, S. D., Duhamel, G., Welsford, D., Koubbi, P., \&
-Johnson, C. R. (2017). Model-based mapping of assemblages for ecology
-and conservation management: a case study of demersal fish on the
-Kerguelen Plateau. \emph{Diversity and Distributions}, \emph{23}(10),
-1216--1230. https://doi.org/\url{https://doi.org/10.1111/ddi.12613}
+Hill, N.
+A., Foster, S.
+D., Duhamel, G., Welsford, D., Koubbi, P., \& Johnson, C.
+R.
+(2017).
+Model-based mapping of assemblages for ecology and conservation management: a case study of demersal fish on the Kerguelen Plateau.
+\emph{Diversity and Distributions}, \emph{23}(10), 1216--1230.
+https://doi.org/\url{https://doi.org/10.1111/ddi.12613}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hill2020}{}}%
-Hill, N. A., Woolley, S. N. C., Foster, S., Dunstan, P. K., McKinlay,
-J., Ovaskainen, O., \& Johnson, C. (2020). Determining marine
-bioregions: a comparison of quantitative approaches. \emph{Methods in
-Ecology and Evolution}, \emph{11}(10), 1258--1272.
+Hill, N.
+A., Woolley, S.
+N.
+C., Foster, S., Dunstan, P.
+K., McKinlay, J., Ovaskainen, O., \& Johnson, C.
+(2020).
+Determining marine bioregions: a comparison of quantitative approaches.
+\emph{Methods in Ecology and Evolution}, \emph{11}(10), 1258--1272.
 \url{https://doi.org/10.1111/2041-210X.13447}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hobday2010}{}}%
-Hobday, A. J. (2010). Ensemble analysis of the future distribution of
-large pelagic fishes off Australia. \emph{Progress in Oceanography},
-\emph{86}(1), 291--301.
+Hobday, A.
+J.
+(2010).
+Ensemble analysis of the future distribution of large pelagic fishes off Australia.
+\emph{Progress in Oceanography}, \emph{86}(1), 291--301.
 \url{https://doi.org/10.1016/j.pocean.2010.04.023}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hsieh2016}{}}%
-Hsieh, T. C., Ma, K. H., \& Chao, A. (2016). iNEXT: an R package for
-rarefaction and extrapolation of species diversity (Hill numbers).
+Hsieh, T.
+C., Ma, K.
+H., \& Chao, A.
+(2016).
+iNEXT: an R package for rarefaction and extrapolation of species diversity (Hill numbers).
 \emph{Methods in Ecology and Evolution}, \emph{7}(12), 1451--1456.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hunt1987}{}}%
-Hunt, G. L., \& Schneider, D. C. (1987). Scale-dependent processes in
-the physical and biological environment of marine birds. In J. P.
-Croxall (Ed.), \emph{Seabirds: feeding ecology and roles in marine
-ecosystem} (pp. 7--41). Cambridge University Press.
+Hunt, G.
+L., \& Schneider, D.
+C.
+(1987).
+Scale-dependent processes in the physical and biological environment of marine birds.
+In J.
+P.
+Croxall (Ed.), \emph{Seabirds: feeding ecology and roles in marine ecosystem} (pp.
+7--41).
+Cambridge University Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-hyrenbach2007}{}}%
-Hyrenbach, K. D., Veit, R. R., Weimerskirch, H., Metzl, N., \& Hunt, G.
-L. (2007). Community structure across a large-scale ocean productivity
-gradient: marine bird assemblages of the Southern Indian Ocean.
-\emph{Deep Sea Research Part I: Oceanographic Research Papers},
-\emph{54}(7), 1129--1145.
+Hyrenbach, K.
+D., Veit, R.
+R., Weimerskirch, H., Metzl, N., \& Hunt, G.
+L.
+(2007).
+Community structure across a large-scale ocean productivity gradient: marine bird assemblages of the Southern Indian Ocean.
+\emph{Deep Sea Research Part I: Oceanographic Research Papers}, \emph{54}(7), 1129--1145.
 \url{https://doi.org/10.1016/j.dsr.2007.05.002}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ismar2011}{}}%
-Ismar, S. M. H., Phillips, R. A., Rayner, M. J., \& Hauber, M. E.
-(2011). Geolocation tracking of the annual migration of adult
-Australasian gannets ({\emph{Morus serrator}}) breeding in New Zealand.
+Ismar, S.
+M.
+H., Phillips, R.
+A., Rayner, M.
+J., \& Hauber, M.
+E.
+(2011).
+Geolocation tracking of the annual migration of adult Australasian gannets ({\emph{Morus serrator}}) breeding in New Zealand.
 \emph{Wilson Journal of Ornithology}, \emph{123}(1), 121--125.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-jaine2014}{}}%
-Jaine, F., Rohner, C., Weeks, S., Couturier, L., Bennett, M., Townsend,
-K., \& Richardson, A. (2014). Movements and habitat use of reef manta
-rays off eastern Australia: offshore excursions, deep diving and eddy
-affinity revealed by satellite telemetry. \emph{Marine Ecology Progress
-Series}, \emph{510}, 73--86. \url{https://doi.org/10.3354/meps10910}
+Jaine, F., Rohner, C., Weeks, S., Couturier, L., Bennett, M., Townsend, K., \& Richardson, A.
+(2014).
+Movements and habitat use of reef manta rays off eastern Australia: offshore excursions, deep diving and eddy affinity revealed by satellite telemetry.
+\emph{Marine Ecology Progress Series}, \emph{510}, 73--86.
+\url{https://doi.org/10.3354/meps10910}
 
 \leavevmode\vadjust pre{\hypertarget{ref-johnson2021}{}}%
-Johnson, B., \& Colombelli-Négrel, D. (2021). Breeding success in
-Southern Australian little penguins is negatively correlated with high
-wind speeds and sea surface temperatures. \emph{Ornithological
-Applications}, \emph{123}(1), duaa062.
+Johnson, B., \& Colombelli-Négrel, D.
+(2021).
+Breeding success in Southern Australian little penguins is negatively correlated with high wind speeds and sea surface temperatures.
+\emph{Ornithological Applications}, \emph{123}(1), duaa062.
 \url{https://doi.org/10.1093/ornithapp/duaa062}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-johnson2011}{}}%
-Johnson, C. R., Banks, S. C., Barrett, N. S., Cazassus, F., Dunstan, P.
-K., Edgar, G. J., Frusher, S. D., Gardner, C., Haddon, M., Helidoniotis,
-F., Hill, K. L., Holbrook, N. J., Hosie, G. W., Last, P. R., Ling, S.
-D., Melbourne-Thomas, J., Miller, K., Pecl, G. T., Richardson, A. J.,
-\ldots{} Taw, N. (2011). Climate change cascades: Shifts in
-oceanography, species' ranges and subtidal marine community dynamics in
-eastern Tasmania. \emph{Journal of Experimental Marine Biology and
-Ecology}, \emph{400}(1), 17--32.
+Johnson, C.
+R., Banks, S.
+C., Barrett, N.
+S., Cazassus, F., Dunstan, P.
+K., Edgar, G.
+J., Frusher, S.
+D., Gardner, C., Haddon, M., Helidoniotis, F., Hill, K.
+L., Holbrook, N.
+J., Hosie, G.
+W., Last, P.
+R., Ling, S.
+D., Melbourne-Thomas, J., Miller, K., Pecl, G.
+T., Richardson, A.
+J., \ldots{} Taw, N.
+(2011).
+Climate change cascades: Shifts in oceanography, species' ranges and subtidal marine community dynamics in eastern Tasmania.
+\emph{Journal of Experimental Marine Biology and Ecology}, \emph{400}(1), 17--32.
 \url{https://doi.org/10.1016/j.jembe.2011.02.032}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-karpouzi2007}{}}%
-Karpouzi, V., Watson, R., \& Pauly, D. (2007). Modelling and mapping
-resource overlap between seabirds and fisheries on a global scale: a
-preliminary assessment. \emph{Marine Ecology Progress Series},
-\emph{343}, 87--99. \url{https://doi.org/10.3354/meps06860}
+Karpouzi, V., Watson, R., \& Pauly, D.
+(2007).
+Modelling and mapping resource overlap between seabirds and fisheries on a global scale: a preliminary assessment.
+\emph{Marine Ecology Progress Series}, \emph{343}, 87--99.
+\url{https://doi.org/10.3354/meps06860}
 
 \leavevmode\vadjust pre{\hypertarget{ref-kessler2013}{}}%
-Kessler, W. S., \& Cravatte, S. (2013). Mean circulation of the Coral
-Sea. \emph{Journal of Geophysical Research: Oceans}, \emph{118}(12),
-6385--6410. \url{https://doi.org/10.1002/2013JC009117}
+Kessler, W. S., \& Cravatte, S. (2013). Mean circulation of the Coral Sea. \emph{Journal of Geophysical Research: Oceans}, \emph{118}(12), 6385--6410. \url{https://doi.org/10.1002/2013JC009117}
 
 \leavevmode\vadjust pre{\hypertarget{ref-kreft2010}{}}%
-Kreft, H., \& Jetz, W. (2010). A framework for delineating
-biogeographical regions based on species distributions. \emph{Journal of
-Biogeography}, \emph{37}(11), 2029--2053.
+Kreft, H., \& Jetz, W.
+(2010).
+A framework for delineating biogeographical regions based on species distributions.
+\emph{Journal of Biogeography}, \emph{37}(11), 2029--2053.
 \url{https://doi.org/10.1111/j.1365-2699.2010.02375.x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lavers2014}{}}%
-Lavers, J. L., Miller, M. G. R., Carter, M. J., Swann, G., \& Clarke, R.
-H. (2014). Predicting the spatial distribution of a seabird community to
-identify priority conservation areas in the Timor Sea.
+Lavers, J.
+L., Miller, M.
+G.
+R., Carter, M.
+J., Swann, G., \& Clarke, R.
+H.
+(2014).
+Predicting the spatial distribution of a seabird community to identify priority conservation areas in the Timor Sea.
 \emph{Conservation Biology}, \emph{28}(6), 1699--1709.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lee2019good}{}}%
-Lee, M. R., Powell, J. R., Oberle, B., Cornwell, W. K., Lyons, M., Rigg,
-J. L., \& Zanne, A. E. (2019). Good neighbors aplenty: Fungal endophytes
-rarely exhibit competitive exclusion patterns across a span of woody
-habitats. \emph{Ecology}, \emph{100}(9), e02790.
+Lee, M.
+R., Powell, J.
+R., Oberle, B., Cornwell, W.
+K., Lyons, M., Rigg, J.
+L., \& Zanne, A.
+E.
+(2019).
+Good neighbors aplenty: Fungal endophytes rarely exhibit competitive exclusion patterns across a span of woody habitats.
+\emph{Ecology}, \emph{100}(9), e02790.
 \url{https://doi.org/10.1002/ecy.2790}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lenoir2020}{}}%
-Lenoir, J., Bertrand, R., Comte, L., Bourgeaud, L., Hattab, T.,
-Murienne, J., \& Grenouillet, G. (2020). Species better track climate
-warming in the oceans than on land. \emph{Nature Ecology \& Evolution},
-\emph{4}(8), 1044--1059. \url{https://doi.org/10.1038/s41559-020-1198-2}
+Lenoir, J., Bertrand, R., Comte, L., Bourgeaud, L., Hattab, T., Murienne, J., \& Grenouillet, G.
+(2020).
+Species better track climate warming in the oceans than on land.
+\emph{Nature Ecology \& Evolution}, \emph{4}(8), 1044--1059.
+\url{https://doi.org/10.1038/s41559-020-1198-2}
 
 \leavevmode\vadjust pre{\hypertarget{ref-li2022}{}}%
-Li, J., Roughan, M., \& Kerry, C. (2022). Drivers of ocean warming in
-the western boundary currents of the Southern Hemisphere. \emph{Nature
-Climate Change}, \emph{12}(10), 901--909.
+Li, J., Roughan, M., \& Kerry, C.
+(2022).
+Drivers of ocean warming in the western boundary currents of the Southern Hemisphere.
+\emph{Nature Climate Change}, \emph{12}(10), 901--909.
 \url{https://doi.org/10.1038/s41558-022-01473-8}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-longhurst2007}{}}%
-Longhurst, A. R. (2007). \emph{Ecological geography of the sea} (2nd
-ed). Academic Press.
+Longhurst, A. R. (2007). \emph{Ecological geography of the sea} (2nd ed). Academic Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-lyons2017}{}}%
-Lyons, M. B., Foster, S. D., \& Keith, D. A. (2017). Simultaneous
-vegetation classification and mapping at large spatial scales.
+Lyons, M.
+B., Foster, S.
+D., \& Keith, D.
+A.
+(2017).
+Simultaneous vegetation classification and mapping at large spatial scales.
 \emph{Journal of Biogeography}, \emph{44}(12), 2891--2902.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-marchant1990}{}}%
-Marchant, R. (1990). Robustness of classification and ordination
-techniques applied to macroinvertebrate communities from the La Trobe
-River, Victoria. \emph{Marine and Freshwater Research}, \emph{41}(4),
-493--504. \url{https://doi.org/10.1071/MF9900493}
+Marchant, R.
+(1990).
+Robustness of classification and ordination techniques applied to macroinvertebrate communities from the La Trobe River, Victoria.
+\emph{Marine and Freshwater Research}, \emph{41}(4), 493--504.
+\url{https://doi.org/10.1071/MF9900493}
 
 \leavevmode\vadjust pre{\hypertarget{ref-marchant1990handbook}{}}%
-Marchant, S., \& Higgins, P. J. (1990). \emph{Handbook of australian,
-new zealand \& antarctic birds. Vol. 1, ratites to ducks, p. AB}. Oxford
-University Press.
+Marchant, S., \& Higgins, P. J. (1990). \emph{Handbook of australian, new zealand \& antarctic birds. Vol. 1, ratites to ducks, p. AB}. Oxford University Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-mason2018}{}}%
-Mason, C., Alderman, R., McGowan, J., Possingham, H. P., Hobday, A. J.,
-Sumner, M., \& Shaw, J. (2018). Telemetry reveals existing marine
-protected areas are worse than random for protecting the foraging
-habitat of threatened shy albatross ({\emph{Thalassarche cauta}}).
+Mason, C., Alderman, R., McGowan, J., Possingham, H.
+P., Hobday, A.
+J., Sumner, M., \& Shaw, J.
+(2018).
+Telemetry reveals existing marine protected areas are worse than random for protecting the foraging habitat of threatened shy albatross ({\emph{Thalassarche cauta}}).
 \emph{Diversity and Distributions}, \emph{24}(12), 1744--1755.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mata2006}{}}%
-Mata, M. M., Wijffels, S. E., Church, J. A., \& Tomczak, M. (2006). Eddy
-shedding and energy conversions in the East Australian Current.
+Mata, M.
+M., Wijffels, S.
+E., Church, J.
+A., \& Tomczak, M.
+(2006).
+Eddy shedding and energy conversions in the East Australian Current.
 \emph{Journal of Geophysical Research: Oceans}, \emph{111}, C09034.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mcduie2015}{}}%
-McDuie, F., Weeks, S. J., Miller, M. G. R., \& Congdon, B. C. (2015).
-Breeding tropical shearwaters use distant foraging sites when
-self-provisioning. \emph{Marine Ornithology}, \emph{43}, 123--129.
+McDuie, F., Weeks, S. J., Miller, M. G. R., \& Congdon, B. C. (2015). Breeding tropical shearwaters use distant foraging sites when self-provisioning. \emph{Marine Ornithology}, \emph{43}, 123--129.
 
 \leavevmode\vadjust pre{\hypertarget{ref-melo2005}{}}%
-Melo, A. S. (2005). Effects of taxonomic and numeric resolution on the
-ability to detect ecological patterns at a local scale using stream
-macroinvertebrates. \emph{Archiv Für Hydrobiologie}, \emph{164}(3),
-309--323. \url{https://doi.org/10.1127/0003-9136/2005/0164-0309}
+Melo, A.
+S.
+(2005).
+Effects of taxonomic and numeric resolution on the ability to detect ecological patterns at a local scale using stream macroinvertebrates.
+\emph{Archiv Für Hydrobiologie}, \emph{164}(3), 309--323.
+\url{https://doi.org/10.1127/0003-9136/2005/0164-0309}
 
 \leavevmode\vadjust pre{\hypertarget{ref-miller2018}{}}%
-Miller, M. G. R., Silva, F. R. O., Machovsky-Capuska, G. E., \& Congdon,
-B. C. (2018). Sexual segregation in tropical seabirds: drivers of
-sex-specific foraging in the brown booby {\emph{Sula leucogaster}}.
+Miller, M.
+G.
+R., Silva, F.
+R.
+O., Machovsky-Capuska, G.
+E., \& Congdon, B.
+C.
+(2018).
+Sexual segregation in tropical seabirds: drivers of sex-specific foraging in the brown booby {\emph{Sula leucogaster}}.
 \emph{Journal of Ornithology}, \emph{159}(2), 425--437.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-morrone2018}{}}%
-Morrone, J. J. (2018). The spectre of biogeographical regionalization.
-\emph{Journal of Biogeography}, \emph{45}(2), 282--288.
-\url{https://doi.org/10.1111/jbi.13135}
+Morrone, J. J. (2018). The spectre of biogeographical regionalization. \emph{Journal of Biogeography}, \emph{45}(2), 282--288. \url{https://doi.org/10.1111/jbi.13135}
 
 \leavevmode\vadjust pre{\hypertarget{ref-morrone2023biogeographical}{}}%
-Morrone, J. J. (2023). Why biogeographical transition zones matter.
-\emph{Journal of Biogeography}. \url{https://doi.org/10.1111/jbi.14632}
+Morrone, J. J. (2023). Why biogeographical transition zones matter. \emph{Journal of Biogeography}. \url{https://doi.org/10.1111/jbi.14632}
 
 \leavevmode\vadjust pre{\hypertarget{ref-mott2018}{}}%
-Mott, R., \& Clarke, R. H. (2018). Systematic review of geographic
-biases in the collection of at-sea distribution data for seabirds.
+Mott, R., \& Clarke, R.
+H.
+(2018).
+Systematic review of geographic biases in the collection of at-sea distribution data for seabirds.
 \emph{Emu - Austral Ornithology}, \emph{118}(3), 235--246.
@@
 Niella, Y., Butcher, P., Holmes, B., Barnett, A., \& Harcourt, R.
-(2022). Forecasting intraspecific changes in distribution of a
-wide-ranging marine predator under climate change. \emph{Oecologia},
-\emph{198}(1), 111--124.
+(2022).
+Forecasting intraspecific changes in distribution of a wide-ranging marine predator under climate change.
+\emph{Oecologia}, \emph{198}(1), 111--124.
 \url{https://doi.org/10.1007/s00442-021-05075-7}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-niella2020}{}}%
-Niella, Y., Smoothey, A., Peddemors, V., \& Harcourt, R. (2020).
-Predicting changes in distribution of a large coastal shark in the face
-of the strengthening East Australian Current. \emph{Marine Ecology
-Progress Series}, \emph{642}, 163--177.
+Niella, Y., Smoothey, A., Peddemors, V., \& Harcourt, R.
+(2020).
+Predicting changes in distribution of a large coastal shark in the face of the strengthening East Australian Current.
+\emph{Marine Ecology Progress Series}, \emph{642}, 163--177.
 \url{https://doi.org/10.3354/meps13322}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-norman1992distribution}{}}%
-Norman, F. (1992). Distribution and abundance of seabirds off {P}hillip
-{I}sland and within {P}ort {P}hillip {B}ay, {V}ictoria, 1986--1988.
-\emph{Emu - Austral Ornithology}, \emph{91}(5), 377--394.
+Norman, F. (1992). Distribution and abundance of seabirds off {P}hillip {I}sland and within {P}ort {P}hillip {B}ay, {V}ictoria, 1986--1988. \emph{Emu - Austral Ornithology}, \emph{91}(5), 377--394.
 
 \leavevmode\vadjust pre{\hypertarget{ref-ohara2020}{}}%
-O'Hara, T. D., Williams, A., Althaus, F., Ross, A. S., \& Bax, N. J.
-(2020). Regional-scale patterns of deep seafloor biodiversity for
-conservation assessment. \emph{Diversity and Distributions},
-\emph{26}(4), 479--494. \url{https://doi.org/10.1111/ddi.13034}
+O'Hara, T.
+D., Williams, A., Althaus, F., Ross, A.
+S., \& Bax, N.
+J.
+(2020).
+Regional-scale patterns of deep seafloor biodiversity for conservation assessment.
+\emph{Diversity and Distributions}, \emph{26}(4), 479--494.
+\url{https://doi.org/10.1111/ddi.13034}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ohara2020a}{}}%
-O'Hara, T. D., Williams, A., Woolley, S. N. C., Nau, A. W., \& Bax, N.
-J. (2020). Deep-sea temperate-tropical faunal transition across uniform
-environmental gradients. \emph{Deep Sea Research Part I: Oceanographic
-Research Papers}, \emph{161}, 103283.
+O'Hara, T.
+D., Williams, A., Woolley, S.
+N.
+C., Nau, A.
+W., \& Bax, N.
+J.
+(2020).
+Deep-sea temperate-tropical faunal transition across uniform environmental gradients.
+\emph{Deep Sea Research Part I: Oceanographic Research Papers}, \emph{161}, 103283.
 \url{https://doi.org/10.1016/j.dsr.2020.103283}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-oke2000}{}}%
-Oke, P. R., \& Middleton, J. H. (2000). Topographically induced
-upwelling off eastern Australia. \emph{Journal of Physical
-Oceanography}, \emph{30}(3), 512--531.
+Oke, P.
+R., \& Middleton, J.
+H.
+(2000).
+Topographically induced upwelling off eastern Australia.
+\emph{Journal of Physical Oceanography}, \emph{30}(3), 512--531.
 \url{https://doi.org/10.1175/1520-0485(2000)030\%3C0512:TIUOEA\%3E2.0.CO;2}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-oke2019a}{}}%
-Oke, P. R., Pilo, G. S., Ridgway, K., Kiss, A., \& Rykova, T. (2019). A
-search for the Tasman Front. \emph{Journal of Marine Systems},
-\emph{199}, 103217. \url{https://doi.org/10.1016/j.jmarsys.2019.103217}
+Oke, P.
+R., Pilo, G.
+S., Ridgway, K., Kiss, A., \& Rykova, T.
+(2019).
+A search for the Tasman Front.
+\emph{Journal of Marine Systems}, \emph{199}, 103217.
+\url{https://doi.org/10.1016/j.jmarsys.2019.103217}
 
 \leavevmode\vadjust pre{\hypertarget{ref-oke2019}{}}%
-Oke, P. R., Roughan, M., Cetina-Heredia, P., Pilo, G. S., Ridgway, K.
-R., Rykova, T., Archer, M. R., Coleman, R. C., Kerry, C. G., Rocha, C.,
-Schaeffer, A., \& Vitarelli, E. (2019). Revisiting the circulation of
-the East Australian Current: its path, separation, and eddy field.
+Oke, P.
+R., Roughan, M., Cetina-Heredia, P., Pilo, G.
+S., Ridgway, K.
+R., Rykova, T., Archer, M.
+R., Coleman, R.
+C., Kerry, C.
+G., Rocha, C., Schaeffer, A., \& Vitarelli, E.
+(2019).
+Revisiting the circulation of the East Australian Current: its path, separation, and eddy field.
 \emph{Progress in Oceanography}, \emph{176}, 102139.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-oliver2017}{}}%
-Oliver, E. C. J., Benthuysen, J. A., Bindoff, N. L., Hobday, A. J.,
-Holbrook, N. J., Mundy, C. N., \& Perkins-Kirkpatrick, S. E. (2017). The
-unprecedented 2015/16 Tasman Sea marine heatwave. \emph{Nature
-Communications}, \emph{8}(1), 16101.
+Oliver, E.
+C.
+J., Benthuysen, J.
+A., Bindoff, N.
+L., Hobday, A.
+J., Holbrook, N.
+J., Mundy, C.
+N., \& Perkins-Kirkpatrick, S.
+E.
+(2017).
+The unprecedented 2015/16 Tasman Sea marine heatwave.
+\emph{Nature Communications}, \emph{8}(1), 16101.
 \url{https://doi.org/10.1038/ncomms16101}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-oliver2018}{}}%
-Oliver, E. C. J., Lago, V., Hobday, A. J., Holbrook, N. J., Ling, S. D.,
-\& Mundy, C. N. (2018). Marine heatwaves off eastern Tasmania: trends,
-interannual variability, and predictability. \emph{Progress in
-Oceanography}, \emph{161}, 116--130.
+Oliver, E.
+C.
+J., Lago, V., Hobday, A.
+J., Holbrook, N.
+J., Ling, S.
+D., \& Mundy, C.
+N.
+(2018).
+Marine heatwaves off eastern Tasmania: trends, interannual variability, and predictability.
+\emph{Progress in Oceanography}, \emph{161}, 116--130.
 \url{https://doi.org/10.1016/j.pocean.2018.02.007}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pantiukhin2021}{}}%
-Pantiukhin, D., Piepenburg, D., Hansen, M. L. S., \& Kraan, C. (2021).
-Data-driven bioregionalization: a seascape-scale study of macrobenthic
-communities in the Eurasian Arctic. \emph{Journal of Biogeography},
-\emph{48}(11), 2877--2890. \url{https://doi.org/10.1111/jbi.14247}
+Pantiukhin, D., Piepenburg, D., Hansen, M.
+L.
+S., \& Kraan, C.
+(2021).
+Data-driven bioregionalization: a seascape-scale study of macrobenthic communities in the Eurasian Arctic.
+\emph{Journal of Biogeography}, \emph{48}(11), 2877--2890.
+\url{https://doi.org/10.1111/jbi.14247}
 
 \leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
-Pebesma, E. (2018). {Simple features for R: standardized support for
-spatial vector data}. \emph{{The R Journal}}, \emph{10}(1), 439--446.
-\url{https://doi.org/10.32614/RJ-2018-009}
+Pebesma, E. (2018). {Simple features for R: standardized support for spatial vector data}. \emph{{The R Journal}}, \emph{10}(1), 439--446. \url{https://doi.org/10.32614/RJ-2018-009}
 
 \leavevmode\vadjust pre{\hypertarget{ref-peck2004}{}}%
-Peck, D., Smithers, B., Krockenberger, A., \& Congdon, B. (2004). Sea
-surface temperature constrains wedge-tailed shearwater foraging success
-within breeding seasons. \emph{Marine Ecology Progress Series},
-\emph{281}, 259--266. \url{https://doi.org/10.3354/meps281259}
+Peck, D., Smithers, B., Krockenberger, A., \& Congdon, B.
+(2004).
+Sea surface temperature constrains wedge-tailed shearwater foraging success within breeding seasons.
+\emph{Marine Ecology Progress Series}, \emph{281}, 259--266.
+\url{https://doi.org/10.3354/meps281259}
 
 \leavevmode\vadjust pre{\hypertarget{ref-puxe9ron2010}{}}%
-Péron, C., Authier, M., Barbraud, C., Delord, K., Besson, D., \&
-Weimerskirch, H. (2010). Interdecadal changes in at-sea distribution and
-abundance of subantarctic seabirds along a latitudinal gradient in the
-Southern Indian Ocean. \emph{Global Change Biology}, \emph{16}(7),
-1895--1909. \url{https://doi.org/10.1111/j.1365-2486.2010.02169.x}
+Péron, C., Authier, M., Barbraud, C., Delord, K., Besson, D., \& Weimerskirch, H.
+(2010).
+Interdecadal changes in at-sea distribution and abundance of subantarctic seabirds along a latitudinal gradient in the Southern Indian Ocean.
+\emph{Global Change Biology}, \emph{16}(7), 1895--1909.
+\url{https://doi.org/10.1111/j.1365-2486.2010.02169.x}
 
 \leavevmode\vadjust pre{\hypertarget{ref-phillips2022}{}}%
-Phillips, L. R., Malan, N., Roughan, M., Harcourt, R., Jonsen, I., Cox,
-M., Brierley, A. S., Slip, D., Wilkins, A., \& Carroll, G. (2022).
-Coastal seascape variability in the intensifying {E}ast {A}ustralian
-{C}urrent southern extension. \emph{Frontiers in Marine Science},
-\emph{9}, 925123.
+Phillips, L.
+R., Malan, N., Roughan, M., Harcourt, R., Jonsen, I., Cox, M., Brierley, A.
+S., Slip, D., Wilkins, A., \& Carroll, G.
+(2022).
+Coastal seascape variability in the intensifying {E}ast {A}ustralian {C}urrent southern extension.
+\emph{Frontiers in Marine Science}, \emph{9}, 925123.
 \url{https://www.frontiersin.org/articles/10.3389/fmars.2022.925123}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pinsky2019}{}}%
-Pinsky, M. L., Eikeset, A. M., McCauley, D. J., Payne, J. L., \& Sunday,
-J. M. (2019). Greater vulnerability to warming of marine versus
-terrestrial ectotherms. \emph{Nature}, \emph{569}(7754), 108--111.
+Pinsky, M.
+L., Eikeset, A.
+M., McCauley, D.
+J., Payne, J.
+L., \& Sunday, J.
+M.
+(2019).
+Greater vulnerability to warming of marine versus terrestrial ectotherms.
+\emph{Nature}, \emph{569}(7754), 108--111.
 \url{https://doi.org/10.1038/s41586-019-1132-4}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pocklington1979}{}}%
-Pocklington, R. (1979). An oceanographic interpretation of seabird
-distributions in the Indian Ocean. \emph{Marine Biology}, \emph{51}(1),
-9--21. \url{https://doi.org/10.1007/BF00389026}
+Pocklington, R. (1979). An oceanographic interpretation of seabird distributions in the Indian Ocean. \emph{Marine Biology}, \emph{51}(1), 9--21. \url{https://doi.org/10.1007/BF00389026}
 
 \leavevmode\vadjust pre{\hypertarget{ref-poloczanska2007}{}}%
-Poloczanska, E. S., Babcock, R. C., Butler, A., Hobday, A. J.,
-Hoegh-Guldberg, O., Kunz, T. J., Matear, R., Milton, D., Okey, T. A., \&
-Richardson, A. J. (2007). Climate change and Australian marine life. In
-R. N. Gibson, R. J. A. Atkinson, \& J. D. M. Gordon (Eds.),
-\emph{Oceanography and marine biology: an annual review} (Vol. 45, pp.
-407--478). CRC Press, Taylor \& Francis Group.
+Poloczanska, E.
+S., Babcock, R.
+C., Butler, A., Hobday, A.
+J., Hoegh-Guldberg, O., Kunz, T.
+J., Matear, R., Milton, D., Okey, T.
+A., \& Richardson, A.
+J.
+(2007).
+Climate change and Australian marine life.
+In R.
+N.
+Gibson, R.
+J.
+A.
+Atkinson, \& J.
+D.
+M.
+Gordon (Eds.), \emph{Oceanography and marine biology: an annual review} (Vol.
+45, pp.
+407--478).
+CRC Press, Taylor \& Francis Group.
 
 \leavevmode\vadjust pre{\hypertarget{ref-poloczanska2013}{}}%
-Poloczanska, E. S., Brown, C. J., Sydeman, W. J., Kiessling, W.,
-Schoeman, D. S., Moore, P. J., Brander, K., Bruno, J. F., Buckley, L.
-B., Burrows, M. T., Duarte, C. M., Halpern, B. S., Holding, J., Kappel,
-C. V., O'Connor, M. I., Pandolfi, J. M., Parmesan, C., Schwing, F.,
-Thompson, S. A., \& Richardson, A. J. (2013). Global imprint of climate
-change on marine life. \emph{Nature Climate Change}, \emph{3}(10),
-919--925. \url{https://doi.org/10.1038/nclimate1958}
+Poloczanska, E.
+S., Brown, C.
+J., Sydeman, W.
+J., Kiessling, W., Schoeman, D.
+S., Moore, P.
+J., Brander, K., Bruno, J.
+F., Buckley, L.
+B., Burrows, M.
+T., Duarte, C.
+M., Halpern, B.
+S., Holding, J., Kappel, C.
+V., O'Connor, M.
+I., Pandolfi, J.
+M., Parmesan, C., Schwing, F., Thompson, S.
+A., \& Richardson, A.
+J.
+(2013).
+Global imprint of climate change on marine life.
+\emph{Nature Climate Change}, \emph{3}(10), 919--925.
+\url{https://doi.org/10.1038/nclimate1958}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ponti2023}{}}%
-Ponti, R., \& Sannolo, M. (2023). The importance of including phenology
-when modelling species ecological niche. \emph{Ecography},
-\emph{2023}(4), e06143. \url{https://doi.org/10.1111/ecog.06143}
+Ponti, R., \& Sannolo, M. (2023). The importance of including phenology when modelling species ecological niche. \emph{Ecography}, \emph{2023}(4), e06143. \url{https://doi.org/10.1111/ecog.06143}
 
 \leavevmode\vadjust pre{\hypertarget{ref-priddel2014}{}}%
-Priddel, D., Carlile, N., Portelli, D., Kim, Y., O'Neill, L.,
-Bretagnolle, V., Ballance, L. T., Phillips, R. A., Pitman, R. L., \&
-Raynei, M. J. (2014). Pelagic distribution of Gould's petrel
-({\emph{Pterodroma leucoptera}}): linking shipboard and onshore
-observations with remote-tracking data. \emph{Emu - Austral
-Ornithology}, \emph{114}(4), 360--370.
+Priddel, D., Carlile, N., Portelli, D., Kim, Y., O'Neill, L., Bretagnolle, V., Ballance, L.
+T., Phillips, R.
+A., Pitman, R.
+L., \& Raynei, M.
+J.
+(2014).
+Pelagic distribution of Gould's petrel ({\emph{Pterodroma leucoptera}}): linking shipboard and onshore observations with remote-tracking data.
+\emph{Emu - Austral Ornithology}, \emph{114}(4), 360--370.
 \url{https://doi.org/10.1071/MU14021}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quillfeldt2015}{}}%
-Quillfeldt, P., Cherel, Y., Delord, K., \& Weimerkirch, H. (2015). Cool,
-cold or colder? Spatial segregation of prions and blue petrels is
-explained by differences in preferred sea surface temperatures.
+Quillfeldt, P., Cherel, Y., Delord, K., \& Weimerkirch, H.
+(2015).
+Cool, cold or colder?
+Spatial segregation of prions and blue petrels is explained by differences in preferred sea surface temperatures.
 \emph{Biology Letters}, \emph{11}(4), 20141090.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-r2022}{}}%
-R Core Team. (2022). \emph{R: A language and environment for statistical
-computing}. R Foundation for Statistical Computing.
-\url{https://www.R-project.org/}
+R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ramuxedrez2017}{}}%
-Ramírez, F., Afán, I., Davis, L. S., \& Chiaradia, A. (2017). Climate
-impacts on global hot spots of marine biodiversity. \emph{Science
-Advances}, \emph{3}(2), e1601198.
+Ramírez, F., Afán, I., Davis, L.
+S., \& Chiaradia, A.
+(2017).
+Climate impacts on global hot spots of marine biodiversity.
+\emph{Science Advances}, \emph{3}(2), e1601198.
 \url{https://doi.org/10.1126/sciadv.1601198}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-raymond2010}{}}%
-Raymond, B., Shaffer, S. A., Sokolov, S., Woehler, E. J., Costa, D. P.,
-Einoder, L., Hindell, M., Hosie, G., Pinkerton, M., Sagar, P. M., Scott,
-D., Smith, A., Thompson, D. R., Vertigan, C., \& Weimerskirch, H.
-(2010). Shearwater foraging in the Southern Ocean: the roles of prey
-availability and winds. \emph{PLoS ONE}, \emph{5}(6), e10960.
+Raymond, B., Shaffer, S.
+A., Sokolov, S., Woehler, E.
+J., Costa, D.
+P., Einoder, L., Hindell, M., Hosie, G., Pinkerton, M., Sagar, P.
+M., Scott, D., Smith, A., Thompson, D.
+R., Vertigan, C., \& Weimerskirch, H.
+(2010).
+Shearwater foraging in the Southern Ocean: the roles of prey availability and winds.
+\emph{PLoS ONE}, \emph{5}(6), e10960.
 \url{https://doi.org/10.1371/journal.pone.0010960}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-receveur2020}{}}%
-Receveur, A., Vourey, E., Lebourges-Dhaussy, A., Menkes, C., Ménard, F.,
-\& Allain, V. (2020). Biogeography of micronekton assemblages in the
-Natural Park of the Coral Sea. \emph{Frontiers in Marine Science},
-\emph{7}, 449. \url{https://doi.org/10.3389/fmars.2020.00449}
+Receveur, A., Vourey, E., Lebourges-Dhaussy, A., Menkes, C., Ménard, F., \& Allain, V.
+(2020).
+Biogeography of micronekton assemblages in the Natural Park of the Coral Sea.
+\emph{Frontiers in Marine Science}, \emph{7}, 449.
+\url{https://doi.org/10.3389/fmars.2020.00449}
 
 \leavevmode\vadjust pre{\hypertarget{ref-reid2023}{}}%
-Reid, K., Baker, G. B., \& Woehler, E. J. (2023). An ecological risk
-assessment for the impacts of offshore wind farms on birds in Australia.
+Reid, K., Baker, G.
+B., \& Woehler, E.
+J.
+(2023).
+An ecological risk assessment for the impacts of offshore wind farms on birds in Australia.
 \emph{Austral Ecology}, \emph{48}(2), 418--439.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-reid2000}{}}%
-Reid, T. A., \& Hindell, M. A. (2000). Coarse-scale relationships
-between seabirds and zooplankton off south-eastern Tasmania.
+Reid, T.
+A., \& Hindell, M.
+A.
+(2000).
+Coarse-scale relationships between seabirds and zooplankton off south-eastern Tasmania.
 \emph{Marine and Freshwater Research}, \emph{51}(8), 789--798.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-reisinger2022}{}}%
-Reisinger, R. R., Corney, S., Raymond, B., Lombard, A. T., Bester, M.
-N., Crawford, R. J. M., Davies, D., de Bruyn, P. J. N., Dilley, B. J.,
-Kirkman, S. P., Makhado, A. B., Ryan, P. G., Schoombie, S., Stevens, K.
-L., Tosh, C. A., Wege, M., Whitehead, T. O., Sumner, M. D., Wotherspoon,
-S., \ldots{} Pistorius, P. A. (2022). Habitat model forecasts suggest
-potential redistribution of marine predators in the southern Indian
-Ocean. \emph{Diversity and Distributions}, \emph{28}(1), 142--159.
+Reisinger, R.
+R., Corney, S., Raymond, B., Lombard, A.
+T., Bester, M.
+N., Crawford, R.
+J.
+M., Davies, D., de Bruyn, P.
+J.
+N., Dilley, B.
+J., Kirkman, S.
+P., Makhado, A.
+B., Ryan, P.
+G., Schoombie, S., Stevens, K.
+L., Tosh, C.
+A., Wege, M., Whitehead, T.
+O., Sumner, M.
+D., Wotherspoon, S., \ldots{} Pistorius, P.
+A.
+(2022).
+Habitat model forecasts suggest potential redistribution of marine predators in the southern Indian Ocean.
+\emph{Diversity and Distributions}, \emph{28}(1), 142--159.
 \url{https://doi.org/10.1111/ddi.13447}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-reygondeau2013}{}}%
-Reygondeau, G., Longhurst, A., Martinez, E., Beaugrand, G., Antoine, D.,
-\& Maury, O. (2013). Dynamic biogeochemical provinces in the global
-ocean. \emph{Global Biogeochemical Cycles}, \emph{27}(4), 1046--1058.
+Reygondeau, G., Longhurst, A., Martinez, E., Beaugrand, G., Antoine, D., \& Maury, O.
+(2013).
+Dynamic biogeochemical provinces in the global ocean.
+\emph{Global Biogeochemical Cycles}, \emph{27}(4), 1046--1058.
 \url{https://doi.org/10.1002/gbc.20089}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-reygondeau2012}{}}%
-Reygondeau, G., Maury, O., Beaugrand, G., Fromentin, J. M., Fonteneau,
-A., \& Cury, P. (2012). Biogeography of tuna and billfish communities.
+Reygondeau, G., Maury, O., Beaugrand, G., Fromentin, J.
+M., Fonteneau, A., \& Cury, P.
+(2012).
+Biogeography of tuna and billfish communities.
 \emph{Journal of Biogeography}, \emph{39}(1), 114--129.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ribic1997}{}}%
-Ribic, C., Ainley, D., \& Spear, L. (1997). Scale-related
-seabird-environmental relationships in Pacific equatorial waters, with
-reference to El Niño-Southern Oscillation events. \emph{Marine Ecology
-Progress Series}, \emph{156}, 183--203.
+Ribic, C., Ainley, D., \& Spear, L.
+(1997).
+Scale-related seabird-environmental relationships in Pacific equatorial waters, with reference to El Niño-Southern Oscillation events.
+\emph{Marine Ecology Progress Series}, \emph{156}, 183--203.
 \url{https://doi.org/10.3354/meps156183}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ridgway2003a}{}}%
-Ridgway, K. R., \& Dunn, J. R. (2003). Mesoscale structure of the mean
-East Australian Current System and its relationship with topography.
+Ridgway, K.
+R., \& Dunn, J.
+R.
+(2003).
+Mesoscale structure of the mean East Australian Current System and its relationship with topography.
 \emph{Progress in Oceanography}, \emph{56}(2), 189--222.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-santora2013}{}}%
-Santora, J., \& Veit, R. (2013). Spatio-temporal persistence of top
-predator hotspots near the Antarctic Peninsula. \emph{Marine Ecology
-Progress Series}, \emph{487}, 287--304.
+Santora, J., \& Veit, R.
+(2013).
+Spatio-temporal persistence of top predator hotspots near the Antarctic Peninsula.
+\emph{Marine Ecology Progress Series}, \emph{487}, 287--304.
 \url{https://doi.org/10.3354/meps10350}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-scales2017}{}}%
-Scales, K. L., Hazen, E. L., Jacox, M. G., Edwards, C. A., Boustany, A.
-M., Oliver, M. J., \& Bograd, S. J. (2017). Scale of inference: on the
-sensitivity of habitat models for wide-ranging marine predators to the
-resolution of environmental data. \emph{Ecography}, \emph{40}(1),
-210--220. \url{https://doi.org/10.1111/ecog.02272}
+Scales, K.
+L., Hazen, E.
+L., Jacox, M.
+G., Edwards, C.
+A., Boustany, A.
+M., Oliver, M.
+J., \& Bograd, S.
+J.
+(2017).
+Scale of inference: on the sensitivity of habitat models for wide-ranging marine predators to the resolution of environmental data.
+\emph{Ecography}, \emph{40}(1), 210--220.
+\url{https://doi.org/10.1111/ecog.02272}
 
 \leavevmode\vadjust pre{\hypertarget{ref-scales2014a}{}}%
-Scales, K. L., Miller, P. I., Embling, C. B., Ingram, S. N., Pirotta,
-E., \& Votier, S. C. (2014). Mesoscale fronts as foraging habitats:
-composite front mapping reveals oceanographic drivers of habitat use for
-a pelagic seabird. \emph{Journal of The Royal Society Interface},
-\emph{11}(100), 20140679. \url{https://doi.org/10.1098/rsif.2014.0679}
+Scales, K.
+L., Miller, P.
+I., Embling, C.
+B., Ingram, S.
+N., Pirotta, E., \& Votier, S.
+C.
+(2014).
+Mesoscale fronts as foraging habitats: composite front mapping reveals oceanographic drivers of habitat use for a pelagic seabird.
+\emph{Journal of The Royal Society Interface}, \emph{11}(100), 20140679.
+\url{https://doi.org/10.1098/rsif.2014.0679}
 
 \leavevmode\vadjust pre{\hypertarget{ref-scales2014}{}}%
-Scales, K. L., Miller, P. I., Hawkes, L. A., Ingram, S. N., Sims, D. W.,
-\& Votier, S. C. (2014). On the front line: frontal zones as priority
-at-sea conservation areas for mobile marine vertebrates. \emph{Journal
-of Applied Ecology}, \emph{51}(6), 1575--1583.
+Scales, K.
+L., Miller, P.
+I., Hawkes, L.
+A., Ingram, S.
+N., Sims, D.
+W., \& Votier, S.
+C.
+(2014).
+On the front line: frontal zones as priority at-sea conservation areas for mobile marine vertebrates.
+\emph{Journal of Applied Ecology}, \emph{51}(6), 1575--1583.
 \url{https://doi.org/10.1111/1365-2664.12330}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-biology2002}{}}%
-Schreiber, E. A., \& Burger, J. (Eds.). (2002). \emph{Biology of marine
-birds}. CRC Press.
+Schreiber, E. A., \& Burger, J. (Eds.). (2002). \emph{Biology of marine birds}. CRC Press.
 
 \leavevmode\vadjust pre{\hypertarget{ref-sherman1994}{}}%
-Sherman, K. (1994). Sustainability, biomass yields, and health of
-coastal ecosystems: an ecological perspective. \emph{Marine Ecology
-Progress Series}, \emph{112}, 277--301.
+Sherman, K.
+(1994).
+Sustainability, biomass yields, and health of coastal ecosystems: an ecological perspective.
+\emph{Marine Ecology Progress Series}, \emph{112}, 277--301.
 \url{https://doi.org/10.3354/meps112277}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sojitra2022}{}}%
-Sojitra, M., Woehler, E. J., Lea, M.-A., \& Wotherspoon, S. (2022).
-Multi-decadal changes in the at-sea distribution and abundance of
-black-browed and light-mantled sooty albatrosses in the {S}outhwest
-{P}acific {O}cean. \emph{ICES Journal of Marine Science}, fsac197.
+Sojitra, M., Woehler, E.
+J., Lea, M.-A., \& Wotherspoon, S.
+(2022).
+Multi-decadal changes in the at-sea distribution and abundance of black-browed and light-mantled sooty albatrosses in the {S}outhwest {P}acific {O}cean.
+\emph{ICES Journal of Marine Science}, fsac197.
 \url{https://doi.org/10.1093/icesjms/fsac197}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-souza2023}{}}%
-Souza, J. S., \& Santos, L. N. (2023). Resident species, not immigrants,
-drive reorganization of estuarine fish assemblages in response to
-warming. \emph{Ecology}, \emph{104}(5), e3987.
+Souza, J.
+S., \& Santos, L.
+N.
+(2023).
+Resident species, not immigrants, drive reorganization of estuarine fish assemblages in response to warming.
+\emph{Ecology}, \emph{104}(5), e3987.
 \url{https://doi.org/10.1002/ecy.3987}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-spalding2007}{}}%
-Spalding, M. D., Fox, H. E., Allen, G. R., Davidson, N., Ferdaña, Z. A.,
-Finlayson, M., Halpern, B. S., Jorge, M. A., Lombana, A., Lourie, S. A.,
-Martin, K. D., McManus, E., Molnar, J., Recchia, C. A., \& Robertson, J.
-(2007). Marine {E}coregions of the {W}orld: A bioregionalization of
-coastal and shelf areas. \emph{BioScience}, \emph{57}(7), 573--583.
+Spalding, M.
+D., Fox, H.
+E., Allen, G.
+R., Davidson, N., Ferdaña, Z.
+A., Finlayson, M., Halpern, B.
+S., Jorge, M.
+A., Lombana, A., Lourie, S.
+A., Martin, K.
+D., McManus, E., Molnar, J., Recchia, C.
+A., \& Robertson, J.
+(2007).
+Marine {E}coregions of the {W}orld: A bioregionalization of coastal and shelf areas.
+\emph{BioScience}, \emph{57}(7), 573--583.
 \url{https://doi.org/10.1641/B570707}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sun2022}{}}%
-Sun, C., Zhang, A., Jin, B., Wang, X., Zhang, X., \& Zhang, L. (2022).
+Sun, C., Zhang, A., Jin, B., Wang, X., Zhang, X., \& Zhang, L.
+(2022).
 Seasonal variability of eddy kinetic energy in the north indian ocean.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-suthers2011}{}}%
-Suthers, I. M., Young, J. W., Baird, M. E., Roughan, M., Everett, J. D.,
-Brassington, G. B., Byrne, M., Condie, S. A., Hartog, J. R., Hassler, C.
-S., Hobday, A. J., Holbrook, N. J., Malcolm, H. A., Oke, P. R.,
-Thompson, P. A., \& Ridgway, K. (2011). The strengthening East
-Australian Current, its eddies and biological effects {\textemdash} an
-introduction and overview. \emph{Deep Sea Research Part II: Topical
-Studies in Oceanography}, \emph{58}(5), 538--546.
+Suthers, I.
+M., Young, J.
+W., Baird, M.
+E., Roughan, M., Everett, J.
+D., Brassington, G.
+B., Byrne, M., Condie, S.
+A., Hartog, J.
+R., Hassler, C.
+S., Hobday, A.
+J., Holbrook, N.
+J., Malcolm, H.
+A., Oke, P.
+R., Thompson, P.
+A., \& Ridgway, K.
+(2011).
+The strengthening East Australian Current, its eddies and biological effects {\textemdash} an introduction and overview.
+\emph{Deep Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5), 538--546.
 \url{https://doi.org/10.1016/j.dsr2.2010.09.029}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sutton2017a}{}}%
-Sutton, T. T., Clark, M. R., Dunn, D. C., Halpin, P. N., Rogers, A. D.,
-Guinotte, J., Bograd, S. J., Angel, M. V., Perez, J. A. A., Wishner, K.,
-Haedrich, R. L., Lindsay, D. J., Drazen, J. C., Vereshchaka, A.,
-Piatkowski, U., Morato, T., Błachowiak-Samołyk, K., Robison, B. H.,
-Gjerde, K. M., \ldots{} Heino, M. (2017). A global biogeographic
-classification of the mesopelagic zone. \emph{Deep Sea Research Part I:
-Oceanographic Research Papers}, \emph{126}, 85--102.
+Sutton, T.
+T., Clark, M.
+R., Dunn, D.
+C., Halpin, P.
+N., Rogers, A.
+D., Guinotte, J., Bograd, S.
+J., Angel, M.
+V., Perez, J.
+A.
+A., Wishner, K., Haedrich, R.
+L., Lindsay, D.
+J., Drazen, J.
+C., Vereshchaka, A., Piatkowski, U., Morato, T., Błachowiak-Samołyk, K., Robison, B.
+H., Gjerde, K.
+M., \ldots{} Heino, M.
+(2017).
+A global biogeographic classification of the mesopelagic zone.
+\emph{Deep Sea Research Part I: Oceanographic Research Papers}, \emph{126}, 85--102.
 \url{https://doi.org/10.1016/j.dsr.2017.05.006}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sydeman2012}{}}%
-Sydeman, W., Thompson, S., \& Kitaysky, A. (2012). Seabirds and climate
-change: roadmap for the future. \emph{Marine Ecology Progress Series},
-\emph{454}, 107--117. \url{https://doi.org/10.3354/meps09806}
+Sydeman, W., Thompson, S., \& Kitaysky, A.
+(2012).
+Seabirds and climate change: roadmap for the future.
+\emph{Marine Ecology Progress Series}, \emph{454}, 107--117.
+\url{https://doi.org/10.3354/meps09806}
 
 \leavevmode\vadjust pre{\hypertarget{ref-taylor2021}{}}%
-Taylor, G., Elliott, G., Walker, K., \& Bose, S. (2021). Year-round
-distribution, breeding cycle, and activity of white-headed petrels
-({\emph{Pterodroma lessonii}}) nesting on adams island, auckland
-islands. \emph{Notornis}, \emph{67}, 369--386.
+Taylor, G., Elliott, G., Walker, K., \& Bose, S.
+(2021).
+Year-round distribution, breeding cycle, and activity of white-headed petrels ({\emph{Pterodroma lessonii}}) nesting on adams island, auckland islands.
+\emph{Notornis}, \emph{67}, 369--386.
 
 \leavevmode\vadjust pre{\hypertarget{ref-tremblay2009}{}}%
-Tremblay, Y., Bertrand, S., Henry, R., Kappes, M., Costa, D., \&
-Shaffer, S. (2009). Analytical approaches to investigating
-seabird--environment interactions: a review. \emph{Marine Ecology
-Progress Series}, \emph{391}, 153--163.
+Tremblay, Y., Bertrand, S., Henry, R., Kappes, M., Costa, D., \& Shaffer, S.
+(2009).
+Analytical approaches to investigating seabird--environment interactions: a review.
+\emph{Marine Ecology Progress Series}, \emph{391}, 153--163.
 \url{https://doi.org/10.3354/meps08146}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-vilhena2015}{}}%
-Vilhena, D. A., \& Antonelli, A. (2015). A network approach for
-identifying and delimiting biogeographical regions. \emph{Nature
-Communications}, \emph{6}(1), 6848.
+Vilhena, D.
+A., \& Antonelli, A.
+(2015).
+A network approach for identifying and delimiting biogeographical regions.
+\emph{Nature Communications}, \emph{6}(1), 6848.
 \url{https://doi.org/10.1038/ncomms7848}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-warton2015}{}}%
-Warton, D. I., Foster, S. D., De'ath, G., Stoklosa, J., \& Dunstan, P.
-K. (2015). Model-based thinking for community ecology. \emph{Plant
-Ecology}, \emph{216}(5), 669--682.
+Warton, D.
+I., Foster, S.
+D., De'ath, G., Stoklosa, J., \& Dunstan, P.
+K.
+(2015).
+Model-based thinking for community ecology.
+\emph{Plant Ecology}, \emph{216}(5), 669--682.
 \url{https://doi.org/10.1007/s11258-014-0366-3}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-waters2008marine}{}}%
-Waters, J. M. (2008). Marine biogeographical disjunction in temperate
-{A}ustralia: Historical landbridge, contemporary currents, or both?
+Waters, J.
+M.
+(2008).
+Marine biogeographical disjunction in temperate {A}ustralia: Historical landbridge, contemporary currents, or both?
 \emph{Diversity and Distributions}, \emph{14}(4), 692--700.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2007}{}}%
-Weimerskirch, H. (2007). Are seabirds foraging for unpredictable
-resources? \emph{Deep Sea Research Part II: Topical Studies in
-Oceanography}, \emph{54}(3-4), 211--223.
+Weimerskirch, H.
+(2007).
+Are seabirds foraging for unpredictable resources?
+\emph{Deep Sea Research Part II: Topical Studies in Oceanography}, \emph{54}(3-4), 211--223.
 \url{https://doi.org/10.1016/j.dsr2.2006.11.013}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2020}{}}%
-Weimerskirch, H., De Grissac, S., Ravache, A., Prudor, A., Corbeau, A.,
-Congdon, B., McDuie, F., Bourgeois, K., Dromzée, S., Butscher, J.,
-Menkes, C., Allain, V., Vidal, E., Jaeger, A., \& Borsa, P. (2020).
-At-sea movements of wedge-tailed shearwaters during and outside the
-breeding season from four colonies in New Caledonia. \emph{Marine
-Ecology Progress Series}, \emph{633}, 225--238.
+Weimerskirch, H., De Grissac, S., Ravache, A., Prudor, A., Corbeau, A., Congdon, B., McDuie, F., Bourgeois, K., Dromzée, S., Butscher, J., Menkes, C., Allain, V., Vidal, E., Jaeger, A., \& Borsa, P.
+(2020).
+At-sea movements of wedge-tailed shearwaters during and outside the breeding season from four colonies in New Caledonia.
+\emph{Marine Ecology Progress Series}, \emph{633}, 225--238.
 \url{https://doi.org/10.3354/meps13171}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-whittaker2005}{}}%
-Whittaker, R. J., Araújo, M. B., Jepson, P., Ladle, R. J., Watson, J. E.
-M., \& Willis, K. J. (2005). Conservation Biogeography: assessment and
-prospect. \emph{Diversity and Distributions}, \emph{11}(1), 3--23.
+Whittaker, R.
+J., Araújo, M.
+B., Jepson, P., Ladle, R.
+J., Watson, J.
+E.
+M., \& Willis, K.
+J.
+(2005).
+Conservation Biogeography: assessment and prospect.
+\emph{Diversity and Distributions}, \emph{11}(1), 3--23.
 \url{https://doi.org/10.1111/j.1366-9516.2005.00143.x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-woehler2023}{}}%
-Woehler, E., \& Hobday, A. (2023). Impacts of marine heatwaves may be
-mediated by seabird life history strategies. \emph{Marine Ecology
-Progress Series}, \emph{HEAT}. \url{https://doi.org/10.3354/meps14333}
+Woehler, E., \& Hobday, A.
+(2023).
+Impacts of marine heatwaves may be mediated by seabird life history strategies.
+\emph{Marine Ecology Progress Series}, \emph{HEAT}.
+\url{https://doi.org/10.3354/meps14333}
 
 \leavevmode\vadjust pre{\hypertarget{ref-woehler2003}{}}%
-Woehler, E., Raymond, B., \& Watts, D. (2003). Decadal-scale seabird
-assemblages in Prydz Bay, East Antarctica. \emph{Marine Ecology Progress
-Series}, \emph{251}, 299--310. \url{https://doi.org/10.3354/meps251299}
+Woehler, E., Raymond, B., \& Watts, D.
+(2003).
+Decadal-scale seabird assemblages in Prydz Bay, East Antarctica.
+\emph{Marine Ecology Progress Series}, \emph{251}, 299--310.
+\url{https://doi.org/10.3354/meps251299}
 
 \leavevmode\vadjust pre{\hypertarget{ref-woolley2020}{}}%
-Woolley, S. N. C., Foster, S. D., Bax, N. J., Currie, J. C., Dunn, D.
-C., Hansen, C., Hill, N. A., O'Hara, T. D., Ovaskainen, O., Sayre, R.,
-Vanhatalo, J. P., \& Dunstan, P. K. (2020). Bioregions in marine
-environments: Combining biological and environmental data for management
-and scientific understanding. \emph{BioScience}, \emph{70}(1), 48--59.
+Woolley, S.
+N.
+C., Foster, S.
+D., Bax, N.
+J., Currie, J.
+C., Dunn, D.
+C., Hansen, C., Hill, N.
+A., O'Hara, T.
+D., Ovaskainen, O., Sayre, R., Vanhatalo, J.
+P., \& Dunstan, P.
+K.
+(2020).
+Bioregions in marine environments: Combining biological and environmental data for management and scientific understanding.
+\emph{BioScience}, \emph{70}(1), 48--59.
 \url{https://doi.org/10.1093/biosci/biz133}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ecomix}{}}%
-Woolley, S., Foster, S., \& Dunstan, P. (2022). \emph{Ecomix: Fitting
-finite mixture models to ecological data}.
-\url{https://github.com/skiptoniam/ecomix}
+Woolley, S., Foster, S., \& Dunstan, P. (2022). \emph{Ecomix: Fitting finite mixture models to ecological data}. \url{https://github.com/skiptoniam/ecomix}
 
 \leavevmode\vadjust pre{\hypertarget{ref-yang2016}{}}%
-Yang, H., Lohmann, G., Wei, W., Dima, M., Ionita, M., \& Liu, J. (2016).
-Intensification and poleward shift of subtropical western boundary
-currents in a warming climate. \emph{Journal of Geophysical Research:
-Oceans}, \emph{121}(7), 4928--4945.
+Yang, H., Lohmann, G., Wei, W., Dima, M., Ionita, M., \& Liu, J.
+(2016).
+Intensification and poleward shift of subtropical western boundary currents in a warming climate.
+\emph{Journal of Geophysical Research: Oceans}, \emph{121}(7), 4928--4945.
 \url{https://doi.org/10.1002/2015JC011513}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-young2011}{}}%
-Young, J. W., Hobday, A. J., Campbell, R. A., Kloser, R. J., Bonham, P.
-I., Clementson, L. A., \& Lansdell, M. J. (2011). The biological
-oceanography of the East Australian Current and surrounding waters in
-relation to tuna and billfish catches off eastern Australia. \emph{Deep
-Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5),
-720--733. \url{https://doi.org/10.1016/j.dsr2.2010.10.005}
+Young, J.
+W., Hobday, A.
+J., Campbell, R.
+A., Kloser, R.
+J., Bonham, P.
+I., Clementson, L.
+A., \& Lansdell, M.
+J.
+(2011).
+The biological oceanography of the East Australian Current and surrounding waters in relation to tuna and billfish catches off eastern Australia.
+\emph{Deep Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5), 720--733.
+\url{https://doi.org/10.1016/j.dsr2.2010.10.005}
 
@@
 \item
-  \textbf{TABLE 1.} Environmental data used in the Region of Common
-  Profile models applied to seabird data off eastern Australia, their
-  spatial/temporal resolutions, units, and sources
+  \textbf{TABLE 1.} Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources
 \item
-  \textbf{TABLE 2.} Full (a) and best (b) Region of Common Profile
-  Models applied to seabird data off eastern Australia, for both data
-  types used (presence-absence, based on a Bernoulli model; abundance,
-  based on a Negative Binomial model). Refer to Table 1 for the
-  environmental data acronyms
+  \textbf{TABLE 2.} Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli model; abundance, based on a Negative Binomial model).
+Refer to Table 1 for the environmental data acronyms
 \item
-  \textbf{FIGURE 1.} Study area showing the main ocean currents (a), and
-  seabird records made by season on top of the 1° latitude × 1°
-  longitude grid cells (b). In (a), the East Australian Current (EAC)
-  system is highlighted with its acronyms in bold and drawn in blue. The
-  South Equatorial Current (SEC) and its branches, the North Caledonian
-  Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red. The EAC
-  `eastern extension' is an eddy field (blue shaded area) often called
-  the `Tasman Front', from which the East Auckland Current originates
-  (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et
-  al.~(2019); for colour references see the online version
+  \textbf{FIGURE 1.} Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
+In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
+The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
+The EAC `eastern extension' is an eddy field (blue shaded area) often called the `Tasman Front', from which the East Auckland Current originates (EAUC, in green).
+Figure (a) is adapted after Oke, Roughan, et al.~(2019); for colour references see the online version
 \item
-  \textbf{FIGURE 2.} Partial plots for the retained covariates in the
-  spring (a) presence-absence and (b) abundance best models. The plot
-  shows the fitted probability of belonging to a Region of Common
-  Profiles (RCP) against the environmental value. All other predictors
-  are held at their mean values. Results for other seasons and models
-  can be found in the Supplementary Material (Fig. S5). Refer to Table 1
-  for the environmental data acronyms
+  \textbf{FIGURE 2.} Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
+All other predictors are held at their mean values.
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S5).
+Refer to Table 1 for the environmental data acronyms
 \item
-  \textbf{FIGURE 3.} Spatialized point predictions (hard-classes) for
-  seabird assemblages (Region of Common Profile; RCP), by season, based
-  on presence-absence (a) and abundance (b) models, off eastern
-  Australia. Each grid was assigned to the RCP group with highest
-  membership probability
+  \textbf{FIGURE 3.} Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+Each grid was assigned to the RCP group with highest membership probability
 \item
-  \textbf{FIGURE 4.} Predicted probability membership of each seabird
-  assemblage (Region of Common Profiles; RCP) and grid, off eastern
-  Australia, from spring presence-absence (a) and abundance (b) models.
-  The central column, `mean', corresponds to the point prediction and
-  Bayesian boot-strapped, lower and upper confidence intervals (CI), on
-  its sides. Note that, in case of two RCPs, RCP2 probablility equals to
-  (1 - RCP1 probability). Results for other seasons and models can be
-  found in the Supplementary Material (Fig. S6)
+  \textbf{FIGURE 4.} Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+The central column, `mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
+Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
+Results for other seasons and models can be found in the Supplementary Material (Fig.
+S6)
 \item
-  \textbf{FIGURE 5.} Species profiles of seabird assemblages (Region of
-  Common Profile; RCP) off eastern Australia in spring. Values are the
-  average and 95\% confidence intervals of probability of occurrence (a)
-  and predicted mean abundance (b) for each species, based on 1000
-  Bayesian bootstraps. Values in (b) were log10-transformed to
-  accommodate the high variation between species. Results for other
-  seasons and models can be found in the Supplementary Material (Fig.
+  \textbf{FIGURE 5.} Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
+Values are the average and 95\% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
+Values in (b) were log10-transformed to accommodate the high variation between species.
+Results for other seasons and models can be found in the Supplementary Material (Fig.
 S7)

### 2023-10-17 ms_preprint/ms_supp_mat.Rmd
ms tidy-up and supplementary material in rmd

---
title: "Supplementary Material"
subtitle: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
# author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    # fig_caption: false
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include = FALSE, echo = FALSE}
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(gt)
```

This document is an accompanying file for **XXXXX et al. (2023) Divers. Distr.**. Briefly, here you will find --

* **List of packages** and version used for data wrangling, visualisation and analyses;
* **Table S1**: Summary of sampling effort by voyage;
* **Figure S1**: Number of occurrence, frequency of occurrence (FO), and numeric frequency (NF) by season;
* **Figure S2**: Species richness and total number of seabirds counted by grid/season;
* **Figure S3**: Plots for choosing best RCP group number (`multifit`);
* **Figure S4**: Residual plots from the best fitted models;
* **Figure S5**: Partial plots for covariates;
* **Figure S6**: Probability maps from seasonal predictions;
* **Figure S7**: Species profiles for each season;
* **Figure S8**: Species-richness and sample-coverage curves.

\newpage

# List of packages

We used the following packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021], and the ones referenced in the main text.

The code is archived in a repository **(Zenodo repo)**, and you can find a detailed walk-through in it.

\newpage

\blandscape

```{r table-S1, echo = FALSE}
read.csv("../results/tableS1-effort-summary.csv")  %>% 
  dplyr::select(-X) %>% 
  gt::gt() %>% 
  # Tidy-up column labels
  gt::cols_label(
    voyage = gt::md("**Voyage**"),
    date_start = gt::md("**Date start**"),
    date_end = gt::md("**Date end**"),
    lat_range = gt::md("**Latitudinal range**"), 
    lon_range = gt::md("**Longitudinal range**"),
    no_records = gt::md("**No. of records**"),
    no_birds = gt::md("**No. of birds**"),
    no_spp = gt::md("**No. of species**")) %>% 
  # Table title
  gt::tab_header(title = gt::md("**TABLE S1.** Summary of seabird sampling effort by voyage, off eastern Australia, during Australasian Seabird Group's ship-based surveys between 2016--2021. Start and finish dates and geographic ranges of each voyage, including the number of seabird records and the number of individuals and species recorded"))
```

\newpage

![**Figure S1.** Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021. The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text). Species are ordered from the largest to the lowest total number of occurrences.](../results/FigS1_spp-nOCC-FO-NF-seasons.png)

\elandscape

\newpage

![**Figure S2.** Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see *Methods* in the main text). Dots represent the grid centroids.](../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png)

\newpage

![**Figure S3.1.** Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_1_multifit-Bernoulli.png)

\newpage

![**Figure S3.2.** Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_2_multifit-NegBin.png)

\newpage

![**Figure S4.1.** Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png)

\newpage

![**Figure S4.2.** Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png)

\newpage

![**Figure S5.1.** Partial plots for the retained covariates in the best seasonal models based on presence-absence data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_1_partial-plots-Bernoulli.png)

\newpage

\blandscape

![**Figure S5.2.** Partial plots for the retained covariates in the best seasonal models based on abundance data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_2_partial-plots-NegBin.png)

\newpage

![**Figure S6.1.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png)

\newpage

![**Figure S6.2.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png)

\newpage

![**Figure S7.1.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model. Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps. ](../results/FigS7_1_spp-profiles-Bernoulli.png)

\newpage

![**Figure S7.2.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model. Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps. Values were log10-transformed to accommodate the high variation between species.](../results/FigS7_2_spp-profiles-NegBin.png)

\elandscape

\newpage

![**Figure S8.** Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png)

\newpage

# References

### 2023-10-17 ms_preprint/ms_supp_mat.tex
ms tidy-up and supplementary material in rmd

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Supplementary Material},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Supplementary Material}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Probabilistic bioregion models suggest seabird assemblages are
linked to the major boundary current off eastern Australia}
\author{}
\date{\vspace{-2.5em}2023-10-17}

\begin{document}
\maketitle

This document is an accompanying file for \textbf{XXXXX et al.~(2023)
Divers. Distr.}. Briefly, here you will find --

\begin{itemize}
\tightlist
\item
  \textbf{List of packages} and version used for data wrangling,
  visualisation and analyses;
\item
  \textbf{Table S1}: Summary of sampling effort by voyage;
\item
  \textbf{Figure S1}: Number of occurrence, frequency of occurrence
  (FO), and numeric frequency (NF) by season;
\item
  \textbf{Figure S2}: Species richness and total number of seabirds
  counted by grid/season;
\item
  \textbf{Figure S3}: Plots for choosing best RCP group number
  (\texttt{multifit});
\item
  \textbf{Figure S4}: Residual plots from the best fitted models;
\item
  \textbf{Figure S5}: Partial plots for covariates;
\item
  \textbf{Figure S6}: Probability maps from seasonal predictions;
\item
  \textbf{Figure S7}: Species profiles for each season;
\item
  \textbf{Figure S8}: Species-richness and sample-coverage curves.
\end{itemize}

\newpage

\hypertarget{list-of-packages}{%
\section{List of packages}\label{list-of-packages}}

We used the following packages `plyr' 1.8.8 (Wickham, 2011), `dplyr'
1.1.2 (Wickham, François, et al., 2023), `tidyr' 1.3.0 (Wickham,
Vaughan, et al., 2023), `readr' 2.1.4 (Wickham, Hester, et al., 2023),
`tibble' 3.2.1 (Müller \& Wickham, 2023), `lubridate' 1.9.2 (Grolemund
\& Wickham, 2011), `stringr' 1.5.0 (Wickham, 2022), `purrr' 1.0.1
(Wickham \& Henry, 2023), `ggplot2' 3.4.2 (Wickham, 2016), `ggspatial'
1.1.7 (Dunnington, 2022), `patchwork' 1.1.2 (Pedersen, 2022),
`RColorBrewer' 1.1-3 (Neuwirth, 2022), `rnaturalearth' 0.3.2 (Massicotte
\& South, 2023), `sp' 1.6-0 (Bivand et al., 2013; E. J. Pebesma \&
Bivand, 2005), `sf' 1.0-8 (E. Pebesma, 2018), `mapview' 2.11.0
(Appelhans et al., 2022), `raster' 3.5-21 (Hijmans, 2022a), `terra'
1.6-7 (Hijmans, 2022b), `rerddap' 1.0.2 (Chamberlain, 2023),
`rerddapXtracto' 1.1.4 (Mendelssohn, 2022), `hadsstR' (Byrnes \& Dunic,
2017), `corrplot' 0.92 (Wei \& Simko, 2021), and the ones referenced in
the main text.

The code is archived in a repository \textbf{(Zenodo repo)}, and you can
find a detailed walk-through in it.

\newpage

\begin{landscape}

\begin{longtable}{lrrllrrr}
\caption*{
{\large \textbf{TABLE S1.} Summary of seabird sampling effort by voyage, off eastern Australia, during Australasian Seabird Group's ship-based surveys between 2016--2021. Start and finish dates and geographic ranges of each voyage, including the number of seabird records and the number of individuals and species recorded}
} \\ 
\toprule
\textbf{Voyage} & \textbf{Date start} & \textbf{Date end} & \textbf{Latitudinal range} & \textbf{Longitudinal range} & \textbf{No. of records} & \textbf{No. of birds} & \textbf{No. of species} \\ 
\midrule
in2016\_t02 & 2016-08-25 & 2016-08-28 & -43 – -34 & 147 – 152 & 344 & 475 & 25 \\ 
in2016\_v06 & 2016-10-29 & 2016-11-12 & -27 – -27 & 153 – 155 & 284 & 2892 & 14 \\ 
in2017\_v02 & 2017-03-16 & 2017-03-27 & -47 – -43 & 142 – 147 & 911 & 7122 & 30 \\ 
in2017\_t01 & 2017-09-24 & 2017-10-01 & -33 – -9 & 143 – 154 & 370 & 4540 & 17 \\ 
in2017\_t02 & 2017-11-24 & 2017-11-25 & -44 – -42 & 141 – 147 & 113 & 11010 & 17 \\ 
in2018\_t02 & 2018-05-14 & 2018-05-20 & -43 – -27 & 148 – 154 & 214 & 6168 & 31 \\ 
in2018\_c01 & 2018-05-28 & 2018-06-07 & -41 – -39 & 146 – 149 & 644 & 2846 & 25 \\ 
in2018\_v04 & 2018-09-11 & 2018-10-07 & -47 – -34 & 141 – 155 & 1136 & 10434 & 36 \\ 
in2018\_v06 & 2018-11-22 & 2018-12-18 & -44 – -41 & 146 – 149 & 1957 & 59628 & 43 \\ 
in2019\_v07 & 2019-04-10 & 2019-04-22 & -43 – -38 & 147 – 150 & 412 & 1472 & 26 \\ 
in2019\_t01 & 2019-04-29 & 2019-05-02 & -44 – -39 & 141 – 148 & 140 & 177 & 18 \\ 
in2019\_v04 & 2019-08-07 & 2019-09-01 & -24 – -10 & 146 – 160 & 1321 & 13383 & 26 \\ 
in2019\_t02 & 2019-10-03 & 2019-10-10 & -27 – -9 & 142 – 154 & 245 & 1895 & 28 \\ 
fk201228 & 2020-12-27 & 2021-01-25 & -27 – -23 & 153 – 157 & 1063 & 17298 & 20 \\ 
fk210206 & 2021-02-06 & 2021-03-05 & -27 – -19 & 153 – 157 & 1107 & 3306 & 14 \\ 
\bottomrule
\end{longtable}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS1_spp-nOCC-FO-NF-seasons.png}
\caption{\textbf{Figure S1.} Number of occurrences (nOCC), frequency of
occurrence (FO) and numeric frequency (NF) of seabirds recorded off
eastern Australia during Australasian Seabird Group's ship-based surveys
between 2016--2021. The dashed line represents the number of occurrence
thresholds (n = 6) each taxon had to match for its inclusion in the
seasonal models (see \emph{Methods} in the main text). Species are
ordered from the largest to the lowest total number of occurrences.}
\end{figure}

\end{landscape}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png}
\caption{\textbf{Figure S2.} Species richness and the total number of
seabirds counted off eastern Australia during Australasian Seabird
Group's ship-based surveys between 2016--2021, by grid/season, after
data aggregation (see \emph{Methods} in the main text). Dots represent
the grid centroids.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS3_1_multifit-Bernoulli.png}
\caption{\textbf{Figure S3.1.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal presence-absence model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS3_2_multifit-NegBin.png}
\caption{\textbf{Figure S3.2.} Multifit plot for Region of Common
Profiles (RCP) for each seasonal abundance (count) model, applied to
seabirds off eastern Australia. The number of groups with the lowest BIC
value indicates the best number of groups (assemblages) that describes
the data. For each number of groups, we ran 100 models with random
starting values to avoid getting stuck in an incorrect `optima' (see
\emph{Methods} in the main text). The resulting plot also shows how many
groups were `empty' (colour scale) with `five or less' or `only one'
sites assigned to an RCP, i.e.~the model was fit with, say, 5 groups,
but 3 of them had `five or less' or `only one' sites (grids) allocated
to an RCP.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS4_1_best-model-residuals-Bernoulli.png}
\caption{\textbf{Figure S4.1.} Best model residuals for each seasonal
presence-absence Region of Common Profile model, applied to seabirds off
eastern Australia.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS4_2_best-model-residuals-NegBin.png}
\caption{\textbf{Figure S4.2.} Best model residuals for each seasonal
abundance (count) Region of Common Profile model, applied to seabirds
off eastern Australia.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS6_1_partial-plots-Bernoulli.png}
\caption{\textbf{Figure S5.1.} Partial plots for the retained covariates
in the best seasonal models based on presence-absence data. The plot
shows the probability of belonging to a Region of Common Profiles (RCP)
against the environmental value. Refer to Table 1 in the main text for
the environmental data acronyms.}
\end{figure}

\newpage

\begin{landscape}

\begin{figure}
\centering
\includegraphics{../results/FigS6_2_partial-plots-NegBin.png}
\caption{\textbf{Figure S5.2.} Partial plots for the retained covariates
in the best seasonal models based on abundance data. The plot shows the
probability of belonging to a Region of Common Profiles (RCP) against
the environmental value. Refer to Table 1 in the main text for the
environmental data acronyms.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS5_1_prob-pred-Bernoulli.png}
\caption{\textbf{Figure S6.1.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from presence-absence models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS5_2_prob-pred-NegBin.png}
\caption{\textbf{Figure S6.2.} Predicted probability membership of for
each seabird assemblage (Region of Common Profiles; RCP) and grid, off
eastern Australia, from abundance (count) models. The central column,
`mean', corresponds to the point prediction and Bayesian bootstraped,
lower and upper confidence intervals (CI), on its sides.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS7_1_spp-profiles-Bernoulli.png}
\caption{\textbf{Figure S7.1.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal presence-absence
model. Values are the average and confidence intervals of probability of
occurrence for each species, based on 1000 Bayesian bootstraps.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS7_2_spp-profiles-NegBin.png}
\caption{\textbf{Figure S7.2.} Species profiles for each assemblage
(Region of Common Profiles; RCP) for each seasonal abundance (count)
model. Values are the average and confidence intervals of predicted mean
abundance for each species, based on 1000 Bayesian bootstraps. Values
were log10-transformed to accommodate the high variation between
species.}
\end{figure}

\end{landscape}

\newpage

\begin{figure}
\centering
\includegraphics{../results/FigS8_iNEXT.png}
\caption{\textbf{Figure S8.} Diversity curve (alpha diversity) and
sample coverage for each assemblage (Region of Common Profile; RCP) from
each presence-absence seasonal model.}
\end{figure}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-mapview}{}}%
Appelhans, T., Detsch, F., Reudenbach, C., \& Woellauer, S. (2022).
\emph{Mapview: Interactive viewing of spatial data in r}.
\url{https://CRAN.R-project.org/package=mapview}

\leavevmode\vadjust pre{\hypertarget{ref-sp2}{}}%
Bivand, R. S., Pebesma, E., \& Gomez-Rubio, V. (2013). \emph{Applied
spatial data analysis with {R}, second edition}. Springer, NY.
\url{https://asdar-book.org/}

\leavevmode\vadjust pre{\hypertarget{ref-hadsstr}{}}%
Byrnes, J., \& Dunic, J. (2017). \emph{hadsstR: R library for working
with HadSST data}. \url{https://github.com/jebyrnes/hadsstR}

\leavevmode\vadjust pre{\hypertarget{ref-rerddap}{}}%
Chamberlain, S. (2023). \emph{Rerddap: General purpose client for
'ERDDAP' servers}. \url{https://CRAN.R-project.org/package=rerddap}

\leavevmode\vadjust pre{\hypertarget{ref-ggspatial}{}}%
Dunnington, D. (2022). \emph{Ggspatial: Spatial data framework for
ggplot2}. \url{https://CRAN.R-project.org/package=ggspatial}

\leavevmode\vadjust pre{\hypertarget{ref-lubridate}{}}%
Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with
{lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3),
1--25. \url{https://www.jstatsoft.org/v40/i03/}

\leavevmode\vadjust pre{\hypertarget{ref-raster}{}}%
Hijmans, R. J. (2022a). \emph{Raster: Geographic data analysis and
modeling}. \url{https://CRAN.R-project.org/package=raster}

\leavevmode\vadjust pre{\hypertarget{ref-terra}{}}%
Hijmans, R. J. (2022b). \emph{Terra: Spatial data analysis}.
\url{https://CRAN.R-project.org/package=terra}

\leavevmode\vadjust pre{\hypertarget{ref-rnaturalearth}{}}%
Massicotte, P., \& South, A. (2023). \emph{Rnaturalearth: World map data
from natural earth}.
\url{https://CRAN.R-project.org/package=rnaturalearth}

\leavevmode\vadjust pre{\hypertarget{ref-rerddapXtracto}{}}%
Mendelssohn, R. (2022). \emph{rerddapXtracto: Extracts environmental
data from 'ERDDAP' web services}.
\url{https://CRAN.R-project.org/package=rerddapXtracto}

\leavevmode\vadjust pre{\hypertarget{ref-tibble}{}}%
Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}.
\url{https://CRAN.R-project.org/package=tibble}

\leavevmode\vadjust pre{\hypertarget{ref-rcolorbrewer}{}}%
Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}.
\url{https://CRAN.R-project.org/package=RColorBrewer}

\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
Pebesma, E. (2018). {Simple features for R: standardized support for
spatial vector data}. \emph{{The R Journal}}, \emph{10}(1), 439--446.
\url{https://doi.org/10.32614/RJ-2018-009}

\leavevmode\vadjust pre{\hypertarget{ref-sp1}{}}%
Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
data in {R}. \emph{R News}, \emph{5}(2), 9--13.
\url{https://CRAN.R-project.org/doc/Rnews/}

\leavevmode\vadjust pre{\hypertarget{ref-patchwork}{}}%
Pedersen, T. L. (2022). \emph{Patchwork: The composer of plots}.
\url{https://CRAN.R-project.org/package=patchwork}

\leavevmode\vadjust pre{\hypertarget{ref-corrplot2021}{}}%
Wei, T., \& Simko, V. (2021). \emph{R package 'corrplot': Visualization
of a correlation matrix}. \url{https://github.com/taiyun/corrplot}

\leavevmode\vadjust pre{\hypertarget{ref-plyr}{}}%
Wickham, H. (2011). The split-apply-combine strategy for data analysis.
\emph{Journal of Statistical Software}, \emph{40}(1), 1--29.
\url{https://www.jstatsoft.org/v40/i01/}

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-stringr}{}}%
Wickham, H. (2022). \emph{Stringr: Simple, consistent wrappers for
common string operations}.
\url{https://CRAN.R-project.org/package=stringr}

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023).
\emph{Dplyr: A grammar of data manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}

\leavevmode\vadjust pre{\hypertarget{ref-purrr}{}}%
Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming
tools}. \url{https://CRAN.R-project.org/package=purrr}

\leavevmode\vadjust pre{\hypertarget{ref-readr}{}}%
Wickham, H., Hester, J., \& Bryan, J. (2023). \emph{Readr: Read
rectangular text data}. \url{https://CRAN.R-project.org/package=readr}

\leavevmode\vadjust pre{\hypertarget{ref-tidyr}{}}%
Wickham, H., Vaughan, D., \& Girlich, M. (2023). \emph{Tidyr: Tidy messy
data}. \url{https://CRAN.R-project.org/package=tidyr}

\end{CSLReferences}

\end{document}

### 2023-10-17 ms_preprint/ms_preprint.Rmd
added specific formatting requirements for submission, such as 'title page' and all statements

@@
 authors:
-  - name: Nicholas W. Daudt
+  - name: Nicholas W. Daudt*
     department: Department of Marine Science
@@
     email: nicholaswdaudt@gmail.com
+    notes: Corresponding author
   - name: Eric J. Woehler
@@
     location: Hobart, Australia
+    email: eric.woehler@gmail.com
   - name: Matthew R. Schofield
@@
     location: Dunedin, Aotearoa New Zealand
+    email: matthew.schofield@otago.ac.nz
   - name: Robert O. Smith
@@
     location: Dunedin, Aotearoa New Zealand
+    email: robert.smith@otago.ac.nz
   - name: Leandro Bugoni
@@
     location: Rio Grande, RS, Brazil
+    email: lbugoni@yahoo.com.br
   - name: William J. Rayment
@@
     location: Dunedin, Aotearoa New Zealand
+    email: will.rayment@otago.ac.nz
 abstract: |
-  **Aim**: To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
-  **Methods**: We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+  **Aim:** To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+  **Location:** East coast of Australia.
+  **Methods:** We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
 These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-  **Results**: Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
+  **Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
 The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
@@
 Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
-  **Conclusion**: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+  **Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
 Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
@@
 csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
-# output: rticles::arxiv_article
 output:
@@
 ```
+
+# ORCID {.unnumbered}
+
+Nicholas W. Daudt <https://orcid.org/0000-0001-8211-8771>
+
+Eric J. Woehler <https://orcid.org/0000-0002-1125-0748>
+
+Matthew R. Schofield <https://orcid.org/0000-0003-1481-2766>
+
+Robert O. Smith <https://orcid.org/0000-0001-5931-3843>
+
+Leandro Bugoni <https://orcid.org/0000-0003-0689-7026>
+
+William J. Rayment -- **do you have ORCID, Will?**
+
+# Biosketch {.unnumbered}
+
+**Nicholas W.
+Daudt (he/him)** is broadly interested in spatial ecology and ecological modelling.
+His research uses data science and modelling to understand biogeographic patterns and ecological processes that build marine top predators' biodiversity patterns across space and time.
+
+# Acknowledgements {.unnumbered}
+
+We thank all observers who assisted in data collection, the crews of RV *Investigator* and RV *Falkor*, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
+The authors would like to thank Dr.
+Skipton N.C.
+Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
+NWD is supported by a Doctoral scholarship from the University of Otago.
+
+***
+
+**Data availability statement:** The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+The raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the raw data, please get in touch with EJW.
+
+**Funding statement:** .........**Eric - any specific funding(s) we need to acknowledge?**
+
+**Conflict of interest disclosure:** The authors declare no conflict of interest.
+
+**Ethics approval statement:** NA.
+
+**Patient consent statement:** NA.
+
+**Permission to reproduce material from other sources:** NA.
+
+**Clinical trial registration:** NA.
+
+\newpage
+
+\begin{center}
+\textbf{Seabird assemblages are linked to the major western boundary current off eastern Australia}
+\end{center}
+
+*Running title:* Seabird assemblages off Eastern Australia
+
+***
+
+**Abstract**
+
+**Aim:** To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+
+**Location:** East coast of Australia.
+
+**Methods:** We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
+These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
+
+**Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
+Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
+
+**Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+
+***Keywords*** Assemblages; Biogeography; Community ecology; East Australian Current; Mixture models; Region of Common Profiles
+
+\newpage
 
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\~18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
-The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.
-\@ref(fig:fig1-study-area)a**; @kessler2013; @ganachaud2014).
-**SST/SSS characteristics?
-ROB?** The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
+The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Figure \@ref(fig:fig1-study-area)a**; @kessler2013; @ganachaud2014).
+The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
 The EAC flow polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.
-\@ref(fig:fig1-study-area)a**), carrying eddy-like features as a transient current [@oke2019].
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure \@ref(fig:fig1-study-area)a**), carrying eddy-like features as a transient current [@oke2019].
 
@@
 
-Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.
-\@ref(fig:fig1-study-area)b**; **Table S1**).
+Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Figure \@ref(fig:fig1-study-area)b**; **Table S1**).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
@@
 We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (c.
-110 $\times$ 110 km) and season (**Fig.
-\@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
+110 $\times$ 110 km) and season (**Figure \@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 sula* (25.4%) and black-browed albatross *T.
-melanophris* (25%) (**Fig.
-S1**).
-An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig.
-S2**).
-Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Fig.
-S2**).
+melanophris* (25%) (**Figure S1**).
+An increase in species richness towards the south of the study region was clear in all but the winter season (**Figure S2**).
+Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Figure S2**).
 However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
@@
 
-For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Fig.
-S3**).
+For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Figure S3**).
 The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table \@ref(tab:table2)**).
-The best models showed a good fit with no apparent deviations from the assumptions (**Fig.
-S4**).
+The best models showed a good fit with no apparent deviations from the assumptions (**Figure S4**).
 All the best models selected 'sst' or 'sss', as essential for characterising water masses.
-The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21 °C, or 'sss' 35.1 PSU for summer (**Fig.
-\@ref(fig:fig2-partial-plots)**, **Fig.
-S5**), similar to the EAC's temperature and salinity thresholds.
-Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Fig.
-S5**).
+The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21 °C, or 'sss' 35.1 PSU for summer (**Figure \@ref(fig:fig2-partial-plots)**, **Figure S5**), similar to the EAC's temperature and salinity thresholds.
+Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Figure S5**).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.
-\@ref(fig:fig3-point-pred)**).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.
-\@ref(fig:fig4-prob-maps)**, **Fig.
-S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
-The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Straight' assemblage appears (**Fig.
-\@ref(fig:fig3-point-pred)**, **Fig.
-S6**).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Figure \@ref(fig:fig3-point-pred)**).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Figure \@ref(fig:fig4-prob-maps)**, **Figure S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
+The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Straight' assemblage appears (**Figure \@ref(fig:fig3-point-pred)**, **Figure S6**).
 
@@
 small albatrosses *Thalassarche* spp., fairy prion *P.
-turtur*) (**Fig.
-\@ref(fig:fig5-spp-profiles)**, **Fig.
-S7**).
+turtur*) (**Figure \@ref(fig:fig5-spp-profiles)**, **Figure S7**).
 The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
@@
 The summer and autumn models reflected the high abundance of short-tailed shearwater before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by the autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
-Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (**Fig.
-S1**; hence they were removed before modelling).
+Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (**Figure S1**; hence they were removed before modelling).
 Additional surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.
@@
 
-# Acknowledgements
-
-We thank all observers who assisted in data collection, the crews of RV *Investigator* and RV *Falkor*, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
-The authors would like to thank Dr.
-Skipton N.C.
-Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago.
-
 # References
@@
 :::
+
+\newpage
+
+# Data Accessibility Statement {.unnumbered}
+
+The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+The raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the raw data, please get in touch with XXX.
 
@@
 All other predictors are held at their mean values.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S5).
+Results for other seasons and models can be found in the Supplementary Material (Figure S5).
 Refer to Table 1 for the environmental data acronyms
@@
 Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S6)
+Results for other seasons and models can be found in the Supplementary Material (Figure S6)
 -   **FIGURE 5.** Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
@@
 Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S7)
+Results for other seasons and models can be found in the Supplementary Material (Figure S7)
 

### 2023-10-17 ms_preprint/ms_preprint.tex
added specific formatting requirements for submission, such as 'title page' and all statements

@@
 \author{
-    Nicholas W. Daudt
+    Nicholas W. Daudt*
    \\
@@
   Hobart, Australia \\
-  \texttt{} \\
+  \texttt{\href{mailto:eric.woehler@gmail.com}{\nolinkurl{eric.woehler@gmail.com}}} \\
    \And
@@
   Dunedin, Aotearoa New Zealand \\
-  \texttt{} \\
+  \texttt{\href{mailto:matthew.schofield@otago.ac.nz}{\nolinkurl{matthew.schofield@otago.ac.nz}}} \\
    \And
@@
   Dunedin, Aotearoa New Zealand \\
-  \texttt{} \\
+  \texttt{\href{mailto:robert.smith@otago.ac.nz}{\nolinkurl{robert.smith@otago.ac.nz}}} \\
    \And
@@
   Rio Grande, RS, Brazil \\
-  \texttt{} \\
+  \texttt{\href{mailto:lbugoni@yahoo.com.br}{\nolinkurl{lbugoni@yahoo.com.br}}} \\
    \And
@@
   Dunedin, Aotearoa New Zealand \\
-  \texttt{} \\
+  \texttt{\href{mailto:will.rayment@otago.ac.nz}{\nolinkurl{will.rayment@otago.ac.nz}}} \\
   }
@@
 \begin{abstract}
-\textbf{Aim}: To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
-\textbf{Methods}: We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+\textbf{Aim:} To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+\textbf{Location:} East coast of Australia.
+\textbf{Methods:} We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
 These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-\textbf{Results}: Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
+\textbf{Results:} Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
 The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
@@
 Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
-\textbf{Conclusion}: RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+\textbf{Main conclusions:} RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
 Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
@@
   }
+
+\hypertarget{orcid}{%
+\section*{ORCID}\label{orcid}}
+\addcontentsline{toc}{section}{ORCID}
+
+Nicholas W. Daudt \url{https://orcid.org/0000-0001-8211-8771}
+
+Eric J. Woehler \url{https://orcid.org/0000-0002-1125-0748}
+
+Matthew R. Schofield \url{https://orcid.org/0000-0003-1481-2766}
+
+Robert O. Smith \url{https://orcid.org/0000-0001-5931-3843}
+
+Leandro Bugoni \url{https://orcid.org/0000-0003-0689-7026}
+
+William J. Rayment -- \textbf{do you have ORCID, Will?}
+
+\hypertarget{biosketch}{%
+\section*{Biosketch}\label{biosketch}}
+\addcontentsline{toc}{section}{Biosketch}
+
+\textbf{Nicholas W.
+Daudt (he/him)} is broadly interested in spatial ecology and ecological modelling.
+His research uses data science and modelling to understand biogeographic patterns and ecological processes that build marine top predators' biodiversity patterns across space and time.
+
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+We thank all observers who assisted in data collection, the crews of RV \emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
+The authors would like to thank Dr.~Skipton N.C.
+Woolley, \texttt{\{ecomix\}} developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
+NWD is supported by a Doctoral scholarship from the University of Otago.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Data availability statement:} The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+The raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the raw data, please get in touch with EJW.
+
+\textbf{Funding statement:} \ldots\ldots\ldots{}\textbf{Eric - any specific funding(s) we need to acknowledge?}
+
+\textbf{Conflict of interest disclosure:} The authors declare no conflict of interest.
+
+\textbf{Ethics approval statement:} NA.
+
+\textbf{Patient consent statement:} NA.
+
+\textbf{Permission to reproduce material from other sources:} NA.
+
+\textbf{Clinical trial registration:} NA.
+
+\newpage
+
+\begin{center}
+\textbf{Seabird assemblages are linked to the major western boundary current off eastern Australia}
+\end{center}
+
+\emph{Running title:} Seabird assemblages off Eastern Australia
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Abstract}
+
+\textbf{Aim:} To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
+
+\textbf{Location:} East coast of Australia.
+
+\textbf{Methods:} We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
+These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
+
+\textbf{Results:} Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
+Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
+
+\textbf{Main conclusions:} RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+
+\textbf{\emph{Keywords}} Assemblages; Biogeography; Community ecology; East Australian Current; Mixture models; Region of Common Profiles
+
+\newpage
 
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte, 2013).
-The South Caledonian Jet completes the EAC system at \textasciitilde22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (\textbf{Fig.
-\ref{fig:fig1-study-area}a}; Kessler \& Cravatte (2013); Ganachaud et al.
+The South Caledonian Jet completes the EAC system at \textasciitilde22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (\textbf{Figure \ref{fig:fig1-study-area}a}; Kessler \& Cravatte (2013); Ganachaud et al.
 (2014)).
-\textbf{SST/SSS characteristics?
-ROB?} The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019; Oke \& Middleton, 2000; Ridgway \& Dunn, 2003).
+The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019; Oke \& Middleton, 2000; Ridgway \& Dunn, 2003).
 The EAC flow polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension' (Oke, Pilo, et al.
 (2019); often termed the `Tasman Front').
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the `EAC southern extension'; \textbf{Fig.
-\ref{fig:fig1-study-area}a}), carrying eddy-like features as a transient current (Oke, Roughan, et al., 2019).
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the `EAC southern extension'; \textbf{Figure \ref{fig:fig1-study-area}a}), carrying eddy-like features as a transient current (Oke, Roughan, et al., 2019).
 
@@
 
-Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV Falkor} on 15 voyages, between 2016/17--2020/21, spanning over \textasciitilde37° of latitude (\textbf{Fig.
-\ref{fig:fig1-study-area}b}; \textbf{Table S1}).
+Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV Falkor} on 15 voyages, between 2016/17--2020/21, spanning over \textasciitilde37° of latitude (\textbf{Figure \ref{fig:fig1-study-area}b}; \textbf{Table S1}).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
@@
 Thus, we built separate modelling dataset for each season.
-We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Fig.
-\ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
+We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Figure \ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \textgreater1 to 1.
@@
 sula} (25.4\%) and black-browed albatross \emph{T.
-melanophris} (25\%) (\textbf{Fig.
-S1}).
-An increase in species richness towards the south of the study region was clear in all but the winter season (\textbf{Fig.
-S2}).
-Noticeably higher counts were made around Tasmania, possibly due to the `colony effect' and/or the presence of Southern Ocean species (\textbf{Fig.
-S2}).
+melanophris} (25\%) (\textbf{Figure S1}).
+An increase in species richness towards the south of the study region was clear in all but the winter season (\textbf{Figure S2}).
+Noticeably higher counts were made around Tasmania, possibly due to the `colony effect' and/or the presence of Southern Ocean species (\textbf{Figure S2}).
 However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
@@
 
-For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (\textbf{Fig.
-S3}).
+For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (\textbf{Figure S3}).
 The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (\textbf{Table \ref{tab:table2}}).
-The best models showed a good fit with no apparent deviations from the assumptions (\textbf{Fig.
-S4}).
+The best models showed a good fit with no apparent deviations from the assumptions (\textbf{Figure S4}).
 All the best models selected `sst' or `sss', as essential for characterising water masses.
-The probability of belonging to one or another RCP varied its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer (\textbf{Fig.
-\ref{fig:fig2-partial-plots}}, \textbf{Fig.
-S5}), similar to the EAC's temperature and salinity thresholds.
-Furthermore, most of the best models selected `bat', suggesting a gradient between `coastal' and `oceanic' taxa based on bathymetry (\textbf{Fig.
-S5}).
+The probability of belonging to one or another RCP varied its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer (\textbf{Figure \ref{fig:fig2-partial-plots}}, \textbf{Figure S5}), similar to the EAC's temperature and salinity thresholds.
+Furthermore, most of the best models selected `bat', suggesting a gradient between `coastal' and `oceanic' taxa based on bathymetry (\textbf{Figure S5}).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the `climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \textasciitilde34°S, with both presence-absence and abundance models pointing to similar results (\textbf{Fig.
-\ref{fig:fig3-point-pred}}).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Fig.
-\ref{fig:fig4-prob-maps}}, \textbf{Fig.
-S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
-The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Fig.
-\ref{fig:fig3-point-pred}}, \textbf{Fig.
-S6}).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \textasciitilde34°S, with both presence-absence and abundance models pointing to similar results (\textbf{Figure \ref{fig:fig3-point-pred}}).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Figure \ref{fig:fig4-prob-maps}}, \textbf{Figure S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
+The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Figure \ref{fig:fig3-point-pred}}, \textbf{Figure S6}).
 
@@
 Although there are overlaps among the seabird groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp., lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche} spp., fairy prion \emph{P.
-turtur}) (\textbf{Fig.
-\ref{fig:fig5-spp-profiles}}, \textbf{Fig.
-S7}).
+turtur}) (\textbf{Figure \ref{fig:fig5-spp-profiles}}, \textbf{Figure S7}).
 The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the `northern' and `southern' RCPs.
@@
 The summer and autumn models reflected the high abundance of short-tailed shearwater before they depart on their trans-equatorial migration (Carey et al., 2014); the presence of the Australasian gannet was picked up by the autumn and winter models (Ismar et al., 2011); and the white-headed petrel \emph{Pterodroma lessonii} movements during chick-rearing and pre-laying exodus (Taylor et al., 2021) were detected in autumn and spring models, respectively.
-Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (\textbf{Fig.
-S1}; hence they were removed before modelling).
+Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (\textbf{Figure S1}; hence they were removed before modelling).
 Additional surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.~albatrosses \emph{Diomedea} spp.
@@
 In particular, the transition zone between assemblages (Morrone, 2023), associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We thank all observers who assisted in data collection, the crews of RV \emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
-The authors would like to thank Dr.~Skipton N.C.
-Woolley, \texttt{\{ecomix\}} developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago.
 
@@
 
+\hypertarget{data-accessibility-statement}{%
+\section*{Data Accessibility Statement}\label{data-accessibility-statement}}
+\addcontentsline{toc}{section}{Data Accessibility Statement}
+
+The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+The raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the raw data, please get in touch with XXX.
+
+\newpage
+
 \hypertarget{tables-and-figures}{%
@@
 Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S6)
+Results for other seasons and models can be found in the Supplementary Material (Figure S6)
 \item
@@
 Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material (Fig.
-S7)
+Results for other seasons and models can be found in the Supplementary Material (Figure S7)
 \end{itemize}

### 2023-10-18 ms_preprint/ms_preprint.Rmd
line numbering and a few corrections after EJWs review, and changed to correct MS title in the supp mat

@@
   **Location:** East coast of Australia.
-  **Methods:** We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+  **Methods:** We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
@@
 - \usepackage{setspace}\doublespacing
+- \usepackage[left]{lineno}
+- \newcommand{\blinenumbers}{\begin{linenumbers}}
+- \newcommand{\elinenumbers}{\end{linenumbers}}
 - \usepackage{pdflscape}
@@
 
-**Data availability statement:** The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+**Data availability statement:** The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
 The raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the raw data, please get in touch with EJW.
+If you are interested in the raw data, please get in touch with Eric J.
+Woehler.
 
-**Funding statement:** .........**Eric - any specific funding(s) we need to acknowledge?**
+**Funding statement:** Parks Australia provided funding support to participants for selected voyages.
 
@@
 \newpage
+
+\blinenumbers
 
@@
 
-**Methods:** We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+**Methods:** We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
@@
 
-The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
 The raw data are currently being used for other projects and thus can not be publicly available.
@@
 Results for other seasons and models can be found in the Supplementary Material (Figure S7)
+
+\elinenumbers
 

### 2023-10-18 ms_preprint/ms_preprint.tex
line numbering and a few corrections after EJWs review, and changed to correct MS title in the supp mat

@@
 \usepackage{setspace}\doublespacing
+\usepackage[left]{lineno}
+\newcommand{\blinenumbers}{\begin{linenumbers}}
+\newcommand{\elinenumbers}{\end{linenumbers}}
 \usepackage{pdflscape}
@@
 \textbf{Location:} East coast of Australia.
-\textbf{Methods:} We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+\textbf{Methods:} We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
@@
 
-\textbf{Data availability statement:} The gridded data used as the basis for all analyses is available in the archived repository along with the code to reproduce the results.
+\textbf{Data availability statement:} The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
 The raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the raw data, please get in touch with EJW.
-
-\textbf{Funding statement:} \ldots\ldots\ldots{}\textbf{Eric - any specific funding(s) we need to acknowledge?}
+If you are interested in the raw data, please get in touch with Eric J.
+Woehler.
+
+\textbf{Funding statement:} Parks Australia provided funding support to participants for selected voyages.
 
@@
 \newpage
+
+\begin{linenumbers}
 
@@
 
-\textbf{Methods:} We used 15 seabird at-sea ship-based survey data collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
+\textbf{Methods:} We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
 RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
@@
 \end{itemize}
+
+\end{linenumbers}
 

### 2023-10-18 ms_preprint/ms_supp_mat.Rmd
line numbering and a few corrections after EJWs review, and changed to correct MS title in the supp mat

@@
 title: "Supplementary Material"
-subtitle: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
+subtitle: "<b>Seabird assemblages are linked to the major western boundary current off eastern Australia</b>"
 # author: "Author"

### 2023-10-18 scripts/00_environment-set-up.R
tidy up comments and anonymized all scripts for peer-review

@@ -1,20 +1,36 @@
 ##
-## ASG environment set-up
+## ASG - environment set-up
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
-## This directory started with ./data (and its sub-directories) & ./scripts
-## If you haven't cloned this repo from GitHub, you need to set up some directories
+## The project started off with the directories "./data" (and its sub-directories) 
+## and "./scripts". If you haven't cloned this repo from GitHub, you need to 
+## create some directories manually
 
 # dir.create("./data_out")
 # dir.create("./EDA")
 # dir.create("./results")
 # dir.create("./results/Bernoulli")
 # dir.create("./results/NegBin")
-# dir.create("./docs")
 
-### Install needed libraries
+# The directory "./ms_preprint" was created when I manually opened an Rmarkdown
+# file using {rticle}'s template for `arXiv` preprints.
+
+## -------------------------------------------------------------------------- ##
+## -------------------------------------------------------------------------- ##
+##
+## NOTE: 
+##
+## Package versions and their dependencies were captured using {renv} and 
+## locked in the `lock.file` file
+##
+## However, below you'll find the code to install them all, although we can not
+## guarantee their versions will be the same if you don't use `renv::restore()`
+##
+## -------------------------------------------------------------------------- ##
+## -------------------------------------------------------------------------- ##
+
+## Install needed libraries
 
 # Install {pacman}, a wrapper for checking/installing packages
 install.packages("pacman")
@@ -30,8 +46,7 @@ pacman::p_install(needed_libraries)
 rm("needed_libraries")
 
 # You will also need {hadsstR} for calculating sea surface temperature gradients,
-# and {ecomix}, which currently is only available on GitHub (14 March 2023), for running RCPs
+# and {ecomix} for running RCP models, which currently (14 March 2023) is only available on GitHub
 pacman::p_install_gh("skiptoniam/ecomix@dev")
 pacman::p_install_gh("jebyrnes/hadsstR")
 
-## INCLUDE {renv} stuff here

### 2023-10-18 scripts/01_data_grooming.R

@@ -1,8 +1,7 @@
 ##
-## ASG data grooming
+## ASG - data grooming
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 ## This code joins data from two ASG's datasets, do some gross wrangling,
 ## and tidy up taxonomy (order, family, species names).
@@ -33,7 +32,7 @@ rm("fk210206")
 
 ## Gross wrangling ####
 
-## Set up right column classes, and create some useful ones *******************#
+## Set up right column classes, and create some useful ones ****************** #
 names(data)
 
 data <- 
@@ -67,7 +66,7 @@ data[numeric_cols] <- lapply(data[numeric_cols], as.numeric)
 
 rm("factor_cols", "numeric_cols")
 
-# Clean some empty columns, columns with no interest for analysis, ************#
+# Clean some empty columns, columns with no interest for analysis, *********** #
 # and missing values in species ID and geographic coordinates
 
 data <- 
@@ -81,7 +80,7 @@ data <-
   dplyr::select(- c(observer, ship_heading, 
                     sitting_on_ice_ct, in_hand_ct, bird_direction))
 
-## Good data - according to Eric **********************************************# **
+## Good data - according to Eric ********************************************* #
 
 # levels(data$voyage)
 
@@ -103,7 +102,7 @@ data <-
     voyage == "fk201228" | voyage == "fk210206") %>% 
   droplevels(.)
 
-# Replace missing values in 'total_ct' column, with the sum of ****************#
+# Replace missing values in 'total_ct' column, with the sum of *************** #
 # other counting columns
 
 # Check for NAs
@@ -123,7 +122,7 @@ plyr::count(is.na(data$total_ct)) # all good
 
 rm("cols_to_sum")
 
-# Exclude rows summing to "0" -- they mean all columns were "NA" or "0" *******#
+# Exclude rows summing to "0" -- they mean all columns were "NA" or "0" ****** #
[…]

### 2023-10-18 scripts/02_data_create-grids.R

@@ -1,8 +1,7 @@
 ##
-## ASG grids East Australia
+## ASG - grids East Australia
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 ## This code builds the East Australia grids on which data will be aggregated. 
 ## Grids are saved under "./data_out/grids_0.5-2.rda"
@@ -20,7 +19,7 @@ library(rnaturalearth)
 data <- readr::read_csv("./data_out/ASG_2016_2021_tidy.csv", 
                         show_col_types = FALSE)
 
-## *********************** Get only 'East Australia' data *********************#
+## *********************** Get only 'East Australia' data ******************** #
 east_aus <- dplyr::filter(data, longitude >= 141)
 
 rm("data")
@@ -70,38 +69,3 @@ rm("grid_sizes", "grid_size")
 ## Save it ####
 
 # save(grids, file = "./data_out/grids_0.5-2.rda")
-
-#### `sessionInfo()` ####
-
-# sessionInfo()
-
-# R version 4.2.0 (2022-04-22)
-# Platform: x86_64-pc-linux-gnu (64-bit)
-# Running under: Ubuntu 20.04.5 LTS
-# 
-# Matrix products: default
-# BLAS:   /usr/lib/x86_64-linux-gnu/blas/libblas.so.3.9.0
-# LAPACK: /usr/lib/x86_64-linux-gnu/lapack/liblapack.so.3.9.0
-# 
-# locale:
-# [1] LC_CTYPE=en_NZ.UTF-8       LC_NUMERIC=C               LC_TIME=en_NZ.UTF-8       
-# [4] LC_COLLATE=en_NZ.UTF-8     LC_MONETARY=en_NZ.UTF-8    LC_MESSAGES=en_NZ.UTF-8   
-# [7] LC_PAPER=en_NZ.UTF-8       LC_NAME=C                  LC_ADDRESS=C              
-# [10] LC_TELEPHONE=C             LC_MEASUREMENT=en_NZ.UTF-8 LC_IDENTIFICATION=C       
-# 
-# attached base packages:
-# [1] stats     graphics  grDevices utils     datasets  methods   base     
-# 
-# other attached packages:
-# [1] rnaturalearth_0.1.0 sf_1.0-8            dplyr_1.0.9         readr_2.1.2        
-# 
[…]

### 2023-10-18 scripts/03_data_extract-env-data.R

@@ -1,15 +1,14 @@
 ##
-## ASG environmental data extraction (for the full dataset)
+## ASG - environmental data extraction (for the full dataset)
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
-## This code extracts all the environmental data that will be used in future analysis.
+## This code extracts all the environmental data that will be used in future analyses.
 ## Data sources are mentioned at the beginning of each section.
 ## As I've run this in my local computer, I did each variable individually, 
 ## due to memory issues. Also, due to the same problem, I saved backups for each step 
 ## to avoid headaches.
-## You can find all backups at "./data/backup-during-env-data-extraction/".
+## I saved backups at "./data/backup-during-env-data-extraction/".
 
 ## Libraries ####
 
@@ -106,8 +105,8 @@ data_sf_envs <-
 ## ----------------------------------------------------------------------------#
 ## Bathymetry (and slope) data consumed a lot of memory from my notebook, so   #
 ## create a backup here, and just read it later on. Additionally, get rid of   # 
-## observations with 'dist_coast' equals to zero. They won't have              #
-## 'oceanographic' data anyways in the following steps                         #
+## observations with 'dist_coast' equals to zero, as they won't have           #
+## 'oceanographic' data anyway in the following steps                          #
 ## ----------------------------------------------------------------------------#
 
 readr::write_csv(
@@ -592,13 +591,13 @@ for (yr in yrs) {
   # plot(r_Grad_stack)
   # length(r_Grad_stack@layers)
   
-  ## Save cropped file ********************************************************#
+  ## Save cropped file ******************************************************* #
   raster::writeRaster(r_Grad_stack,
                       filename = paste0(stringr::str_sub(
                         file_dirs[stringr::str_detect(file_dirs, as.character(yr))], end = -9), 
                         "_SSTgrad.nc"), 
                       format = "CDF", overwrite = TRUE)
[…]

### 2023-10-18 scripts/04_data_aggregate-by-grids-and-tidy-it-up.R

@@ -1,8 +1,7 @@
 ##
-## ASG aggregate East Australia data by grids
+## ASG - aggregate East Australia data by grids
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 ## This code aggregate raw data by grid sizes/season,
 ## summarises environmental and seabird data by grid/season,
@@ -106,7 +105,8 @@ for (i in 1:length(grids_EAus)) {
 ## Labels for plot
 # how many different grids populated by season, by grid size
 txt_hist <- 
-  df_grid_summary %>% dplyr::group_by(grid, season) %>% 
+  df_grid_summary %>% 
+  dplyr::group_by(grid, season) %>% 
   dplyr::summarise(n_grids = paste0("(n = ", as.character(n_distinct(IDgrid)), ")")) %>%
   dplyr::ungroup(.) %>%
   dplyr::mutate(n_grids = paste(as.character(season), n_grids))

### 2023-10-18 scripts/05_data_EDA.R

@@ -1,10 +1,9 @@
 ##
-## ASG -- Region of Common Profile models
+## ASG - exploratory data analyses
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
-## This code makes some Exploratory Data Analysis (plots & maps)
+## This code makes some Exploratory Data Analyses (plots & maps)
 
 ## Libraries ####
 
@@ -206,7 +205,7 @@ ggsave(plot_nOCC_FO_NF,
        filename = "./results/FigS1_spp-nOCC-FO-NF-seasons.png",
        height = 14, width = 20, units = "cm", dpi = 300)
 
-## String vector w/ spp names to remove from models & how many spp/season was modelled -----------####
+## String vector w/ spp names to remove from models & how many spp/season was modelled ---------- ####
 ## based on <6 occurrences
 
 df_season_nOCC <-
@@ -247,7 +246,7 @@ for (i in 1:length(remove_spp_vec)) {
 
 save("remove_spp_vec", file = "./data_out/rm_spp-season_vec.rda")
 
-## Where the effort was -- individual records from raw data --------------------####
+## Where the effort was -- individual records from raw data ------------------- ####
 
 aus_sf <- rnaturalearth::ne_countries(country = "australia", returnclass = "sf")
 world_sf <- ne_countries(returnclass = "sf")
@@ -307,7 +306,7 @@ ggsave((ggplot() +
        filename = "./EDA/EDA_map-oceanogr.svg", ## png
        height = 10, width = 8, units = "cm", dpi = 300)
 
-## Where the effort was -- richness, by grid ------------------------------####
+## Where the effort was -- richness, by grid ----------------------------- ####
 
 effort_richness <-
   ggplot(aus_sf) + 
@@ -329,7 +328,7 @@ ggsave(effort_richness,
        filename = "./EDA/EDA_map-effort-richness-seasons.png",
        height = 10, width = 20, units = "cm", dpi = 300)
 
-## Where the effort was -- total numbers, by grid -------------------------####
+## Where the effort was -- total numbers, by grid ------------------------ ####
 
 effort_sum.all <-
   ggplot(aus_sf) + 
@@ -360,7 +359,3 @@ SuppMat_Rich.Number <-
 ggsave(SuppMat_Rich.Number, 
        filename = "./results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png",
        height = 20, width = 20, units = "cm", dpi = 300)
-
[…]

### 2023-10-18 scripts/06_data_corr-env-vars_prep-to-fit-RCPs.R

@@ -1,8 +1,7 @@
 ##
-## ASG -- Data prep for running the models
+## ASG -- data prep for running the models
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 ## This code prepares data for specifying RCPs.
 
@@ -11,7 +10,6 @@
 library(dplyr)
 library(tidyr)
 library(corrplot)
-# library(PerformanceAnalytics) # at the end, didn't use it
 
 ## Data #### 
 
@@ -103,6 +101,7 @@ for (season_vec in seasons_vec) {
   
   png(file = file_name, height = 12, width = 12, units = "cm", res = 300)
   
+  ## If plotting VIF data, uncomment `par()`, `barplot() + abline()`, and `par()` ['back to default']
   # par(mfrow = c(1, 2))
   
   corrplot::corrplot.mixed(
@@ -124,7 +123,7 @@ for (season_vec in seasons_vec) {
   rm("season_vec", "rcp_data", "file_name")
 }
 
-## Visually inspecting the plots, I've decided to remove some variables 
+## Visually inspecting the plots, I've decided to remove variables 
 ## to avoid collinearity (r > |0.7|)
 
 env_cols_models_season <- list(

### 2023-10-18 scripts/07_data_env-data-for-prediction.R

@@ -1,8 +1,7 @@
 ##
-## ASG environmental average layers and create prediction data.frame
+## ASG - average environmental layers to create prediction data.frame
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 # This code creates average ('climatic'), static layers for all environmental 
 # data by season and, finally, average values by grid (1 x 1 degree) to be 
@@ -315,7 +314,7 @@ file_dirs <- file_dirs[grepl(x = file_dirs,
 
 ## pseudo-code ---------------------------------------#
 # select one season
-# open file & # crop to 'g1' extent
+# open file & crop to 'g1' extent
 # extract mean value for each grid (remember: clim_EKEmean == eke, clim_SSTgrad == SSTgrad)
 # Scale environmental data
 # store as data.frame

### 2023-10-18 scripts/08_1_modelling_fit-RCPs-Bernoulli.R

@@ -1,11 +1,10 @@
 ##
-## ASG -- Region of Common Profile (Bernoulli)
-## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## ASG -- Region of Common Profile models (Bernoulli)
+##
+## *************************************************************************** #
 
 ## This code implements RCP models, 
-## using presence/absence data based on 'Bernoulli' as error distribution.
+## using presence/absence data based on 'Bernoulli' models.
 
 ## Libraries ####
 
@@ -53,7 +52,7 @@ spp_form <- as.formula(~ 1)
 control <- list(penalty = 0.01, penalty.tau = 10, penalty.gamma = 10,
                 penalty.disp = c(10, sqrt(10)), quiet = TRUE)
 
-## How many random starts should 'multifit' fit?
+## How many random starts should 'multifit' fit
 nstarts <- 100
 
 ## Loop through to get the results
@@ -835,7 +834,7 @@ for (season_vec in seasons_vec) {
   if(season_vec == "summer") {
     
     # As RCP group colours are being depicted from 'south' (RCP1) to 'north' (RCP2),
-    # we had to do a hack here for the probability maps visually agree with 
+    # I had to do a hack here for the probability maps visually agree with 
     # point-predictions in a meaningful way (basically, swap "RCP1" to "RCP2", and vice-versa)
     
     new_col_names <- c("IDgrid", "lon", "lat", 

### 2023-10-18 scripts/08_2_modelling_fit-RCPs-NegBin.R

@@ -1,11 +1,11 @@
 ##
-## ASG -- Region of Common Profile (NegBin)
+## ASG -- Region of Common Profile models (NegBin)
 ## 
 ## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
 ## This code implements RCP models, 
-## using count data based on 'Negative Binomial' error distribution.
+## using count data based on 'Negative Binomial' models.
 
 ## Note: To keep colour congruence between all visualizations, 
 ## for the *spring* results, I had to swap 'RCP1' / 'RCP2' labels.
@@ -54,7 +54,7 @@ spp_form <- as.formula(~ 1)
 control <- list(penalty = 0.01, penalty.tau = 10, penalty.gamma = 10,
                 penalty.disp = c(10, sqrt(10)), quiet = TRUE)
 
-## How many random starts should 'multifit' fit?
+## How many random starts should 'multifit' fit
 nstarts <- 100
 
 ## Loop through to get the results

### 2023-10-18 scripts/09_figs-patchwork.R

@@ -1,10 +1,9 @@
 ##
-## ASG -- Figures & Tables
+## ASG -- {patchwork} figures for final MS & Supp Mat
 ## 
-## Nicholas W Daudt
-## ****************************************************************************#
+## *************************************************************************** #
 
-## This code put together some Figs & tables for the manuscript.
+## This code put together the Figs for the manuscript.
 
 ## Libraries ####
 
@@ -14,7 +13,7 @@ library(magick)
 
 ## Figures main text #### 
 
-## FIGURE 1 ------------------------------------------------------------------ #
+### FIGURE 1 -------------------------------------------------------------- ####
 
 # Based on:
 # https://stackoverflow.com/questions/76079316/including-vector-graphics-file-svg-in-a-figure-generated-with-the-patchwork-pa
@@ -36,7 +35,7 @@ ggsave(fig1_study_area,
 ## The quality of this plot was not great. So I ended up using a .svg version
 ## that I put both maps together by hand in Inkscape software
 
-## FIGURE 2 ------------------------------------------------------------------ #
+### FIGURE 2 -------------------------------------------------------------- ####
 
 load("./results/Bernoulli/Bernoulli_07_RCP-point-predictions-ggplot.rda")
 # As both files were originally names the same ('map_point_pred'), 
@@ -61,7 +60,7 @@ ggsave(fig2_point_pred,
 
 rm("map_point_pred_Ber", "map_point_pred", "fig2_point_pred")
 
-## FIGURE 3 ------------------------------------------------------------------ #
+### FIGURE 3 --------------------------------------------------------------- ####
 
 load("./results/Bernoulli/Bernoulli_spring_07_RCP-prob-predictions-ggplot.rda")
 # As both files were originally names the same ('plotRCPs'), 
@@ -82,7 +81,7 @@ ggsave(fig3_prob_maps_spring,
 
 rm("plotRCPs_Ber", "plotRCPs", "fig3_prob_maps_spring")
 
-## FIGURE 4 ------------------------------------------------------------------ #
+### FIGURE 4 -------------------------------------------------------------- ####
 
 B_spr_bat <- 
   magick::image_read("./results/Bernoulli/Bernoulli_spring_07_partial-plot-bat.png") %>% 
@@ -113,7 +112,7 @@ ggsave(fig4_partial_plots_spring,
[…]

### 2023-10-18 scripts/source.R

@@ -1,8 +1,11 @@
-## Nicholas W Daudt - 2022
+##
+## ASG -- helper functions for plotting RCP outputs
+## 
+## *************************************************************************** #
 
-#### ----------------- Helper functions for plotting RCPs -------------------- #
+## Functions were all named starting with "FUN_"
 
-# FUN: max value in row to 1 and others to 0
+## FUN: max value in row to 1 and others to 0 
 FUN_max1else0 <- function(x) {
   # All values other than the max to 0
   x <- replace(x, x < max(x), 0)
@@ -114,8 +117,8 @@ FUN_prob_maps <- function(df = pred_data,
 }
 
 ## FUN: Partial plots 
-## ----- Function written by Matthew R. Schofield
-## ----- Department of Mathematics and Statistics, University of Otago
+## ----- Function written by XXXX
+
 FUN_partial_plot = function(beta.mat, 
                             predplot = c(NA,rep(0,ncol(beta.mat)-2)), 
                             xlm.real = c(-2,2), 
@@ -125,6 +128,7 @@ FUN_partial_plot = function(beta.mat,
                             ylb = "Probability", 
                             location = "topright", 
                             coluse = palette()[1:(nrow(beta.mat)+1)],
+                            # These next two args are only needed because of the colour swap explained in the modelling scripts
                             model = c("pres_abs", "abund"), 
                             season = c("summer", "autumn", "winter", "spring")){
   ## beta.mat is a matrix of beta values.  This is obtained from coef(model)

### 2023-10-19 ms_preprint/ms_preprint.Rmd
added Wills corrections to the ms

@@
 
-William J. Rayment -- **do you have ORCID, Will?**
+William J. Rayment -- <https://orcid.org/0000-0001-9558-5109>
 
@@
 The delimitation of bioregions is made through identifying sets of species that are endemic or have high levels of co-occurrence, and are distinct enough from other regions.
-In this sense, bioregions are interpreted as a natural representations of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010].
+In this sense, bioregions are interpreted as natural representations of common historical and/or ecological processes shaped by physical and biological forces [@kreft2010].
 In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space, and relating these regions to environmental characteristics.
@@
 Considering the hierarchical nature of oceanographic processes and their resulting features [@haury1978patterns], it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales [@haury1978patterns; @hunt1987].
-In the marine environment, large scale (\>100 km length scale) oceanographic features (e.g.
+In the marine environment, large scale (\>100 km length) oceanographic features (e.g.
 boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.
@@
 In contrast, there are few studies at large spatial and temporal scales off eastern Australia.
-@reid2000 and @evans2021a analysed seabird assemblages and their relationships with oceanographic and prey data at coarse scales (tens of km) off southeast Tasmania.
+@reid2000 and @evans2021a analysed seabird assemblages and their relationships with oceanographic and prey data at mesoscale off southeast Tasmania.
 The remaining studies are mostly descriptive accounts locally [@norman1992distribution; @blaber1994distribution], or focused on a single species [@priddel2014].
@@
 The East Australian Current (EAC) is the major western boundary current in the South Pacific Ocean, and the dominant boundary current off eastern Australia [@ganachaud2014].
-The EAC has been extensively studied, although research into the links between the EAC's oceanography and marine predators are not well known.
+The EAC has been extensively studied, although the links between the EAC's oceanography and marine predators are not well known.
 The dynamic features associated with the EAC, such as mesoscale eddies and upwelling, are likely to influence marine biota [@suthers2011; @scales2014].
@@
 Top predators, for instance, may change their core habitats [@hazen2013a; @reisinger2022], with unprecedented population consequences [@woehler2023].
-For seabirds, @péron2010 and @sojitra2022 showed that sea surface temperature anomalies affected the occurrence and abundances of albatrosses and petrels species with latitude, thus affecting assemblage structure.
+For seabirds, @péron2010 and @sojitra2022 showed that sea surface temperature anomalies affected the occurrence and abundances of albatross and petrel species with latitude, thus affecting assemblage structure.
 In eastern Australia, changing EAC conditions have affected a range of marine species.
@@
 
-In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
+In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea survey data and a state-of-the-art modelling technique.
 We use the Region of Common Profiles [RCP; @foster2013], a mixture model regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
@@
 In addition, presence-absence and abundance data were used to test if the different numerical resolutions would result in the identification of similar assemblages.
-This study provides a repeatable framework on which future studies can compare to assess how future changes in the EAC may influence the distribution of seabird assemblages.
+This study provides a repeatable framework on which future studies can assess how future changes in the EAC may influence the distribution of seabird assemblages.
 
@@
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
-The EAC flow polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
+The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
 Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure \@ref(fig:fig1-study-area)a**), carrying eddy-like features as a transient current [@oke2019].
@@
 @reid2023], and in the probabilities of occurrence and in total numbers.
-Thus, we built separate modelling dataset for each season.
-We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (c.
-110 $\times$ 110 km) and season (**Figure \@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
+Thus, we built separate modelling datasets for each season.
+We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Figure \@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 RCP is a one-stage bioregionalization approach, as it identifies bioregions by specifying a statistical model that simultaneously relates response (multi-species data) and predictor (environmental data) [@woolley2020; @hill2020].
-Given a set of environmental variables, the model can predict assemblage probabilities of sites where no biotic (observational) data exist.
+Given a set of environmental variables, the model can predict assemblage probabilities for sites where no biotic (observational) data exist.
 As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals [@warton2015; @hill2020].
@@
 There are $K$ assemblages (i.e.
-RCPs) that are common across the study area, $K = (1, ..., k)$.
+RCPs), $K = (1, ..., k)$, that are common across the study area.
 The outcome variable is a vector $y_{i} = (y_{i1}, ...
-y_{iS})$, that give the count or presence-absence of each of the $S$ species at site $i$.
+y_{iS})$, that gives the count or presence-absence of each of the $S$ species at site $i$.
 Depending on whether we model the counts or presence-absence data, the assemblages are defined in terms of the expected frequency or relative frequency of the species, respectively.
@@
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) [@foster2013].
-However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the $K$ assemblage [@foster2013].
+However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the $K$ assemblages [@foster2013].
 As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
@@
 averaged values of environmental variables in each cell).
-Presence-absence data were modelled based on a Bernoulli model, whereas abundance data based on a Negative Binomial model.
-RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($K$) that describes the data, (ii) re-fit the model by fixing $K$ and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a Negative Binomial model.
+RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($K$) that describes the data, (ii) re-fit the model by fixing $K$ and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
 
@@
 For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Figure S3**).
-The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (**Table \@ref(tab:table2)**).
+The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (**Table \@ref(tab:table2)**).
 The best models showed a good fit with no apparent deviations from the assumptions (**Figure S4**).
 All the best models selected 'sst' or 'sss', as essential for characterising water masses.
-The probability of belonging to one or another RCP varied its 'sst' threshold between 18--21 °C, or 'sss' 35.1 PSU for summer (**Figure \@ref(fig:fig2-partial-plots)**, **Figure S5**), similar to the EAC's temperature and salinity thresholds.
+The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Figure \@ref(fig:fig2-partial-plots)**, **Figure S5**), similar to the characteristics used to to define the EAC.
 Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Figure S5**).
@@
 These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Figure \@ref(fig:fig4-prob-maps)**, **Figure S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
-The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Straight' assemblage appears (**Figure \@ref(fig:fig3-point-pred)**, **Figure S6**).
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Figure \@ref(fig:fig3-point-pred)**, **Figure S6**).
 
@@
 turtur*) (**Figure \@ref(fig:fig5-spp-profiles)**, **Figure S7**).
-The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
-However, the intermediate 'Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross and Australasian gannet *Morus serrator* prominent with high probabilities of occurrence.
+The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
+However, the intermediate 'Tasmania/Bass Strait' assemblage showed a mixture of both groups with shy albatross and Australasian gannet *Morus serrator* prominent with high probabilities of occurrence.
 
@@
 It is worth noting that the Longhurst Provinces approach was updated to include SST, SSS and bathymetry (the same variables retained by our best RCP models) and a temporal component to the classification [@reygondeau2013].
-The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage at EAC region includes several long-distance migratory species.
+The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage in the EAC region includes several long-distance migratory species.
 Australia's bioregionalization has a long story of disagreements [@ebach2012history], however, the inherent link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
 
-## Fitting RCP models to high mobile animals
+## Fitting RCP models to highly mobile animals
 
@@
 While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009], they fail to perceive how species assemblages interact with the environment and each other.
-Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
+Therefore, using at-sea counts of seabirds is a well established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
 Further, the spatial scales on which assemblages relate to their environment is variable and species-specific [@hunt1987; @ribic1997; @scales2017].
@@
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere [@harris2007; @santora2013; @sojitra2022].
-Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to had been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
+Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to have been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
 
@@
 The added information contained in abundance data make them inherently more powerful than binary presence-absence data.
-However, the use of different data inputs in multivariate, biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
+However, the use of different data inputs in multivariate biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
 At larger spatial and temporal scales, presence-absence data describe similar patterns to abundance data [@anderson2005], as we found.
@@
 Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore 'new bioregions') likely nested within the bioregions described at large scales.
-Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect not exist at present.
+Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect do not exist at present.
 
@@
 Southern Ocean albatrosses off southeast Australia, for instance, have shown 'early stages of a southward range shift' due to environmental variability [@sojitra2022].
-Resident species such as the little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher [@carroll2016; @peck2004].
-Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species' profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
+Resident species such as little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher [@carroll2016; @peck2004].
+Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
 In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
@@
 (2019); for colour references see the online version
--   **FIGURE 2.** Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+-   **FIGURE 2.** Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
 The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
-Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
+Note that, in case of two RCPs, RCP2 probability equals (1 - RCP1 probability).
 Results for other seasons and models can be found in the Supplementary Material (Figure S6)

### 2023-10-19 ms_preprint/ms_preprint.tex
added Wills corrections to the ms

@@
 
-William J. Rayment -- \textbf{do you have ORCID, Will?}
+William J. Rayment -- \url{https://orcid.org/0000-0001-9558-5109}
 
@@
 The delimitation of bioregions is made through identifying sets of species that are endemic or have high levels of co-occurrence, and are distinct enough from other regions.
-In this sense, bioregions are interpreted as a natural representations of common historical and/or ecological processes shaped by physical and biological forces (Kreft \& Jetz, 2010).
+In this sense, bioregions are interpreted as natural representations of common historical and/or ecological processes shaped by physical and biological forces (Kreft \& Jetz, 2010).
 In particular, the identification of bioregions allows representation of the distribution of multiple species at a given time and space, and relating these regions to environmental characteristics.
@@
 Considering the hierarchical nature of oceanographic processes and their resulting features (Haury et al., 1978), it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales (Haury et al., 1978; Hunt \& Schneider, 1987).
-In the marine environment, large scale (\textgreater100 km length scale) oceanographic features (e.g.~boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.~the occurrence of species and their relative numerical composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al., 1997).
+In the marine environment, large scale (\textgreater100 km length) oceanographic features (e.g.~boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.~the occurrence of species and their relative numerical composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al., 1997).
 In contrast, local abundance is related to meso- (10--100 km) and sub-mesoscale (1--10 km) processes that aggregate prey and therefore predators, such as fronts and eddies (Scales, Miller, Hawkes, et al., 2014; Weimerskirch, 2007).
@@
 Reid \& Hindell (2000) and Evans et al.
-(2021) analysed seabird assemblages and their relationships with oceanographic and prey data at coarse scales (tens of km) off southeast Tasmania.
+(2021) analysed seabird assemblages and their relationships with oceanographic and prey data at mesoscale off southeast Tasmania.
 The remaining studies are mostly descriptive accounts locally (Blaber \& Milton, 1994; Norman, 1992), or focused on a single species (Priddel et al., 2014).
@@
 The East Australian Current (EAC) is the major western boundary current in the South Pacific Ocean, and the dominant boundary current off eastern Australia (Ganachaud et al., 2014).
-The EAC has been extensively studied, although research into the links between the EAC's oceanography and marine predators are not well known.
+The EAC has been extensively studied, although the links between the EAC's oceanography and marine predators are not well known.
 The dynamic features associated with the EAC, such as mesoscale eddies and upwelling, are likely to influence marine biota (Scales, Miller, Hawkes, et al., 2014; Suthers et al., 2011).
@@
 (2010) and Sojitra et al.
-(2022) showed that sea surface temperature anomalies affected the occurrence and abundances of albatrosses and petrels species with latitude, thus affecting assemblage structure.
+(2022) showed that sea surface temperature anomalies affected the occurrence and abundances of albatross and petrel species with latitude, thus affecting assemblage structure.
 In eastern Australia, changing EAC conditions have affected a range of marine species.
@@
 
-In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea surveys data and a state-of-the-art modelling technique.
+In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea survey data and a state-of-the-art modelling technique.
 We use the Region of Common Profiles {[}RCP; Foster et al.
@@
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019; Oke \& Middleton, 2000; Ridgway \& Dunn, 2003).
-The EAC flow polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension' (Oke, Pilo, et al.
+The EAC flows polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension' (Oke, Pilo, et al.
 (2019); often termed the `Tasman Front').
@@
 Reid et al., 2023), and in the probabilities of occurrence and in total numbers.
-Thus, we built separate modelling dataset for each season.
-We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Figure \ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
+Thus, we built separate modelling datasets for each season.
+We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (\emph{c.} 110 \(\times\) 110 km) and season (\textbf{Figure \ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \textgreater1 to 1.
@@
 Woolley et al., 2020).
-Given a set of environmental variables, the model can predict assemblage probabilities of sites where no biotic (observational) data exist.
+Given a set of environmental variables, the model can predict assemblage probabilities for sites where no biotic (observational) data exist.
 As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals (N.
@@
 The model is specified in terms of assemblages and site-specific probabilities.
-There are \(K\) assemblages (i.e.~RCPs) that are common across the study area, \(K = (1, ..., k)\).
+There are \(K\) assemblages (i.e.~RCPs), \(K = (1, ..., k)\), that are common across the study area.
 The outcome variable is a vector \(y_{i} = (y_{i1}, ...
-y_{iS})\), that give the count or presence-absence of each of the \(S\) species at site \(i\).
+y_{iS})\), that gives the count or presence-absence of each of the \(S\) species at site \(i\).
 Depending on whether we model the counts or presence-absence data, the assemblages are defined in terms of the expected frequency or relative frequency of the species, respectively.
@@
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) (Foster et al., 2013).
-However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the \(K\) assemblage (Foster et al., 2013).
+However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the \(K\) assemblages (Foster et al., 2013).
 As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
@@
 RCP models were used to relate the response variables (i.e.~presence-absence or abundance of each seabird species in each grid cell) to the independent predictor covariates (i.e.~averaged values of environmental variables in each cell).
-Presence-absence data were modelled based on a Bernoulli model, whereas abundance data based on a Negative Binomial model.
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a Negative Binomial model.
 RCPs were fitted using the `ecomix' 1.0.0 package (S.
-Woolley et al., 2022), with the following workflow: (i) find the best number of groups (\(K\)) that describes the data, (ii) re-fit the model by fixing \(K\) and selecting optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+Woolley et al., 2022), with the following workflow: (i) find the best number of groups (\(K\)) that describes the data, (ii) re-fit the model by fixing \(K\) and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
 
@@
 For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (\textbf{Figure S3}).
-The best fitting models generally had 1 or 2 covariates selected, although the abundance-autumn model had five covariates (\textbf{Table \ref{tab:table2}}).
+The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (\textbf{Table \ref{tab:table2}}).
 The best models showed a good fit with no apparent deviations from the assumptions (\textbf{Figure S4}).
 All the best models selected `sst' or `sss', as essential for characterising water masses.
-The probability of belonging to one or another RCP varied its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer (\textbf{Figure \ref{fig:fig2-partial-plots}}, \textbf{Figure S5}), similar to the EAC's temperature and salinity thresholds.
+The thresholds for defining RCP membership were 18--21°C for `sst' and 35.1 PSU for `sss' (\textbf{Figure \ref{fig:fig2-partial-plots}}, \textbf{Figure S5}), similar to the characteristics used to to define the EAC.
 Furthermore, most of the best models selected `bat', suggesting a gradient between `coastal' and `oceanic' taxa based on bathymetry (\textbf{Figure S5}).
@@
 These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Figure \ref{fig:fig4-prob-maps}}, \textbf{Figure S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
-The presence-absence-autumn model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Figure \ref{fig:fig3-point-pred}}, \textbf{Figure S6}).
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Strait' assemblage appears (\textbf{Figure \ref{fig:fig3-point-pred}}, \textbf{Figure S6}).
 
@@
 turtur}) (\textbf{Figure \ref{fig:fig5-spp-profiles}}, \textbf{Figure S7}).
-The presence-absence-autumn model, which had three RCPs, also showed the characteristic taxa representing the `northern' and `southern' RCPs.
-However, the intermediate `Tasmania/Bass Straight' assemblage showed a mixture of both groups with shy albatross and Australasian gannet \emph{Morus serrator} prominent with high probabilities of occurrence.
+The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the `northern' and `southern' RCPs.
+However, the intermediate `Tasmania/Bass Strait' assemblage showed a mixture of both groups with shy albatross and Australasian gannet \emph{Morus serrator} prominent with high probabilities of occurrence.
 
@@
 It is worth noting that the Longhurst Provinces approach was updated to include SST, SSS and bathymetry (the same variables retained by our best RCP models) and a temporal component to the classification (Reygondeau et al., 2013).
-The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage at EAC region includes several long-distance migratory species.
+The updated provinces are more similar to the seabird assemblages described here, although still not a perfect match, which could be partially explained because the seabird assemblage in the EAC region includes several long-distance migratory species.
 Australia's bioregionalization has a long story of disagreements (Ebach, 2012), however, the inherent link that RCPs develop between environment and biological data can be a strong tool to support marine bioregionalization.
 
-\hypertarget{fitting-rcp-models-to-high-mobile-animals}{%
-\subsection{Fitting RCP models to high mobile animals}\label{fitting-rcp-models-to-high-mobile-animals}}
+\hypertarget{fitting-rcp-models-to-highly-mobile-animals}{%
+\subsection{Fitting RCP models to highly mobile animals}\label{fitting-rcp-models-to-highly-mobile-animals}}
 
@@
 While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales (Tremblay et al., 2009), they fail to perceive how species assemblages interact with the environment and each other.
-Therefore, using at-sea counts of seabirds is a well-established method to capture these relationships (Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
+Therefore, using at-sea counts of seabirds is a well established method to capture these relationships (Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
 Further, the spatial scales on which assemblages relate to their environment is variable and species-specific (Hunt \& Schneider, 1987; Ribic et al., 1997; Scales et al., 2017).
@@
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere (Harris et al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022).
-Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to had been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
+Therefore, for describing seabird assemblages at a macro-scale, the chosen resolution seems to have been a good fit, accords with studies elsewhere, and the resultant models described patterns of seabird species' composition consistent with our current understanding of species' ecologies and life history strategies.
 
@@
 The added information contained in abundance data make them inherently more powerful than binary presence-absence data.
-However, the use of different data inputs in multivariate, biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
+However, the use of different data inputs in multivariate biodiversity assessments and assemblage descriptions has shown contrasting patterns at different spatial scales.
 At larger spatial and temporal scales, presence-absence data describe similar patterns to abundance data (Anderson et al., 2005), as we found.
@@
 Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore `new bioregions') likely nested within the bioregions described at large scales.
-Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect not exist at present.
+Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect do not exist at present.
 
@@
 Southern Ocean albatrosses off southeast Australia, for instance, have shown `early stages of a southward range shift' due to environmental variability (Sojitra et al., 2022).
-Resident species such as the little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher (Carroll et al., 2016; Peck et al., 2004).
-Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species' profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
+Resident species such as little penguins and wedge-tailed shearwater change their behaviour and have decreased foraging success when SST is higher (Carroll et al., 2016; Peck et al., 2004).
+Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
 In particular, the transition zone between assemblages (Morrone, 2023), associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
@@
 \item
-  \textbf{FIGURE 2.} Partial plots for the retained covariates in the spring (a) presence-absence and (b) abundance best models.
+  \textbf{FIGURE 2.} Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
 The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 The central column, `mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
-Note that, in case of two RCPs, RCP2 probablility equals to (1 - RCP1 probability).
+Note that, in case of two RCPs, RCP2 probability equals (1 - RCP1 probability).
 Results for other seasons and models can be found in the Supplementary Material (Figure S6)

### 2023-10-20 ms_preprint/ms_preprint.Rmd
added Wills corrections to the ms

@@
 
-***
+------------------------------------------------------------------------
 
@@
 
+```{=tex}
 \begin{center}
@@
 \end{center}
-
+```
 *Running title:* Seabird assemblages off Eastern Australia
 
-***
+------------------------------------------------------------------------
 
@@
 
-**Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+**Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages ('northern' and 'southern'), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\~18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
-The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Figure \@ref(fig:fig1-study-area)a**; @kessler2013; @ganachaud2014).
+The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Figure** \@ref(fig:fig1-study-area)**a**; @kessler2013; @ganachaud2014).
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
 The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure \@ref(fig:fig1-study-area)a**), carrying eddy-like features as a transient current [@oke2019].
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure** \@ref(fig:fig1-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
 
@@
 
-Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Figure \@ref(fig:fig1-study-area)b**; **Table S1**).
+Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Figure** \@ref(fig:fig1-study-area)**b**; **Table S1**).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
@@
 To investigate and predict species' distributions and relationships with environmental variables, a considered choice of broad-scale climatic and oceanographic covariates is important [@barbet-massin2014].
-We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table \@ref(tab:table1)**).
+We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (**Table** \@ref(tab:table1)).
 These variables are the most frequently used in seabird-environment relationship studies [@tremblay2009], capturing processes at a range of spatial and temporal scales from macro-scale water-mass properties to prey availability in meso- to submesoscale oceanographic features [@haury1978patterns; @weimerskirch2007; @scales2014].
@@
 Thus, 'eke' and 'sst_grad' environmental layers were calculated before extracting environmental data for each seabird record.
-For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table \@ref(tab:table1)**) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
+For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table** \@ref(tab:table1)) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
 Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ [@sun2022].
 The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
-In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table \@ref(tab:table1)**).
+In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table** \@ref(tab:table1)).
 Seabed slope was calculated using 'raster::terrain()' [@raster], and distance from the coast was calculated through 'sf::st_distance()' [@sf] using Australia and Tasmania as the mainland (i.e.
@@
 Thus, we built separate modelling datasets for each season.
-We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Figure \@ref(fig:fig1-study-area)b**), by summing seabirds from the same species.
+We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Figure** \@ref(fig:fig1-study-area)**b**), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Figure S3**).
-The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (**Table \@ref(tab:table2)**).
+The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (**Table** \@ref(tab:table2)).
 The best models showed a good fit with no apparent deviations from the assumptions (**Figure S4**).
 All the best models selected 'sst' or 'sss', as essential for characterising water masses.
-The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Figure \@ref(fig:fig2-partial-plots)**, **Figure S5**), similar to the characteristics used to to define the EAC.
+The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Figure** \@ref(fig:fig2-partial-plots), **Figure S5**), similar to the characteristics used to to define the EAC.
 Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Figure S5**).
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Figure \@ref(fig:fig3-point-pred)**).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Figure \@ref(fig:fig4-prob-maps)**, **Figure S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
-The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Figure \@ref(fig:fig3-point-pred)**, **Figure S6**).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Figure** \@ref(fig:fig3-point-pred)).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Figure** \@ref(fig:fig4-prob-maps), **Figure S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Figure** \@ref(fig:fig3-point-pred), **Figure S6**).
 
@@
 small albatrosses *Thalassarche* spp., fairy prion *P.
-turtur*) (**Figure \@ref(fig:fig5-spp-profiles)**, **Figure S7**).
+turtur*) (**Figure** \@ref(fig:fig5-spp-profiles), **Figure S7**).
 The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.

### 2023-10-20 scripts/00_environment-set-up.R
update script '00' with 'renv::restore'

@@ -3,18 +3,8 @@
 ## 
 ## *************************************************************************** #
 
-## The project started off with the directories "./data" (and its sub-directories) 
-## and "./scripts". If you haven't cloned this repo from GitHub, you need to 
-## create some directories manually
-
-# dir.create("./data_out")
-# dir.create("./EDA")
-# dir.create("./results")
-# dir.create("./results/Bernoulli")
-# dir.create("./results/NegBin")
-
-# The directory "./ms_preprint" was created when I manually opened an Rmarkdown
-# file using {rticle}'s template for `arXiv` preprints.
+# install.packages("renv") # uncomment if you need to install {renv}
+renv::restore()
 
 ## -------------------------------------------------------------------------- ##
 ## -------------------------------------------------------------------------- ##
@@ -30,7 +20,21 @@
 ## -------------------------------------------------------------------------- ##
 ## -------------------------------------------------------------------------- ##
 
-## Install needed libraries
+## The project started off with the directories "./data" (and its sub-directories) 
+## and "./scripts". If you haven't cloned this repo from GitHub, you need to 
+## create some directories manually
+
+# dir.create("./data_out")
+# dir.create("./EDA")
+# dir.create("./results")
+# dir.create("./results/Bernoulli")
+# dir.create("./results/NegBin")
+
+# The directory "./ms_preprint" was created when I manually opened an Rmarkdown
+# file using {rticle}'s template for `arXiv` preprints.
+
+# ---------------------------------------------------------------------------- #
+## Install needed libraries -------------------------------------------------- #
 
 # Install {pacman}, a wrapper for checking/installing packages
 install.packages("pacman")

### 2023-10-24 ms_preprint/ms_preprint.Rmd
ms adjustments after LB RS and MS reviews pre-submission

@@
 authors:
-  - name: Nicholas W. Daudt*
+  - name: Nicholas W. Daudt
     department: Department of Marine Science
@@
     email: nicholaswdaudt@gmail.com
-    notes: Corresponding author
+    thanks: Corresponding author
   - name: Eric J. Woehler
@@
     department: Institute of Biological Sciences
-    affiliation: Universidade Federal do Rio Grande--FURG
+    affiliation: Universidade Federal do Rio Grande---FURG
     location: Rio Grande, RS, Brazil
@@
 These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-  **Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+  **Results:** Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels,  prions, shearwaters, boobies, and terns, among other taxa.
+All models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
@@
   **Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 keywords:
@@
 - \usepackage{pdflscape}
-- \DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}} # needed to understand unicode for greek Delta in Table 1
+- \DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}} # Needed to understand unicode for greek Delta in Table 1
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
+# As Table 1 was build enterely within this .Rmd file, I need the following hack to make sure the reference shows up in the References section:
+nocite: |
+  @huang2021
 ---
 
-<!-- Running title: Seabird assemblages off Eastern Australia -->
-
-<!-- Note: I've changed the 'arxiv.sty' source file to include the running title in the PDF -->
+<!-- Note: I've changed the 'arxiv.sty' source file to include the running title in the PDF, and added commands in the YAML header too -->
 
@@
 
-William J. Rayment -- <https://orcid.org/0000-0001-9558-5109>
+William J. Rayment <https://orcid.org/0000-0001-9558-5109>
 
@@
 Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago.
+NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
 
@@
 **Data availability statement:** The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
-The raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the raw data, please get in touch with Eric J.
+The seabird raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the seabird raw data, please contact Eric J.
 Woehler.
+The environmental data used in this study are freely available from the source (see Table 1 in the main text).
 
@@
 
-**Results:** Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages ('northern' and 'southern'), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at \~34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+**Results:** Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels,  prions, shearwaters, boobies, and terns, among other taxa.
+All models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \~34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
@@
 **Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 
@@
 Considering the hierarchical nature of oceanographic processes and their resulting features [@haury1978patterns], it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales [@haury1978patterns; @hunt1987].
-In the marine environment, large scale (\>100 km length) oceanographic features (e.g.
-boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.
+In the marine environment, large-scale oceanographic features (e.g.
+boundary currents and water masses) act similarly to terrestrial biomes (based on vegetation), and influence seabird assemblage structure---i.e.
 the occurrence of species and their relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014].
@@
 In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea survey data and a state-of-the-art modelling technique.
-We use the Region of Common Profiles [RCP; @foster2013], a mixture model regarded as one of the best available tools for describing bioregions [@hill2020; @woolley2020].
+We use the Region of Common Profiles -- RCP [@foster2013], a mixture model capable of describing bioregions [@hill2020; @woolley2020].
 RCP is a model-based approach, and has a range of advantages over traditional algorithm-based statistics [@warton2015; @hill2020].
@@
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
-The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension' (@oke2019a; often termed the 'Tasman Front').
+The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension', often termed the 'Tasman Front' [@oke2019a].
 Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure** \@ref(fig:fig1-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
@@
 Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material.
-Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in **(Zenodo repo)**.
+Code and the modelling dataset (see Section \ref{sec:modellingdatasets}) are available in **(Zenodo)**.
 
@@
 For both 'eke' and 'sst_grad', we used their original cell spatial resolution (**Table** \@ref(tab:table1)) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
-Eddy kinetic energy was calculated as $0.5 (u'² + v'²)$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ [@sun2022].
-The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS² + WE²$ [@burrows2011].
+Eddy kinetic energy was calculated as $0.5 (u'^{2} + v'^{2})$, where $u' = u - \bar{u}$ and $v' = v - \bar{v}$ are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as $\bar{u}$ and $\bar{v}$ [@sun2022].
+The spatial 'sst_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as $NS^{2} + WE^{2}$ [@burrows2011].
 In addition, we calculated climatic seasonal average layers for 'eke' and 'sst_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the 'eke' standard deviation over the same period (named with the prefix 'clim\_' in **Table** \@ref(tab:table1)).
@@
 Given a set of environmental variables, the model can predict assemblage probabilities for sites where no biotic (observational) data exist.
-As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals [@warton2015; @hill2020].
+As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing residuals [@warton2015; @hill2020].
 
@@
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) [@foster2013].
-However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the $K$ assemblages [@foster2013].
-As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
+RCPs can be directly interpreted as assemblages; the assemblage structure is given by the species profiles.
 
@@
 Due to seabird migration and dispersion, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
-In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
+In light of the difference among seasons in the numbers of species that met our occurrence threshold (see Section \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
 RCP models were used to relate the response variables (i.e.
@@
 averaged values of environmental variables in each cell).
-Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a Negative Binomial model.
-RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($K$) that describes the data, (ii) re-fit the model by fixing $K$ and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a negative binomial model.
+RCPs were fitted using the 'ecomix' 1.0.0 package [@ecomix], with the following workflow: (i) find the best number of groups ($K$) that describes the data, (ii) re-fit the model by fixing $K$ and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty of parameter estimates, and (v) produce results.
 
@@
 Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Figure S2**).
-However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
+However, in summer, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania.
 
@@
 The best models showed a good fit with no apparent deviations from the assumptions (**Figure S4**).
-All the best models selected 'sst' or 'sss', as essential for characterising water masses.
+All the best models selected sea surface temperature ('sst') or salinity ('sss'), as essential for characterising water masses.
 The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Figure** \@ref(fig:fig2-partial-plots), **Figure S5**), similar to the characteristics used to to define the EAC.
-Furthermore, most of the best models selected 'bat', suggesting a gradient between 'coastal' and 'oceanic' taxa based on bathymetry (**Figure S5**).
+Furthermore, most of the best models selected bathymetry ('bat'), suggesting a gradient between 'coastal' and 'oceanic' taxa (**Figure S5**).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
@@
 Our analyses clearly revealed a 'northern' and a 'southern' assemblage of seabirds off eastern Australia within the EAC.
-There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \~34°S, near the average separation latitude (\~32°S) of the EAC and its eastward extension into the Tasman Sea [@oke2019].
+There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \~34°S, near the average separation latitude (\~32°S; 25--38°S) of the EAC and its eastward extension into the Tasman Sea [@oke2019].
 However, the 'northern/southern' assemblage pattern identified herein does not agree with previously-proposed large-scale regionalizations, such as the pelagic Longhurst Provinces [@longhurst2007], the Marine Ecoregions of the World [MEOW; @spalding2007], or the Large Marine Ecosystems [LME; @sherman1994].
@@
 Understanding the biology of the studied group is thus essential to set the appropriate scales of analyses.
-Our choice of 1° latitude $\times$ 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \ref{sec:limitations}).
+Our choice of 1° latitude $\times$ 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also Section \ref{sec:limitations}).
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere [@harris2007; @santora2013; @sojitra2022].
@@
 In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
-Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Fig S8**).
+Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Figure S8**).
 
@@
 The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
-The raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the raw data, please get in touch with XXX.
+The seabird raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the seabird raw data, please contact XXX.
+The environmental data used in this study are freely available from the source (see Table 1 in the main text).
 

### 2023-10-24 ms_preprint/ms_preprint.tex
ms adjustments after LB RS and MS reviews pre-submission

@@
 \author{
-    Nicholas W. Daudt*
+    Nicholas W. Daudt
+    \thanks{Corresponding author}
    \\
@@
     Institute of Biological Sciences \\
-    Universidade Federal do Rio Grande--FURG \\
+    Universidade Federal do Rio Grande---FURG \\
   Rio Grande, RS, Brazil \\
@@
 These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-\textbf{Results:} Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+\textbf{Results:} Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels, prions, shearwaters, boobies, and terns, among other taxa.
+All models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
@@
 \textbf{Main conclusions:} RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 \end{abstract}
@@
 
-William J. Rayment -- \url{https://orcid.org/0000-0001-9558-5109}
+William J. Rayment \url{https://orcid.org/0000-0001-9558-5109}
 
@@
 Woolley, \texttt{\{ecomix\}} developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago.
+NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
 
@@
 \textbf{Data availability statement:} The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
-The raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the raw data, please get in touch with Eric J.
+The seabird raw data are currently being used for other projects and thus can not be publicly available.
+If you are interested in the seabird raw data, please contact Eric J.
 Woehler.
+The environmental data used in this study are freely available from the source (see Table 1 in the main text).
 
@@
 
-\textbf{Results:} Based on 142,646 seabirds recorded, from 80 species, all models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current separates from the Australian continental slope.
+\textbf{Results:} Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels, prions, shearwaters, boobies, and terns, among other taxa.
+All models suggested two macro-scale assemblages (`northern' and `southern'), except for the autumn presence-absence model that identified three groups.
+The model results consistently show a biogeographic transition at \textasciitilde34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
@@
 \textbf{Main conclusions:} RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given the rapid climatic changes that the East Australian Current is experiencing and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
 
@@
 Considering the hierarchical nature of oceanographic processes and their resulting features (Haury et al., 1978), it is not surprising that the distribution of highly mobile seabirds are linked differently across temporal and spatial scales (Haury et al., 1978; Hunt \& Schneider, 1987).
-In the marine environment, large scale (\textgreater100 km length) oceanographic features (e.g.~boundary currents and water masses) act as terrestrial biomes (based on vegetation), and influence seabird assemblage structure--i.e.~the occurrence of species and their relative numerical composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al., 1997).
+In the marine environment, large-scale oceanographic features (e.g.~boundary currents and water masses) act similarly to terrestrial biomes (based on vegetation), and influence seabird assemblage structure---i.e.~the occurrence of species and their relative numerical composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al., 1997).
 In contrast, local abundance is related to meso- (10--100 km) and sub-mesoscale (1--10 km) processes that aggregate prey and therefore predators, such as fronts and eddies (Scales, Miller, Hawkes, et al., 2014; Weimerskirch, 2007).
@@
 In this study, we aim to identify assemblages of seabird species within the EAC and adjacent areas, and relate these assemblages to environmental conditions, by using at-sea survey data and a state-of-the-art modelling technique.
-We use the Region of Common Profiles {[}RCP; Foster et al.
-(2013){]}, a mixture model regarded as one of the best available tools for describing bioregions (N.
+We use the Region of Common Profiles -- RCP (Foster et al., 2013), a mixture model capable of describing bioregions (N.
 A.
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte, 2013).
-The South Caledonian Jet completes the EAC system at \textasciitilde22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (\textbf{Figure \ref{fig:fig1-study-area}a}; Kessler \& Cravatte (2013); Ganachaud et al.
+The South Caledonian Jet completes the EAC system at \textasciitilde22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (\textbf{Figure} \ref{fig:fig1-study-area}\textbf{a}; Kessler \& Cravatte (2013); Ganachaud et al.
 (2014)).
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019; Oke \& Middleton, 2000; Ridgway \& Dunn, 2003).
-The EAC flows polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension' (Oke, Pilo, et al.
-(2019); often termed the `Tasman Front').
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the `EAC southern extension'; \textbf{Figure \ref{fig:fig1-study-area}a}), carrying eddy-like features as a transient current (Oke, Roughan, et al., 2019).
+The EAC flows polewards and is primarily associated with the shelf break, but at \textasciitilde32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea (Mata et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension', often termed the `Tasman Front' (Oke, Pilo, et al., 2019).
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the `EAC southern extension'; \textbf{Figure} \ref{fig:fig1-study-area}\textbf{a}), carrying eddy-like features as a transient current (Oke, Roughan, et al., 2019).
 
@@
 
-Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV Falkor} on 15 voyages, between 2016/17--2020/21, spanning over \textasciitilde37° of latitude (\textbf{Figure \ref{fig:fig1-study-area}b}; \textbf{Table S1}).
+Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV Falkor} on 15 voyages, between 2016/17--2020/21, spanning over \textasciitilde37° of latitude (\textbf{Figure} \ref{fig:fig1-study-area}\textbf{b}; \textbf{Table S1}).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
@@
 To investigate and predict species' distributions and relationships with environmental variables, a considered choice of broad-scale climatic and oceanographic covariates is important (Barbet-Massin \& Jetz, 2014).
-We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (\textbf{Table \ref{tab:table1}}).
+We used three physiographic variables and seven oceanographic variables, from which a further three climatic layers were calculated (\textbf{Table} \ref{tab:table1}).
 These variables are the most frequently used in seabird-environment relationship studies (Tremblay et al., 2009), capturing processes at a range of spatial and temporal scales from macro-scale water-mass properties to prey availability in meso- to submesoscale oceanographic features (Haury et al., 1978; Scales, Miller, Hawkes, et al., 2014; Weimerskirch, 2007).
@@
 Data wrangling, modelling and visualization were done in R 4.2.0 (R Core Team, 2022); the full list of packages can be found in the Supplementary Material.
-Code and the modelling dataset (see \ref{sec:modellingdatasets}) are available in \textbf{(Zenodo repo)}.
+Code and the modelling dataset (see Section \ref{sec:modellingdatasets}) are available in \textbf{(Zenodo)}.
 
@@
 Thus, `eke' and `sst\_grad' environmental layers were calculated before extracting environmental data for each seabird record.
-For both `eke' and `sst\_grad', we used their original cell spatial resolution (\textbf{Table \ref{tab:table1}}) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
-Eddy kinetic energy was calculated as \(0.5 (u'² + v'²)\), where \(u' = u - \bar{u}\) and \(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022).
-The spatial `sst\_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as \(NS² + WE²\) (Burrows et al., 2011).
-In addition, we calculated climatic seasonal average layers for `eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the `eke' standard deviation over the same period (named with the prefix `clim\_' in \textbf{Table \ref{tab:table1}}).
+For both `eke' and `sst\_grad', we used their original cell spatial resolution (\textbf{Table} \ref{tab:table1}) and calculations were done layer-by-layer to match seabird data on the finest spatial scale possible.
+Eddy kinetic energy was calculated as \(0.5 (u'^{2} + v'^{2})\), where \(u' = u - \bar{u}\) and \(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian (longitudinal) geostrophic currents, respectively, using a rolling mean of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022).
+The spatial `sst\_grad' for each cell was calculated by averaging the maximum value for each neighbouring cell (average neighbourhood maximum), from north-to-south (NS) and west-to-east (WE); then, the magnitude of the gradient was defined as \(NS^{2} + WE^{2}\) (Burrows et al., 2011).
+In addition, we calculated climatic seasonal average layers for `eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a proxy for persistent oceanographic features, and the `eke' standard deviation over the same period (named with the prefix `clim\_' in \textbf{Table} \ref{tab:table1}).
 Seabed slope was calculated using `raster::terrain()' (Hijmans, 2022a), and distance from the coast was calculated through `sf::st\_distance()' (Pebesma, 2018) using Australia and Tasmania as the mainland (i.e.~ignoring smaller islands).
@@
 Thus, we built separate modelling datasets for each season.
-We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (\emph{c.} 110 \(\times\) 110 km) and season (\textbf{Figure \ref{fig:fig1-study-area}b}), by summing seabirds from the same species.
+We aggregated seabird data by 1° latitude \(\times\) 1° longitude grid cells (\emph{c.} 110 \(\times\) 110 km) and season (\textbf{Figure} \ref{fig:fig1-study-area}\textbf{b}), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \textgreater1 to 1.
@@
 Given a set of environmental variables, the model can predict assemblage probabilities for sites where no biotic (observational) data exist.
-As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing distribution of residuals (N.
+As a model-based approach, RCPs can be compared by information criteria, present uncertainty around the results, and check for model assumptions through assessing residuals (N.
 A.
@@
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) (Foster et al., 2013).
-However, the final estimated assemblage structure is assumed to be constant for all sites belonging to the \(K\) assemblages (Foster et al., 2013).
-As such, RCPs can be directly interpreted as assemblages, and their species profiles as the assemblage structure.
+RCPs can be directly interpreted as assemblages; the assemblage structure is given by the species profiles.
 
@@
 Due to seabird migration and dispersion, seasons could act as `sampling artifacts' (Foster et al., 2017), thus biasing results (Ponti \& Sannolo, 2023).
-In light of the difference among seasons in the numbers of species that met our occurrence threshold (see \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
+In light of the difference among seasons in the numbers of species that met our occurrence threshold (see Section \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
 RCP models were used to relate the response variables (i.e.~presence-absence or abundance of each seabird species in each grid cell) to the independent predictor covariates (i.e.~averaged values of environmental variables in each cell).
-Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a Negative Binomial model.
+Presence-absence data were modelled based on a Bernoulli model, whereas abundance data were based on a negative binomial model.
 RCPs were fitted using the `ecomix' 1.0.0 package (S.
-Woolley et al., 2022), with the following workflow: (i) find the best number of groups (\(K\)) that describes the data, (ii) re-fit the model by fixing \(K\) and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty around parameter estimates, and (v) produce results.
+Woolley et al., 2022), with the following workflow: (i) find the best number of groups (\(K\)) that describes the data, (ii) re-fit the model by fixing \(K\) and selecting the optimal combination of covariates, (iii) check residuals, (iv) bootstrap the best model to estimate uncertainty of parameter estimates, and (v) produce results.
 
@@
 Noticeably higher counts were made around Tasmania, possibly due to the `colony effect' and/or the presence of Southern Ocean species (\textbf{Figure S2}).
-However, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania in summer.
+However, in summer, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania.
 
@@
 For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (\textbf{Figure S3}).
-The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (\textbf{Table \ref{tab:table2}}).
+The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (\textbf{Table} \ref{tab:table2}).
 The best models showed a good fit with no apparent deviations from the assumptions (\textbf{Figure S4}).
-All the best models selected `sst' or `sss', as essential for characterising water masses.
-The thresholds for defining RCP membership were 18--21°C for `sst' and 35.1 PSU for `sss' (\textbf{Figure \ref{fig:fig2-partial-plots}}, \textbf{Figure S5}), similar to the characteristics used to to define the EAC.
-Furthermore, most of the best models selected `bat', suggesting a gradient between `coastal' and `oceanic' taxa based on bathymetry (\textbf{Figure S5}).
+All the best models selected sea surface temperature (`sst') or salinity (`sss'), as essential for characterising water masses.
+The thresholds for defining RCP membership were 18--21°C for `sst' and 35.1 PSU for `sss' (\textbf{Figure} \ref{fig:fig2-partial-plots}, \textbf{Figure S5}), similar to the characteristics used to to define the EAC.
+Furthermore, most of the best models selected bathymetry (`bat'), suggesting a gradient between `coastal' and `oceanic' taxa (\textbf{Figure S5}).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the `climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \textasciitilde34°S, with both presence-absence and abundance models pointing to similar results (\textbf{Figure \ref{fig:fig3-point-pred}}).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Figure \ref{fig:fig4-prob-maps}}, \textbf{Figure S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
-The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Strait' assemblage appears (\textbf{Figure \ref{fig:fig3-point-pred}}, \textbf{Figure S6}).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \textasciitilde34°S, with both presence-absence and abundance models pointing to similar results (\textbf{Figure} \ref{fig:fig3-point-pred}).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (\textbf{Figure} \ref{fig:fig4-prob-maps}, \textbf{Figure S6}), showing that around \textasciitilde34°S, the `northern' RCP tended to have higher probabilities in coastal regions, whereas the `southern' RCP spread northward offshore.
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong `Tasmania/Bass Strait' assemblage appears (\textbf{Figure} \ref{fig:fig3-point-pred}, \textbf{Figure S6}).
 
@@
 Although there are overlaps among the seabird groups, both presence-absence and abundance models captured typical tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp., lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche} spp., fairy prion \emph{P.
-turtur}) (\textbf{Figure \ref{fig:fig5-spp-profiles}}, \textbf{Figure S7}).
+turtur}) (\textbf{Figure} \ref{fig:fig5-spp-profiles}, \textbf{Figure S7}).
 The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the `northern' and `southern' RCPs.
@@
 Our analyses clearly revealed a `northern' and a `southern' assemblage of seabirds off eastern Australia within the EAC.
-There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \textasciitilde34°S, near the average separation latitude (\textasciitilde32°S) of the EAC and its eastward extension into the Tasman Sea (Oke, Roughan, et al., 2019).
+There is a noticeable link with the EAC dynamics: the boundaries of RCPs occurred at \textasciitilde34°S, near the average separation latitude (\textasciitilde32°S; 25--38°S) of the EAC and its eastward extension into the Tasman Sea (Oke, Roughan, et al., 2019).
 However, the `northern/southern' assemblage pattern identified herein does not agree with previously-proposed large-scale regionalizations, such as the pelagic Longhurst Provinces (Longhurst, 2007), the Marine Ecoregions of the World {[}MEOW; Spalding et al.
@@
 Understanding the biology of the studied group is thus essential to set the appropriate scales of analyses.
-Our choice of 1° latitude \(\times\) 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also \ref{sec:limitations}).
+Our choice of 1° latitude \(\times\) 1° longitude was mainly determined by data availability reflecting survey effort, and the trade-off between the number of species per grid/season (see also Section \ref{sec:limitations}).
 Nonetheless, the chosen spatial scale is similar to other seabird distribution analyses conducted elsewhere (Harris et al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022).
@@
 In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
-Nonetheless, rarefaction curves show that the sample coverage (see Chao \& Jost, 2012) was high at the RCP level, although there is room for detecting more species in most of them (\textbf{Fig S8}).
+Nonetheless, rarefaction curves show that the sample coverage (see Chao \& Jost, 2012) was high at the RCP level, although there is room for detecting more species in most of them (\textbf{Figure S8}).
 
@@
 https://doi.org/\url{https://doi.org/10.1111/2041-210X.12613}
+
+\leavevmode\vadjust pre{\hypertarget{ref-huang2021}{}}%
+Huang, B., Liu, C., Banzon, V., Freeman, E., Graham, G., Hankins, B., Smith, T., \& Zhang, H.-M.
+(2021).
+Improvements of the daily optimum interpolation sea surface temperature (DOISST) version 2.1.
+\emph{Journal of Climate}, \emph{34}(8), 2923--2939.
+\url{https://doi.org/10.1175/JCLI-D-20-0166.1}
 

### 2023-10-24 ms_preprint/ms_supp_mat.Rmd
ms adjustments after LB RS and MS reviews pre-submission

@@
 title: "Supplementary Material"
-subtitle: "<b>Seabird assemblages are linked to the major western boundary current off eastern Australia</b>"
+subtitle: "Seabird assemblages are linked to the major western boundary current off eastern Australia"
 # author: "Author"
@@
 
-The code is archived in a repository **(Zenodo repo)**, and you can find a detailed walk-through in it.
+The code is archived in a repository **(Zenodo)**, where you can find a detailed walk-through.
 
@@
 
-![**Figure S1.** Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021.
+![**FIGURE S1.** Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021.
 The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text).
@@
 
-![**Figure S2.** Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see *Methods* in the main text).
+![**FIGURE S2.** Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see *Methods* in the main text).
 Dots represent the grid centroids.](../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png)
@@
 
-![**Figure S3.1.** Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia.
+![**FIGURE S3.1.** Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia.
 The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data.
@@
 
-![**Figure S3.2.** Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia.
+![**FIGURE S3.2.** Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia.
 The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data.
@@
 
-![**Figure S4.1.** Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png)
+![**FIGURE S4.1.** Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png)
 
@@
 
-![**Figure S4.2.** Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png)
+![**FIGURE S4.2.** Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png)
 
@@
 
-![**Figure S5.1.** Partial plots for the retained covariates in the best seasonal models based on presence-absence data.
+![**FIGURE S5.1.** Partial plots for the retained covariates in the best seasonal models based on presence-absence data.
 The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 
-![**Figure S5.2.** Partial plots for the retained covariates in the best seasonal models based on abundance data.
+![**FIGURE S5.2.** Partial plots for the retained covariates in the best seasonal models based on abundance data.
 The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 
-![**Figure S6.1.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models.
+![**FIGURE S6.1.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models.
 The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png)
@@
 
-![**Figure S6.2.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models.
+![**FIGURE S6.2.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models.
 The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png)
@@
 
-![**Figure S7.1.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model.
+![**FIGURE S7.1.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model.
 Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps.
@@
 
-![**Figure S7.2.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model.
+![**FIGURE S7.2.** Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model.
 Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps.
@@
 
-![**Figure S8.** Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png)
+![**FIGURE S8.** Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png)
 

### 2023-10-24 ms_preprint/ms_supp_mat.tex
line numbering and a few corrections after EJWs review, and changed to correct MS title in the supp mat
ms adjustments after LB RS and MS reviews pre-submission

@@
 \makeatother
-\subtitle{Probabilistic bioregion models suggest seabird assemblages are
-linked to the major boundary current off eastern Australia}
+\subtitle{Seabird assemblages are linked to the major western boundary
+current off eastern Australia}
 \author{}
-\date{\vspace{-2.5em}2023-10-17}
+\date{\vspace{-2.5em}2023-10-24}
 
@@
 
-The code is archived in a repository \textbf{(Zenodo repo)}, and you can
-find a detailed walk-through in it.
+The code is archived in a repository \textbf{(Zenodo)}, where you can
+find a detailed walk-through.
 
@@
 \includegraphics{../results/FigS1_spp-nOCC-FO-NF-seasons.png}
-\caption{\textbf{Figure S1.} Number of occurrences (nOCC), frequency of
+\caption{\textbf{FIGURE S1.} Number of occurrences (nOCC), frequency of
 occurrence (FO) and numeric frequency (NF) of seabirds recorded off
@@
 \includegraphics{../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png}
-\caption{\textbf{Figure S2.} Species richness and the total number of
+\caption{\textbf{FIGURE S2.} Species richness and the total number of
 seabirds counted off eastern Australia during Australasian Seabird
@@
 \includegraphics{../results/FigS3_1_multifit-Bernoulli.png}
-\caption{\textbf{Figure S3.1.} Multifit plot for Region of Common
+\caption{\textbf{FIGURE S3.1.} Multifit plot for Region of Common
 Profiles (RCP) for each seasonal presence-absence model, applied to
@@
 \includegraphics{../results/FigS3_2_multifit-NegBin.png}
-\caption{\textbf{Figure S3.2.} Multifit plot for Region of Common
+\caption{\textbf{FIGURE S3.2.} Multifit plot for Region of Common
 Profiles (RCP) for each seasonal abundance (count) model, applied to
@@
 \includegraphics{../results/FigS4_1_best-model-residuals-Bernoulli.png}
-\caption{\textbf{Figure S4.1.} Best model residuals for each seasonal
+\caption{\textbf{FIGURE S4.1.} Best model residuals for each seasonal
 presence-absence Region of Common Profile model, applied to seabirds off
@@
 \includegraphics{../results/FigS4_2_best-model-residuals-NegBin.png}
-\caption{\textbf{Figure S4.2.} Best model residuals for each seasonal
+\caption{\textbf{FIGURE S4.2.} Best model residuals for each seasonal
 abundance (count) Region of Common Profile model, applied to seabirds
@@
 \includegraphics{../results/FigS6_1_partial-plots-Bernoulli.png}
-\caption{\textbf{Figure S5.1.} Partial plots for the retained covariates
+\caption{\textbf{FIGURE S5.1.} Partial plots for the retained covariates
 in the best seasonal models based on presence-absence data. The plot
@@
 \includegraphics{../results/FigS6_2_partial-plots-NegBin.png}
-\caption{\textbf{Figure S5.2.} Partial plots for the retained covariates
+\caption{\textbf{FIGURE S5.2.} Partial plots for the retained covariates
 in the best seasonal models based on abundance data. The plot shows the
@@
 \includegraphics{../results/FigS5_1_prob-pred-Bernoulli.png}
-\caption{\textbf{Figure S6.1.} Predicted probability membership of for
+\caption{\textbf{FIGURE S6.1.} Predicted probability membership of for
 each seabird assemblage (Region of Common Profiles; RCP) and grid, off
@@
 \includegraphics{../results/FigS5_2_prob-pred-NegBin.png}
-\caption{\textbf{Figure S6.2.} Predicted probability membership of for
+\caption{\textbf{FIGURE S6.2.} Predicted probability membership of for
 each seabird assemblage (Region of Common Profiles; RCP) and grid, off
@@
 \includegraphics{../results/FigS7_1_spp-profiles-Bernoulli.png}
-\caption{\textbf{Figure S7.1.} Species profiles for each assemblage
+\caption{\textbf{FIGURE S7.1.} Species profiles for each assemblage
 (Region of Common Profiles; RCP) for each seasonal presence-absence
@@
 \includegraphics{../results/FigS7_2_spp-profiles-NegBin.png}
-\caption{\textbf{Figure S7.2.} Species profiles for each assemblage
+\caption{\textbf{FIGURE S7.2.} Species profiles for each assemblage
 (Region of Common Profiles; RCP) for each seasonal abundance (count)
@@
 \includegraphics{../results/FigS8_iNEXT.png}
-\caption{\textbf{Figure S8.} Diversity curve (alpha diversity) and
+\caption{\textbf{FIGURE S8.} Diversity curve (alpha diversity) and
 sample coverage for each assemblage (Region of Common Profile; RCP) from

### 2024-03-06 ms_PROOCE/ms_PROOCE_review1_response-to-reviewers.Rmd
update README, update scripts-workflow figure, updated references inlcuding Daudt et al 2024 and 'corrected' refs for R pckgs, plus all source RMD files

---
title: ""
# author: "Author"
# date: ""
output: pdf_document
---
`r Sys.Date()` \newline

\begin{flushright}
\textit{Submission}: \textbf{PROOCE-D-23-00336} \newline
\textit{Seabird assemblages are linked to the major western boundary current off eastern Australia} \newline
\end{flushright}

---

\begin{flushright}
\underline{\textbf{Response to Reviewer's questions comments}} \newline
\end{flushright}

\textcolor{blue}{Most of the comments from Reviewer 2 in this section are also under their detailed comments. Therefore, those were omitted from this section and answered/commented below.} \newline

*Question 5*

**Reviewer 1**

*Line 8 if the Highlights are those that the study has uncovered, the first ‘highlight’, southward movement of the EAC, is not that which is determined in this study.*

\textcolor{blue}{That is right, thanks for noticing it. We have now removed this sentence from the Highlight session.} \newline

*Line 341 'While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales (Tremblay et al., 2009)', this study as an understanding of how much tracking studies advance the understanding of seabird interaction and their environment fails to look at more current papers. A good example (with referenced other examples within) is: Trevail et al (2023) Tracking seabird migration in the tropical Indian Ocean reveals basin-scale conservation need. <https://doi.org/10.1016/j.cub.2023.10.060>*

\textcolor{blue}{Thanks for suggesting Trevail et al. (2023), which we have included alongside Tremblay et al. (2009).} \newline

**Reviewer 2**

\textcolor{blue}{See comments below.}

---

\begin{flushright}
\underline{\textbf{Response to Reviewer's 2 detailed comments}} \newline
\end{flushright}

*This study uses Region of Common Profile mixture models to describe seasonal seabird distributions with respect to environmental parameters, using at-sea survey data and oceanographic, climatic and physiographic factors. The authors found two macro-scale seabird assemblages, defined as 'northern' and 'southern' assemblages separated roughly near the latitude where the East Australian Current (EAC) moves off the Australian continental slope (~34 S). The authors state that the two seabird assemblages were mostly persistent across four seasons. The authors suggest that the predicted poleward intensification of the EAC makes it imperative that changes to seabird distribution in this important region be monitored, using this study as a baseline.*

*This paper provides new information about an important seabird 'hot spot' in the South Pacific Ocean, and it is timely in that climate-driven changes to the region have implications to the seabird community. In addition, the results describe physical drivers that influence seabird distribution and assemblages in this marine system. The paper is generally well written and easy to follow.*

\textcolor{blue}{Thanks for your thorough, positive review of our manuscript. We greatly appreciate your time and dedication to commenting and suggesting points that could be clarified/better explored, which have undoubtedly improved the manuscript. Please find below replies to the main 'general aspects' you stressed out, followed by detailed replies for each specific topic.} \newline

*There are four general aspects for which I recommend clarification or changes.*

*First, unless this journal has a strict limit on length of the paper, I think the scope and integration of the various components of this study would be better served by moving some of the detail available in Supplemental material into the main paper. Or, perhaps as an attached appendix. In particular, I suggest including the results of species richness analysis, which seems relevant to the focus of the paper. I hope more about the assemblages themselves (species involved and what they are responding to) are covered in subsequent papers.*

\textcolor{blue}{Thanks for suggesting the inclusion of supplementary figures in the main paper. The species richness analysis was moved to the main text, while other figures, as explained below, were kept in the supplemental material.}

\textcolor{blue}{Also, thanks for encouraging a separate publication on the assemblages (a likely following paper on the way).} \newline

*Second, it's not clear why only the spring is presented for some results (Figures 2, 4, 5) in the main paper (except perhaps for brevity and simplicity). If the authors provided justification for why spring results were highlighted, I could not find it. Furthermore, Suppl. figures 9 and 10 appear to show considerable differences among seasons in the spatial distribution of the two assemblages. Although the Discussion also mentions the influence of summer/autumn migrants (lines 293-299), there are statements about the seasonal persistence of assemblages distributions, which is confusing. More explanation of the observed seasonal differences may be warranted.*

\textcolor{blue}{Although \textit{Progress in Oceanography} does not limit figures, we decided to keep supplementary figures as they were. We chose to present only spring results for the sake of the brevity of the main text. Readers interested in specific maps and graphs can refer to the supplementary material at their leisure. From our perspective, this does not influence the reader's understanding of the main messages of the paper.}

\textcolor{blue}{To clarify this in the main paper, we added the following sentence under the \textit{RCP models} heading:}

\textcolor{blue}{"In this section, we will present only the main results for the spring season when delving into specifics about each model and its results. The results of spring models summarise the key findings of our analysis, but we encourage the reader to look at the Supplementary Material for other seasons."}

\textcolor{blue}{We believe the differences in the spatial patterns of assemblages shown in the (original) supplementary figures 9 and 10 (predicted probability maps) are due to the lack of comprehensive sampling in all seasons. Given the strong support for a 'southern' and a 'northern' assemblage, we note that both summer and autumn fail to have observations in both southern/northern areas. The rarefaction curves also show this by lower values on 'sample coverage' for RCP1 (northern) in summer and RCP3 (northern) in autumn. Further, the point predictions assigned higher probabilities of 'southern' assemblage in the 'northern' area for summer and the opposite in autumn, suggesting some limitations about these results. We agree we could have explored these incongruences further in the main text, so we have now expanded this explanation on specific sections of the discussion, as detailed in the replies below.} \newline

*Third, I could not find any definition of the seasonal breakdowns - which months comprise summer, autumn, winter, spring? It may be clear to the authors, but studies vary widely in those seasonal categories. Also, please make it clear to northern readers that you are describing austral seasons.*

\textcolor{blue}{Thanks for noticing this. We have now added a sentence to include which months comprise each season (see below).} \newline

*Fourth, I suggest the authors make more effort to compare the patterns found in this study region with results from studies in other regions, including the northern hemisphere and sub-Arctic and Arctic.*

\textcolor{blue}{Thanks for your comment. This is the first study applying RCP models to highly mobile animals, more specifically to seabirds. Although there is an obvious component of seabirds throughout the paper, the key points in the discussion are broader--i.e. the application of RCP to highly mobile animals and marine biogeography. However, we believe these comparisons asked by the reviewer are among a sound background Intro section (L 29--43 and 50--56) and later subtly addressed in L 286--287 and 378--379 in the Discussion, according to the references provided. Further, the different oceanography dynamics, seabird families, and foraging strategies for the seabird species compared in a sub-Arctic/Arctic study with a tropical-cool temperate study do not seem adequate to us. Therefore, we would prefer not to add another paragraph to the main text.} \newline

*Specifics (by Line; L)*

*L 41-43: Consider including publications from other bioregions (some more recent), which also show SST and water masses as key drivers of seabird distributions. There are many; a few are: Gall et al. 2013, 2022; Nur et al. 2011, Serratosa et al. 2020.*

\textcolor{blue}{Thanks for the suggestion. We have included Serratosa et al. (2020) and Gall et al. (2022).} \newline

*L 90 - not sure a blanket statement like 'seabird distribution is under-represented in climate change studies' is the case elsewhere (e.g, studies in the North Atlantic, and to less extent in the North Pacific).*

\textcolor{blue}{Even though there are studies in the North Atlantic/Pacific, these are mainly around the USA; therefore, we believe seabirds are still under-represented in this field of research. However, we do acknowledge these efforts and studies. Thus, we have changed the sentence to reflect that the lack of these studies in the Australasia region is particularly noticeable.} \newline

*L 127-128 - Briefly describe the main features of this protocol, even though details can be obtained from the citation given. How were flying (vs on water or foraging) birds treated; how was distance (within 300m) determined; when were surveys halted due to conditions? Since 'one or more' observers recorded birds, how did you adjust for differences in effort?*

\textcolor{blue}{The protocol does not record distance from the ship for each individual record (i.e. it is a \textit{strip-transect} protocol); we have added this info in the section 2.2. We also added the behaviours, \textit{"(i.e. flying, feeding, sitting on the water)"}, in the section 2.2 and 2.3.2, as we summed all behaviours for our aggregated data set. In addition, we added the conditions when surveys were halted (\textit{"Seabird surveys were halted when sea conditions exceeded 6 m wave height and 40 kn of wind or visibility was under 500 m."}).}

\textcolor{blue}{Sorry, the issue raised about having 'one or more' observers was a misunderstanding between the authors. \textit{Two} observers continuously recorded seabirds throughout the day; single observers were only present during brief meal and comfort [toilet] breaks (5--15 min). So, there was no significant difference in effort, since the observations were confined to forward quarter 300 m and there was never the density of birds that reduced data entry or impacted on survey efficacy. This was fixed in the main text by changing from \textit{"One or more observers"} to \textit{"Two observers"}.} \newline

*L 137 - 'data wrangling' is jargon; revise to something like 'Data management', or 'compilation'.*

\textcolor{blue}{Data 'wrangling' is an actual concept in the field of Data Science, which means 'tidying and transforming' your data. See well-known data science book "R for Data Science", which has a whole section on 'wrangle' in its 1st edition (at \url{https://r4ds.had.co.nz/wrangle-intro.html}), and properly define the concept in its 2nd edition (at \url{https://r4ds.hadley.nz/intro}).} \newline

*Section 2.3.2 - somewhere in this section, describe what months constitute each season.*

\textcolor{blue}{Done. We have included a sentence that reads:}

\textcolor{blue}{"Seasons were defined as summer (Dec--Feb), autumn (Mar--May), winter (Jun--Aug), and spring (Sep--Nov)."} \newline

*L 187 - not clear what the 'species profiles' are here. Please clarify what this refers to.*

\textcolor{blue}{Done. We have ammended the sentence, which now reads:}

\textcolor{blue}{"RCPs can be directly interpreted as assemblages; the assemblage structure is given by the species profiles (i.e. the proportion of each species within the assemblage given by their probability of occurrence or predicted abundance)."} \newline

*L 231-233 - I think Supple Fig.2 (species richness maps) would be best in the main paper. While the rarefaction curves (Supple Fig. 13) serve to determine if sampling was adequate, the results in Fig. 2 seem directly relevant to the goal of describing the distributions of seabird assemblages.*

\textcolor{blue}{Thanks for pointing this out---we moved species richness analysis into the main paper.} \newline

*L 245-247 - authors note that in particular, 'proxies for persistent oceanographic features that may aggregate food…were selected only in the autumn and spring abundance models'. This is interesting, but I don't see in the Discussion where the authors propose why this may be. Somewhere in Discussion, it would also be helpful to indicate what prey are being affected, and where, by the oceanographic features and SST. Alternatively, the Methods section could include information on prey consumed by regional seabirds in the region.*

\textcolor{blue}{Thanks for your comment. We did expand on this interesting finding of only abundance models selecting variables summarising persistent oceanographic features (see the new sentences below). However, we think it is out-of-scope for our study to delve into prey and prey distribution. This would require new sections, from Methods to Discussion, and very likely new analyses, which may deserve a separate publication.}

\textcolor{blue}{The new sentences were added under section 4.2, splitting the (original) third paragraph into two. The sentences read as follows:}

\textcolor{blue}{"[...]. Although we have not approached the data in a multi-scale analysis, our results partially support this. Presence-absence models retained static variables and oceanographic variables that define water masses (i.e. SST and salinity). In addition to those, the autumn and spring abundance models also retained variables associated with persistent oceanographic features (e.g. climatic variables related to eddies and fronts), as well as mixed layer depth and chlorophyll in the autumn model. Eddies, fronts, high chlorophyll and shallower mixed layer depths are all related to increased seabird foraging success (Weimerskirch, 2007; Scales et al., 2014; De Pascalis et al., 2021), and therefore likely to influence seabird numbers. [...]"} \newline

*Fig. 5 - Might note in caption that the species along the Y-axis are arranged taxonomically.*

\textcolor{blue}{The species are ordered alphabetically (from bottom up) as the default output from the \{ecomix\} package. Nevertheless, we noted this in the figure caption to avoid doubts (in both the main paper and supplementary material).} \newline

*L 266-275 - Very nice opening paragraph. It summarizes key findings and implications to monitoring and conservation.*

\textcolor{blue}{Many thanks!} \newline

*L 277 - Suggest rewording to '..and water masses has been well established'. Could also include northern studies (e.g., Gall et al 2013 or 2022).*

\textcolor{blue}{Thanks, we have accepted your suggestion and included Gall et al. (2022) in the reference list.} \newline

*L 279 - as noted earlier, Supple Fig. 9 & 10 appear to show considerable spatial shifts in latitude, even if still two assemblages - thus the 'spatial consistency of these patterns throughout seasons' isn't obvious to me. They show similar patterns in winter and spring, but summer and autumn each have different patterns. A difference in pattern during autumn is mentioned in lines 254-255, but not addressed later. Please explain, or if you're basing this conclusion on another set of figures, cite those.*

\textcolor{blue}{We have included "(but see Section 4.3)" [limitations] after the L 279 sentence, and expanded on that in the proper section by writing a new paragraph:}

\textcolor{blue}{"These limitations are illustrated in the results for the summer and autumn seasons. Considering the strong support for a southern and a northern assemblage, we note that both summer and autumn lack samples in representative southern and northern (summer) and northern (autumn) areas. Albeit the sample coverage values were overall high, compared to winter and spring, summer and autumn had lower values (in particular, for the southern assemblage 'RCP1' in summer and for the northern assemblage 'RCP3' in autumn; Supplementary Fig. 12). Further, based on point predictions, higher probabilities of the southern assemblage in the northern area in summer (both presence-absence and abundance models) and of the northern assemblage in the southern area in autumn (abundance model), were assigned to a few grids (Fig. 4). The latter reinforces the possible weakness of these models to detect stronger relationships between the identified assemblages and the predictors. We also acknowledge the sharp spatial change in probabilities between assemblages in summer models (i.e. without a proper 'transition zone'), which may be related to the previous point. However, both summer and autumn models detected higher probabilities of expected taxa in their species profiles, suggesting that the make-up of the assemblage was well distinguished by the models. In addition, these models also seem to have captured the appropriate number of assemblages, agreeing with other models/seasons. Despite their limitations, we believe these models agree with the big picture of our results, therefore being a valid resource as a baseline."}

\textcolor{blue}{Regarding the 'three assemblage pattern' for autumn (L 254--255) mentioned by the reviewer, we believe we have touched on it in the second paragraph of section 4.1 and when concluding it by saying "[...] In turn, including these taxa in the models could result in a more stable 'three assemblage pattern' (as revealed by the autumn presence-absence model) in all but the summer season, when the majority of Antarctic species are breeding (Marchant and Higgins, 1990)."} \newline

*L 306-332 - I like this section describing the results in context of other descriptions of biogeographic units and LMEs. It's clearly written and is relevant to management issues.*

\textcolor{blue}{Thank you!} \newline

*L 340 - Suggest mentioning the positive aspects of studying seabirds in context of environmental parameters - i.e., seabirds are highly visible and therefore it's possible to obtain good sample sizes over large areas, and have high encounter rates in some locations (easy to identify hotspots). (there are multiple citations stating these points).*

\textcolor{blue}{We understand that for a broader readership, your suggestion might be valid (e.g. people not used to seabird surveys). Therefore, we have included the sentences below:}

\textcolor{blue}{"Among marine megafauna, seabirds are relatively easy to detect visually, enabling suitable sample sizes to be gathered over large areas (Ballance, 2007). Thus, seabirds can be used as models to study the distribution of highly mobile animals in relation to environmental characteristics in the marine environment."} \newline

*L365-367 - Could add some recent studies*

\textcolor{blue}{Done. We have added Commins et al. (2014) and Gall et al. (2022) in the first part of the sentence, and De Pascalis et al. (2021) on the second part of the sentece.} \newline

*L 373 - remove 'which we suspect do not exist….'. Replace with something like 'which to our knowledge are not generally available at present, ". You might check whether this is in fact the case, at least for your region.*

\textcolor{blue}{Thanks for the suggestion. To the best of our knowledge, no such data is available for our study area. So, we slightly modified the previous sentence to:}

\textcolor{blue}{"Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which, to our knowledge, do not exist for our study area."} \newline

*L 381-384 - While I don't disagree with these statements on 'need for more seabird data', these are generally known concepts, and not directly relevant to scope of this journal. I would remove or reduce this paragraph to keep focus on the important findings.*

\textcolor{blue}{Thanks for your comment. The (original) paragraph was about our study's limitations and how the analyses could benefit from more data. As you stressed before (mainly regarding summer/autumn), this also has implications for the described patterns, which we have expanded the text on 'limitations' in the current version of the manuscript. Therefore we would prefer not to remove this paragraph.} \newline

*L 392 - Can authors provide information on why SST is a key variable discriminating the two seabird assemblages? The Discussion could provide brief examination of what SST is linked to: e.g,, the biological (prey type and distribution?) and physical (more on fronts and upwelling?) features birds are likely responding to. How do prey differ north and south of the 'transitional zone'?*

\textcolor{blue}{Our analysis does not allow us to dive into the 'whys' as a correlative method. However, as shown elsewhere, SST likely influences prey distribution, affecting predator distribution. We briefly mentioned that while presenting biogeographic schema that agreed with our findings, such as the mesopelagic biogeographic provinces from Sutton et al. (2017). As mentioned previously, we think it is out-of-scope to go further down the 'prey distribution' topic. Nonetheless, we have highlighted this in a new sentence in the Conclusions:}

\textcolor{blue}{"The mechanistic links between SST and seabird assemblages deserve more studies but may be due to the distribution of seabird prey, which at the biogeographic level seems to correlate with our findings (Sutton et al., 2017)."} \newline

*L 402 - might the 'transition zone' also shift poleward? Or might it expand in size? That would have implications for future seabird studies attempting to replicate your study.*

\textcolor{blue}{Good point. We believe the 'transition zone' might shift polewards following EAC's strengthening. We have modified the original sentence to reflect this; it now reads:} \newline

\textcolor{blue}{"In particular, the transition zone between assemblages might be pushed southwards along with the expected trend of the EAC. The transition zone between assemblages (Morrone, 2023), associated with the eastern extension of the EAC, may be fundamental to effectively monitoring the strengthening of the EAC and the subsequent impact on marine biota."} \newline

### 2024-03-06 ms_PROOCE/ms_PROOCE.Rmd
update README, update scripts-workflow figure, updated references inlcuding Daudt et al 2024 and 'corrected' refs for R pckgs, plus all source RMD files

--- 2023-10-24 ms_preprint/ms_preprint.Rmd
@@
 title: Seabird assemblages are linked to the major western boundary current off eastern Australia
-authors:
+author:
   - name: Nicholas W. Daudt
-    department: Department of Marine Science
-    affiliation: University of Otago
-    location: Dunedin, Aotearoa New Zealand
     email: nicholaswdaudt@gmail.com
-    thanks: Corresponding author
+    affiliation: MarSci-Otago
+    correspondingauthor: true
   - name: Eric J. Woehler
-    department: Australasian Seabird Group
-    affiliation: BirdLife Australia
-    location: Hobart, Australia
-    email: eric.woehler@gmail.com
+    # email: eric.woehler@gmail.com
+    affiliation: ASG
   - name: Matthew R. Schofield
-    department: Department of Mathematics and Statistics
-    affiliation: University of Otago
-    location: Dunedin, Aotearoa New Zealand
-    email: matthew.schofield@otago.ac.nz
+    # email: matthew.schofield@otago.ac.nz
+    affiliation: MatStat-Otago
   - name: Robert O. Smith
-    department: Department of Marine Science
-    affiliation: University of Otago
-    location: Dunedin, Aotearoa New Zealand
-    email: robert.smith@otago.ac.nz
+    # email: robert.smith@otago.ac.nz
+    affiliation: MarSci-Otago
   - name: Leandro Bugoni
-    department: Institute of Biological Sciences
-    affiliation: Universidade Federal do Rio Grande---FURG
-    location: Rio Grande, RS, Brazil
-    email: lbugoni@yahoo.com.br
+    # email: lbugoni@yahoo.com.br
+    affiliation: FURG
   - name: William J. Rayment
-    department: Department of Marine Science
-    affiliation: University of Otago
-    location: Dunedin, Aotearoa New Zealand
-    email: will.rayment@otago.ac.nz
+    # email: will.rayment@otago.ac.nz
+    affiliation: MarSci-Otago
+address:
+  - code: MarSci-Otago
+    address: Department of Marine Science, University of Otago, Dunedin, Aotearoa New Zealand
+  - code: ASG
+    address: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
+  - code: MatStat-Otago
+    address: Department of Mathematics and Statistics, University of Otago, Dunedin, Aotearoa New Zealand
+  - code: FURG
+    address: Institute of Biological Sciences, Universidade Federal do Rio Grande---FURG, Rio Grande, RS, Brazil
 abstract: |
-  **Aim:** To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
-  **Location:** East coast of Australia.
-  **Methods:** We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
-RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
+  Identifying species assemblages helps understand the relationship between organisms and their environment.
+Assemblages can be used to predict biological changes caused by environmental perturbations, and are thus essential surrogates to monitor biodiversity.
+In this study, to identify and describe seabird assemblages, we used 15 at-sea ship-based survey data sets collected over 37° of latitude off eastern Australia, from 2016 to 2021.
+We fitted seasonal Region of Common Profile (RCP) mixture models, for two types of data (presence-absence and abundance).
+RCP groups are defined as regions where the probability of encountering a particular species profile is constant within regions, but different amongst them.
 These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-  **Results:** Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels,  prions, shearwaters, boobies, and terns, among other taxa.
+Results were based on 142,646 seabirds recorded from 80 species, including albatrosses, petrels, prions, shearwaters, boobies, and terns, among other taxa.
 All models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at ~34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
+The model results consistently show a biogeographic transition at \~34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
 Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
 Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
-  **Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
+RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
+Given the ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing the distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages and how they are affected by changing environmental conditions.
 keywords: 
@@
   - Region of Common Profiles
-bibliography: ../references.bib
-csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
+journal: "Progress in Oceanography"
+date: "2023-11-02"
+classoption: preprint, 3p, authoryear
+linenumbers: true
+numbersections: true
 output:
   bookdown::pdf_book:
-    base_format: rticles::arxiv_article
+    base_format: rticles::elsevier_article
+    keep_tex: true
+    citation_package: default # Changing to 'default' allows specifying the CSL
+bibliography: ../references.bib
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/elsevier-harvard.csl
 header-includes:
 - \usepackage{setspace}\doublespacing
-- \usepackage[left]{lineno}
-- \newcommand{\blinenumbers}{\begin{linenumbers}}
-- \newcommand{\elinenumbers}{\end{linenumbers}}
 - \usepackage{pdflscape}
@@
   @huang2021
+  @hu2012
 ---
 
-<!-- Note: I've changed the 'arxiv.sty' source file to include the running title in the PDF, and added commands in the YAML header too -->
-
-```{r setup, include = FALSE, echo = FALSE}
+```{r r-pckgs, include = FALSE, echo = FALSE}
 library(readr)
@@
 library(kableExtra)
-
-### If using {gt}:
-# library(gt)
-# options(gt.html_tag_check = FALSE)
 ```
 
-# ORCID {.unnumbered}
+# Highlights {.unnumbered}
 
-Nicholas W. Daudt <https://orcid.org/0000-0001-8211-8771>
-
-Eric J. Woehler <https://orcid.org/0000-0002-1125-0748>
-
-Matthew R. Schofield <https://orcid.org/0000-0003-1481-2766>
-
-Robert O. Smith <https://orcid.org/0000-0001-5931-3843>
-
-Leandro Bugoni <https://orcid.org/0000-0003-0689-7026>
-
-William J. Rayment <https://orcid.org/0000-0001-9558-5109>
-
-# Biosketch {.unnumbered}
-
-**Nicholas W.
-Daudt (he/him)** is broadly interested in spatial ecology and ecological modelling.
-His research uses data science and modelling to understand biogeographic patterns and ecological processes that build marine top predators' biodiversity patterns across space and time.
-
-# Acknowledgements {.unnumbered}
-
-We thank all observers who assisted in data collection, the crews of RV *Investigator* and RV *Falkor*, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
-The authors would like to thank Dr.
-Skipton N.C.
-Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
-
-------------------------------------------------------------------------
-
-**Data availability statement:** The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
-The seabird raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the seabird raw data, please contact Eric J.
-Woehler.
-The environmental data used in this study are freely available from the source (see Table 1 in the main text).
-
-**Funding statement:** Parks Australia provided funding support to participants for selected voyages.
-
-**Conflict of interest disclosure:** The authors declare no conflict of interest.
-
-**Ethics approval statement:** NA.
-
-**Patient consent statement:** NA.
-
-**Permission to reproduce material from other sources:** NA.
-
-**Clinical trial registration:** NA.
-
-\newpage
-
-\blinenumbers
-
-```{=tex}
-\begin{center}
-\textbf{Seabird assemblages are linked to the major western boundary current off eastern Australia}
-\end{center}
-```
-*Running title:* Seabird assemblages off Eastern Australia
-
-------------------------------------------------------------------------
-
-**Abstract**
-
-**Aim:** To identify assemblages of seabird species and relate them to environmental conditions off eastern Australia, a world seabird hotspot.
-
-**Location:** East coast of Australia.
-
-**Methods:** We used 15 seabird at-sea ship-based survey data sets collected over 37° of latitude, from 2016 to 2021, and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and abundance).
-RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within regions, but different amongst them.
-These groups also vary according to covariates, which in our case included oceanographic, climatic, and physiographic parameters.
-
-**Results:** Results were based on 142,646 seabirds recorded, from 80 species, including albatrosses, petrels,  prions, shearwaters, boobies, and terns, among other taxa.
-All models suggested two macro-scale assemblages (‘northern’ and ‘southern’), except for the autumn presence-absence model that identified three groups.
-The model results consistently show a biogeographic transition at \~34°S, near the latitude at which the East Australian Current (EAC) separates from the Australian continental slope.
-Sea surface temperatures or sea surface salinities were selected in all final models, further indicating a close relationship between seabird assemblages and water masses.
-Results from both data types, presence-absence and abundance, resulted in similar spatial and species profile patterns.
-
-**Main conclusions:** RCP models clearly identified two seabird assemblages off the east coast of Australia, suggesting the persistence of these groups at seasonal and macro spatial scales.
-Given ongoing poleward intensification that the EAC is experiencing, which is projected to continue over the next century, and its importance in influencing distributions of seabirds, the methods applied in our study could be replicated to assess possible changes in seabird assemblages, and how they are affected by changing environmental conditions.
-
-***Keywords*** Assemblages; Biogeography; Community ecology; East Australian Current; Mixture models; Region of Common Profiles
+* The East Australian Current (EAC) is experiencing an ongoing poleward intensification
+* We described seabird assemblages based on the Region of Common Profiles (RCP) method
+* Assemblages of seabird species were correlated with the EAC
+* Our analysis can be replicated to assess possible changes in seabird distribution
+* The RCP method works well at the macro spatial scale for highly mobile animals
 
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\~18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
-The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Figure** \@ref(fig:fig1-study-area)**a**; @kessler2013; @ganachaud2014).
+The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.** \@ref(fig:fig1-study-area)**a**; @kessler2013; @ganachaud2014).
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
 The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension', often termed the 'Tasman Front' [@oke2019a].
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Figure** \@ref(fig:fig1-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.** \@ref(fig:fig1-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
+
+\blandscape
+
+```{r fig1-study-area, fig.cap = "\\label{fig1-study-area} Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
+In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
+The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
+The EAC 'eastern extension' is an eddy field (blue shaded area) often called the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
+Figure (a) is adapted after Oke, Roughan, et al.
+(2019); for colour references see the online version", fig.align = "center", out.width = "100%", echo = FALSE}
+
+knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
+```
+
+\elandscape
 
@@
 
-Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Figure** \@ref(fig:fig1-study-area)**b**; **Table S1**).
+Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.** \@ref(fig:fig1-study-area)**b**; **Supplementary Table 1**).
 One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
@@
 
+\blandscape
+
+```{r table1, echo = FALSE}
+data.frame(
+  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
+  # Abbreviation needs the `str_replace_all` because I'm using `escape=F` in `kbl()`
+  Abbreviation = stringr::str_replace_all(c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"), "_", "\\\\_"),
+  Units = c("°Celsius", paste0(parse(text = "\u0394"), "°Celsius"), "PSU", "m", "mg/m${^3}$", "$log_{10}$(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", paste0("mean(", paste0(parse(text = "\u0394"), "°Celsius"), ")/season/10 years"), "m", "degrees (°)", "km"),
+  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
+  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
+  Source = c("Huang et al.
+(2021)${^a}$", "Derived from Huang et al.
+(2021)${^a}$", "E.U.
+Copernicus Marine Service Information${^b}{^c}$", "E.U.
+Copernicus Marine Service Information${^b}{^c}$", "Hu et al.
+(2012)${^d}$", "Derived from Hu et al.
+(2012)${^d}$", "IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from Huang et al.
+(2021)${^a}$", "GEBCO${^f}$", "Derived from GEBCO${^f}$", "Calculated in R (see main text)")
+) %>% 
+  kableExtra::kbl(caption = "\\label{table1}Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources", booktabs = TRUE, escape = FALSE)  %>%
+  kableExtra::row_spec(0, bold = TRUE) %>% 
+  kableExtra::column_spec(1, width = "4cm") %>% 
+  kableExtra::column_spec(6, width = "3cm") %>% 
+  kableExtra::footnote(
+    alphabet = c(
+    "NOAA OISST: (https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html); ", 
+    "2011--2019: GLOBAL_MULTIYEAR_PHY_001_030 (https://doi.org/10.48670/moi-00021); ", 
+    "2020--2021: GLOBAL_ANALYSIS_FORECAST_PHY_001_024 (https://doi.org/10.48670/moi-00016); ",
+    "ERDDAP/NOAA (https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html); ",
+    "IMOS/AODN (https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set; ",
+    "GEBCO (https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"))
+```
+
+\elandscape
+
 ## Data treatment
@@
 Data wrangling, modelling and visualization were done in R 4.2.0 [@r2022]; the full list of packages can be found in the Supplementary Material.
-Code and the modelling dataset (see Section \ref{sec:modellingdatasets}) are available in **(Zenodo)**.
+Code and the modelling dataset (see Section \ref{sec:modellingdatasets}) are available in [@daudt_data_2023].
 
@@
 Thus, we built separate modelling datasets for each season.
-We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Figure** \@ref(fig:fig1-study-area)**b**), by summing seabirds from the same species.
+We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Fig.** \@ref(fig:fig1-study-area)**b**), by summing seabirds from the same species.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 
-Due to seabird migration and dispersion, seasons could act as 'sampling artifacts' [@foster2017], thus biasing results [@ponti2023].
+Due to seabird migration and dispersion, seasons could act as 'sampling artefacts' [@foster2017], thus biasing results [@ponti2023].
 In light of the difference among seasons in the numbers of species that met our occurrence threshold (see Section \ref{sec:modellingdatasets}), we decided to use RCP models by season and data type (presence-absence and abundance).
@@
 sula* (25.4%) and black-browed albatross *T.
-melanophris* (25%) (**Figure S1**).
-An increase in species richness towards the south of the study region was clear in all but the winter season (**Figure S2**).
-Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Figure S2**).
+melanophris* (25%) (**Supplementary Fig.
+1**).
+An increase in species richness towards the south of the study region was clear in all but the winter season (**Supplementary Fig.
+2**).
+Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Supplementary Fig.
+2**).
 However, in summer, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania.
@@
 
-For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Figure S3**).
+For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Supplementary Figs.
+3, 4**).
 The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (**Table** \@ref(tab:table2)).
-The best models showed a good fit with no apparent deviations from the assumptions (**Figure S4**).
+The best models showed a good fit with no apparent deviations from the assumptions (**Supplementary Figs.
+5, 6**).
 All the best models selected sea surface temperature ('sst') or salinity ('sss'), as essential for characterising water masses.
-The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Figure** \@ref(fig:fig2-partial-plots), **Figure S5**), similar to the characteristics used to to define the EAC.
-Furthermore, most of the best models selected bathymetry ('bat'), suggesting a gradient between 'coastal' and 'oceanic' taxa (**Figure S5**).
+The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Fig.** \@ref(fig:fig2-partial-plots), **Supplementary Figs.
+7, 8**), similar to the characteristics used to to define the EAC.
+Furthermore, most of the best models selected bathymetry ('bat'), suggesting a gradient between 'coastal' and 'oceanic' taxa (**Supplementary Figs.
+7, 8**).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
 
+```{r table2, echo = FALSE, message = FALSE}
+df <- 
+  rbind(
+    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
+    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
+  purrr::map_df(~readr::read_csv(.)) %>% 
+  dplyr::select(- `...1`) %>% 
+  dplyr::mutate(data_type = factor(data_type, 
+                                   levels = c("bernoulli", "neg_bin"),
+                                   labels = c("Bernoulli", "Negative Binomial")),
+                season = factor(season, 
+                                levels = c("summer", "autumn", "winter", "spring"),
+                                labels = c("Summer", "Autumn", "Winter", "Spring")),
+                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
+                bic = round(bic, digits = 2)) %>%
+  dplyr::group_by(data_type) %>% 
+  dplyr::arrange(data_type, season) %>% 
+  as.data.frame(.) %>% 
+  dplyr::ungroup(.)
+
+# Correct a value in a cell
+# Full abundance-autumn model did not fit properly, returning a value of -Inf
+df[df$data_type=='Negative Binomial' & df$season=='Autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"
+
+# A few adjustments for nicer output
+# df$model_specification <- stringr::str_replace_all(df$model_specification, " ", "")
+colnames(df)[3] <- 'Model'
+colnames(df)[4] <- 'Model specification'
+colnames(df)[5] <- 'BIC'
+
+kableExtra::kbl(df, caption = "\\label{table2}Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on Bernoulli models; abundance, based on negative binomial models).
+Refer to Table 1 for the environmental data acronyms", booktabs = TRUE) %>%
+  kableExtra::column_spec(0, bold = TRUE) %>%
+  kableExtra::column_spec(1, bold = TRUE) %>%
+  kableExtra::column_spec(4, width = "11cm") %>% 
+  kableExtra::collapse_rows(columns = c(1:3),
+                            latex_hline = 'custom',
+                            custom_latex_hline = 1:3,
+                            row_group_label_position = 'stack')
+
+rm("df")
+```
+
+```{r fig2-partial-plots, fig.cap = "\\label{fig2-partial-plots}Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
+The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
+All other predictors are held at their mean values.
+Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
+7, 8}.
+Refer to Table 1 for the environmental data acronyms", fig.align = "center", out.width = "100%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
+```
+
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Figure** \@ref(fig:fig3-point-pred)).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Figure** \@ref(fig:fig4-prob-maps), **Figure S6**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
-The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Figure** \@ref(fig:fig3-point-pred), **Figure S6**).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.** \@ref(fig:fig3-point-pred)).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.** \@ref(fig:fig4-prob-maps), **Supplementary Figs.
+9, 10**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Fig.** \@ref(fig:fig3-point-pred), **Supplementary Figs.
+9, 10**).
+
+```{r fig3-point-pred, fig.cap = "\\label{fig3-point-pred}Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+Each grid was assigned to the RCP group with highest membership probability", fig.align = "center", out.width = "110%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig2_point-predictions.png")
+```
+
+\newpage
+
+\blandscape
+
+```{r fig4-prob-maps, fig.cap = "\\label{fig4-prob-maps}Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
+Note that, in case of two RCPs, RCP2 probability equals (1 - RCP1 probability).
+Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
+9, 10}", fig.align = "center", out.width = "110%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig3_probability-maps-spring.png")
+```
+
+\elandscape
+
+\newpage
 
@@
 small albatrosses *Thalassarche* spp., fairy prion *P.
-turtur*) (**Figure** \@ref(fig:fig5-spp-profiles), **Figure S7**).
+turtur*) (**Fig.** \@ref(fig:fig5-spp-profiles), **Supplementary Figs.
+11, 12**).
 The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
 However, the intermediate 'Tasmania/Bass Strait' assemblage showed a mixture of both groups with shy albatross and Australasian gannet *Morus serrator* prominent with high probabilities of occurrence.
+
+\blandscape
+
+```{r fig5-spp-profiles, fig.cap = "\\label{fig5-spp-profiles}Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
+Values are the average and 95\\% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
+Values in (b) were log10-transformed to accommodate the high variation between species.
+Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
+11, 12}", fig.align = "center", out.width = "80%", echo = FALSE}
+
+knitr::include_graphics("../results/Fig5_spp-profiles-spring.png")
+```
+
+\elandscape
+
+\newpage
 
@@
 The summer and autumn models reflected the high abundance of short-tailed shearwater before they depart on their trans-equatorial migration [@carey2014]; the presence of the Australasian gannet was picked up by the autumn and winter models [@ismar2011]; and the white-headed petrel *Pterodroma lessonii* movements during chick-rearing and pre-laying exodus [@taylor2021] were detected in autumn and spring models, respectively.
-Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (**Figure S1**; hence they were removed before modelling).
+Given the geographic location of the southern assemblage, however, we believe sub-Antarctic taxa were not well represented in our models due to their low number of occurrences and relative paucity of survey effort to date (**Supplementary Fig.
+1**; hence they were removed before modelling).
 Additional surveys may reveal the seasonal northward pulse of these taxa into the southern assemblage (e.g.
@@
 In addition, ideally, we would have had more information within each grid cell by season (our sampling units).
-Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Figure S8**).
+Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Supplementary Fig.
+13**).
 
@@
 
-# References
+# Acknowledgements {.unnumbered}
 
-::: {#refs}
-:::
+We thank all observers who assisted in data collection, the crews of RV *Investigator* and RV *Falkor*, and the Australasian Seabird Group (BirdLife Australia) for logistical support.
+The authors would like to thank Dr.
+Skipton N.C.
+Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
+NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
+
+# References {-}
+
+<div id="refs"></div>
+
+\newpage
+
+# Authors’ contribution (CRediT) {.unnumbered}
+
+**Nicholas W.
+Daudt:** Conceptualization; Methodology; Software; Formal analysis; Data Curation; Visualization; Writing - Original Draft.
+**Eric J.
+Woehler:** Conceptualization; Methodology; Investigation; Resources; Data Curation; Writing - Review & Editing; Supervision.
+**Matthew R.
+Schofield:** Conceptualization; Software; Writing - Review & Editing; Supervision.
+**Robert O.
+Smith:** Writing - Review & Editing; Supervision.
+**Leandro Bugoni:** Writing - Review & Editing; Supervision.
+**William J.
+Rayment:** Conceptualization; Writing - Review & Editing; Supervision.
+
+# ORCID {.unnumbered}
+
+Nicholas W. Daudt <https://orcid.org/0000-0001-8211-8771>
+
+Eric J. Woehler <https://orcid.org/0000-0002-1125-0748>
+
+Matthew R. Schofield <https://orcid.org/0000-0003-1481-2766>
+
+Robert O. Smith <https://orcid.org/0000-0001-5931-3843>
+
+Leandro Bugoni <https://orcid.org/0000-0003-0689-7026>
+
+William J. Rayment <https://orcid.org/0000-0001-9558-5109>
 
@@
 
-The gridded data used as the basis for all analyses are available in the archived repository along with the code to reproduce the results.
-The seabird raw data are currently being used for other projects and thus can not be publicly available.
-If you are interested in the seabird raw data, please contact XXX.
-The environmental data used in this study are freely available from the source (see Table 1 in the main text).
+The data that support the findings of this study are openly available in Daudt *et al.* (2023), from Open Science Framework (OSF) at <https://doi.org/10.17605/OSF.IO/N582D>.
+The seabird raw data are currently being used for other projects and thus can not be publicly available---if you are interested in it, please contact Eric J.
+Woehler.
+The environmental data used in this study are freely available in their source: (i) GEBCO Compilation Group provided detailed bathymetric data (<doi:10.5285/a29c5465-b138-234d-e053-6c86abc040b9>); (ii) Eddy kinetic energy was derived from data sourced from Australia’s Integrated Marine Observing System (IMOS), enabled by the National Collaborative Research Infrastructure Strategy (NCRIS); (iii) Sea surface salinity and mixed layer depth were sourced from E.U.
+Copernicus Marine Service Information; (iv) Sea surface temperature was sourced from NOAA OI SST V2 High Resolution Dataset provided by the NOAA PSL, Boulder, Colorado, USA; and (v) Chlorophyll-*a* was sourced from ERDDAP/NOAA, which uses NASA's Aqua-MODIS satellite data.
+See detailed information regarding environmental layers in the main text (Table 1).
 
-\newpage
+# Funding {.unnumbered}
 
-# Tables and Figures {.unnumbered}
+Parks Australia provided funding support to participants for selected voyages.
 
--   **TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources
--   **TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli model; abundance, based on a Negative Binomial model).
-Refer to Table 1 for the environmental data acronyms
--   **FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
-In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
-The South Equatorial Current (SEC) and its branches, the North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red.
-The EAC 'eastern extension' is an eddy field (blue shaded area) often called the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green).
-Figure (a) is adapted after Oke, Roughan, et al.
-(2019); for colour references see the online version
--   **FIGURE 2.** Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
-The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
-All other predictors are held at their mean values.
-Results for other seasons and models can be found in the Supplementary Material (Figure S5).
-Refer to Table 1 for the environmental data acronyms
--   **FIGURE 3.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
-Each grid was assigned to the RCP group with highest membership probability
--   **FIGURE 4.** Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
-The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
-Note that, in case of two RCPs, RCP2 probability equals (1 - RCP1 probability).
-Results for other seasons and models can be found in the Supplementary Material (Figure S6)
--   **FIGURE 5.** Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
-Values are the average and 95% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
-Values in (b) were log10-transformed to accommodate the high variation between species.
-Results for other seasons and models can be found in the Supplementary Material (Figure S7)
+# Conflict of interest disclosure {.unnumbered}
 
-\elinenumbers
-
-\blandscape
-
-```{r table1, echo = FALSE}
-
-data.frame(
-  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
-  # Abbreviation needs the `str_replace_all` because I'm using `escape=F` in `kbl()`
-  Abbreviation = stringr::str_replace_all(c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"), "_", "\\\\_"),
-  Units = c("°Celsius", paste0(parse(text = "\u0394"), "°Celsius"), "PSU", "m", "mg/m${^3}$", "$log_{10}$(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", paste0("mean(", paste0(parse(text = "\u0394"), "°Celsius"), ")/season/10 years"), "m", "degrees (°)", "km"),
-  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
-  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
-  Source = c("Huang et al.
-(2021)${^a}$", "Derived from Huang et al.
-(2021)${^a}$", "E.U.
-Copernicus Marine Service Information${^b}{^c}$", "E.U.
-Copernicus Marine Service Information${^b}{^c}$", "ERDDAP/NOAA${^d}$", "Derived from ERDDAP/NOAA${^d}$", "IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from IMOS/AODN${^e}$", "Derived from Huang et al.
-(2021)${^a}$", "GEBCO${^f}$", "Derived from GEBCO${^f}$", "Calculated in R (see main text)")
-) %>% 
-  kableExtra::kbl(caption = " ", booktabs = TRUE, escape = FALSE)  %>% 
-  kableExtra::row_spec(0, bold = TRUE) %>% 
-  kableExtra::column_spec(1, width = "4cm") %>% 
-  kableExtra::column_spec(6, width = "3cm") %>% 
-  kableExtra::footnote(
-    alphabet = c(
-    "NOAA OISST: (https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html); ", 
-    "2011--2019: GLOBAL_MULTIYEAR_PHY_001_030 (https://doi.org/10.48670/moi-00021); ", 
-    "2020--2021: GLOBAL_ANALYSIS_FORECAST_PHY_001_024 (https://doi.org/10.48670/moi-00016); ",
-    "ERDDAP/NOAA (https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html); ",
-    "IMOS/AODN (https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set; ",
-    "GEBCO (https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"))
-
-### Using {gt} -- just pipe this onto the data.frame
-  # gt::gt() |>
-  # ## SST & SSTgrad
-  # gt::tab_footnote(footnote = 
-  #                    gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
-  #                  locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
-  # ## SSS & MLD
-  # gt::tab_footnote(footnote = 
-  #                    gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
-  #                  locations = gt::cells_body(columns = Source, rows = 3:4)) |>
-  # ## CHL8 & log10-CHL8
-  # gt::tab_footnote(footnote = 
-  #                    gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
-  #                  locations = gt::cells_body(columns = Source, rows = 5:6)) |>
-  # ## EKE and derivations
-  # gt::tab_footnote(footnote = 
-  #                    gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
-  #                  locations = gt::cells_body(columns = Source, rows = 7:9)) |>
-  # ## GEBCO & slope
-  # gt::tab_footnote(footnote = 
-  #                    gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
-  #                  locations = gt::cells_body(columns = Source, rows = 11:12)) |>
-  # ## Improve cols names
-  # gt::cols_label(
-  #   Variable = gt::md("**Variable**"),
-  #   Abbreviation = gt::md("**Abbreviation**"),
-  #   Units = gt::md("**Units**"),
-  #   `Spatial.resolution` = gt::md("**Spatial resolution**"), 
-  #   `Temporal.resolution` = gt::md("**Temporal resolution**"),
-  #   Source = gt::md("**Source**")) |>
-  # ## Table title
-  # gt::tab_header(title = gt::md("Table 1.")) |>
-  # gt::fmt_markdown(columns = gt::starts_with("Var")) #|> #columns = c("Variable", "Source")
-  # # gt::tab_options(table.width = gt::pct(100))
-  # # gt::tab_options(container.width = 300, container.height = 300)
-```
-
-\elandscape
-
-\newpage
-
-```{r table2, echo = FALSE, message = FALSE}
-
-df <- 
-  rbind(
-    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
-    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
-  purrr::map_df(~readr::read_csv(.)) %>% 
-  dplyr::select(- `...1`) %>% 
-  dplyr::mutate(data_type = factor(data_type, 
-                                   levels = c("bernoulli", "neg_bin"),
-                                   labels = c("Bernoulli", "Negative Binomial")),
-                season = factor(season, 
-                                levels = c("summer", "autumn", "winter", "spring"),
-                                labels = c("Summer", "Autumn", "Winter", "Spring")),
-                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
-                bic = round(bic, digits = 2)) %>%
-  dplyr::group_by(data_type) %>% 
-  dplyr::arrange(data_type, season) %>% 
-  as.data.frame(.) %>% 
-  dplyr::ungroup(.)
-
-# Correct a value in a cell
-# Full abundance-autumn model did not fit properly, returning a value of -Inf
-df[df$data_type=='Negative Binomial' & df$season=='Autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"
-
-# A few adjustments for nicer output
-# df$model_specification <- stringr::str_replace_all(df$model_specification, " ", "")
-colnames(df)[3] <- 'Model'
-colnames(df)[4] <- 'Model specification'
-colnames(df)[5] <- 'BIC'
-
-kableExtra::kbl(df, caption = " ", booktabs = TRUE) %>% 
-  kableExtra::column_spec(0, bold = TRUE) %>%
-  kableExtra::column_spec(1, bold = TRUE) %>%
-  kableExtra::column_spec(4, width = "11cm") %>% 
-  kableExtra::collapse_rows(columns = c(1:3),
-                            latex_hline = 'custom',
-                            custom_latex_hline = 1:3,
-                            row_group_label_position = 'stack')
-
-### Using {gt}:
-# A few tweaks to make the table look better
-# df$data_type <- c("Bernoulli", rep("", 7), "Negative Binomial", rep("", 7))
-# df$season <- rep(c("Summer", "", "Autumn", "", "Winter", "", "Spring", ""), 2)
-
-# gt::gt(df)|>
-#   ## Improve cols names
-#   gt::cols_label(
-#     data_type = gt::md("**Data type**"),
-#     season = gt::md("**Season**"),
-#     model = gt::md("**Model**"),
-#     model_specification = gt::md("**Model specification**"), 
-#     bic = gt::md("**BIC**")) |>
-#   ## Table title
-#   gt::tab_header(title = gt::md("Table 2."))
-
-rm("df")
-```
-
-\newpage
-
-\blandscape
-
-```{r fig1-study-area, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
-
-# knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
-
-knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
-```
-
-\elandscape
-
-\newpage
-
-```{r fig2-partial-plots, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
-
-knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
-```
-
-\newpage
-
-```{r fig3-point-pred, fig.cap = " ", fig.align = "center", out.width = "110%", echo = FALSE}
-
-knitr::include_graphics("../results/Fig2_point-predictions.png")
-```
-
-\newpage
-
-\blandscape
-
-```{r fig4-prob-maps, fig.cap = " ", fig.align = "center", out.width = "110%", echo = FALSE}
-
-knitr::include_graphics("../results/Fig3_probability-maps-spring.png")
-```
-
-\newpage
-
-```{r fig5-spp-profiles, fig.cap = " ", fig.align = "center", out.width = "100%", echo = FALSE}
-
-knitr::include_graphics("../results/Fig5_spp-profiles-spring.png")
-```
-
-\elandscape
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

### 2024-03-06 ms_PROOCE/ms_PROOCE_revised-markup.Rmd
update README, update scripts-workflow figure, updated references inlcuding Daudt et al 2024 and 'corrected' refs for R pckgs, plus all source RMD files

--- 2024-03-06 ms_PROOCE/ms_PROOCE.Rmd
@@
 journal: "Progress in Oceanography"
-date: "2023-11-02"
+date: "`r Sys.Date()`"
 classoption: preprint, 3p, authoryear
@@
 
-* The East Australian Current (EAC) is experiencing an ongoing poleward intensification
-* We described seabird assemblages based on the Region of Common Profiles (RCP) method
-* Assemblages of seabird species were correlated with the EAC
-* Our analysis can be replicated to assess possible changes in seabird distribution
-* The RCP method works well at the macro spatial scale for highly mobile animals
+-   \textcolor{red}{\sout{The East Australian Current (EAC) is experiencing an ongoing poleward intensification}}
+
+-   We described seabird assemblages based on the Region of Common Profiles (RCP) method
+
+-   Assemblages of seabird species were correlated with the \textcolor{red}{East Australian Current}
+
+-   Our analysis can be replicated to assess possible changes in seabird distribution
+
+-   The RCP method works well at the macro spatial scale for highly mobile animals
 
@@
 water masses) to fine (e.g.
-fronts) [@dunlop1988; @ribic1997; @quillfeldt2015].
+fronts) [@dunlop1988; @ribic1997; @quillfeldt2015; @gall2022; @serratosa2020].
 Despite extensive surveys [e.g.
@@
 @sydeman2012].
-Surprisingly though, seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic [@grémillet2009; @sydeman2012].
+Surprisingly though, \textcolor{red}{particularly in the Australasian region,} seabird distribution is under-represented in climate change studies, despite being highlighted as a priority topic [@grémillet2009; @sydeman2012].
 
@@
 When the SEC's North Caledonian Jet meets the Queensland Plateau (\~18°S), it splits into a north-flowing branch towards the Gulf of Papua and a south-flowing branch, which marks the start of the EAC [@kessler2013; @ganachaud2014].
-The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.** \@ref(fig:fig1-study-area)**a**; @kessler2013; @ganachaud2014).
+The South Caledonian Jet completes the EAC system at \~22°S after encountering the shelf break in the southern portion of the Great Barrier Reef (**Fig.** \@ref(fig:fig-study-area)**a**; @kessler2013; @ganachaud2014).
 The EAC interacts with the bathymetry of the seafloor, giving rise to oceanographic features and processes, such as smaller currents, upwellings, and persistent dynamic eddies [@oke2000; @ridgway2003a; @oke2019].
 The EAC flows polewards and is primarily associated with the shelf break, but at \~32°S it detaches from the shelf and spawns an area of eddy generation that dissipates throughout the western Tasman Sea [@mata2006; @oke2019], the 'EAC eastern extension', often termed the 'Tasman Front' [@oke2019a].
-Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.** \@ref(fig:fig1-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
+Although most EAC energy is converted into eddies, the EAC still runs southwards following the shelf break towards Tasmania (the 'EAC southern extension'; **Fig.** \@ref(fig:fig-study-area)**a**), carrying eddy-like features as a transient current [@oke2019].
 
@@
 
-```{r fig1-study-area, fig.cap = "\\label{fig1-study-area} Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
+```{r fig-study-area, fig.cap = "\\label{fig-study-area} Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b).
 In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue.
@@
 
-Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.** \@ref(fig:fig1-study-area)**b**; **Supplementary Table 1**).
-One or more observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway.
-The species, the numbers of birds, and their behaviours were logged in real time into a ship-based web server.
-Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.
+Seabird data were gathered aboard *RV Investigator* and *RV Falkor* on 15 voyages, between 2016/17--2020/21, spanning over \~37° of latitude (**Fig.** \@ref(fig:fig-study-area)**b**; **Supplementary Table 1**).
+\textcolor{red}{Two} observers continuously recorded seabirds within a 300 m forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours while the ship was underway\textcolor{red}{, following a strip-transect protocol}.
+The species, the numbers of birds, and their behaviours \textcolor{red}{(i.e.
+flying, feeding, sitting on the water)} were logged in real time into a ship-based web server.
+\textcolor{red}{Seabird surveys were halted when sea conditions exceeded 6 m wave height and 40 kn of wind or visibility was under 500 m.} Environmental parameters for atmospheric and oceanographic conditions were logged continuously by automated-ship based sensors, with seabird observations supplementing relevant environmental conditions such as visibility or associations with marine mammals.
 Individual records received a date/time/geographic coordinate label (multi-species records have the same label but are logged as separate entries for each species).
@@
   Abbreviation = stringr::str_replace_all(c("sst", "sst_grad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke_mean", "clim_eke_sd", "clim_sst_grad", "bat", "slope", "dist_coast"), "_", "\\\\_"),
-  Units = c("°Celsius", paste0(parse(text = "\u0394"), "°Celsius"), "PSU", "m", "mg/m${^3}$", "$log_{10}$(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", paste0("mean(", paste0(parse(text = "\u0394"), "°Celsius"), ")/season/10 years"), "m", "degrees (°)", "km"),
+  Units = c("°Celsius", paste0(parse(text = "\u0394"), "°Celsius"), "PSU", "m", "mg/m${^3}$", "$log_{10}$(chl8)", "m/s", "mean(m/s)/season/11 years", "sd(m/s)/season/11 years", paste0("mean(", paste0(parse(text = "\u0394"), "°Celsius"), ")/season/11 years"), "m", "degrees (°)", "km"),
   `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
@@
 Thus, we built separate modelling datasets for each season.
-We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Fig.** \@ref(fig:fig1-study-area)**b**), by summing seabirds from the same species.
+\textcolor{red}{Seasons were defined as summer (Dec--Feb), autumn (Mar--May), winter (Jun--Aug), and spring (Sep--Nov).} We aggregated seabird data by 1° latitude $\times$ 1° longitude grid cells (*c.* 110 $\times$ 110 km) and season (**Fig.** \@ref(fig:fig-study-area)**b**), by summing seabirds from the same species \textcolor{red}{(flying, feeding, and sitting on the water)}.
 To test if the presence-absence and abundance (raw counts) data would result in similar assemblage patterns, we built two species matrices, where the presence-absence matrix transformed values \>1 to 1.
@@
 RCPs are a multivariate adaptation of a mixture-of-experts model; as such, the mixing proportions (assemblage structure) vary according to covariates (predictors) [@foster2013].
-RCPs can be directly interpreted as assemblages; the assemblage structure is given by the species profiles.
+RCPs can be directly interpreted as assemblages; the assemblage structure is given by the species profiles \textcolor{red}{(i.e.
+the proportion of each species within the assemblage given by their probability of occurrence or predicted abundance)}.
 
@@
 1**).
-An increase in species richness towards the south of the study region was clear in all but the winter season (**Supplementary Fig.
-2**).
-Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Supplementary Fig.
-2**).
+An increase in species richness towards the south of the study region was clear in all but the winter season (**Fig.** \@ref(fig:fig-spprich-nbirds)).
+Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' and/or the presence of Southern Ocean species (**Fig.** \@ref(fig:fig-spprich-nbirds)).
 However, in summer, one cell in the Coral Sea had counts of the same order of magnitude as off Tasmania.
+
+```{r fig-spprich-nbirds, fig.cap = "\\label{fig-spprich-nbirds}Species richness and the total number of seabirds counted off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021, by grid/season, after data aggregation (see Methods).
+Dots represent the grid centroids.", fig.align = "center", out.width = "100%", echo = FALSE}
+
+knitr::include_graphics("../results/FigS2_map-effort-spp-richness-and-total-birds-seasons.png")
+```
 
@@
 
+\textcolor{red}{In this section, we will present only the main results for the spring season when delving into specifics about each model and its results.
+The results of spring models summarise the key findings of our analysis, but we encourage the reader to look at the Supplementary Material for other seasons.}
+
 For all models, two RCPs was the optimal number of seabird groups, except for the presence-absence-autumn model, which indicated three RCPs as the best fit (**Supplementary Figs.
-3, 4**).
+2, 3**).
 The best fitting models generally had one or two covariates selected, although the abundance-autumn model had five covariates (**Table** \@ref(tab:table2)).
 The best models showed a good fit with no apparent deviations from the assumptions (**Supplementary Figs.
-5, 6**).
+4, 5**).
 All the best models selected sea surface temperature ('sst') or salinity ('sss'), as essential for characterising water masses.
-The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Fig.** \@ref(fig:fig2-partial-plots), **Supplementary Figs.
-7, 8**), similar to the characteristics used to to define the EAC.
+The thresholds for defining RCP membership were 18--21°C for 'sst' and 35.1 PSU for 'sss' (**Fig.** \@ref(fig:fig-partial-plots), **Supplementary Figs.
+6, 7**), similar to the characteristics used to to define the EAC.
 Furthermore, most of the best models selected bathymetry ('bat'), suggesting a gradient between 'coastal' and 'oceanic' taxa (**Supplementary Figs.
-7, 8**).
+6, 7**).
 Of particular note, proxies for persistent oceanographic features that may aggregate food resources (the 'climatic' variables) for seabirds were selected only in the autumn and spring abundance models.
@@
 
-```{r fig2-partial-plots, fig.cap = "\\label{fig2-partial-plots}Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
+```{r fig-partial-plots, fig.cap = "\\label{fig-partial-plots}Partial plots for the retained covariates in the best spring (a) presence-absence and (b) abundance models.
 The plot shows the fitted probability of belonging to a Region of Common Profiles (RCP) against the environmental value.
@@
 Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
-7, 8}.
+6, 7}.
 Refer to Table 1 for the environmental data acronyms", fig.align = "center", out.width = "100%", echo = FALSE}
@@
 Grid point-predictions are the assignment of the RCP with the highest probability value for a given cell.
-When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.** \@ref(fig:fig3-point-pred)).
-These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.** \@ref(fig:fig4-prob-maps), **Supplementary Figs.
-9, 10**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
-The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Fig.** \@ref(fig:fig3-point-pred), **Supplementary Figs.
-9, 10**).
+When predicting RCP onto space, patterns of point-predictions suggest a biogeographic transition in seabird species around \~34°S, with both presence-absence and abundance models pointing to similar results (**Fig.** \@ref(fig:fig-point-pred)).
+These patterns are further reinforced when looking at the mean probability along with the lower and upper confidence intervals of predictions (**Fig.** \@ref(fig:fig-prob-maps), **Supplementary Figs.
+8, 9**), showing that around \~34°S, the 'northern' RCP tended to have higher probabilities in coastal regions, whereas the 'southern' RCP spread northward offshore.
+The autumn presence-absence model (three RCPs) had a similar broad-scale pattern, although the southern RCP is farther south, and a strong 'Tasmania/Bass Strait' assemblage appears (**Fig.** \@ref(fig:fig-point-pred), **Supplementary Figs.
+8, 9**).
 
-```{r fig3-point-pred, fig.cap = "\\label{fig3-point-pred}Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
+```{r fig-point-pred, fig.cap = "\\label{fig-point-pred}Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia.
 Each grid was assigned to the RCP group with highest membership probability", fig.align = "center", out.width = "110%", echo = FALSE}
@@
 
-```{r fig4-prob-maps, fig.cap = "\\label{fig4-prob-maps}Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
+```{r fig-prob-maps, fig.cap = "\\label{fig-prob-maps}Predicted probability membership of each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models.
 The central column, 'mean', corresponds to the point prediction and Bayesian boot-strapped, lower and upper confidence intervals (CI), on its sides.
@@
 Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
-9, 10}", fig.align = "center", out.width = "110%", echo = FALSE}
+8, 9}", fig.align = "center", out.width = "110%", echo = FALSE}
 
@@
 small albatrosses *Thalassarche* spp., fairy prion *P.
-turtur*) (**Fig.** \@ref(fig:fig5-spp-profiles), **Supplementary Figs.
-11, 12**).
+turtur*) (**Fig.** \@ref(fig:fig-spp-profiles), **Supplementary Figs.
+10, 11**).
 The autumn presence-absence model, which had three RCPs, also showed the characteristic taxa representing the 'northern' and 'southern' RCPs.
@@
 
-```{r fig5-spp-profiles, fig.cap = "\\label{fig5-spp-profiles}Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
+```{r fig-spp-profiles, fig.cap = "\\label{fig-spp-profiles}Species profiles of seabird assemblages (Region of Common Profile; RCP) off eastern Australia in spring.
 Values are the average and 95\\% confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps.
@@
 Results for other seasons and models can be found in the \\textbf{Supplementary Figs.
-11, 12}", fig.align = "center", out.width = "80%", echo = FALSE}
+10, 11}.
+Note the species on the Y-axis are ordered alphabetically, from bottom up.", fig.align = "center", out.width = "80%", echo = FALSE}
 
@@
 
-The relationships between seabird assemblages and water masses has been known for a long time [@pocklington1979; @elphick1993; @ribic1997].
+The relationships between seabird assemblages and water masses has been \textcolor{red}{well established} [@pocklington1979; @elphick1993; @ribic1997; @gall2022].
 Thus, it is not surprising that two seabird assemblages were identified off eastern Australia.
-However, the spatial consistency of these patterns throughout seasons is surprising and remarkable.
+However, the spatial consistency of these patterns throughout seasons is surprising and remarkable \textcolor{red}{(but see Section \ref{sec:limitations})}.
 Resident species seem to characterise these assemblages.
@@
 As such, capturing the distribution of highly mobile species, and their relationships with environmental parameters is not a trivial task, and much harder than for sessile organisms.
+\textcolor{red}{Among marine megafauna, seabirds are relatively easy to detect visually, enabling suitable sample sizes to be gathered over large areas} [@ballance2007]\textcolor{red}{.
+Thus, seabirds can be used as models to study the distribution of highly mobile animals in relation to environmental characteristics in the marine environment.}
 
-While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009], they fail to perceive how species assemblages interact with the environment and each other.
+While remote tracking studies are advancing our understanding on how individual seabirds use their environment at fine scales [@tremblay2009; @trevail2023], they fail to perceive how species assemblages interact with the environment and each other.
 Therefore, using at-sea counts of seabirds is a well established method to capture these relationships [@ainley2012; @woehler2003; @hyrenbach2007].
@@
 If combined with ecological knowledge of a particular group, the information contained in both presence-absence and abundance data could be leveraged to better interpret the resulting patterns.
-For example, seabirds will likely occur in specific water masses or currents [@hunt1987; @ribic1997], whereas their abundance is related to coarse- to fine-scale processes that enhance prey availability [@hunt1987; @weimerskirch2007].
-Future studies assessing bioregions of highly mobile animals could describe the compositions of assemblages using smaller resolution grid cells and presence-absence models.
+
+For example, seabirds will likely occur in specific water masses or currents [@hunt1987; @ribic1997; @commins2014; @gall2022], whereas their abundance is related to coarse- to fine-scale processes that enhance prey availability [@hunt1987; @weimerskirch2007; @depascalis2021].
+\textcolor{red}{Although we have not approached the data in a multi-scale analysis, our results partially support this.
+Presence-absence models retained static variables and oceanographic variables that define water masses (i.e.
+SST and salinity).
+In addition to those, the autumn and spring abundance models also retained variables associated with persistent oceanographic features (e.g.
+climatic variables related to eddies and fronts), as well as mixed layer depth and chlorophyll in the autumn model.
+Eddies, fronts, high chlorophyll and shallower mixed layer depths are all related to increased seabird foraging success} [@weimerskirch2007; @scales2014; @depascalis2021]\textcolor{red}{, and therefore likely to influence seabird numbers.
+Thus,} future studies assessing bioregions of highly mobile animals could describe the compositions of assemblages using smaller resolution grid cells and presence-absence models.
 Then, abundance models using higher resolution grid cells could improve our understanding of how these animals relate to their environment and possibly reveal multi-species aggregation areas (therefore 'new bioregions') likely nested within the bioregions described at large scales.
-Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which we suspect do not exist at present.
+Such a hierarchical approach would require substantial concurrent observational and environmental data sets, which\textcolor{red}{, to our knowledge, do not exist for our study area}.
 
@@
 Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species in most of them (**Supplementary Fig.
-13**).
+12**).
+
+\textcolor{red}{These limitations are illustrated in the results for the summer and autumn seasons.
+Considering the strong support for a southern and a northern assemblage, we note that both summer and autumn lack samples in representative southern and northern (summer) and northern (autumn) areas.
+Albeit the sample coverage values were overall high, compared to winter and spring, summer and autumn had lower values (in particular, for the southern assemblage 'RCP1' in summer and for the northern assemblage 'RCP3' in autumn; \textbf{Supplementary Fig.
+12}).
+Further, based on point predictions, higher probabilities of the southern assemblage in the northern area in summer (both presence-absence and abundance models) and of the northern assemblage in the southern area in autumn (abundance model), were assigned to a few grids (\textbf{Fig.} 4).
+The latter reinforces the possible weakness of these models to detect stronger relationships between the identified assemblages and the predictors.
+We also acknowledge the sharp spatial change in probabilities between assemblages in summer models (i.e.
+without a proper 'transition zone'), which may be related to the previous point.
+However, both summer and autumn models detected higher probabilities of expected taxa in their species profiles, suggesting that the make-up of the assemblage was well distinguished by the models.
+In addition, these models also seem to have captured the appropriate number of assemblages, agreeing with other models/seasons.
+Despite their limitations, we believe these models agree with the big picture of our results, therefore being a valid resource as a baseline.}
 
@@
 The models point toward a consistent pattern of a 'northern' and a 'southern' seabird assemblage, where the spatial division is attributable to the eastward separation of the EAC from the coast, and that SST seems to be the key variable discriminating these groups.
+\textcolor{red}{The mechanistic links between SST and seabird assemblages deserve more studies but may be due to the distribution of seabird prey, which at the biogeographic level seems to correlate with our findings} [@sutton2017a].
 The EAC has penetrated poleward over the past half-century [@li2022; @hill2008], a trend which is anticipated to continue into the future [@yang2016], and extreme weather events such as marine heatwaves to happen more often [@oliver2018].
@@
 Given that eastern Australia is a biodiversity and numerical hotspot for seabirds, it is imperative to establish a monitoring program allowing the assemblages and species profiles presented here to be used as the baseline for understanding possible effects of a warming ocean on seabird distribution.
-In particular, the transition zone between assemblages [@morrone2023biogeographical], associated with the eastern extension of the EAC, may be fundamental to effectively monitoring strengthening of the EAC and the subsequent impact on marine biota.
+\textcolor{red}{In particular, the transition zone between assemblages might be pushed southwards along with the expected trend of the EAC.
+The transition zone between assemblages} [@morrone2023biogeographical], associated with the eastern extension of the EAC, may be fundamental to effectively monitoring the strengthening of the EAC and the subsequent impact on marine biota.
 
@@
 Woolley, `{ecomix}` developer, who kindly helped NWD build preliminary models and gave advice in several stages of model construction.
-NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
+\textcolor{red}{NWD thanks the Department of Marine Science (University of Otago) for supporting a PhD writing retreat, during which most of this manuscript was drafted.
+Thanks also to two anonymous reviewers whose comments helped improve our manuscript.} NWD is supported by a Doctoral scholarship from the University of Otago; LB is a research fellow from CNPq (310145/2022-8).
 
-# References {-}
+# References {.unnumbered}
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-# Authors’ contribution (CRediT) {.unnumbered}
+# Authors' contribution (CRediT) {.unnumbered}
 
@@
 Woehler.
-The environmental data used in this study are freely available in their source: (i) GEBCO Compilation Group provided detailed bathymetric data (<doi:10.5285/a29c5465-b138-234d-e053-6c86abc040b9>); (ii) Eddy kinetic energy was derived from data sourced from Australia’s Integrated Marine Observing System (IMOS), enabled by the National Collaborative Research Infrastructure Strategy (NCRIS); (iii) Sea surface salinity and mixed layer depth were sourced from E.U.
+The environmental data used in this study are freely available in their source: (i) GEBCO Compilation Group provided detailed bathymetric data (<doi:10.5285/a29c5465-b138-234d-e053-6c86abc040b9>); (ii) Eddy kinetic energy was derived from data sourced from Australia's Integrated Marine Observing System (IMOS), enabled by the National Collaborative Research Infrastructure Strategy (NCRIS); (iii) Sea surface salinity and mixed layer depth were sourced from E.U.
 Copernicus Marine Service Information; (iv) Sea surface temperature was sourced from NOAA OI SST V2 High Resolution Dataset provided by the NOAA PSL, Boulder, Colorado, USA; and (v) Chlorophyll-*a* was sourced from ERDDAP/NOAA, which uses NASA's Aqua-MODIS satellite data.

### 2024-03-06 ms_PROOCE/ms_PROOCE_supplementary-material.Rmd
update README, update scripts-workflow figure, updated references inlcuding Daudt et al 2024 and 'corrected' refs for R pckgs, plus all source RMD files

---
title: "Supplementary Material"
subtitle: "**Seabird assemblages are linked to the major western boundary current off eastern Australia**"
author: "Daudt, N.W.; Woehler, E.J.; Schofield, M.R.; Smith, R.O.; Bugoni, L.; Rayment, W.J."
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/elsevier-harvard.csl
classoption: a4paper
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include = FALSE, echo = FALSE}
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(gt)
```

This document is a supplementary file for @daudt2024. Briefly, here you will find:

-   **List of packages** (and version) used for data wrangling, visualisation and analyses, and their full references;
-   **Table 1**: Summary of sampling effort by voyage;
-   **Fig. 1**: Number of occurrences, frequency of occurrence (FO), and numeric frequency (NF) for each recorded species by season;
-   **Figs. 2, 3**: Plots for choosing best RCP group number (`multifit`);
-   **Figs. 4, 5**: Residual plots from the best fitted models;
-   **Figs. 6, 7**: Partial plots for covariates;
-   **Figs. 8, 9**: Probability maps from seasonal predictions;
-   **Figs. 10, 11**: Species profiles for each season;
-   **Fig. 12**: Species-richness and sample-coverage curves.

\newpage

# List of packages

We used the following packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021], and the ones referenced in the main text.

The code is archived in an Open Science Framework repository in [@daudt_data_2023], where you can find a detailed walk-through.

# References {.unnumbered}

::: {#refs}
:::

\newpage

\blandscape

```{r table-S1, echo = FALSE}
read.csv("../results/tableS1-effort-summary.csv")  %>% 
  dplyr::select(-X) %>% 
  gt::gt() %>% 
  # Tidy-up column labels
  gt::cols_label(
    voyage = gt::md("**Voyage**"),
    date_start = gt::md("**Date start**"),
    date_end = gt::md("**Date end**"),
    lat_range = gt::md("**Latitudinal range**"), 
    lon_range = gt::md("**Longitudinal range**"),
    no_records = gt::md("**No. of records**"),
    no_birds = gt::md("**No. of birds**"),
    no_spp = gt::md("**No. of species**")) %>% 
  # Table title
  gt::tab_header(title = gt::md("Table 1. Summary of seabird sampling effort by voyage, off eastern Australia, during Australasian Seabird Group's ship-based surveys between 2016--2021. Start and finish dates and geographic ranges of each voyage, including the number of seabird records and the number of individuals and species recorded"))
```

\newpage

![Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021. The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text). Species are ordered from the largest to the lowest total number of occurrences.](../results/FigS1_spp-nOCC-FO-NF-seasons.png){width="180%"}

\elandscape

\newpage

![Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_1_multifit-Bernoulli.png){width="100%"}

\newpage

![Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_2_multifit-NegBin.png){width="100%"}

\newpage

![Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png){width="100%"}

\newpage

![Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png){width="100%"}

\newpage

![Partial plots for the retained covariates in the best seasonal models based on presence-absence data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_1_partial-plots-Bernoulli.png){width="90%"}

\newpage

\blandscape

![Partial plots for the retained covariates in the best seasonal models based on abundance data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_2_partial-plots-NegBin.png){width="150%"}

\newpage

![Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png){width="165%"}

\newpage

![Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png){width="165%"}

\newpage

![Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model. Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps. Note the species on the Y-axis are ordered alphabetically, from bottom up.](../results/FigS7_1_spp-profiles-Bernoulli.png){width="145%"}

\newpage

![Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model. Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps. Values were log10-transformed to accommodate the high variation between species. Note the species on the Y-axis are ordered alphabetically, from bottom up.](../results/FigS7_2_spp-profiles-NegBin.png){width="145%"}

\elandscape

\newpage

![Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png){width="90%"}

\newpage
